---
title: "04_O2_Statistical_Analysis"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : **2 Objective**, Statistical Analysis
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Statistical Analysis
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(purrr) # reduce()
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gt)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
library(chromote)
library(car)
library(mediation)
library(shiny)
library(shinydashboard)
library(data.table)
library(DT)
library(mice)
library(dagitty)
library(ggdag)
library(gridExtra)
library(grid)
library(dotwhisker)
library(pROC)
library(pscl)
library(ResourceSelection)
library(car)
library(writexl)
```

# 0. Import the imputed data 

```{r}
imputed_data_O2 <- read.csv("imputed_data_O2_V2.csv") %>%
  dplyr::select(-X, -ends_with("_participated"),
                -OMA, -SPE, -ADENO,
                -comorbidite_bin,
                -delai_diag_cat
                )

#----------------------
#  Imputed Dataframe
#----------------------
imputed_data_O2 <- imputed_data_O2 %>%
  mutate(across(where(is.character), as.factor))%>%
  mutate(
    live_birth_bin_atleast_once_all = factor(live_birth_bin_atleast_once_all, 
                                             levels = c("Non", "Oui")),
    
    diplome_ini2 = factor(
      diplome_ini2,
      levels = c("<=Baccalauréat/Aucun/Autre", "Bac+2", ">=Bac+3"),
      ordered = FALSE
    ),
    
    situation_financiere   = factor(
      situation_financiere,
      levels = c(
        "Balanced",
        "Difficult/In debt",
        "Comfortable/Very comfortable"
      ),
      ordered = FALSE
    ),
    
    
    endo_ade_categorie = factor(
      endo_ade_categorie,
      levels = c("Only endometriosis", "Only adenomyosis", "Both")
    ),
    
    type_endo =
      factor(
        type_endo,
        levels = c(
          "DIE only",
          "SPE only",
          "OMA only",
          "ADENO only",
          "Multiple types",
          "Aucun mentionné/nsp"
        )
      ),
    
    first_preg_age_cat = factor(
      first_preg_age_cat,
      levels = c("<30", "30-34", ">=34"),
      ordered = FALSE
    ),
    
    age_symptome_cat = factor(
      age_symptome_cat,
      levels = c(
        "<= 13",
        "(13;15]",
        "(15;22]",
        "22+",
        "No symptoms"
      ),
      ordered = FALSE
    ))
```


# I. Univariable analysis adjusted for age 
## 1. Results Table 

```{r}
#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables 
outcome_variable <- "live_birth_bin_atleast_once_all"
adjustment_variable <- "age"

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure outcome variable is suitable for logistic regression (factor or 0/1 numeric)
if (!is.factor(imputed_data_O2[[outcome_variable]]) && !all(imputed_data_O2[[outcome_variable]] %in% c(0, 1, NA))) {
  stop(paste("Outcome variable '", outcome_variable, "' must be a factor or numeric with values 0/1."))
  }

# Convert outcome to factor if it's 0/1 numeric, as tbl_summary needs factor for 'by'
if (is.numeric(imputed_data_O2[[outcome_variable]])) {
    imputed_data_O2 <- imputed_data_O2 %>%
        mutate(!!sym(outcome_variable) := factor(!!sym(outcome_variable)))
    print(paste("INFO: Converted numeric outcome variable '",
                outcome_variable, "' to factor."))
    }

# Ensure adjustment variable exists
if (!all(adjustment_variable %in% names(imputed_data_O2))) {
  stop(paste("Some adjustment variables not found in the dataframe:", 
             paste(setdiff(adjustment_variable, names(imputed_data_O2)), collapse = ", ")))
}
# if (!adjustment_variable %in% names(imputed_data_O2)) {
#   stop(paste("Adjustment variable '", adjustment_variable, "' not found in the dataframe."))
#   }

missing_vars <- setdiff(adjustment_variable, names(imputed_data_O2))
if (length(missing_vars) > 0) {
  stop(paste("Adjustment variable(s) not found in the dataframe:", paste(missing_vars, collapse = ", ")))
  }

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(imputed_data_O2), 
                               c(outcome_variable, adjustment_variable))

if (length(predictor_variables) == 0) {
    stop("No predictor variables found after excluding outcome and adjustment variables.")
  }

print(paste("Outcome:", outcome_variable))
print(paste("Adjustment:", paste(adjustment_variable, collapse=", ")))
print(paste("Predictors:", paste(predictor_variables, collapse=", ")))


#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- imputed_data_O2 %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>% 
  tbl_summary(
    by = all_of(outcome_variable),           # Stratify by the outcome
    statistic = list(                         # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",                           
    digits = list(all_continuous() ~ 1, 
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%     # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>% 
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted Multivariate Logistic Regression Table
#-------------------------------------------------------------
# This function fits glm(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_glm <- function(predictor, outcome_var, adjust_var, data) {
  # Use backticks for safety with non-standard names
  adjust_str <- paste(sprintf("`%s`", adjustment_variable), collapse = " + ")
    # adjust_vars는 벡터
  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)
  
  # Fit the model
  model <- tryCatch(
    glm(as.formula(formula_str), data = data, family = binomial(link = "logit")),
    error = function(e) {
      warning(paste("Could not fit model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )
  
  # If model failed, return NULL
  if (is.null(model)) return(NULL)
  
  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int=TRUE) %>%
    # Keep rows where 'variable' matches the predictor name
    # This correctly handles factor variables which have multiple rows
    modify_table_body(~ dplyr::filter(.x, variable == predictor)) 
  }

# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_glm(.x, outcome_variable,
                                            adjustment_variable,
                                            imputed_data_O2))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables, 
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
            )
          )
      ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted OR (95% CI)**"
    )
  print("Step 2: Adjusted Odds Ratios Table Created.")
} else {
  warning("No successful models were fitted for adjusted ORs.", 
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
adjust_str <- paste(adjustment_variable, collapse = " + ")

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      "**Age-Adjusted Model**")
    ) %>%
      modify_caption(
        paste0(
          "**Multivariate Logistic Regression Analysis for '",
          outcome_variable, 
          "' (Adjusted for '", adjust_str, "')**"))

    print("Step 3: Merged Table Created.")
    
    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '", 
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results 
#------------------------------------------------------------
table_res2_final <- as_tibble(tbl_final) %>% 
  dplyr::select(1,2,3,7,6) 

table_res2_final %>% View

write_xlsx(table_res2_final, "Table_O2_univariable_adjusted_final.xlsx")
```

## 2. Preselection of variables with p-value < 0.2
[1] "statut_marital_ini_cat"     "situation_financiere"      
[3] "endo_ade_categorie"         "age_symptome_cat"          
[5] "complication_operation_mod" "first_preg_age_cat" 

```{r}
ls_var_uni_results <- tbl_final$table_body %>% 
  filter(p.value_2<=0.2) %>%
  distinct(variable) %>% pull()
ls_var_uni_results
```



3. P-trend table 
ordinal variables -> as.numeric :  p-trend ? 
- if the modality "Not_answered" exists,"Not_answered" to NA and then check the p trend 

Question : Differences between results : 
- ordinal variable to numeric variable VS glm(outcome ~ ordinal exposure variable + age, data, family=binomial) 
- .L : linear tendance 

```{r}
# imputed_data_O2_2 <- imputed_data_O2 %>%
#   mutate(
#     diplome_ini2 = factor(
#       diplome_ini2,
#       levels= c("<=Baccalauréat/Aucun/Autre", "Bac+2", ">=Bac+3"),
#       ordered = TRUE
#     ) %>% 
#       as.numeric(),
#     
#     situation_financiere = factor(
#       situation_financiere,
#       levels = c(
#         "Difficult/In debt",
#         "Balanced",
#         "Comfortable/Very comfortable"
#         ),
#       ordered = TRUE
#     ) %>% 
#       as.numeric(),
#     
#     first_preg_age_cat = factor(
#       first_preg_age_cat, 
#       levels= c("<30","30-34",">=34"), 
#       ordered = TRUE) %>% 
#       as.numeric(),
#     
#   age_symptome_cat = factor(
#     age_symptome_cat,
#     levels = c("No symptoms", "<= 13", "(13;15]", "(15;22]", "22+"),
#     ordered = FALSE
#     )%>%
#     as.numeric(),
# 
#   age_firstOP_endoadeno_cat = factor(
#     age_firstOP_endoadeno_cat,
#     levels = c("No operation","<27", "27-31", "31>="),
#     ordered = TRUE
#     )%>%
#     as.numeric(),
#   
#   nb_mal_sansEndoAde_cat= factor(
#     nb_mal_sansEndoAde_cat,
#     levels= c("0","1","2 ou plus"),
#     ordered = TRUE
#   ) %>% 
#     as.numeric(),
#   
#   delai_diag_cat = factor(
#     delai_diag_cat,
#     levels = c("No symptoms","<= 5", "(5;10]", "(10;15]", "15+"),
#     ordered = FALSE
#     )%>% 
#     as.numeric()
#   )
# 
# tbl_descriptive2 <- imputed_data_O2_2 %>%
#   dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>% 
#   tbl_summary(
#     by = all_of(outcome_variable),           
#     statistic = list(
#       all_continuous() ~ "{mean} ({sd})",
#       all_categorical() ~ "{n} ({p}%)"
#     ),
#     missing = "no",                           
#     digits = list(all_continuous() ~ 1, 
#                   all_categorical() ~ c(0, 1))
#   ) %>%
#   modify_header(label ~ "**Characteristic**") %>% 
#   bold_labels()
# 
# list_of_adj_tables2 <- 
#   map(predictor_variables,
#       ~fit_adjusted_glm(.x,outcome_variable,
#                         adjustment_variable,
#                         imputed_data_O2_2))
# 
# # Remove any NULL results (from failed models)
# list_of_adj_tables2 <-
#   list_of_adj_tables2[!sapply(list_of_adj_tables2,is.null)]
# 
# # Stack the individual predictor tables into one
# if (length(list_of_adj_tables2) > 0) {
#   tbl_adjusted_ors2 <- tbl_stack(list_of_adj_tables2, 
#                                 group_header = NULL) %>%
#     modify_table_body(
#       ~ .x %>%
#         mutate(
#           est_ci = if_else(
#             !is.na(estimate),
#             sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
#             NA_character_
#             )
#           )
#       ) %>%
#     modify_header(
#       label ~"**Characteristic**",
#       est_ci ~"**Adjusted OR (95% CI)**"
#     )
#   print("Step 2: Adjusted Odds Ratios Table Created.")
# } else {
#   warning("No successful models were fitted for adjusted ORs.", 
#           call. = FALSE)
#   tbl_adjusted_ors2 <- NULL
# }
# 
# adjust_str <- paste(adjustment_variable, collapse = " + ")
# 
# if (!is.null(tbl_adjusted_ors2)) {
#     tbl_final2 <- tbl_merge(
#       tbls = list(tbl_descriptive2, tbl_adjusted_ors2),
#       tab_spanner = c("**Descriptive Statistics by Outcome**",
#                       "**Age-Adjusted Model**")
#     ) %>%
#       modify_caption(
#         paste0(
#           "**Multivariate Logistic Regression Analysis for '",
#           outcome_variable, 
#           "' (Adjusted for '", adjust_str, "')**"))
# 
#     print("Step 3: Merged Table Created.")
#     
#     # --- Display the final table ---
#     tbl_final2
#     } else {
#     print("Final merged table could not be created as the adjusted OR table is empty.")
#     print("Displaying only the descriptive statistics table:")
#     tbl_descriptive2 %>%
#         modify_caption(
#           paste0(
#             "**Descriptive Statistics for Predictors of '", 
#             outcome_variable,"'**"))
#     }
# 
# tbl_final2

```


**Results** 
- statut_marital_ini : En couple (2.02(1.28,3.15)) , Mariee(2.61(1.57,4.32)) -> CI large
statut_marital_ini_cat :  Single/divorced (0.50 (0.33, 0.76))
- type_endo : OMA only (2.10(1.04,4.25)) -> CI large 
- first_preg_age_cat : >=34 (0.48(0.3,0.76))

# II. Multivariate logistic regression

## 1. DAG
```{r}
# DAG : 
library(dagitty)
library(ggdag)
dag_code <- "dag {

  age -> live_birth_bin_atleast_once_all
  age -> statut_marital_ini_cat
  age -> diplome_ini2
  age -> situation_financiere
  age -> situation_emploi_ini2
  age -> operation_endo_bin
  age -> nb_mal_sansEndoAde_cat
  
  nb_mal_sansEndoAde_cat -> live_birth_bin_atleast_once_all
  
  type_endo ->Endo_ade_categorie
  type_endo -> operation_endo_bin
  type_endo -> live_birth_bin_atleast_once_all
  type_endo -> stade_dvpt_legere_severe
  
  operation_endo_bin -> complication_operation
  operation_endo_bin -> live_birth_bin_atleast_once_all
  
  stade_dvpt_legere_severe -> live_birth_bin_atleast_once_all
  stade_dvpt_legere_severe -> operation_endo_bin
  
  first_preg_age_cat -> live_birth_bin_atleast_once_all
  
}"

dag_model <- dagitty(dag_code)

ggdag(dag_model, text_size = 
        3.5, text_col = "darkred", 
      node_size = 5) +
  theme_dag(base_size = 8, base_family = "Nunito") +
  theme(legend.position = "none") +
  scale_fill_manual(values = rep("white", 100)) +  
  scale_color_manual(values = rep("gray", 100))  


adjustmentSets(dag_model, 
               exposure ="operation_endo_bin", 
               outcome = "live_birth_bin_atleast_once_all",
               type= "canonical")
```

## 2. Final Model and Results
### 1) Final model (ref)
[1] "statut_marital_ini_cat"     "situation_financiere"      
[3] "endo_ade_categorie"         "age_symptome_cat"          
[5] "complication_operation_mod" "first_preg_age_cat" 

```{r}
data%>%sapply(function(x) sum(is.na(x)))
```
#### - Ref model 

stade (multicollinearity)
age_symptome_cat + delai_diag_cat_simple (multicollinearity)

```{r}
mod_glm_multivariable_0 <-
  glm(
    live_birth_bin_atleast_once_all ~
      age +
      diplome_ini2 +
      statut_marital_ini_cat +
      situation_financiere +
      nb_mal_sansEndoAde_cat +
      
      endo_ade_categorie +
      type_endo +
      #age_symptome_cat +
      
      #delai_diag_cat_simple+
      op_avant_1_preg +
      operation_endo_bin +
      
      first_preg_age_cat,
    data = imputed_data_O2,
    family = binomial
  )

# summary(mod_glm_multivariable_0)
AIC(mod_glm_multivariable_0)
# vif(mod_glm_multivariable_0)
```

#### - Final model 
- Variables selected by Marina

```{r}
# Final model
mod_glm_multivariable_final <- glm(
  live_birth_bin_atleast_once_all ~
    
    age +
    statut_marital_ini_cat +
    situation_financiere +
    
    nb_mal_sansEndoAde_cat +
    
    first_preg_age_cat +
    
    type_endo +
    operation_endo_bin,
  
  data = imputed_data_O2,
  family = binomial
)
summary(mod_glm_multivariable_final)
AIC(mod_glm_multivariable_final)
```


### 2) Assumption test
#### - Log linearity: OK
```{r}
data <- imputed_data_O2

# In R
df <- imputed_data_O2
df$log_x <- log(df$age)
df$x_logx <- df$age * log(df$age)

model <- glm(live_birth_bin_atleast_once_all ~ age + x_logx, family = binomial, data = df)
summary(model)  # Check p-value for x_logx
```

#### - Hosmer-Lemeshow : OK  p > 0.05
```{r}
# Hosmer-Lemeshow p-value> 0.05
y_obs <- ifelse(imputed_data_O2$live_birth_bin_atleast_once_all == "Oui", 1, 0)
hoslem <- hoslem.test(y_obs, fitted(mod_glm_multivariable_final), g = 10)
hoslem_p <- hoslem$p.value; hoslem_p
```

#### - VIF : OK (<5)
```{r}
# Multicollinearity Check : <5
vif(mod_glm_multivariable_final)
```

## 3) Results

```{r}
# Results 
# 1) summary()
summary(mod_glm_multivariable_final)

# 2) tidy(exponentiate=TRUE) results table 
res_multivar_glm <- tidy(mod_glm_multivariable_final, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_mod <- res_multivar_glm[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_mod <- res_multivar_glm_mod %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)

# Remove the intercept result
res_multivar_glm_mod <-res_multivar_glm_mod[-1,]

# Check only significants variables
res_multivar_glm_mod%>% filter(p.value<0.05)
```

#### - Save the results table

```{r}

# 1. Preparation 
data <- imputed_data_O2
outcome_variable <- "live_birth_bin_atleast_once_all"
covariables <- c(
  "age",
  "statut_marital_ini_cat",
  "situation_financiere",
  "nb_mal_sansEndoAde_cat",
  
  "first_preg_age_cat",
  
  "type_endo" ,
  "operation_endo_bin"
)

# 2. Descriptive table
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(covariables)) %>%
  tbl_summary(
    by = all_of(outcome_variable),         
    statistic = list(                     
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) 
  ) %>%
  modify_header(label ~ "**Characteristic**") %>%
  bold_labels()

# 3. Multivariable analysis results table
tbl_glm_multivariable <- tbl_regression(mod_glm_multivariable_final, exponentiate = TRUE, conf.int = TRUE)

# 4. Combined table
tbl_combined <- tbl_merge(
  tbls = list(tbl_descriptive, tbl_glm_multivariable),
  tab_spanner = c("**Descriptive Statistics**", "**Multivariable Logistic Regression**")
);tbl_combined

# 5. tbl_merge to dataframe
df_combined <- as_tibble(tbl_combined)

# 6. Save as Excel : update: 250722
#write_xlsx(df_combined, path = "O2_glm_multivar_LiveBirth.xlsx")

```


## 3) Visualization of results

```{r}
#------------
#  gt table
#------------
library(gt)
res_multivar_glm_mod %>%
  gt() %>%
  tab_header(
    title = md("**Multivariate Logistic Regression: **  
    *Factors Associated with Live Birth*")
  ) %>%
  cols_label(
    term = "Variables",
    estimate = "Adjusted OR",
    IC95 = "95% CI",
    p.value = "p-value"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = p.value,
      rows = as.numeric(gsub("<", "", p.value)) < 0.05
    )
  ) %>%
  cols_align(
    align = "left",
    columns = term
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95, p.value)
  )

#---------------
#  Forest plot 
#---------------
res_multivar_glm$term <- factor(res_multivar_glm$term, 
                                levels = rev(res_multivar_glm$term))

ggplot(res_multivar_glm, aes(x = term, y = estimate)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray") +
  coord_flip() +  
  labs(
    title = "Résultats du modèle de régression logistique multivariée",
    x = NULL,
    y = "OR (Odds Ratio avec IC 95%)"
  ) +
  theme_minimal()

#------------------------------
#  Forest plot with the ORs
#------------------------------
library(dotwhisker)
library(grid)

res_multivar_glm_mod$term <- c(
  "Age",
  "Maried/Couple vs Single/divorced",
  "Financial situation:  Balanced vs Difficult/In debt",
  "Balanced vs Comfortable/Very comfortable",
  "Nb_comorbidity: 0 vs 1",
  "0 vs 2+",
  "Age_1stPregnancy: <30 vs 30-34",
  "<30 vs >=34",
  "type_endo:  vs DIE only vs SPE only",
  "DIE only vs OMA only",
  "DIE only vs ADENO only",
  "DIE only vs Multiple",
  "DIE only vs No mention",
  "Endo surgery: No vs Yes"
)

res_multivar_glm_mod <- res_multivar_glm_mod %>%
  mutate(
    label_ci = sprintf("%.2f (%.2f–%.2f)", estimate, conf.low, conf.high)
  )

dwplot(res_multivar_glm_mod, show_intercept = TRUE) + 
  geom_text(aes(x = 3.2, y = term, label = label_ci), #paste(estimate, "(", IC95, ")")
            hjust = -0.1, size = 3) +
  theme_minimal() +
  labs(title = "Results of Multivariable Logistic Regression",
       subtitle ="- Experience of giving a Live birth : Yes/No",
       x= "OR with 95% CI") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0, size = 12,face = "bold"),
    plot.subtitle = element_text(hjust = 0, size = 10,face = "bold"), 
    axis.title.x = element_text(hjust = 0.3, vjust = 0),
    panel.grid = element_blank(),                
    panel.background = element_blank(),           
    axis.line.x =element_blank(), 
    axis.line.y = element_blank(),                
    #axis.ticks = element_line(color = "black"),   
    axis.text = element_text(color = "black"),
    axis.text.y = element_text(size = 10)
  ) +
  scale_x_continuous(breaks = 0:3) +              
  coord_cartesian(xlim = c(0, 5), ylim = c(1, 15)) +
  annotation_custom(
    grob = textGrob("OR (95%CI)", 
                    gp = gpar(fontsize = 10, col = "black",fontface = "bold")),
    xmin = 2, ymin = 14.5)+
  geom_segment(aes(x = -0.3, xend = 3.2, y = 0.5, yend = 0.5),  
             color = "black", size = 0.5)+
  geom_linerange(aes(xmin = conf.low, xmax = conf.high, y = term),
                 color = "red", size = 0.5, linetype=1)+
  geom_vline(xintercept = c(0,2,3), linetype = "dashed", color = "gray90", size = 0.5)+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", size = 0.5)->
  O2_Forest_Plot


O2_Forest_Plot
```

#### - Save Plot
```{r}
# Save  : update 250722
ggsave("O2_Forest_Plot.jpg", O2_Forest_Plot,
       width = 10, height = 5, units = "in", dpi = 300)
```

