---
title: "02_merge_cleaned_data"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : Merge all cleaned dataset
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Merge all cleaned dataset 
*Date of creation* : 2025-04-09
*Date of update* : 2025-06-02
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

```{r, setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(purrr) # reduce()
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gt)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
library(chromote)
library(car)
library(mediation)
```

# 0. Import all cleaned datasets

```{r}
# 1) socio demographique
pop_demo_final <-
  read.csv("C:/Users/nmj93/OneDrive/Master1_Biostats/Stage_M2_Complic_Gross/pop_demo_inclusion_final.csv") %>%
  dplyr::select(-X)
#pop_demo_final %>% glimpse

# 2) comorbidité
pop_comorb_final <-
  read.csv("pop_comorbidite_final.csv",
           stringsAsFactors = TRUE) %>%
  dplyr::select(-X)
#pop_comorb_final%>% glimpse


# 3) Grossesse sans grossesses en cours
SansGrossesseEnCours_final <-
  read.csv("sansgrossesseencours_final.csv",
           stringsAsFactors = TRUE) %>%
  dplyr::select(-X)
# # 1113 id_unik, 1575 pregnancies
# SansGrossesseEnCours_final %>% distinct(id_unik) %>% n_distinct()
# SansGrossesseEnCours_final %>% distinct(IdPregnancy) %>% n_distinct()

# All participants
# grossesse_all <-
#   read.csv("grossesse_mod_final.csv", stringsAsFactors = TRUE)%>%
#   dplyr::select(-X)
# #grossesse_all %>% glimpse

# 4) Santé des femmes
SanteFemmes_final <-
  read.csv("Sante_femmes_final.csv",
           stringsAsFactors = TRUE) %>%
  dplyr::select(-X)
#SanteFemmes_final%>% glimpse

# 5) Endo/Adenomyosis
EndoAdeno_final <-
  read.csv("Endo_Adeno_initial_final.csv",
           stringsAsFactors = TRUE) %>%
  dplyr::select(-X)
#EndoAdeno_final%>% glimpse

# 6) Endo/Adenomyosis Operation
EndoAdeno_op_examen_final <-
  read.csv(
    "Endo_parcours_chir_examens_final.csv",
    stringsAsFactors = TRUE
  ) %>%
  dplyr::select(-X)
#EndoAdeno_op_examen_final%>% glimpse

# 7) Dysmenorrhea
EndoAdeno_dysmenor_final <-
  read.csv("Endo_douleur_pelvic_abd_final.csv",
           stringsAsFactors = TRUE) %>%
  dplyr::select(-X)
#EndoAdeno_dysmenor_final%>% glimpse

# 8) Dyspareunia, dysuria and dyschezia
EndoAdeno_douleurs_final <-
  read.csv("Endo_dyspareunie_final.csv",
           stringsAsFactors = TRUE) %>%
  dplyr::select(-X)
#EndoAdeno_douleurs_final%>% glimpse

# 9) Financial situation
situation_finan_final <-
  read.csv("situation_finan_final.csv",
           stringsAsFactors = TRUE) %>%
  dplyr::select(-X)

# 10) PHQ9_GAD7
PHQ9_GAD7_final <-  read.csv("PHQ9_GAD7_final.csv",
                             stringsAsFactors = TRUE) %>%
  dplyr::select(-X)

# 11) PSQI
PSQI_final <-  read.csv("PSQI_final.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-X)

# 11) Treatment Burden Questionnaire
fardeau_trt_final <-  read.csv("fardeau_trt_final.csv",
                               stringsAsFactors = TRUE) %>%
  dplyr::select(-X)

# 12) Physical activity
activite_physique_final <- read.csv("activite_physique_final.csv",
                                    stringsAsFactors = TRUE) %>%
  dplyr::select(-X)

# 13) Quality of life
qualite_vie_final <-  read.csv("qualite_vie_final.csv",
                               stringsAsFactors = TRUE) %>%
  dplyr::select(-X)

# 14) Neuropathic pain
Doul_neuro_final <- read.csv("doul_neuro_final.csv",
                             stringsAsFactors = TRUE) %>%
  dplyr::select(-X)
```


# I. Merge final datasets
- All final datasets with 'SansGrossesseEnCours_final' -> df_merged_filterd

```{r}
df_list_filtered <- list(
  SansGrossesseEnCours_final,
  pop_demo_final,
  situation_finan_final,
  PHQ9_GAD7_final,
  PSQI_final,
  qualite_vie_final,
  activite_physique_final,
  fardeau_trt_final,
  pop_comorb_final,
  SanteFemmes_final,
  EndoAdeno_final,
  EndoAdeno_op_examen_final,
  EndoAdeno_dysmenor_final,
  EndoAdeno_douleurs_final,
  Doul_neuro_final
)

df_merged_filtered <- reduce(df_list_filtered, 
                             left_join, 
                             by="id_unik")

df_merged_filtered <- df_merged_filtered %>%
  mutate(across(c(starts_with("date"), 
                  starts_with("Date"), 
                  merged_DateDebut), 
                ~ as.Date(., format = "%Y-%m-%d")))


#test <- df_merged_filtered
```

- Check 
```{r}
df_merged_filtered %>% distinct(id_unik) %>% n_distinct()
df_merged_filtered %>% distinct(IdPregnancy) %>% n_distinct()
```


# II. Create new variables 

```{r}
#-----------------------------------
# Age of first endo/Adeno operation : age_firstOP_endoadeno
#-----------------------------------
# 1. Calculate the first age of operation related to Endo/Adeno
df_merged_filtered <- df_merged_filtered %>%
  mutate(
    age_firstOP_endoadeno =
      as.character(round((date_operation1 - Date_naissance) / 365, 1)) %>%
      as.numeric())

# 2. Create a categorical variable 
quantiles_agefirstOP <- round(
  quantile(df_merged_filtered$age_firstOP_endoadeno, 
                                probs = c(1/3,2/3), 
           na.rm = TRUE))

df_merged_filtered <- df_merged_filtered %>% 
  mutate(
    age_firstOP_endoadeno_cat = 
      case_when(
        age_firstOP_endoadeno < quantiles_agefirstOP[[1]] ~
          paste0("<",quantiles_agefirstOP[[1]]),
        age_firstOP_endoadeno >= quantiles_agefirstOP[[1]]& 
          age_firstOP_endoadeno < quantiles_agefirstOP[[2]] ~
          paste0(quantiles_agefirstOP[[1]],"-",quantiles_agefirstOP[[2]]),
        age_firstOP_endoadeno >= quantiles_agefirstOP[[2]] ~
          paste0(quantiles_agefirstOP[[2]],">="),
        TRUE ~ NA_character_
      ) %>% as.factor()
  ) 

# 3. Add new modalities to the categorical variable
df_merged_filtered <- df_merged_filtered%>% 
  mutate(
    age_firstOP_endoadeno_cat=
      case_when(
        operation_endo_bin=="Non"& is.na(age_firstOP_endoadeno_cat) ~
          "No operation",
        operation_endo_bin=="Not_answered" & 
          is.na(age_firstOP_endoadeno_cat) ~ "Not_answered",
        TRUE ~ age_firstOP_endoadeno_cat
           ))

#------------------------------------------------------------
#   Number of different types of pain
# : how many different type of pain they suffered in their life
#------------------------------------------------------------
# 1. Create a temporary dataframe by making a new variable 
new_dataframe <- df_merged_filtered %>%
  distinct(
    id_unik,
    doul_pelvienne_regle,
    doul_rapport_sexuel,
    douleur_abdo,
    douleur_miction,
    trouble_digestif, 
    doul_neuro_bin
  ) %>%
  mutate(
    nb_types_doul = case_when(
      rowSums(across(c(starts_with("doul"), trouble_digestif),
                     ~ . %in% c("Not_answered","not_answered")), na.rm = TRUE) == 6 ~
        "Not_answered",
      TRUE ~ 
        as.character(rowSums(across(c(starts_with("doul"), trouble_digestif), 
                                    ~ . == "Oui"), 
                             na.rm = TRUE)
                     )
      ) %>% as.factor()
    ) %>% 
  distinct(id_unik, nb_types_doul)

# 2. Add this new variable to the main dataframe
df_merged_filtered <-  df_merged_filtered %>% 
  left_join(new_dataframe, by="id_unik")

# 3. Delete the temporary dataframe
rm(new_dataframe)



#----------------------------
# Infertility (diagnostics)  : IRM, echo, ART or OP for the infertility
#----------------------------
new_dataframe <- df_merged_filtered %>%
  group_by(id_unik) %>%
  summarise(
    methode_perID_IUI = any(methode_perID_IUI == TRUE, 
                            na.rm = TRUE),
    methode_perID_IVF= any(methode_perID_IVF == TRUE, 
                            na.rm = TRUE)
  ) %>% ungroup()

df_merged_filtered <- df_merged_filtered %>% 
  dplyr::select(-methode_perID_IUI,-methode_perID_IVF) %>% 
  left_join(new_dataframe, by="id_unik")

df_merged_filtered <- df_merged_filtered %>%
  rowwise() %>%
  mutate(
    infertilite_diag = case_when(
      all(
        is.na(methode_perID_IUI),
        is.na(methode_perID_IVF),
        is.na(indic_motif_1_infertilite),
        is.na(indic_motif_2_infertilite),
        is.na(nb_echo_infertilite),
        is.na(nb_IRM_infertilite)
      ) ~ NA_character_,
      any(
        methode_perID_IUI == TRUE,
        methode_perID_IVF == TRUE,
        infertility_comorbidite == "Oui",
        indic_motif_1_infertilite == "Opérée pour ce motif",
        indic_motif_2_infertilite == "Opérée pour ce motif",
        nb_echo_infertilite %in% c("1-5","6-10","11-20","Plus de 20"),
        nb_IRM_infertilite %in% c("1-5","6-10")
      ) ~ "Oui",

      TRUE ~ "Non"
    )
  ) %>%
  ungroup()

rm(new_dataframe)

# df_merged_filtered %>% distinct(id_unik, infertilite_diag, PMA_perID,infertility_comorbidite, nb_echo_infertilite, nb_IRM_infertilite, indic_motif_1_infertilite, indic_motif_2_infertilite) %>% View

#==================================
#    ADD SEVERAL NEW VARIABLES
#==================================
df_merged_filtered <- df_merged_filtered %>%
  distinct() %>%
  mutate(
    #---------------------------------------------------------------------
    # Endometriosis operation before the first pregnancy registered by participants
    #---------------------------------------------------------------------
    op_avant_1_preg =
      case_when(
        age_firstOP_endoadeno <= first_preg_age ~ "Oui",
        age_firstOP_endoadeno > first_preg_age ~ "Non",
        age_firstOP_endoadeno_cat == "No operation" ~ "Non",
        age_firstOP_endoadeno_cat == "Not_answered" ~
          "Not_answered",
        TRUE ~ NA_character_
      ),
    
    #---------------------------------------------------------------------
    # Endometriosis operation before the last pregnancy registered by participants
    #---------------------------------------------------------------------
    op_avant_preg_perID =
      case_when(
        age_firstOP_endoadeno <= first_preg_age ~ "Before",
        age_firstOP_endoadeno > first_preg_age &
          is.na(last_preg_age_mod) ~
          "After",
        age_firstOP_endoadeno > first_preg_age &
          !is.na(last_preg_age_mod) &
          age_firstOP_endoadeno <= last_preg_age_mod ~
          "Before",
        age_firstOP_endoadeno > first_preg_age &
          !is.na(last_preg_age_mod) &
          age_firstOP_endoadeno > last_preg_age_mod ~
          "After",
        age_firstOP_endoadeno_cat == "No operation" ~ "No Operation",
        age_firstOP_endoadeno_cat == "Not_answered" ~ "nsp",
        TRUE ~ NA_character_
      ),
    #---------------------------------------------------------------
    #  Endometriosis operation for the treatement of infertility
    #--------------------------------------------------------------
    op_infertility_perID = 
      case_when(
        operation_endo_bin=="Oui"&
        (indic_motif_1_infertilite =="Opérée pour ce motif" | 
          indic_motif_2_infertilite =="Opérée pour ce motif") ~ "for_infertility",
        
        operation_endo_bin == "Oui" ~ "for_others",
        
        operation_endo_bin=="Non" ~ "No operation",
        operation_endo_bin == "Not_answered" ~ "Not_answered",
        TRUE ~NA_character_
      ),
    
    #---------------------------------------------------------------------
    # Add the information on operations related to infertility into conception methode per id
    #---------------------------------------------------------------------
    methode_conception_perID_cat_simple_mod = case_when(
      op_avant_preg_perID != "Before"  ~ 
        methode_conception_perID_cat_simple,
      op_avant_preg_perID =="Before" & op_infertility_perID =="for_others" ~ 
        methode_conception_perID_cat_simple,
      op_avant_preg_perID == "Before" & op_infertility_perID =="for_infertility" &
        is.na(methode_conception_perID_cat_simple) ~ 
        "Operation",
      op_avant_preg_perID == "Before" & op_infertility_perID =="for_infertility" &
        methode_conception_perID_cat_simple %in% c("Naturelle seule", "Operation") ~ 
        "Operation",
      op_avant_preg_perID == "Before" & op_infertility_perID =="for_infertility" &
        methode_conception_perID_cat_simple %in% c("PMA/Medicaments", "Operation+PMA") ~ 
        "Operation+PMA",
      TRUE~ NA_character_
      ),
    
    Methode_conception_bin_perID_mod = 
      ifelse(methode_conception_perID_cat_simple_mod =="Naturelle seule",
             "Naturelle", 
             "Avec méthodes"),
      
    # methode_conception_perID_cat_simple_mod = case_when(
    #   op_avant_preg_perID != "Before" ~ methode_conception_perID_cat_simple,
    #   op_avant_preg_perID == "Before" & is.na(methode_conception_perID_cat_simple) ~ 
    #     "Operation",
    #   op_avant_preg_perID == "Before" &
    #     methode_conception_perID_cat_simple %in% c("PMA/Medicaments", "Operation+PMA") ~
    #     "Operation+PMA",
    #   op_avant_preg_perID == "Before" &
    #     methode_conception_perID_cat_simple %in% c("Naturelle seule", "Operation") ~
    #     "Operation",
    #   TRUE ~ NA_character_
    # ),
    # Methode_conception_bin_perID_mod = 
    #   ifelse(methode_conception_perID_cat_simple_mod=="Naturelle seule","Naturelle","Avec méthodes")
    
    complication_operation_mod = case_when(
      complication_operation=="Oui"~ "Oui",
      complication_operation=="No surgery" ~ "No surgery",
      complication_operation=="Not_answered" ~ "Not_answered",
      complication_operation=="Non" & 
        indic_motif_2_complication =="Opérée pour ce motif" ~ "Oui",
      complication_operation=="Non" &
        indic_motif_2_complication !="Opérée pour ce motif" ~ "Non",
      
      TRUE ~NA_character_
      )
  )


#------------------------
# OP before the pregnancy : Before : Yes/No
#-------------------------
df_merged_filtered <- df_merged_filtered %>% 
  mutate(
    op_before_preg_perPreg = 
      case_when(
        operation_endo_bin=="Non" ~ "No operation",
        operation_endo_bin =="Not_answered" ~ "nsp",
        operation_endo_bin=="Oui"&
          age_firstOP_endoadeno <= age_maternal ~ "Yes",
        operation_endo_bin=="Oui"&
          age_firstOP_endoadeno > age_maternal ~ "No",
        TRUE ~ NA_character_
      )
  )


#--------------------------
# Regrouping the modality : status_marital_ini and situation_emploi_ini2
#--------------------------
df_merged_filtered <- df_merged_filtered %>% 
  mutate(
    statut_marital_ini_cat =
      case_when(
        statut_marital_ini %in% c("Marié(e)", "En couple (concubinage, PACS)") ~
          "Maried/Couple(PACS,concubinage)",
        statut_marital_ini %in% c("Célibataire", "Divorcé(e)") ~
          "Single/divorced"
      ) %>% as.factor(),
    
    situation_emploi_ini2_cat =
      case_when(
        situation_emploi_ini2 %in% c(
          "Sans Emploi",
          "En invalidité / en longue maladie",
          "Etudiant",
          "Autre"
        ) ~ "Pas active",
        situation_emploi_ini2 == "Employee" ~ "Employee"
      ) %>% as.factor()
  )

#test <- df_merged_filtered


#------------------------------------------------------------
# Conception method + endometriosis operation per pregnancy 
#------------------------------------------------------------
# 1. Create new variables : Age at each operation and Age at each operation only related to infertility
df_merged_filtered <- df_merged_filtered %>% 
  # Age at each operation
  mutate(
    # age_firstOP_endoadeno
    age_2ndOP_endoadeno=
      round(as.numeric(as.character(date_operation2-Date_naissance))/365,1),
    age_3rdOP_endoadeno=
      round(as.numeric(as.character(date_operation3-Date_naissance))/365,1),
    age_4thOP_endoadeno=
      round(as.numeric(as.character(date_operation4-Date_naissance))/365,1),
    age_5thOP_endoadeno=
      round(as.numeric(as.character(date_operation5-Date_naissance))/365,1),
    age_6thOP_endoadeno=
      round(as.numeric(as.character(date_operation6-Date_naissance))/365,1)
    ) %>% 
  # Age at each operation only related to infertility
  mutate(
    age_OP_infertility = case_when(
      indic_motif_1_infertilite == "Opérée pour ce motif" ~ 
        age_firstOP_endoadeno,
      indic_motif_2ndOP_infertilite == "Opérée pour ce motif" ~
        age_2ndOP_endoadeno,
      indic_motif_3rdOP_infertilite == "Opérée pour ce motif" ~ 
        age_3rdOP_endoadeno, 
      indic_motif_4thOP_infertilite == "Opérée pour ce motif" ~
        age_4thOP_endoadeno,
      indic_motif_5thOP_infertilite == "Opérée pour ce motif" ~
        age_5thOP_endoadeno,
      indic_motif_6thOP_infertilite == "Opérée pour ce motif" ~
        age_6thOP_endoadeno,
      TRUE ~ NA_real_
  )
)

# 2. Create a temporary dataframe 
new_dataframe <- df_merged_filtered %>%
  mutate(
    op_infertility_before_preg = case_when(
      op_infertility_perID == "for_infertility" &
        age_OP_infertility <= age_maternal ~
        "Yes",
      op_infertility_perID == "for_infertility" &
        age_OP_infertility > age_maternal ~
        "No",
      op_infertility_perID != "for_infertility" ~
        op_infertility_perID,
      TRUE ~ NA_character_
    ),
    closest_op_time_diff = case_when(
      op_infertility_before_preg == "Yes" ~ age_maternal - age_OP_infertility,
      TRUE ~ NA_real_
    )
  ) %>%
  group_by(id_unik, IdPregnancy) %>%
  mutate(
    min_diff = ifelse(all(is.na(
      closest_op_time_diff
    )), NA_real_, min(closest_op_time_diff, na.rm = TRUE)),
    
    op_infertility_1yr_before_each_preg = case_when(
      !is.na(closest_op_time_diff) &
        closest_op_time_diff == min_diff &
        closest_op_time_diff <= 1 ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  ungroup() %>%
  distinct(
    id_unik,
    IdPregnancy,
    min_diff,
    op_infertility_1yr_before_each_preg,
    methode_conception_perPreg_cat_simple,
    op_infertility_perID
  )

#-----------------
# Justification  |
#--------------------------------------------------------------------------------------------------------
# The severity of endometriosis negatively correlates with IVF pregnancy rates, with the optimal time for IVF being 7-25 months post-surgery (Alkudmani et al., 2018). 
# Laparoscopic treatment of endometriosis in infertile women led to 65% achieving pregnancy within 8.5 months on average, with shorter time to spontaneous pregnancy for earlier stage disease.(https://doi.org/10.1016/j.rbmo.2009.12.010) 
# Most infertile women with endometriosis conceived within 6 months after laparoscopic surgery.(https://doi.org/10.5653/cerm.2013.40.1.29)
# Laparoscopic surgery for endometriosis can improve fertility, with most spontaneous pregnancies occurring within 12 months postoperatively.(https://doi.org/10.1016/j.gyobfe.2011.07.037)
#----------------------------------------------------------------------------------------------------------
# 3. Add the new endometriosis operation information to each pregnancy in order to create a new variable
new_dataframe <- new_dataframe %>% 
  mutate(
    methode_conception_perPreg_cat_simple_mod =
      case_when(
        op_infertility_1yr_before_each_preg==FALSE ~ 
          methode_conception_perPreg_cat_simple,
        
        op_infertility_1yr_before_each_preg==TRUE & 
          methode_conception_perPreg_cat_simple =="PMA/Medicaments" ~ 
          "Operation+PMA",
        
        op_infertility_1yr_before_each_preg==TRUE & 
          methode_conception_perPreg_cat_simple %in% c("Operation","Naturelle seule")|
          is.na(methode_conception_perPreg_cat_simple) ~ 
          "Operation",
        TRUE ~ NA_character_
      )
  )

# 4. Merge
df_merged_filtered <- df_merged_filtered %>% 
  left_join(new_dataframe %>% distinct(IdPregnancy, min_diff, methode_conception_perPreg_cat_simple_mod), 
            by="IdPregnancy") #%>% 
  #dplyr::select(-c(age_2ndOP_endoadeno:age_6thOP_endoadeno))

# 5. Delete the temporary dataframe 
rm(new_dataframe)

#---------------------------------------
# Conception method in dummy variables
#---------------------------------------
df_merged_filtered <- df_merged_filtered%>%
  mutate(
    # ART conception method
    conception_PMA_perPreg = 
      ifelse(methode_conception_perPreg_cat_simple_mod %in% 
                           c("PMA/Medicaments","Operation+PMA"), "Oui","Non"),
    
    conception_PMA_perID= 
      ifelse(methode_conception_perID_cat_simple_mod %in% 
                           c("PMA/Medicaments","Operation+PMA"), "Oui","Non"),
    
    # Natural conception
    conception_natural_perPreg = 
      ifelse(methode_conception_perPreg_cat_simple_mod =="Naturelle seule",
             "Oui","Non"),
    conception_natural_perID= 
      ifelse(methode_conception_perID_cat_simple_mod =="Naturelle seule",
             "Oui","Non"),
    
    # Surgery taken coception method
    conception_operation_perPreg = 
      ifelse(methode_conception_perPreg_cat_simple_mod %in% 
               c("Operation","Operation+PMA"),"Oui","Non"),
    conception_operation_perID= 
      ifelse(methode_conception_perID_cat_simple_mod %in% 
               c("Operation","Operation+PMA"),"Oui","Non")
  )

#---------------------------------------------------------
# Temps en moyenne entre
# - la 1ère grossesse et le diagnostic d’endométriose
# - le 1er accouchement et le diagnostic d’endométriose
#---------------------------------------------------------
df_merged_filtered <- df_merged_filtered %>% 
  mutate(
    time_entre_preg_diag = 
      case_when(
        !is.na(first_preg_age)&!is.na(Age_premier_diag_endo_ade_cat) ~
          first_preg_age-Age_premier_diag_endo_ade,
        TRUE~ NA_real_
      ),
    
    time_entre_accouche_diag= 
      case_when(
        !is.na(first_preg_age)&!is.na(Age_premier_diag_endo_ade_cat) & 
          status_terme_all== "Terme" ~
          first_preg_age-Age_premier_diag_endo_ade,
        TRUE~NA_real_
      )
    )

#-------------------------------------
# Diagnostic before the 1st pregnancy
#-------------------------------------
df_merged_filtered <- df_merged_filtered %>% 
  mutate(
    diagnostic_avant_1_preg =
      case_when(
        time_entre_preg_diag < 0  ~ "Apres",
        time_entre_preg_diag >= 0  ~ "Avant/même_moment",
        TRUE~ NA_character_
      )
    )

#-------------------------------------------
#   Endometriosis stage in dummy variables 
#-------------------------------------------
df_merged_filtered <- df_merged_filtered %>% 
  mutate(
    stade_dvpt_I_II = case_when(
      stade_dvpt_legere_severe =="I-II" ~ "Oui",
      stade_dvpt_legere_severe !="I-II" ~ "Non",
      TRUE ~ NA_character_
    ),
    stade_dvpt_III_IV = case_when(
      stade_dvpt_legere_severe =="III-IV" ~ "Oui",
      stade_dvpt_legere_severe !="III-IV" ~ "Non",
      TRUE ~ NA_character_
    )
  )


#----------------------------------------------
# History of F/C, Preterm, IVG, IMG 
#----------------------------------------------
df_merged_filtered  <- df_merged_filtered %>% 
  mutate(
    history_FC_preterm_IVGIMG_perPreg = 
      case_when(
        type_preg_all_simple %in% c("Fausse-couche","IVG/IMG","Preterm") ~ "Oui",
        TRUE ~ "Non")) %>% 
  group_by(id_unik) %>%
  mutate(
    history_FC_preterm_IVGIMG_perID =
      case_when(
        any(history_FC_preterm_IVGIMG_perPreg=="Oui") ~ "Oui",
        all(history_FC_preterm_IVGIMG_perPreg=="Non") ~ "Non",
        TRUE ~ NA_character_
    )) %>% ungroup()
```


# III. Save the final merged dataframe 

```{r}
# update : 2025-06-20
write.csv(df_merged_filtered, "df_merged_filtered.csv")
```


# IV. Final dataframe for the analysis
## Objective 2 : Live birth (Yes/No)

```{r}
df_final_LB_1 <- df_merged_filtered %>%
  distinct(
    id_unik,
    live_birth_bin_atleast_once_all, # outcome
    
    # - Socio-demographic and helath condition
    age,
    statut_marital_ini_cat,
    diplome_ini2,
    situation_emploi_ini2_cat,
    situation_financiere,
    
    # - Menstrual and Reproductive history 
    family_atcd_EndoAdeno_cat,
    family_atcd_doul_pelv,
    first_preg_age,
    
    # - Comordibidty
    nb_mal_sansEndoAde_cat,
    comorbidite_bin,
    #gynecologique_diseases,
    
    # - Endo/Adeno 
    endo_ade_categorie,
    type_endo,
    delai_diag_cat,
    age_symptome_cat,
    age_firstOP_endoadeno_cat,
    
    stade_dvpt_legere_severe,
    
    # - Operation
    operation_endo_bin,
    op_avant_1_preg,
    complication_operation_mod,
    #trompe_anomalie,
    
    # # - Pain
    # dysmenorrhea_avant_1_preg,
    # doul_abdo_avant_1_preg,
    # dyspareunia_avant_1_preg,
    # dyschezia_avant_1_preg,
    # dysuria_avant_1_preg
  ) %>% 
  mutate(across(where(is.character), as.factor))

##############################################
#              Minji version
##############################################
df_final_LB_2 <- df_merged_filtered %>%
  dplyr::select(
    id_unik,
    live_birth_bin_atleast_once_all, # outcome
    
    # - Socio-demographic and helath condition
    age,
    statut_marital_ini_cat,
    diplome_ini2,
    situation_emploi_ini2_cat,
    situation_financiere,
    
    # - Menstrual and Reproductive history 
    family_atcd_EndoAdeno_cat,
    family_atcd_doul_pelv,
    first_preg_age,
    
    # - Comordibidty
    nb_mal_sansEndoAde_cat,
    comorbidite_bin,
    gynecologique_diseases,
    autoimmune_diseases,
    metabolic_syndrome,
    
    # - Endo/Adeno 
    endo_ade_categorie,
    type_endo,
    DIE,OMA,SPE,ADENO,
    
    delai_diag_cat,    
    delai_diag_cat_simple,

    age_symptome_cat,
    age_firstOP_endoadeno_cat,
    stade_dvpt_legere_severe,

    
    # - Operation
    operation_endo_bin,
    op_avant_1_preg,
    complication_operation_mod,
    #trompe_anomalie,
    
    
    ends_with("_participated")
  ) %>% 
  distinct() %>%
  mutate(across(where(is.character), as.factor))

```

```{r}
##############################################
#              Minji version
##############################################
df_final_LB_2 <- df_merged_filtered %>%
  dplyr::select(
    id_unik,
    live_birth_bin_atleast_once_all, # outcome
    
    # - Socio-demographic and helath condition
    age,
    statut_marital_ini_cat,
    diplome_ini2,
    situation_emploi_ini2_cat,
    situation_financiere,
    
    # - Menstrual and Reproductive history 
    family_atcd_EndoAdeno_cat,
    family_atcd_doul_pelv,
    first_preg_age,
    
    # - Comordibidty
    nb_mal_sansEndoAde_cat,
    comorbidite_bin,
    gynecologique_diseases,
    autoimmune_diseases,
    metabolic_syndrome,
    
    # - Endo/Adeno 
    endo_ade_categorie,
    type_endo,
    DIE,OMA,SPE,ADENO,
    
    delai_diag_cat,    
    delai_diag_cat_simple,

    age_symptome_cat,
    age_firstOP_endoadeno_cat,
    stade_dvpt_legere_severe,

    
    # - Operation
    operation_endo_bin,
    op_avant_1_preg,
    complication_operation_mod,
    #trompe_anomalie,
    
    
    ends_with("_participated")
  ) %>% 
  distinct() %>%
  mutate(across(where(is.character), as.factor))
```



### - Save the final dataframe for the analysis

```{r}
# update : 25-07-11
write.csv(df_final_LB_1, "df_final_LB_1.csv")
write.csv(df_final_LB_2, "df_final_LB_2.csv")

```


## Objective 3 : Pregnancy Complications 

Risk factors  : 
Maternal age, nulliparity, history of miscarriage/stillbirth, and lack of pregnancy wantedness (https://doi.org/10.1186/1471-2393-14-347)

Obesity is a major risk factor for a wide range of adverse obstetric outcomes. (https://doi.org/10.1016/j.siny.2009.10.001)

Preeclampsia :
maternal age, multiple pregnancy, nulliparity, chronic hypertension, inadequate antenatal care, obesity, and diabetes. (https://doi.org/10.30574/wjarr.2023.17.1.0012)

*Metabolic syndrome (MS)* :is increasingly prevalent in pregnant women, with rates ranging from 3% to 42% depending on various factors (Hamilton dos Prazeres Tavares et al., 2015). MS during pregnancy is associated with numerous obstetric complications, including gestational diabetes, preeclampsia, fetal growth retardation, macrosomia, preterm delivery, and stillbirth (А. В. Хромылев, 2014; Hamilton dos Prazeres Tavares et al., 2015; A. Zakharko et al., 2018). These complications are often referred to as "great obstetrical syndromes" due to their shared pathogenesis (L. Shelestova & L.F. Ushullu, 2021). Women with MS are also at higher risk for venous thrombosis during pregnancy (А. В. Хромылев, 2014). Newborns of mothers with MS tend to have higher birth weights and are more likely to experience hypoxic and ischemic central nervous system damage (A. Zakharko et al., 2018). To mitigate these risks, lifestyle modifications, including balanced low-calorie diets and exercise, are recommended during the preconception period (L. Shelestova & L.F. Ushullu, 2021).

*ART* : Assisted reproductive techniques (ART) have been associated with an increased risk of obstetric and perinatal complications compared to spontaneous pregnancies (Bănică et al., 2021; Madrazo-Cabo et al., 2020). These complications include low birth weight, prematurity, small for gestational age infants, admission to neonatal intensive care units, congenital malformations, cesarean sections, and premature rupture of membranes (Madrazo-Cabo et al., 2020). Multiple pregnancies, which are more common in ART, are an independent risk factor for complications (Bănică et al., 2021). However, even singleton ART pregnancies demonstrate higher rates of perinatal complications and maternal issues such as preeclampsia and gestational diabetes (Reddy et al., 2007). The increased risk is observed in both in vitro fertilization (IVF) and intracytoplasmic sperm injection (ICSI) procedures (Madrazo-Cabo et al., 2020). While it's challenging to separate ART-related risks from underlying reproductive pathology, healthcare providers should discuss these potential complications with couples considering ART (Reddy et al., 2007).

*Earlier Age at Menarche*
- associated with obstetric outcomes in nulliparous women.(https://doi.org/10.1111/jog.15057)
- Early menarche is a small risk factor for endometriosis.(https://doi.org/10.1016/j.fertnstert.2012.05.035 )
- Early menarche (before age 12) is associated with increased risk of cardiovascular disease events, cardiovascular mortality, and overall mortality in women.(https://doi.org/10.1210/jc.2009-1789)
- Earlier age at menarche is associated with increased risk of gestational diabetes in pregnancy. (https://doi.org/10.1007/s00592-018-1214-z)
- Early menarche is associated with increased risk of type 2 diabetes, potentially mediated through excessive adult adiposity.(doi: 10.1093/aje/kwp372; https://pmc-ncbi-nlm-nih-gov.ezproxy.u-paris.fr/articles/PMC2842205/, https://doi.org/10.1016/j.diabet.2017.01.002 ) 
- Early menarche is associated with increased risk of metabolic disorders, female-specific cancers, and mortality in women.(https://doi.org/10.1093/qjmed/hcac187)
- Early menarche (≤11 years) is associated with a higher risk of developing gestational diabetes mellitus.(https://doi.org/10.1093/aje/kww201)
- Early menarche is associated with higher risk of preeclampsia, and prepregnancy weight may modify the association between long menstrual cycles and preeclampsia risk.(https://doi.org/10.5402/2011/472083)
- Early menarche age is associated with higher general and abdominal obesity in young women, while late menarche age is associated with higher probability of underweight.(https://doi.org/10.3390/medicina60101711 ;2024)
- Earlier menarche age is associated with increased risks of gestational hypertension and preeclampsia.(https://doi.org/10.1111/jch.14737 ;2023)
- Earlier menarche (≤11 years) is associated with an increased prevalence of preterm birth.(https://doi.org/10.1038/s41598-017-12817-2 ; 2017)


*Age at the first pregnancy* 
- Early age at first childbirth is associated with increased risk of poor pregnancy outcomes, including low birthweight, preterm birth, and maternal anemia.


### - One row = One participant (n=378)
#### * DF V1 : Original
```{r}
# Filter only the pregnancies covered by the new module
df_final_complic_perID <- 
  df_merged_filtered %>% 
  filter(module=="Nouveau") %>% 
  dplyr::select(
    id_unik, 
    #IdPregnancy,
    
    # Outcome
    total_complications_bin_perID,
    total_complications_cat_perID,
    Complications_cat_perID_grouped,
    
    # - Socio-demographic and helath condition
    age,
    statut_marital_ini_cat,
    diplome_ini2,
    situation_emploi_ini2_cat,
    #Statut_tabac_ini,
    #freq_alcool_ini,
    situation_financiere,
    
    # - Comordibidty
    nb_mal_sansEndoAde_cat,
    #comorbidite_bin,
    gynecologique_diseases,
    metabolic_syndrome,
    infertilite_diag,
    autoimmune_diseases,
    
    # - Menstrual and Reproductive history 
    #age_regle_cat,
    age_regle_cat_simple,

    # - Endo/Adeno 
    family_atcd_EndoAdeno_cat,
    family_atcd_doul_pelv,
    endo_ade_categorie,
    type_endo,
    ADENO,
    OMA,
    SPE,
    DIE,
    
    stade_dvpt_legere_severe,
    stade_dvpt_I_II,
    stade_dvpt_III_IV,
    
    #delai_diag,
    delai_diag_cat,
    delai_diag_cat_simple,
    age_symptome_cat,
    #Age_premier_diag_endo_ade,
    #Age_premier_diag_endo_ade_cat,
    
    # - Operation
    operation_endo_bin,
    #nb_operation_endo_cat2,
    complication_operation,
    age_firstOP_endoadeno_cat,
    #num_localisation_endo_cat,
    #niveaux_adherences_bin,
    op_avant_1_preg,
    indic_ovaire_touche,
    trompe_anomalie,
    
    # - Pregnancy 
    nb_total_gross,
    #live_birth_bin_atleast_once_all, # livebirth yes/no
    nb_live_births_all,
    first_preg_age,
    # type_preg_all_simple,
    # PoidInf2,
    # PoidSup2,
    history_FC_preterm_IVGIMG_perID,
    
    methode_conception_perID_cat_simple_mod,
    conception_PMA_perID, 
    conception_natural_perID, 
    conception_operation_perID,
    
    Mode_accouchement_perID,
    VD_perID,
    Csection_perID,
    
    Travail_accouchement_perID,
    labour_natural_perID,
    labour_induced_perID,
    
    PriseenCharge_cat_perID,
    # nb_allaitement,
    # allaitement_YN_atleast,
    
    # - Pain 
    doul_abdo_avant_1_preg,
    dysmenorrhea_avant_1_preg,
    dyspareunia_avant_1_preg,
    dyschezia_avant_1_preg,
    dysuria_avant_1_preg,
    
    Endo_Adeno_initial_participated,
    Endo_chir_examens_participated,
    Endo_dysmenorrhea_abd_participated,
    Endo_douleurs_participated,
    sante_femmes_participated

  ) %>% 
  distinct()

df_final_complic_perID %>%dim
```

#### * DF V2: with all age related variable in continuous 

```{r}
# Filter only the pregnancies covered by the new module
df_final_complic_perID_V2 <- 
  df_merged_filtered %>% 
  filter(module=="Nouveau") %>% 
  dplyr::select(
    id_unik, 
    #IdPregnancy,
    
    # Outcome
    total_complications_bin_perID,
    total_complications_cat_perID,
    Complications_cat_perID_grouped,
    
    # - Socio-demographic and helath condition
    age,
    statut_marital_ini_cat,
    diplome_ini2,
    situation_emploi_ini2_cat,
    #Statut_tabac_ini,
    #freq_alcool_ini,
    situation_financiere,
    
    # - Comordibidty
    nb_mal_sansEndoAde_cat,
    #comorbidite_bin,
    gynecologique_diseases,
    metabolic_syndrome,
    infertilite_diag,
    autoimmune_diseases,
    
    # - Menstrual and Reproductive history 
    age_regles_ini,
    age_regle_cat,
    age_regle_cat_simple,

    # - Endo/Adeno 
    family_atcd_EndoAdeno_cat,
    family_atcd_doul_pelv,
    endo_ade_categorie,
    type_endo,
    ADENO,
    OMA,
    SPE,
    DIE,
    
    stade_dvpt_legere_severe,
    stade_dvpt_I_II,
    stade_dvpt_III_IV,
    
    delai_diag,
    delai_diag_cat,
    delai_diag_cat_simple,
    age_symptome_cat,
    Age_premier_diag_endo_ade,
    Age_premier_diag_endo_ade_cat,
    
    # - Operation
    operation_endo_bin,
    complication_operation,
    age_firstOP_endoadeno,
    age_firstOP_endoadeno_cat,
    op_avant_1_preg,
    indic_ovaire_touche,
    trompe_anomalie,
    
    # - Pregnancy 
    nb_total_gross,
    nb_live_births_all,
    first_preg_age,
    history_FC_preterm_IVGIMG_perID,
    
    methode_conception_perID_cat_simple_mod,
    conception_PMA_perID, 
    conception_natural_perID, 
    conception_operation_perID,
    
    Mode_accouchement_perID,
    VD_perID,
    Csection_perID,
    
    Travail_accouchement_perID,
    labour_natural_perID,
    labour_induced_perID,
    
    PriseenCharge_cat_perID,
    
    # - Pain 
    doul_abdo_avant_1_preg,
    dysmenorrhea_avant_1_preg,
    dyspareunia_avant_1_preg,
    dyschezia_avant_1_preg,
    dysuria_avant_1_preg,
    
    Endo_Adeno_initial_participated,
    Endo_chir_examens_participated,
    Endo_dysmenorrhea_abd_participated,
    Endo_douleurs_participated,
    sante_femmes_participated

  ) %>% 
  distinct()

df_final_complic_perID_V2 %>%dim
```

#### => Save the final dataframe for the analysis

```{r}
# Update : 2025-06-20
write.csv(df_final_complic_perID, "df_final_complic_perID.csv")
write.csv(df_final_complic_perID_V2, "df_final_complic_perID_V2.csv")
```

### - One row = One pregnancy
n= 468

```{r}
df_final_complic_perPreg <- 
  df_merged_filtered %>% 
  filter(module=="Nouveau") %>% 
  dplyr::select(
    id_unik, 
    IdPregnancy,
    
    # Outcome
    total_complications_bin_perPreg,
    total_complications_cat_perPreg,
    Complications_cat_perPreg_grouped,
    
    # - Socio-demographic and helath condition
    age,
    statut_marital_ini_cat,
    diplome_ini2,
    situation_emploi_ini2_cat,
    situation_financiere,
    
    # - Comordibidty
    nb_mal_sansEndoAde_cat,
    gynecologique_diseases,
    metabolic_syndrome,
    infertilite_diag,
    autoimmune_diseases,
    
    # - Menstrual and Reproductive history 
    #age_regle_cat,
    age_regle_cat_simple,
    #nb_total_gross,
    #nb_live_births_all,
    #nb_incomplete_preg_pt_onlynew,
    
    # - Endo/Adeno 
    family_atcd_EndoAdeno_cat,
    family_atcd_doul_pelv,
    endo_ade_categorie,
    type_endo,
    ADENO,
    OMA,
    SPE,
    DIE,
    
    stade_dvpt_legere_severe,
    stade_dvpt_III_IV,
    # delai_diag,
    # delai_diag_cat,
    delai_diag_cat_simple,
    #Age_premier_diag_endo_ade,
    age_symptome_cat,
    #Age_premier_diag_endo_ade_cat,
    
    # - Operation
    #operation_endo_bin,
    nb_operation_endo_cat2,
    complication_operation,
    age_firstOP_endoadeno_cat,
    #num_localisation_endo_cat,
    #niveaux_adherences_bin,
    op_avant_1_preg,
    indic_ovaire_touche, 
    trompe_anomalie,
    
    # - Pain 
    doul_abdo_avant_1_preg,
    dysmenorrhea_avant_1_preg,
    dyspareunia_avant_1_preg,
    dyschezia_avant_1_preg,
    dysuria_avant_1_preg,
    
    # - Pregnancy
    #nb_total_gross,
    age_maternal,
    live_birth_bin_all, # Live birth Oui/Non per preg
    type_preg_all_simple, # type of pregnancies
    gravida,
    para,
    PoidInf2, # per preg
    PoidSup2, # per preg
    methode_conception_perPreg_cat_simple_mod,
    conception_PMA_perPreg,
    Mode_accouchement,
    Travail_accouchement,
    PriseenCharge_bin,
    PriseenCharge_cat_perPreg,
    Allaitement2, # per perg
    
    history_FC_preterm_IVGIMG_perID,
    
    Endo_Adeno_initial_participated,
    Endo_chir_examens_participated,
    Endo_dysmenorrhea_abd_participated,
    Endo_douleurs_participated,
    sante_femmes_participated
    
  ) %>% 
  distinct()
df_final_complic_perPreg %>% dim
```

#### =>  Save the final dataframe 

```{r}
# update : 2025-06-16
write.csv(df_final_complic_perPreg, "df_final_complic_perPreg.csv")
```


GEE TEST
```{r}
#install.packages("geepack")
library(geepack)

# outcome : binary
df_final_complic_perPreg <- df_final_complic_perPreg %>% 
  mutate(live_birth_bin_all=ifelse(live_birth_bin_all=="Oui",1,0))

gee <- geeglm(live_birth_bin_all ~ PMA_perPreg, 
              id = id_unik, 
              data = df_final_complic_perPreg, 
              family = binomial, 
              corstr = "exchangeable")
summary(gee)

library(geepack)

gee_data <- df_final_complic_perPreg %>%
  mutate(
    live_birth_bin_all = ifelse(live_birth_bin_all == "Oui", 1, 0)
  ) %>%
  filter(!is.na(live_birth_bin_all), !is.na(situation_financiere), !is.na(id_unik)) %>%
  mutate(situation_financiere = factor(situation_financiere))

gee <- geeglm(
  live_birth_bin_all ~ situation_financiere,
  id = id_unik,
  data = gee_data,
  family = binomial,
  corstr = "exchangeable"
)

summary(gee)


table(df_final_complic_perPreg$live_birth_bin_all, useNA = "always")
table(df_final_complic_perPreg$situation_financiere, useNA = "always")

```


#### - Save the final dataframe for the analysis

```{r}

```


Sequence analysis TEST
```{r}
library(TraMineR)

#-------------------------------------
#     Age at each operation
#-------------------------------------
# 1. operation and age at the operation
operation_cols <- paste0("date_operation", 1:13)
df_long_op <- df_merged_filtered %>%
  dplyr::select(id_unik, all_of(operation_cols), Date_naissance) %>%
  pivot_longer(cols = starts_with("date_operation"),
               names_to = "operation_num",
               values_to = "date_op") %>%
  filter(!is.na(date_op)) %>%
  mutate(age = round(interval(Date_naissance, ymd(date_op)) / years(1),1),
         event = "O") %>%
  dplyr::select(id_unik, age, event) %>% 
  distinct()

# 2. Onset Age of Symptoms and Age at First Diagnosis
df_symptom_diag <- df_merged_filtered %>%
  left_join(Endo_Adeno_initial %>% 
              distinct(id_unik, age_symptome), 
            by ="id_unik") %>%
  dplyr::select(id_unik, age_symptome, Age_premier_diag_endo_ade) %>%
  pivot_longer(
    cols = c(age_symptome, Age_premier_diag_endo_ade),
    names_to = "event_type",
    values_to = "age"
  ) %>%
  mutate(
    event = case_when(
      event_type == "age_symptome" ~ "S",
      event_type == "Age_premier_diag_endo_ade" ~ "D"
    )
  ) %>%
  dplyr::select(id_unik, age, event)

# 3. pregnancy start and end date and pregnancy results
df_preg <- df_merged_filtered %>%
  mutate(
    age_preg = round(
      interval(Date_naissance, ymd(merged_DateDebut)) / years(1), 3),
    age_outcome = round(interval(Date_naissance, ymd(DateFin)) / years(1), 3),
    event_preg = "P",
    event_outcome = ifelse(status_terme_all == "Terme", "B", "I")
  ) %>%
  dplyr::select(id_unik, age_preg, event_preg, age_outcome, event_outcome) %>%
  pivot_longer(
    cols = c(age_preg, age_outcome),
    names_to = "what",
    values_to = "age"
  ) %>%
  mutate(event = ifelse(what == "age_preg", event_preg, event_outcome)) %>%
  dplyr::select(id_unik, age, event)

# 4. age menarche and event 
df_age_menarche <- df_merged_filtered %>% 
  distinct(id_unik, age_regles_ini) %>% 
  mutate(
    event = ifelse(!is.na(age_regles_ini), "M", NA),
    age = age_regles_ini) %>% 
  dplyr::select(- age_regles_ini)


# 4. Determine the Most Recent Date of the Last Observation
df_Endo_Adeno_initial <- Endo_Adeno_initial %>%
  distinct(id_unik, DateValidation) %>%
  mutate(DateValidation_EndoAdenoinit =
           as.Date(DateValidation, format = "%d/%m/%Y")) %>%
  dplyr::select(-DateValidation)
df_Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  distinct(id_unik, DateValidation) %>%
  mutate(DateValidation_operation = 
           as.Date(DateValidation, format = "%d/%m/%Y")) %>%
  dplyr::select(-DateValidation)
df_grossesse_mod2 <- grossesse_mod2 %>%
  distinct(id_unik, DateModification, Date_naissance) %>%
  group_by(id_unik) %>%
  summarise(
    DateModification_gross = max(DateModification, na.rm = TRUE),  
    Date_naissance = first(Date_naissance)  
  ) %>%
  ungroup()

df <- df_Endo_Adeno_initial %>%
  left_join(df_Endo_parcours_chir_examens, by = "id_unik") %>%
  left_join(df_grossesse_mod2, by = "id_unik") %>% 
  left_join(df_merged_filtered %>% distinct(id_unik,age), by="id_unik")

df_age <- df %>%
  mutate(
    recent_DateValidation = pmax(
      DateValidation_EndoAdenoinit,
      DateValidation_operation,
      DateModification_gross,
      na.rm = TRUE
    ),
    age_at_last_obs = 
      round(as.numeric((recent_DateValidation - Date_naissance) / 365), 1),
    age_final = pmax(age, age_at_last_obs, na.rm = TRUE)
  ) %>%
  distinct(id_unik, age_final) %>%
  mutate(age = age_final, event = "END") %>%
  dplyr::select(-age_final) %>% 
  filter(!is.na(age)) %>%  # Remove rows with NA in age
  group_by(id_unik) %>%
  slice_max(age) %>%  # Get the row with the maximum age
  ungroup() 
  
# MERGE ALL dataframes 
event_df <- bind_rows(
  df_symptom_diag,
  df_long_op,
  df_preg, 
  df_age_menarche,
  df_age) %>%
  filter(!is.na(age)) %>%
  arrange(id_unik, age) %>% 
  distinct()

# Add 0.1 if some ages are equal
event_df <- event_df %>%
  group_by(id_unik, age) %>%
  mutate(
    age = if_else(
      "M" %in% event & "S" %in% event & event == "S",
      age + 0.1,
      age
    ),
    age = if_else(
      "P" %in% event & "I" %in% event & event=="I",
      age + 0.1,
      age
    ),
    age= if_else(
      "S" %in% event & "D" %in% event & event=="D",
      age + 0.1,
      age
    )
  ) %>%
  ungroup()

# Arrange data selon age
event_df <- event_df %>%
  group_by(id_unik) %>%
  arrange(age) %>%
  mutate(time = row_number()) %>%
  ungroup()

##############TEST##################
# Without Pregnancy information 
event_df_mod <-  event_df %>% 
  filter(event!="P") %>%
  group_by(id_unik) %>% 
  mutate(time= row_number()) %>% 
  ungroup()
#event_df_mod %>% View
seq_data_mod <- event_df_mod %>%
  pivot_wider(id_cols = id_unik,
              names_from = time, 
              values_from = event, 
              names_prefix = "T") %>%
  replace(is.na(.), "-")

seq_data_mod %>% View

seq_obj_mod <- seqdef(seq_data_mod %>%
                        dplyr::select(-id_unik))

seqstatf(seq_obj_mod)
seqdplot(seq_obj_mod, border = NA, with.legend = TRUE)
seqIplot(seq_obj_mod,
         border = NA,
         with.legend = TRUE,
         main = "Reproductive Trajectories (Individual)")


#################################
#------------------------------
# Tranformation in wide format
#------------------------------
seq_data <- event_df %>%
  pivot_wider(id_cols = id_unik,
              names_from = time, 
              values_from = event, 
              names_prefix = "T") %>%
  replace(is.na(.), "-")


seq_data %>% View
#----------------------------
#      Sequence Analysis 
#----------------------------
#install.packages("TraMineR")
library(TraMineR)

seq_obj <- seqdef(seq_data %>% dplyr::select(-id_unik))

seqstatf(seq_obj)
seqdplot(seq_obj, border = NA, with.legend = TRUE)
seqIplot(seq_obj,
         border = NA,
         with.legend = TRUE,
         main = "Reproductive Trajectories (Individual)")


# 트랜지션 레이트 매트릭스 시각화
seqtrate(seq_obj)
```



```{r}
dist_matrix <- seqdist(seq_obj_mod, method = "OM", sm = "CONSTANT")

dist_matrix <- as.dist(dist_matrix)
#class(dist_matrix)
hc <- hclust(dist_matrix)

clusters <- cutree(hc, k = 3)

seqdplot(seq_obj_mod, group = clusters,
         with.legend = TRUE,
         main = "Sequence Distribution by Cluster")

# -------------------
# silhouette 
#----------------



```

 CoxPH, Kaplan-Meier
```{r}
diagnosis_data <- event_df[event_df$event %in% c("I"), ]
diagnosis_data$event_bin <- ifelse(diagnosis_data$event == "I", 1, 0)

library(survival)
library(survminer)
surv_obj <- Surv(time = diagnosis_data$age, event = diagnosis_data$event_bin)

# Kaplan-Meier 
km_fit <- survfit(surv_obj ~1)

plot(km_fit, xlab = "Age", ylab = "Probability", 
     main = "Kaplan-Meier Curve for -")

# coxph
cox_fit <- coxph(surv_obj ~ first_preg_age, 
                 data = diagnosis_data %>% 
                   left_join(df_merged_filtered %>% 
                               distinct(id_unik, first_preg_age),
                             by="id_unik"))
summary(cox_fit)



###############################
#   Competing risks : B / I
###############################
library(cmprsk)

# event: "B" (Birth), "I" (incomplete) 
# time: age at the pregnancy outcome
preg_data <- event_df[event_df$event %in% c("B", "I"), ]
preg_data$status <- ifelse(preg_data$event == "B", 1, 2) # 1: birth, 2: incomplete
preg_data <- preg_data %>% 
  left_join(df_merged_filtered %>% 
                distinct(id_unik, PMA_perID),
            by="id_unik")
# competing risks
cr_fit <- cuminc(ftime = preg_data$age, 
                 fstatus = preg_data$status, 
                 group = preg_data$PMA_perID)
plot(cr_fit, color = 1:4, lty = 1:4)
```

```{r}
df_noNA <- df_final_complic_perPreg %>%
  dplyr::select(Complications_cat_perPreg_grouped,
                age, 
                age_maternal,
         statut_marital_ini_cat,
         situation_financiere,
         diplome_ini2,
         endo_ade_categorie,
         age_regle_cat_simple,
         type_endo,
         age_firstOP_endoadeno_cat,
         operation_endo_bin,
         methode_conception_perPreg_cat_simple_mod,
         Mode_accouchement,
         op_avant_1_preg,
         Travail_accouchement) %>%
  na.omit()

# 이걸 기준으로 model.matrix 만들기
df_dummy <- model.matrix(~ . -1, data = df_noNA)
df_dummy <- as.data.frame(df_dummy)
df_dummy$complication <- df_noNA$Complications_cat_perPreg_grouped


# df_dummy <- model.matrix( ~ live_birth_bin_all + 
#                             statut_marital_ini_cat +
#                             situation_financiere +
#                             # nb_mal_sansEndoAde_cat+
#                             gynecologique_diseases+
#                             endo_ade_categorie +
#                             age_regle_cat_simple +
#                             type_endo +
#                             age_firstOP_endoadeno_cat +
#                             # operation_endo_bin +
#                             methode_conception_perPreg_cat_simple_mod +
#                             Mode_accouchement +
#                             op_avant_1_preg +
#                             Travail_accouchement
#                           -1,
#                           data = df_final_complic_perPreg)
# 
# # 2. 필요한 변수 추가
# df_dummy <- as.data.frame(df_dummy)
# df_dummy$complication <-
#   df_final_complic_perPreg$Complications_cat_perPreg_grouped 
# df_dummy$id_unik <- df_final_complic_perPreg$id_unik

# 3. PCA
pca_res <- prcomp(df_dummy[, !names(df_dummy) %in% 
                             c("complication", "id_unik")], 
                  scale. = TRUE)

# numeric 열만 필터링
df_numeric <- df_dummy[sapply(df_dummy, is.numeric)]
# PCA 실행
pca_res <- prcomp(df_numeric, scale. = TRUE)

# 3. 계층적 클러스터링
dist_mat <- dist(pca_res$x[, 1:5]) 
hc <- hclust(dist_mat)

# 4. 덴드로그램 확인 및 군집 수 결정
plot(hc)
rect.hclust(hc, k=3,  border = "red") 

#install.packages("factoextra")
library(factoextra)
fviz_nbclust(df_dummy, kmeans, method = "wss")

# 5. 군집 할당
df_dummy$cluster <- cutree(hc, k = 3)

# 6. 군집별 합병증 분포 확인
table(df_dummy$cluster, df_dummy$Complications_cat_perPreg_groupedMultiple_complications)

```

