---
title: "04_O3_Statistical_Analysis_perPreg"
author: "Minji NAM"
date: "2025-06-02"
output: html_document
---

*Name* : **3 Objective**, Statistical Analysis for the repeated pregnancy data
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Statistical Analysis
*Date of creation* : 2025-06-02
*Date of update* : 2025-06-02
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(purrr) # reduce()
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gt)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
library(chromote)
library(car)
library(mediation)
library(shiny)
library(shinydashboard)
library(data.table)
library(DT)
library(mice)
library(dagitty)
library(ggdag)
library(gridExtra)
library(grid)
library(dotwhisker)
```

# 0. Import the imputed data 
## 0-1) dataset for the univariable analysis

```{r}
imputed_data_O3_perPreg_imp1 <-
  read.csv("imputed_data_O3_perPreg_imp1.csv")

imputed_data_O3_perPreg_imp1
```

## 0-2) dataset for the multivariable analysis 

```{r}
imputed_data_O3_perPreg_all_imputations <-
  read.csv("imputed_data_O3_perPreg_all_imputations.csv")

imputed_data_O3_perPreg_all_imputations
```

# I. Data Preparation 
## 1) dataset for the univariable analysis

```{r}
#---------------------------------------------
# Remove unnecessary variables for the model
#---------------------------------------------
imputed_data_O3_perPreg_imp1 <- imputed_data_O3_perPreg_imp1 %>% 
  dplyr:: select(-ends_with("_participated"))

# Check NA
sapply(imputed_data_O3_perPreg_imp1, function(x) sum(is.na(x)))

#------------------------------------
#  Prepare the data for the analysis
#------------------------------------
imputed_data_O3_perPreg_imp1 <- imputed_data_O3_perPreg_imp1 %>%
  mutate(across(where(~ is.character(.) | is.factor(.)), 
                ~ as.factor(.))) %>% 
  mutate(
    diplome_ini2 = factor(
      diplome_ini2,
      levels = c("<=Baccalauréat/Aucun/Autre", "Bac+2", ">=Bac+3"),
      ordered = FALSE
    ),
    
    situation_financiere = factor(
      situation_financiere,
      levels = c(
        "Difficult/In debt",
        "Balanced",
        "Comfortable/Very comfortable"
      ),
      ordered = FALSE
    ),
    
    type_endo =
      factor(
        type_endo,
        levels = c(
          "DIE only",
          "SPE only",
          "OMA only",
          "ADENO only",
          "Multiple types",
          "Aucun mentionné/nsp"
        )
      ),
    
    delai_diag_cat_simple = factor(
      delai_diag_cat_simple,
      levels = c("<= 7", "8-13", "13+", "No symptoms"),
      ordered = FALSE
    ),
    
    age_symptome_cat = factor(
      age_symptome_cat,
      levels = c(
        "<= 13",
        "(13;15]",
        "(15;22]",
        "22+",
        "No symptoms"
      ),
      ordered = FALSE
    ),
    
    nb_operation_endo_cat2= factor(
      nb_operation_endo_cat2,
      levels= c("0","1","2",">2")
    ),
    
    indic_ovaire_touche_mod = case_when(
      indic_ovaire_touche %in% c("Ne sait pas", "No surgery") ~
        "No OP/nsp",
      TRUE ~ indic_ovaire_touche
    ),

    age_firstOP_endoadeno_cat = factor(
      age_firstOP_endoadeno_cat,
      levels = c("<28", "28-31", "31>=", "No operation"),
      ordered = FALSE
    ),
    
    nb_mal_sansEndoAde_cat = factor(
      nb_mal_sansEndoAde_cat,
      levels = c("0", "1", "2 ou plus"),
      ordered = FALSE
    )
  )
```

## 2) dataset for the multivariable analysis 

```{r}
#----------------------
#  Imputed Dataframe
#----------------------
imputed_data_O3_perPreg_imp1 <- imputed_data_O3_perPreg_imp1 %>%
  mutate(across(where(~ is.character(.) | is.factor(.)), 
                ~ as.factor(.))) %>% 
  mutate(
    diplome_ini2 = factor(
      diplome_ini2,
      levels = c("<=Baccalauréat/Aucun/Autre", "Bac+2", ">=Bac+3"),
      ordered = FALSE
    ),
    
    situation_financiere = factor(
      situation_financiere,
      levels = c(
        "Difficult/In debt",
        "Balanced",
        "Comfortable/Very comfortable"
      ),
      ordered = FALSE
    ),
    
    type_endo =
      factor(
        type_endo,
        levels = c(
          "SPE only",
          "DIE only",
          "OMA only",
          "ADENO only",
          "Multiple types",
          "Aucun mentionné/nsp"
        )
      ),
    
    stade_dvpt_legere_severe =
      factor(
        stade_dvpt_legere_severe,
        levels = c("legere", "severe", "No surgery", "Non et nsp")
      ),
    
    delai_diag_cat_simple = factor(
      delai_diag_cat_simple,
      levels = c("<= 7", "8-13", "13+", "No symptoms"),
      ordered = FALSE
    ),
    
    age_symptome_cat = factor(
      age_symptome_cat,
      levels = c(
        "<= 13",
        "(13;15]",
        "(15;22]",
        "22+",
        "No symptoms"
      ),
      ordered = FALSE
    ),
    
    nb_operation_endo_cat2= factor(
      nb_operation_endo_cat2,
      levels= c("0","1","2",">2")
    ),
    
    indic_ovaire_touche_mod = case_when(
      indic_ovaire_touche %in% c("Ne sait pas", "No surgery") ~
        "No OP/nsp",
      TRUE ~ indic_ovaire_touche
    ),

    age_firstOP_endoadeno_cat = factor(
      age_firstOP_endoadeno_cat,
      levels = c("No operation","<28", "28-31", "31>="),
      ordered = FALSE
    ),
    
    nb_mal_sansEndoAde_cat = factor(
      nb_mal_sansEndoAde_cat,
      levels = c("0", "1", "2 ou plus"),
      ordered = FALSE
    ),
    
    comorbidities_bin = case_when(
      nb_mal_sansEndoAde_cat =="0" ~ "Non",
      TRUE~ "Oui"
    )
  )


imputed_data_O3_perPreg_imp1 <- imputed_data_O3_perPreg_imp1 %>%
  group_by(id_unik) %>%
  mutate(preg_number = row_number()) %>%
  ungroup()
```

# II. Univariable analysis adjusted for age 

## 1) Model
```{r}
formula_model <- as.formula("as.factor(Complications_cat_perPreg_grouped) ~ 
                            diplome_ini2 +
                            age +
                            gynecologique_diseases +
                            DIE +
                            age_firstOP_endoadeno_cat +
                            indic_ovaire_touche +
                            age_regle_cat_simple +
                            type_preg_all_simple +
                            age_maternal +
                            para +
                            conception_PMA_perPreg +
                            Mode_accouchement")

X <- model.matrix(formula_model, data = imputed_data_O3_perPreg_imp1)


fit <- nomLORgee(
  formula = formula_model,
  id = imputed_data_O3_perPreg_imp1$id_unik,
  repeated = imputed_data_O3_perPreg_imp1$preg_number,
  data = imputed_data_O3_perPreg_imp1
)

fit <- nomLORgee(
   formula = as.factor(Complications_cat_perPreg_grouped) ~ 
     age +  para + diplome_ini2,
   id = imputed_data_O3_perPreg_imp1$id_unik,
   repeated = imputed_data_O3_perPreg_imp1$preg_number,
   data = imputed_data_O3_perPreg_imp1
 )

fit <- nomLORgee(
  formula = as.factor(Complications_cat_perPreg_grouped) ~  diplome_ini2,
  id = imputed_data_O3_perPreg_imp1$id_unik,
  repeated = imputed_data_O3_perPreg_imp1$preg_number,
  data = imputed_data_O3_perPreg_imp1,
  LORstr = "uniform"   # local odds ratio 구조 꼭 지정
)


data(housing)
housing %>% glimpse
```


```{r}
library(multgee)
formula_uni <- as.formula("Complications_cat_perPreg_grouped ~
    diplome_ini2 +
    age +
    
    gynecologique_diseases +
    infertilite_diag +
    autoimmune_diseases +
    nb_mal_sansEndoAde_cat +
    
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    
    age_regle_cat_simple +
    type_preg_all_simple +
    #gravida +
    para +
    conception_PMA_perPreg +
    Mode_accouchement")
fit <- nomLORgee(
  as.factor(Complications_cat_perPreg_grouped) ~
    diplome_ini2 +
    # age +
    
    gynecologique_diseases +
    #infertilite_diag +
    #autoimmune_diseases +
    nb_mal_sansEndoAde_cat +
    
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    age_regle_cat_simple +
    type_preg_all_simple +
    age_maternal +
    #gravida +
    para +
    conception_PMA_perPreg +
    Mode_accouchement,
  
  id=imputed_data_O3_perPreg_imp1$id_unik, 
  data=imputed_data_O3_perPreg_imp1, 
  repeated= imputed_data_O3_perPreg_imp1$preg_number)
```


```{r}
library(sandwich)
#TEST
data <- imputed_data_O3_perPreg_imp1

data$Complications_cat_perPreg_grouped

model_test <- multinom(
  Complications_cat_perPreg_grouped ~
    diplome_ini2 +
    age +
    
    gynecologique_diseases +
    infertilite_diag +
    autoimmune_diseases +
    nb_mal_sansEndoAde_cat +
    
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    
    age_regle_cat_simple +
    type_preg_all_simple +
    gravida +
    para +
    conception_PMA_perPreg +
    Mode_accouchement,
  
  data = data
)


# Robust SE 
robust_vcov <- vcovCL(model_test, cluster = data$id_unik)
robust_results <- coeftest(model_test, vcov = robust_vcov)
?vcovCL
```


## 2) Results Table 

```{r}
data <- imputed_data_O3_perPreg_imp1

# test

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables
outcome_variable <- "Complications_cat_perPreg_grouped"
adjustment_variable <- "age" 

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(data), 
                               c(outcome_variable, adjustment_variable))

#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>%
  tbl_summary(
    by = all_of(outcome_variable),         # Stratify by the outcome
    statistic = list(                     # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%   # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>%
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted multivariable multinomial Regression Table
#-------------------------------------------------------------
# This function fits multinom(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_gee_multinom <- function(predictor, outcome_var, adjust_var, data) {
  adjust_str <- sprintf("`%s`", adjust_var) # If adjust_var is a single string

  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)

  # Fit the model
  model <- tryCatch(
    nnet::multinom(as.formula(formula_str), data = data), 
    error = function(e) {
      warning(paste("Could not fit ordinal regression model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )

  # If model failed, return NULL
  if (is.null(model)) return(NULL)

}





# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_multinom(.x, outcome_variable,
                                             adjustment_variable,
                                             data))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables,
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
          )
        )
    ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted Proportional OR (95% CI)**" # Clarify as Proportional OR
    )
  print("Step 2: Adjusted Proportional Odds Ratios Table Created.")
} else {
  warning("No successful ordinal regression models were fitted for adjusted ORs.",
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
# The 'adjustment_variable' is a single string in this setup, so no paste collapse needed
adjust_str <- adjustment_variable

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      paste0("**Age-Adjusted Proportional Odds Model**")) # Update spanner for clarity
    ) %>%
      modify_caption(
        paste0(
          "**multivariable Ordinal Logistic Regression Analysis for '",
          outcome_variable,
          "' (Adjusted for '", adjust_str, "')**")) # Update caption for clarity

    print("Step 3: Merged Table Created.")

    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '",
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results
#------------------------------------------------------------
O3_tab_multinom_univariable_final <- as_tibble(tbl_final) %>%
  dplyr::select(-"**OR**",-"**95% CI**")  
```



```{r}
data <- imputed_data_O3_perPreg_imp1

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables 
outcome_variable <- "live_birth_bin_atleast_once_all"
adjustment_variable <- "age"

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure outcome variable is suitable for logistic regression (factor or 0/1 numeric)
if (!is.factor(imputed_data_O2[[outcome_variable]]) && !all(imputed_data_O2[[outcome_variable]] %in% c(0, 1, NA))) {
  stop(paste("Outcome variable '", outcome_variable, "' must be a factor or numeric with values 0/1."))
  }

# Convert outcome to factor if it's 0/1 numeric, as tbl_summary needs factor for 'by'
if (is.numeric(imputed_data_O2[[outcome_variable]])) {
    imputed_data_O2 <- imputed_data_O2 %>%
        mutate(!!sym(outcome_variable) := factor(!!sym(outcome_variable)))
    print(paste("INFO: Converted numeric outcome variable '",
                outcome_variable, "' to factor."))
    }

# Ensure adjustment variable exists
if (!all(adjustment_variable %in% names(imputed_data_O2))) {
  stop(paste("Some adjustment variables not found in the dataframe:", 
             paste(setdiff(adjustment_variable, names(imputed_data_O2)), collapse = ", ")))
}
# if (!adjustment_variable %in% names(imputed_data_O2)) {
#   stop(paste("Adjustment variable '", adjustment_variable, "' not found in the dataframe."))
#   }

missing_vars <- setdiff(adjustment_variable, names(imputed_data_O2))
if (length(missing_vars) > 0) {
  stop(paste("Adjustment variable(s) not found in the dataframe:", paste(missing_vars, collapse = ", ")))
  }

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(imputed_data_O2), 
                               c(outcome_variable, adjustment_variable))

if (length(predictor_variables) == 0) {
    stop("No predictor variables found after excluding outcome and adjustment variables.")
  }

print(paste("Outcome:", outcome_variable))
print(paste("Adjustment:", paste(adjustment_variable, collapse=", ")))
print(paste("Predictors:", paste(predictor_variables, collapse=", ")))


#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- imputed_data_O2 %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>% 
  tbl_summary(
    by = all_of(outcome_variable),           # Stratify by the outcome
    statistic = list(                         # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",                           
    digits = list(all_continuous() ~ 1, 
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%     # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>% 
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted Multivariable Logistic Regression Table
#-------------------------------------------------------------
# This function fits glm(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_glm <- function(predictor, outcome_var, adjust_var, data) {
  # Use backticks for safety with non-standard names
  adjust_str <- paste(sprintf("`%s`", adjustment_variable), collapse = " + ")
    # adjust_vars는 벡터
  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)
  
  # Fit the model
  model <- tryCatch(
    glm(as.formula(formula_str), data = data, family = binomial(link = "logit")),
    error = function(e) {
      warning(paste("Could not fit model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )
  
  # If model failed, return NULL
  if (is.null(model)) return(NULL)
  
  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int=TRUE) %>%
    # Keep rows where 'variable' matches the predictor name
    # This correctly handles factor variables which have multiple rows
    modify_table_body(~ dplyr::filter(.x, variable == predictor)) 
  }

# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_glm(.x, outcome_variable,
                                            adjustment_variable,
                                            imputed_data_O2))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables, 
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
            )
          )
      ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted OR (95% CI)**"
    )
  print("Step 2: Adjusted Odds Ratios Table Created.")
} else {
  warning("No successful models were fitted for adjusted ORs.", 
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
adjust_str <- paste(adjustment_variable, collapse = " + ")

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      "**Age-Adjusted Model**")
    ) %>%
      modify_caption(
        paste0(
          "**multivariable Logistic Regression Analysis for '",
          outcome_variable, 
          "' (Adjusted for '", adjust_str, "')**"))

    print("Step 3: Merged Table Created.")
    
    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '", 
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results 
#------------------------------------------------------------
table_res2_final <- as_tibble(tbl_final) %>% 
  dplyr::select(-"**OR**",-"**95% CI**",-"**p-value**")

table_res2_final %>% View
```

## 2) P-trend table 
ordinal variables -> as.numeric :  p-trend ? 
- if the modality "Not_answered" exists,"Not_answered" to NA and then check the p trend 

Question : Differences between results : 
- ordinal variable to numeric variable VS glm(outcome ~ ordinal exposure variable + age, data, family=binomial) 
- .L : linear tendance 

```{r}
imputed_data_O2_2 <- imputed_data_O2 %>%
  mutate(
    diplome_ini2 = factor(
      diplome_ini2,
      levels= c("<=Baccalauréat/Aucun/Autre", "Bac+2", ">=Bac+3"),
      ordered = TRUE
    ) %>% 
      as.numeric(),
    
    situation_financiere = factor(
      situation_financiere,
      levels = c(
        "Difficult/In debt",
        "Balanced",
        "Comfortable/Very comfortable"
        ),
      ordered = TRUE
    ) %>% 
      as.numeric(),
    
    first_preg_age_cat = factor(
      first_preg_age_cat, 
      levels= c("<30","30-34",">=34"), 
      ordered = TRUE) %>% 
      as.numeric(),
    
  age_symptome_cat = factor(
    age_symptome_cat,
    levels = c("No symptoms", "<= 13", "(13;15]", "(15;22]", "22+"),
    ordered = FALSE
    )%>%
    as.numeric(),

  age_firstOP_endoadeno_cat = factor(
    age_firstOP_endoadeno_cat,
    levels = c("No operation","<27", "27-31", "31>="),
    ordered = TRUE
    )%>%
    as.numeric(),
  
  nb_mal_sansEndoAde_cat= factor(
    nb_mal_sansEndoAde_cat,
    levels= c("0","1","2 ou plus"),
    ordered = TRUE
  ) %>% 
    as.numeric(),
  
  delai_diag_cat = factor(
    delai_diag_cat,
    levels = c("No symptoms","<= 5", "(5;10]", "(10;15]", "15+"),
    ordered = FALSE
    )%>% 
    as.numeric()
  )

tbl_descriptive2 <- imputed_data_O2_2 %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>% 
  tbl_summary(
    by = all_of(outcome_variable),           
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",                           
    digits = list(all_continuous() ~ 1, 
                  all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(label ~ "**Characteristic**") %>% 
  bold_labels()

list_of_adj_tables2 <- 
  map(predictor_variables,
      ~fit_adjusted_glm(.x,outcome_variable,
                        adjustment_variable,
                        imputed_data_O2_2))

# Remove any NULL results (from failed models)
list_of_adj_tables2 <-
  list_of_adj_tables2[!sapply(list_of_adj_tables2,is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables2) > 0) {
  tbl_adjusted_ors2 <- tbl_stack(list_of_adj_tables2, 
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
            )
          )
      ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted OR (95% CI)**"
    )
  print("Step 2: Adjusted Odds Ratios Table Created.")
} else {
  warning("No successful models were fitted for adjusted ORs.", 
          call. = FALSE)
  tbl_adjusted_ors2 <- NULL
}

adjust_str <- paste(adjustment_variable, collapse = " + ")

if (!is.null(tbl_adjusted_ors2)) {
    tbl_final2 <- tbl_merge(
      tbls = list(tbl_descriptive2, tbl_adjusted_ors2),
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      "**Age-Adjusted Model**")
    ) %>%
      modify_caption(
        paste0(
          "**multivariable Logistic Regression Analysis for '",
          outcome_variable, 
          "' (Adjusted for '", adjust_str, "')**"))

    print("Step 3: Merged Table Created.")
    
    # --- Display the final table ---
    tbl_final2
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive2 %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '", 
            outcome_variable,"'**"))
    }

tbl_final2

```


**Results** 
- statut_marital_ini : En couple (2.02(1.28,3.15)) , Mariee(2.61(1.57,4.32)) -> CI large
statut_marital_ini_cat :  Single/divorced (0.50 (0.33, 0.76))
- type_endo : OMA only (2.10(1.04,4.25)) -> CI large 
- first_preg_age_cat : >=34 (0.48(0.3,0.76))

# III. multivariable logistic regression

## 1) DAG
```{r}
# DAG : 
library(dagitty)
library(ggdag)
dag_code <- "dag {

  age -> live_birth_bin_atleast_once_all
  age -> statut_marital_ini_cat
  age -> diplome_ini2
  age -> situation_financiere
  age -> situation_emploi_ini2
  age -> operation_endo_bin
  age -> nb_mal_sansEndoAde_cat
  
  nb_mal_sansEndoAde_cat -> live_birth_bin_atleast_once_all
  
  type_endo ->Endo_ade_categorie
  type_endo -> operation_endo_bin
  type_endo -> live_birth_bin_atleast_once_all
  type_endo -> stade_dvpt_legere_severe
  
  operation_endo_bin -> complication_operation
  operation_endo_bin -> live_birth_bin_atleast_once_all
  
  stade_dvpt_legere_severe -> live_birth_bin_atleast_once_all
  stade_dvpt_legere_severe -> operation_endo_bin
  
  first_preg_age_cat -> live_birth_bin_atleast_once_all
  
}"

dag_model <- dagitty(dag_code)

ggdag(dag_model, text_size = 
        3.5, text_col = "darkred", 
      node_size = 5) +
  theme_dag(base_size = 8, base_family = "Nunito") +
  theme(legend.position = "none") +
  scale_fill_manual(values = rep("white", 100)) +  
  scale_color_manual(values = rep("gray", 100))  


adjustmentSets(dag_model, 
               exposure ="statut_marital_ini_cat", 
               outcome = "live_birth_bin_atleast_once_all",
               type= "canonical")
```

## 2) Final Model and Results
```{r}
mod_multinom_multivariable_Complic <-
  nnet::multinom(
      Complications_cat_perID_grouped_mod ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin + 
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_symptome_cat+
      age_regle_cat_simple +
      nb_live_births_all_cat +

      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Csection_perID +

      methode_conception_perID_cat_simple_mod+
      Travail_accouchement_perID+
      
      doul_abdo_avant_1_preg,
    
    data = imputed_data_O3_perPreg_imp1)



# Multicollinearity Check : <2
vif(mod)

# Results 
# 1) summary()
summary(mod)
# 2) tidy(exponentiate=TRUE) 
res_multivar_glm <- tidy(mod, exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_mod <- res_multivar_glm[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_mod <- res_multivar_glm_mod %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_mod
```

 Remove the results for the modality Not_answered
```{r}
res_multivar_glm_mod2 <- res_multivar_glm_mod[-c(10,17),]
res_multivar_glm_mod2
```


## 3) Visualization of results

```{r}
#------------
#  gt table
#------------
library(gt)
res_multivar_glm_mod %>%
  gt() %>%
  tab_header(
    title = md("**multivariable Logistic Regression: **  
    *Factors Associated with Live Birth*")
  ) %>%
  cols_label(
    term = "Variables",
    estimate = "Adjusted OR",
    IC95 = "95% CI",
    p.value = "p-value"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = p.value,
      rows = as.numeric(gsub("<", "", p.value)) < 0.05
    )
  ) %>%
  cols_align(
    align = "left",
    columns = term
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95, p.value)
  )

#---------------
#  Forest plot 
#---------------
res_multivar_glm$term <- factor(res_multivar_glm$term, 
                                levels = rev(res_multivar_glm$term))

ggplot(res_multivar_glm, aes(x = term, y = estimate)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray") +
  coord_flip() +  
  labs(
    title = "Résultats du modèle de régression logistique multivariée",
    x = NULL,
    y = "OR (Odds Ratio avec IC 95%)"
  ) +
  theme_minimal()

#------------------------------
#  Forest plot with the ORs
#------------------------------
library(dotwhisker)
library(grid)

res_multivar_glm_mod$term <- c(
  "Intercept",
  "Single/divorced vs Maried/Couple",
  "Balanced vs Difficult/In debt",
  "Comfortable/Very comfortable vs Difficult/In debt",
  "Age_1stPregnancy: 30-34 vs <30",
  ">=34 vs <30",
  "Age",
  "Nb_comorbidity: 1 vs 0",
  "2+ vs 0",
  "Endo surgery: Not_answered vs No",
  "Yes vs No",
  "type_endo: DIE only vs SPE only",
  "OMA only vs SPE only",
  "ADENO only vs SPE only",
  "Multiple vs SPE only",
  "Aucun mentionné/nsp vs SPE only",
  "Not_answered vs SPE only"
)

dwplot(res_multivar_glm_mod, show_intercept = TRUE) + 
  geom_text(aes(x = 5.2, y = term, label = paste(estimate, "(", IC95, ")")),
            hjust = -0.1, size = 3) +
  theme_minimal() +
  labs(title = "Results of multivariable Logistic Regression",
       x= "OR with 95% CI") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0, size = 12,face = "bold"),
    axis.title.x = element_text(hjust = 0.3, vjust = 0),
    panel.grid = element_blank(),                
    panel.background = element_blank(),           
    axis.line.x =element_blank(), 
    axis.line.y = element_blank(),                
    axis.ticks = element_line(color = "black"),   
    axis.text = element_text(color = "black")     
  ) +
  scale_x_continuous(breaks = 0:5) +              
  coord_cartesian(xlim = c(0, 7), ylim = c(1, 18)) +
  geom_segment(aes(x = 1, xend = 1, y = 0.8, yend = 18),
             color = "black", size = 0.5)+
  annotation_custom(
    grob = textGrob("OR (95%CI)", gp = gpar(fontsize = 10, col = "black",fontface = "bold")),
    xmin = 5, ymin = 17.5)+
  geom_segment(aes(x = -0.3, xend = 5.2, y = 0.5, yend = 0.5),  
             color = "black", size = 0.5)+
  geom_linerange(aes(xmin = conf.low, xmax = conf.high, y = term),
                 color = "red", size = 0.5, linetype=1)
```

 - Forest plot without the modality Not_answered
```{r}
#------------------------------
#  Forest plot with the ORs
#------------------------------
library(dotwhisker)
library(grid)

res_multivar_glm_mod2$term <- c(
  "Intercept",
  "Single/divorced vs Maried/Couple",
  "Balanced vs Difficult/In debt",
  "Comfortable/Very comfortable vs Difficult/In debt",
  "Age_1stPregnancy: 30-34 vs <30",
  ">=34 vs <30",
  "Age",
  "Nb_comorbidity: 1 vs 0",
  "2+ vs 0",
  "Endo surgery: Yes vs No",
  "type_endo: DIE only vs SPE only",
  "OMA only vs SPE only",
  "ADENO only vs SPE only",
  "Multiple vs SPE only",
  "Aucun mentionné/nsp vs SPE only"
)

dwplot(res_multivar_glm_mod2, show_intercept = TRUE) + 
  geom_text(aes(x = 5.2, y = term, label = paste(estimate, "(", IC95, ")")),
            hjust = -0.1, size = 3) +
  theme_minimal() +
  labs(title = "Results of multivariable Logistic Regression",
       x= "OR with 95% CI") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0, size = 12,face = "bold"),
    axis.title.x = element_text(hjust = 0.3, vjust = 0),
    panel.grid = element_blank(),                
    panel.background = element_blank(),           
    axis.line.x =element_blank(), 
    axis.line.y = element_blank(),                
    axis.ticks = element_line(color = "black"),   
    axis.text = element_text(color = "black")     
  ) +
  scale_x_continuous(breaks = 0:5) +              
  coord_cartesian(xlim = c(0, 7), ylim = c(1, 16)) +
  geom_segment(aes(x = 1, xend = 1, y = 0.5, yend = 16),
             color = "black", size = 0.5)+
  annotation_custom(
    grob = textGrob("OR (95%CI)", gp = gpar(fontsize = 10, col = "black",fontface = "bold")), xmin = 5, ymin = 15.4)+
  geom_segment(aes(x = -0.3, xend = 5.2, y = 0.5, yend = 0.5),  
             color = "black", size = 0.5)+
  geom_linerange(aes(xmin = conf.low, xmax = conf.high, y = term),
                 color = "red", size = 0.5, linetype=1)
```


```{r}
library(ggplot2)
library(dplyr)
library(patchwork)
library(gridExtra)


df <- res_multivar_glm_mod %>%
  mutate(
    or_ci = paste0(estimate, " [", IC95, "]"),
    pval_fmt = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value)),
    term = factor(term, levels = rev(term))  # y축 반전
  )

# ▶ Forest plot
p_forest <- ggplot(df, aes(x = estimate, y = term)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray") +
  scale_x_continuous(trans = "log10", limits = c(0.1, 10)) +
  labs(x = "OR (log scale)", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    panel.grid.major.y = element_blank()
  )

df_text <- df %>%
  mutate(y = as.numeric(term)) 

p_table <- ggplot(df_text, aes(y = y)) +
  geom_text(aes(x = 1, label = or_ci), hjust = 0, size = 3) +
  geom_text(aes(x = 3, label = pval_fmt), hjust = 0, size = 3) +
  scale_y_continuous(breaks = df_text$y, labels = NULL) +
  xlim(0.8, 3.8) +
  theme_void() +
  theme(
    plot.margin = margin(5, 20, 5, 0)
  ) +
  annotate("text", x = 1, y = max_y + 1.5, label = "OR [95% CI]", fontface = "bold", hjust = 0, size = 3.5) +
  annotate("text", x = 3, y = max_y + 1.5, label = "p-value", fontface = "bold", hjust = 0, size = 3.5)


# ▶ Combine with patchwork
combined_plot <- p_forest + p_table + plot_layout(widths = c(2.5, 1))
combined_plot

```


