---
title: "03_O1_descriptive_analysis"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 03_O1_descriptive_analysis
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Objective 1, Descriptive analysis 
*Date of creation* : 2025-04-09
*Date of update* : 2025-05-13
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

```{r}
library(tidyverse)
library(gt)
library(gtsummary)
library(readr)
library(writexl)
```

# 0. Import data

```{r}
df_merged_filtered <- read.csv("df_merged_filtered.csv") %>%
  dplyr::select(-X) %>%
  distinct()

df_merged_filtered %>% dim
```

# I. Characterisitics of participants

```{r}
#------------------------
# Principale Population 
#------------------------
library(dplyr)

df_description_population <- df_merged_filtered %>%
  distinct(
    id_unik,
    age,
    
    statut_marital_ini_cat,
    diplome_ini2,
    situation_emploi_ini2,
    situation_financiere,
    
    IMC2,
    Statut_tabac_ini,
    freq_alcool_ini,
    
    score_GAD7_cat,
    score_PHQ9_cat,
    psqif,
    eval_sante_cat,
    IPAQ,
    sc_TBQ_cat,
    
    endo_ade_categorie,
    type_endo,
    stade_dvpt_legere_severe,
    
    nb_mal_sansEndoAde_cat,
    metabolic_syndrome,
    gynecologique_diseases,
    infertility_comorbidite,
    autoimmune_diseases,
    
    age_regle_cat_simple
  ) %>%
  mutate(across(
    .cols = where(is.factor), # 또는 특정 변수들을 선택할 수도 있어
    .fns = ~na_if(as.character(.), "Not_answered")
  )) %>%
  mutate(across(
    .cols = where(is.character),
    .fns = ~na_if(., "Not_answered")
  ))


tab_description_population <- df_description_population %>% 
  dplyr::select(-id_unik) %>%
  tbl_summary(
    by = NULL,
    digits =  ~ 1,
    statistic = list(
      all_continuous() ~"{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  # add_n(
  #   statistic = "{N_nonmiss} ({p_nonmiss}%)",
  #   col_label = "**Number of non-missing observations**",
  #   footnote = FALSE,
  #   last = FALSE
  # ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(id_unik) %>% n_distinct()})**") %>%
  bold_labels() %>% as.tibble()

# Save as Excel file : update 2025-07-18
write_xlsx(tab_description_population, 
           "C:/Stage_Master2/Stage_M2_Complic_Gross/Table_description_population.xlsx")
```

```{r}
#------------------------
#     Sub-population 
#------------------------
data <- df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  filter(nb_live_births_all!=0 & !is.na(total_complications_perID)) %>% 
  distinct(
        id_unik,
        age,

        statut_marital_ini_cat,
        diplome_ini2,
        situation_emploi_ini2,
        situation_financiere,
        
        IMC2,
        Statut_tabac_ini,
        freq_alcool_ini,
        
        score_GAD7_cat,
        score_PHQ9_cat,
        psqif,
        eval_sante_cat,
        IPAQ,
        sc_TBQ_cat,
        
        endo_ade_categorie,
        type_endo,
        stade_dvpt_legere_severe,

        nb_mal_sansEndoAde_cat,
        metabolic_syndrome,
        gynecologique_diseases,
        infertility_comorbidite,
        autoimmune_diseases,

        age_regle_cat_simple
      ) %>%
  mutate(across(
    .cols = where(is.factor), # 또는 특정 변수들을 선택할 수도 있어
    .fns = ~na_if(as.character(.), "Not_answered")
  )) %>%
  mutate(across(
    .cols = where(is.character),
    .fns = ~na_if(., "Not_answered")
  ))

tab_description_subpopulation <- data %>% 
  dplyr::select(-id_unik) %>%
  tbl_summary(
    by = NULL,
    digits =  ~ 1,
    statistic = list(
      all_continuous() ~"{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  # add_n(
  #   statistic = "{N_nonmiss} ({p_nonmiss}%)",
  #   col_label = "**Number of non-missing observations**",
  #   footnote = FALSE,
  #   last = FALSE
  # ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(id_unik) %>% n_distinct()})**") %>%
  bold_labels() %>% as.tibble()

# Save as Excel file : update 2025-07-18
write_xlsx(
  tab_description_subpopulation,
  "C:/Stage_Master2/Stage_M2_Complic_Gross/Table_description_subpopulation.xlsx")
```


# II. Reproductive health
## Per ID
### -Two populations
Live birth at least once : yes/no
Number of pregnancies
Number of live births
Maternal age at the 1st preg and the last preg

```{r}
library(gtsummary);library(gt)

#------------------------------------
#   gtsummary table	  : per patient 
#------------------------------------
# 1. ALL
tbl_grosse_all <- df_merged_filtered %>%
  distinct(
    id_unik,
  
    live_birth_bin_atleast_once_all,

    nb_total_gross,
    nb_live_births_all,
    
    first_preg_age, 
    last_preg_age_mod,
    # Methode_conception_bin_perID,
    # methode_conception_perID_cat_simple_mod,
    # 
    # Mode_accouchement_perID, 
    # Travail_accouchement_perID,
    # 
    # allaitement_YN_atleast,
    # 
    # PriseenCharge_cat_perID, 
    # 
    # total_complications_bin_perID,
    # total_complications_cat_perID,
    # Complications_cat_perID,
    # Complications_cat_perID_grouped
    
  ) %>%
  # mutate(
  #   Complications_cat_perID_grouped_mod =
  #     ifelse(
  #       Complications_cat_perID_grouped == "Hemorr_postpartum_single",
  #       "Other_single",
  #       Complications_cat_perID_grouped
  #     )
  # ) %>% 
  dplyr::select(-id_unik) %>%
  tbl_summary(
    by = NULL,
    digits =  ~ 1,
    statistic = list(
      all_continuous() ~"{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  # add_n(
  #   statistic = "{N_nonmiss} ({p_nonmiss}%)",
  #   col_label = "**Number of non-missing observations**",
  #   footnote = FALSE,
  #   last = FALSE
  # ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(id_unik) %>% n_distinct()})**") %>%
  bold_labels()

# 2. Per modules
tbl_grosse_modules <- df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  filter(nb_live_births_all!=0 & !is.na(total_complications_perID)) %>%
  dplyr::distinct(
    id_unik,
    module,
  
    live_birth_bin_atleast_once_all,

    nb_total_gross,
    nb_live_births_all,
    
    first_preg_age, 
    last_preg_age_mod,
    
    # Methode_conception_bin_perID,
    # methode_conception_perID_cat_simple_mod,
    # 
    # Mode_accouchement_perID, 
    # Travail_accouchement_perID,
    # 
    # allaitement_YN_atleast,
    # 
    # PriseenCharge_cat_perID, 
    # 
    # total_complications_bin_perID,
    # total_complications_cat_perID,
    # Complications_cat_perID,
    # Complications_cat_perID_grouped
    
  ) %>%
  # mutate(
  #   Complications_cat_perID_grouped_mod =
  #     ifelse(
  #       Complications_cat_perID_grouped == "Hemorr_postpartum_single",
  #       "Other_single",
  #       Complications_cat_perID_grouped
  #     )
  # ) %>% 
  dplyr::select(-id_unik) %>%
  tbl_summary(
    by = module,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(id_unik) %>% n_distinct()})**") %>%
  bold_labels()

# 3. Merge 
tbl_twopops_combined <- tbl_merge(
  tbls = list(tbl_grosse_all, tbl_grosse_modules),
  tab_spanner = c("**Overall Population**", "**By Module**")
); tbl_twopops_combined
```
### - Sous-population, n=310 
methode de conception 
mode d'accouchement
phase de travail
allaitement
Prise en charge
Complications 

```{r}
tbl_souspop_modules <- df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  filter(nb_live_births_all!=0 & !is.na(total_complications_perID)) %>%
  dplyr::distinct(
    id_unik,

    Methode_conception_bin_perID,
    methode_conception_perID_cat_simple_mod,
    
    Mode_accouchement_perID, 
    Travail_accouchement_perID,
    
    allaitement_YN_atleast,
    
    PriseenCharge_cat_perID, 
    PriseenCharge_bin,
    
    total_complications_bin_perID,
    total_complications_cat_perID,
    Complications_cat_perID,
    Complications_cat_perID_grouped
  ) %>%
  mutate(
    Complications_cat_perID_grouped_mod =
      ifelse(
        Complications_cat_perID_grouped == "Hemorr_postpartum_single",
        "Other_single",
        Complications_cat_perID_grouped
      )
  ) %>% 
  dplyr::select(-id_unik) %>%
  tbl_summary(
    by = NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  modify_header(label = "**Variable**") %>% 
  bold_labels()

tbl_souspop_modules
```




## Per Preg
### - Two populations
age_maternal
type_preg_all
live_birth_bin_all

```{r}
tab1 <- df_merged_filtered %>%
  distinct(
    IdPregnancy,
    age_maternal,
    type_preg_all,
    type_preg_all_simple,
    live_birth_bin_all
  ) %>%
  dplyr::select(-IdPregnancy) %>%
  tbl_summary(
    by = NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list()
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Reproductive history in two populations**") %>% 
  bold_labels()

tab2 <-  df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  filter(nb_live_births_all!=0 & !is.na(total_complications_perID)) %>%  
  distinct(
    IdPregnancy,
    age_maternal,
    type_preg_all,
    type_preg_all_simple,
    live_birth_bin_all
  ) %>%
dplyr::select(-IdPregnancy) %>%
  tbl_summary(
    by = NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list()
  ) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels()

# 3. Merge 
tbl_twopops_preg_combined <- tbl_merge(
  tbls = list(tab1, tab2),
  tab_spanner = c("**Overall Population**", "**By Module**")
); tbl_twopops_preg_combined
```

### - Sous-population, n= 383 pregnancies
Methode de conception
Mode d'accouchement
Phase de travail
Allaitement
Prise en charge
Complications obstétricales

```{r}
tbl_souspop_preg_modules<- df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  filter(nb_live_births_all!=0 & !is.na(total_complications_perID)) %>%
  distinct(
    IdPregnancy,
    
    Methode_conception_bin_perPreg,
    methode_conception_perPreg_cat_simple_mod,
    
    Mode_accouchement,
    Travail_accouchement,
    
    Allaitement2,
    mois_allaitement,
    PriseenCharge_cat_perPreg,
    
    total_complications_bin_perPreg,
    total_complications_cat_perPreg,
    Complications_cat_perPreg,
    Complications_cat_perPreg_grouped
  ) %>%
  mutate(
    Complications_cat_perPreg_grouped_mod =
      ifelse(
        Complications_cat_perPreg_grouped == "Hemorr_postpartum_single",
        "Other_single",
        Complications_cat_perPreg_grouped
      )
  ) %>% 
dplyr::select(-IdPregnancy) %>%
  tbl_summary(
    by = NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list()
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Reproductive history (pregnancy) among the sub-population**") %>% 
  bold_labels()

tbl_souspop_preg_modules

# Median

new_df <- df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  filter(nb_live_births_all!=0 & !is.na(total_complications_perID)) %>% 
  distinct(IdPregnancy, mois_allaitement)

summary(new_df$mois_allaitement)
mean(new_df$mois_allaitement, na.rm=TRUE) %>% round(1) # 4.9
median(new_df$mois_allaitement, na.rm=TRUE) # 2
min(new_df$mois_allaitement, na.rm=TRUE) # 0
max(new_df$mois_allaitement, na.rm=TRUE) # 54
sd(new_df$mois_allaitement, na.rm = TRUE) %>% round(1) # 7.9


df_merged_filtered$operation_endo_bin %>% table(useNA = "a") %>% prop.table() %>% addmargins
```

```{r}
df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  filter(nb_live_births_all!=0 & !is.na(total_complications_perID)) %>%
  distinct(id_unik,PriseenCharge_bin) %>% 
  group_by(PriseenCharge_bin) %>% 
  summarise(n=n()) %>% 
  mutate(somme= sum(n),
         prop= round(n/somme,2)) %>% 
  ungroup()
```



## **Common variables for two populations 
### => Annexe : Combined Table for the two populations

```{r}
final_two_tbl <- tbl_merge(
  tbls = list(tbl_twopops_combined, tbl_twopops_preg_combined),
  tab_spanner = c("**Par Femme**", "**Par Grossesse**")
)
final_two_tbl_df <- final_two_tbl %>%  as.tibble()


# Save : update 250718
write_xlsx(final_two_tbl_df,"C:/Stage_Master2/Stage_M2_Complic_Gross/O1_twopopulations_table.xlsx")
```


## **Sous popoulation for preg + poulation 

```{r}
final_souspop_combined_tbl <- tbl_merge(
  tbls = list(tbl_souspop_modules, tbl_souspop_preg_modules),
  tab_spanner = c("**Par Femme**", "**Par Grossesse**")
)
final_souspop_combined_tbl_df <- final_souspop_combined_tbl %>%  as.tibble()


# Save : update 250718
write_xlsx(final_souspop_combined_tbl_df,"C:/Stage_Master2/Stage_M2_Complic_Gross/O1_souspop_combinded_table.xlsx")
```


# ----------------------------------------------------
# IV.  Descriptive analysis

## 1. Number of participants and pregnancies according to pregnancy modules

  Participants : 1113 for the live birth outcome 
  Participants in new module : 378 
  
  Pregnancies : 1575 (new 468/1575= 29.7%, old 1107/1575= 70.3%)
  Pregnancies in new module : 468
  
```{r}
# Participants : 1113
df_merged_filtered %>% distinct(id_unik) %>% n_distinct()

# Participants in new module : 378
df_merged_filtered %>%filter(module=="Nouveau") %>% distinct(id_unik) %>% n_distinct()

# Pregnancies : 1575
df_merged_filtered %>% distinct(IdPregnancy) %>% n_distinct()

# Pregnancies in new module : 468
df_merged_filtered %>% filter(module=="Nouveau") %>%  distinct(IdPregnancy) %>% n_distinct()




# Pregnancies
df_merged_filtered %>% 
  distinct(IdPregnancy, module) %>% 
  count(module) %>% 
  mutate(total = sum(n), 
         prop = round(n / total * 100, 1))
```



### - Table 2 : Pregnancy characteristics 

#### 1) Total number of pregnancies
- Total = 1575
- New module = 468
- Old module = 1107
```{r}
df_merged_filtered %>% distinct(id_unik,nb_total_gross, nb_total_gross_NewModule, nb_total_gross_OldModule) %>% 
  summarise(total_gross = sum(nb_total_gross),
            newmodule_total_gross = sum(nb_total_gross_NewModule, na.rm = TRUE),
            oldmodule_total_gross = sum(nb_total_gross_OldModule, na.rm=TRUE))
```

#### 2) Frequency : term, incomplete pregnancy

Per participant (2nd Objective's outcome)
- At least once, live birth  : Yes (902/1113,81%), No (211/1113,19%)

```{r}
df_merged_filtered %>% 
  distinct(id_unik,live_birth_bin_atleast_once_all) %>% 
  group_by(live_birth_bin_atleast_once_all) %>% 
  summarise(n=n()) %>% 
  mutate(somme= sum(n),
         proportion= round((n/somme)*100,1))
```

Per pregnancy
-     All modules  : Term = 474(/1575, 70%), Incomplete = 1101(/1575, 30%)
- Only new modules : Term = 369(/468, 79%), Incomplete = 99(/468, 21%)

```{r}
df_merged_filtered %>% 
  distinct(id_unik,IdPregnancy,live_birth_bin_all) %>% 
  group_by(live_birth_bin_all) %>% 
  summarise(n=n(),.groups = "drop") %>% 
  mutate(somme= sum(n),
         proportion= round((n/somme)*100,1))
  
df_merged_filtered %>% 
  distinct(id_unik,IdPregnancy,live_birth_bin_onlynew) %>%
  group_by(live_birth_bin_onlynew) %>% 
  summarise(n=n(),.groups = "drop") %>%
  filter(!is.na(live_birth_bin_onlynew)) %>% 
  mutate(somme= sum(n),
         proportion= round((n/somme)*100,1))
```

#### 3) maternal age at the first and last pregnancy 

- mean maternal age at the first pregnancy = 31.23
- mean maternal age at the last pregnancy = 33.9

- mean maternal age : 32.14584
```{r}
df_merged_filtered %>% distinct(id_unik, first_preg_age, last_preg_age_mod) %>% 
  summarise(mean_first_age= mean(first_preg_age),
            mean_last_age= mean(last_preg_age_mod, na.rm=TRUE))


df_merged_filtered %>% distinct(IdPregnancy, age_maternal) %>% 
  summarise(mean_age_maternal = mean(age_maternal))
```

#### 3) For Each conception method

- From 'methode_conception_perID_cat_simple_mod' including two informations : conception method and the operation for the infertility, so even old module can have conception method information, in terms of only operation. 


```{r}
# df_merged_filtered %>% 
#   distinct(id_unik, methode_conception_perID_cat_simple_mod) %>% 
#   group_by(methode_conception_perID_cat_simple_mod) %>% 
#   summarise(n=n(), .groups = "drop") %>% 
#   mutate(somme=sum(n),
#          proportion= paste0(round(n/somme*100)," %"))

df_merged_filtered %>% 
  filter(module=="Nouveau") %>% 
  distinct(id_unik, methode_conception_perID_cat_simple_mod) %>% 
  group_by(methode_conception_perID_cat_simple_mod) %>% 
  summarise(n=n(), .groups = "drop") %>% 
  mutate(somme=sum(n),
         proportion= paste0(round(n/somme*100)," %"))
```


- Operation : methode_perID_Operation_mod (in wide format)
1) All:  224 / 1113 = 20.1 %
2) New module:  72 / 378 = 19 %
3) Non-NA only:  224 / 530 = 42.3 %

```{r}
test <- df_merged_filtered %>% 
  left_join(
    grossesse_mod2 %>%
      distinct(id_unik, .keep_all = TRUE) %>%
      dplyr::select(id_unik, methode_perID_Naturelle, methode_perID_Operation,
             methode_perID_Medicament_sans_PMA, methode_conception_long_perID),
    by = "id_unik"
  ) %>%
  dplyr::select(id_unik, module, methode_perID_IUI, methode_perID_Naturelle, 
         methode_perID_IVF, methode_perID_Operation,
         methode_perID_Medicament_sans_PMA, methode_conception_long_perID,
         op_infertility_perID, methode_conception_perID_cat_simple_mod) %>%
  
  # Add the information of operation for the infertility to the conception method
  mutate(
    methode_perID_Operation_mod = case_when(
      (is.na(methode_perID_Operation) | methode_perID_Operation == FALSE) & 
        op_infertility_perID == "for_infertility" ~ TRUE,
      methode_perID_Operation == TRUE ~ TRUE,
      op_infertility_perID != "for_infertility" ~ methode_perID_Operation,
      TRUE ~ NA 
    )
  )

n_total <- n_distinct(test$id_unik)
n_new <- n_distinct(test %>% filter(module == "Nouveau") %>% pull(id_unik))

n_operation_mod <- test %>% 
  distinct(id_unik, methode_perID_Operation_mod) %>% 
  filter(methode_perID_Operation_mod == TRUE) %>% 
  nrow()

n_operation_mod_new <- test %>%
  filter(module == "Nouveau") %>%
  distinct(id_unik, methode_perID_Operation_mod) %>%
  filter(methode_perID_Operation_mod == TRUE) %>%
  nrow()

n_operation_mod_nonNA <- test %>% 
  distinct(id_unik, methode_perID_Operation_mod) %>% 
  filter(!is.na(methode_perID_Operation_mod)) %>%
  nrow()

cat("1) All: ", n_operation_mod, "/", n_total, "=", 
    round(100 * n_operation_mod / n_total, 1), "%\n")
cat("2) New module: ", n_operation_mod_new, "/", n_new, "=", 
    round(100 * n_operation_mod_new / n_new, 1), "%\n")
cat("3) Non-NA only: ", n_operation_mod, "/", n_operation_mod_nonNA, "=", 
    round(100 * n_operation_mod / n_operation_mod_nonNA, 1), "%\n")

```




```{r}


# After adding the information : 224 participants among 1113  
# ---> IF only new module, 72 participants among 378 
#%>%filter(module=="Nouveau") 
test %>% distinct(id_unik, methode_perID_Operation_mod) %>%
  dplyr::select(methode_perID_Operation_mod) %>% table(useNA = "a") %>% addmargins

# Just using the new module's information : 14 participants among 1113
# ---> IF only new module, 14 participants among 378 
test %>% distinct(id_unik, methode_perID_Operation) %>%
  dplyr::select(methode_perID_Operation) %>% table(useNA = "a") %>% addmargins

# Question : for calculating the proportion data, le total number would be 1113(all population) or 378 (sub population who answered to the new module)

# 1) Adding the information of operation from chirurgie dataset
# - 224/1113  ou 72/378
# - or excluding all NA (respectively 584 and 735) : 224/(1113-583) ou 14/(1113-735)

# 2) Just using the new module's information : 14 participants among 1113
# - 14/1113 ou 14/378
# - or excluding all NA (735 and 0) : 14/(1113-735) 


##########################################################################
test %>%
  filter(module=="Nouveau") %>%
  dplyr::select(id_unik, methode_conception_long_perID) %>%
  distinct() %>% count(methode_conception_long_perID, name = "nb_patients") %>%
  mutate(somme= sum(nb_patients),
         proportion= round(nb_patients/somme*100,1))

test %>%
  filter(module=="Nouveau") %>%
  dplyr::select(id_unik,op_infertility_perID) %>%
  group_by(op_infertility_perID) %>%
  summarise(n=n(), .groups = "drop")

df_merged_filtered %>% distinct(id_unik,op_infertility_perID) %>% dplyr::select()
test %>% distinct(id_unik, methode_conception_perID_cat_simple_mod) %>%
  group_by(methode_conception_perID_cat_simple_mod) %>%
  summarise(n=n())
###################################################################################
```

#### Type of pregnancy outcome in all population : Fullterm, Preterm, F-C, IVG/IMG, Autre 

All population 
- Fullterm : 62.3
- Preterm : 7.6
- Fausse-couche : 20.1
- IVG/IMG : 7.6
- Autre : 2.3


```{r}
df_merged_filtered %>% distinct(id_unik, IdPregnancy,type_preg_all_simple) %>% 
  group_by(type_preg_all_simple) %>%
  summarise(n=n(), .groups = "drop") %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))

# # New module
# df_merged_filtered %>% 
#   filter(module=="Nouveau") %>% 
#   distinct(id_unik, IdPregnancy,type_preg_all_simple) %>% 
#   group_by(type_preg_all_simple) %>%
#   summarise(n=n(), .groups = "drop") %>% 
#   mutate(somme=sum(n),
#          proportion = paste0(round(n/somme*100,1)," %"))
```

##### Delivery mode in the sub population

- Per participant : Exprience of delivery mode for the participants who gave birth an infant at least once

  Both : 7/321 (2.2%)
  Only cesarean : 82/321 (25.5%)
  Only VD : 225/321 (70.1%)
```{r}
df_merged_filtered %>%
  filter(module=="Nouveau")%>% 
  distinct(id_unik,Mode_accouchement_perID) %>% 
  count(Mode_accouchement_perID) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))

df_merged_filtered %>%
  filter(module=="Nouveau"& live_birth_bin_atleast_once_all=="Oui")%>% 
  distinct(id_unik,Mode_accouchement_perID) %>% 
  count(Mode_accouchement_perID) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))
```

- Per pregnancy :
 the number and proportion of delivery mode used among the term pregnancies
 
  Cesarean : 99/369 (26.3%)
  VD : 278/369 (73.7%)
```{r}

df_merged_filtered %>%
  filter(module=="Nouveau"& status_terme_all == "Terme") %>%  
  count(Mode_accouchement) %>%
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))


df_merged_filtered %>%
  filter(module=="Nouveau") %>%  
  count(Mode_accouchement) %>%
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))
```

#### Type of labor
- Per participant : Exprience of delivery mode for the *participants who gave birth an infant at least once*

  No : 35/321 (10.9%)
  Induced labour : 84/321 (26.2%)
  Natural labour : 173/321 (53.9%)
  Mixed : 7/321 (2.2%)
  
```{r}
df_merged_filtered %>%
  filter(module=="Nouveau")%>% 
  distinct(id_unik,Travail_accouchement_perID) %>% 
  count(Travail_accouchement_perID) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))


df_merged_filtered %>%
  filter(module=="Nouveau"& live_birth_bin_atleast_once_all=="Oui")%>% 
  distinct(id_unik,Travail_accouchement_perID) %>% 
  count(Travail_accouchement_perID) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))
```

- Per pregnancy :
 the number and proportion of delivery mode used *among the term pregnancies*
 
  Pas de travail	42/369	(11.4 %)	
  Travail provoqué	110/369	(29.8 %)	
  Travail spontané	217/369	(58.8 %)
  
```{r}
df_merged_filtered %>%
  filter(module=="Nouveau") %>%  
  count(Travail_accouchement) %>%
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))


df_merged_filtered %>%
  filter(module=="Nouveau"& status_terme_all == "Terme") %>%  
  count(Travail_accouchement) %>%
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))
```

#### Care taken among women who exprienced the incomplete pregnancy in new module

  Absence : 17/81 (21%)
  Multiple : 19/81 (7.4%)
  Only surgical : 16/81(19.8%)
  Only medications : 20/81(24.7%)
  Only psychological : 2/81 (2.5%)
  NA : 1/81 (1.2%)

```{r}
# per pregnancy 
df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  distinct(id_unik, PriseenCharge_bin) %>% 
  count(PriseenCharge_bin) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))

# Per pregnancy, incomplete pregnancy and complete pregnancy 
df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  group_by(status_terme_all) %>% 
  distinct(id_unik,PriseenCharge_bin) %>% 
  count(PriseenCharge_bin) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))

df_merged_filtered %>%
  filter(module=="Nouveau"& status_terme_all=="Incomplete")%>% 
  distinct(id_unik,PriseenCharge_cat_perPreg) %>% 
  count(PriseenCharge_cat_perPreg) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))

# # In all modules
# df_merged_filtered %>%
#   filter(status_terme_all=="Incomplete")%>% 
#   distinct(id_unik,PriseenCharge_cat_perPreg) %>% 
#   count(PriseenCharge_cat_perPreg) %>% 
#   mutate(somme=sum(n),
#          proportion = paste0(round(n/somme*100,1)," %"))
```

#### Frequency of obstetric complications exprienced in new module

-  Number of experienced complications per participant
  0<1<2+
```{r}
df_merged_filtered %>%
  filter(module=="Nouveau")%>% 
  distinct(id_unik,total_complications_cat_perID) %>% 
  count(total_complications_cat_perID) %>% 
  mutate(somme=sum(n),
         proportion = paste0(round(n/somme*100,1)," %"))
```

-  Number of experienced complications per participant
  0<1<2+
  
```{r}

```
  
  


#### ) nombre d’enfants nés vivants
Per participant 

```{r}
df_merged_filtered %>%
  distinct(id_unik, nb_enfts_vivants_ini_cat) %>%
  group_by(nb_enfts_vivants_ini_cat) %>%
  summarise(n = n(),.groups="drop") %>% 
  mutate(
    total= sum(n),
    proportions= round(n/total,3)*100
  )
```



각 수태 방법별 총 임신율 및 만삭 임신율도 계산

```{r}
# per preg n= 468
df_merged_filtered %>%
  filter(module=="Nouveau") %>% 
  distinct(IdPregnancy, status_terme_all,methode_conception_perPreg_cat_simple) %>% 
  group_by(methode_conception_perPreg_cat_simple) %>% 
  summarise(n = n(),.groups="drop") %>% 
  mutate(
    total= sum(n),
    proportions= round(n/total,3)*100
  )
```




#### Croisement 

```{r}
tab <- table(df_merged_filtered$Travail_accouchement_perID,
      df_merged_filtered$age_regle_cat)

chisq.test(tab, correct = TRUE)
```














```{r}
#devtools::install_github("ddsjoberg/gtsummary")
#library(gtsummary)

#--------------------------------------------------------------
# Per module  (All pregnancies)
# Total number of live births/Total number of all pregnancies
#--------------------------------------------------------------
df_merged_filtered %>%
  mutate(
    nb_live_births_old = 
      ifelse(module == "Ancien", 
             nb_live_births_all - nb_live_births_onlynew, 
             NA)
  ) %>%
  distinct(
    id_unik,
    module,
    nb_total_gross,
    nb_total_gross_NewModule,
    nb_total_gross_OldModule,
    nb_live_births_all,
    nb_live_births_onlynew,
    nb_live_births_old
  ) %>%  
  group_by(module) %>%
  summarise(
    somme_total_gross = sum(
      case_when(
        module == "Nouveau" ~ nb_total_gross_NewModule,
        module == "Ancien" ~ nb_total_gross_OldModule,
        TRUE ~ nb_total_gross
      ), na.rm = TRUE
    ),
    somme_liveBirths = sum(
      case_when(
        module == "Nouveau" ~ nb_live_births_onlynew,
        module == "Ancien" ~ nb_live_births_old,
        TRUE ~ nb_live_births_all
      ), na.rm = TRUE
    ),
    prop = round(somme_liveBirths / somme_total_gross * 100, 2)
  ) %>%
  bind_rows(
    df_merged_filtered %>%
      distinct(id_unik, nb_total_gross, nb_live_births_all) %>%
      summarise(
        somme_total_gross = sum(nb_total_gross, na.rm = TRUE),
        somme_liveBirths = sum(nb_live_births_all, na.rm = TRUE),
        prop = round(somme_liveBirths / somme_total_gross * 100, 2)
      ) %>%
      mutate(module = "Overall")
  ) 

# QUESTIONS :  all population VS population who responded to the new module
# - MAIS THERE ARE patients who responded to two modules 

#------------------------------------------
#  Per modules (all pregnancies)
# Incomplete pregnancy rate
#------------------------------------------
df_merged_filtered %>%
  distinct(id_unik, IdPregnancy, module, status_terme_all) %>%
  select(-IdPregnancy) %>%
  group_by(module) %>%
  summarise(
    total_pregnancies = n(),
    incomplete_pregnancies = sum(status_terme_all == "Incomplete"),
    incomplete_rate = incomplete_pregnancies / total_pregnancies * 100
  ) %>%
  mutate(incomplete_rate = round(incomplete_rate, 2)) %>%
  bind_rows(
    df_merged_filtered %>%
      distinct(id_unik, IdPregnancy, status_terme_all) %>%
      select(-IdPregnancy) %>%
      summarise(
        total_pregnancies = n(),
        incomplete_pregnancies = sum(
          status_terme_all == "Incomplete"),
        incomplete_rate = 
          incomplete_pregnancies / total_pregnancies * 100
      ) %>%
      mutate(
        module = "Overall",
        incomplete_rate = round(incomplete_rate, 2)
      )
  ) %>%
  mutate(module = factor(
    module, levels = c("Nouveau", "Ancien", "Overall"))) %>%
  arrange(module, desc(incomplete_rate))

  
#------------------------------------
#   gtsummary table	  : per patient 
#------------------------------------
# 1. ALL
tbl_grosse_all <- df_merged_filtered %>%
  distinct(
    id_unik,
  
    #status_terme_all, # incomplete et terme
    #type_preg_all,
    
    nb_total_gross,
    #nb_live_births_all,
    taux_LB_ALL_all,
    
    #age_maternal, 
    first_preg_age, 
    last_preg_age_mod,
    
    Methode_conception_bin_perID,
    methode_conception_perID_cat_simple,
    
    Mode_accouchement_perID, 
    Travail_accouchement_perID,
  
   # nb_nevivants_parpt,
    
    allaitement_YN_atleast,
    
    #PriseenCharge_bin,
    PriseenCharge_cat_perID, 
    
    Complications_cat_perID,
    #total_complications,
    total_complications_bin,
    total_complications_cat 
    
  ) %>%
  mutate(nb_total_gross = nb_total_gross %>% as.character() %>% as.numeric()) %>%
  select(-id_unik) %>%
  tbl_summary(
    by = NULL,
    digits =  ~ 1,
    statistic = list(
      all_continuous() ~"{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  add_n(
    statistic = "{N_nonmiss} ({p_nonmiss}%)",
    col_label = "**Number of non-missing observations**",
    footnote = FALSE,
    last = FALSE
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(id_unik) %>% n_distinct()})**") %>%
  bold_labels()

# 2. Per modules
tbl_grosse_modules <- df_merged_filtered %>%
  distinct(
    id_unik,
    module,
    #status_terme_all, # incomplete et terme
    #type_preg_all,
    
    nb_total_gross,
    #nb_live_births_all,
    taux_LB_ALL_all,
    
    #age_maternal, 
    first_preg_age, 
    last_preg_age_mod,
    
    Methode_conception_bin_perID,
    methode_conception_perID_cat_simple,
    
    Mode_accouchement_perID, 
    Travail_accouchement_perID,
  
    #nb_nevivants_parpt,
    
    allaitement_YN_atleast,
    
    PriseenCharge_bin,
    PriseenCharge_cat_perID, 
    
    Complications_cat_perID,
    #total_complications,
    total_complications_bin,
    total_complications_cat 
    
  ) %>%
  mutate(nb_total_gross = nb_total_gross %>% as.character() %>% as.numeric()) %>% 
  select(-id_unik) %>%
  tbl_summary(
    by = module,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(id_unik) %>% n_distinct()})**") %>%
  bold_labels()

# 3. Merge 
tbl_grosse_combined <- tbl_merge(
  tbls = list(tbl_grosse_all, tbl_grosse_modules),
  tab_spanner = c("**Overall Population**", "**By Module**")
); tbl_grosse_combined


#----------------------------------------
#   gtsummary table	  : per pregnancy 
#----------------------------------------
tbl_grosse_all_perPreg <- df_merged_filtered %>%
  distinct(
    IdPregnancy,
  
    status_terme_all,
    type_preg_all,
    
    age_maternal, 
    
    Methode_conception_bin_perPreg,
    methode_conception_perPreg_cat, 
    methode_conception_perPreg_cat_simple,
    
    Mode_accouchement,

    Travail_accouchement,

    mois_allaitement,

    PriseenCharge_cat_perPreg,
    
    Complications_cat_perPreg
    ) %>%
  select(-IdPregnancy) %>%
  tbl_summary(
    by = NULL,
    digits =  ~ 1,
    statistic = list(
      all_continuous() ~"{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(IdPregnancy) %>% n_distinct()})**") %>%
  bold_labels()

tbl_grosse_modules_perPreg <- df_merged_filtered %>%
  distinct(
    IdPregnancy,
    module,

    status_terme_all,
    type_preg_all,
    
    age_maternal, 
    
    Methode_conception_bin_perPreg,
    methode_conception_perPreg_cat, 
    methode_conception_perPreg_cat_simple,
    
    Mode_accouchement,

    Travail_accouchement,

    mois_allaitement,

    PriseenCharge_cat_perPreg,
    
    Complications_cat_perPreg
    ) %>%
  select(-IdPregnancy) %>%
  tbl_summary(
    by = module,
    digits =  ~ 1,
    statistic = list(
      all_continuous() ~"{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    missing_stat = "{N_miss} ({p_miss}%)",
    label = list(
    )
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Characteristics of Patients Related to Pregnancy (n={df_merged_filtered %>% distinct(IdPregnancy) %>% n_distinct()})**") %>%
  bold_labels()


# 3. Merge 
tbl_grosse_combined_perPreg <- tbl_merge(
  tbls = list(tbl_grosse_all_perPreg, tbl_grosse_modules_perPreg),
  tab_spanner = c("**Overall Population**", "**By Module**")
); tbl_grosse_combined_perPreg


#------------------------------------------------------------
#   LBR per different types of conception method
#
# Because one pregnancy can have several conception methods
#------------------------------------------------------------
# Question : all modules vs only new module(Reselect the population or just ignore the informations of conception method)
# - ALL MODULES
df_merged_filtered %>%
  distinct(id_unik,
           nb_total_gross,
           nb_live_births_all,
           methode_conception_perID_cat_simple) %>%
  group_by(methode_conception_perID_cat_simple) %>%
  summarise(
    somme_total_gross = sum(nb_total_gross),
    # total number of pregnancies
    somme_liveBirths = sum(nb_live_births_all),
    # total number of Live birth
    LBR_per_ConceptionMethod = round(somme_liveBirths / somme_total_gross, 2)  # LBR
  ) %>%
  mutate(
    AUTOTAL_Gross = sum(somme_total_gross),
    AUTOTAL_LB= sum(somme_liveBirths),
    prop_grossesse_par_methode = round(somme_total_gross / AUTOTAL_Gross, 3),
    LBR_par_methode= round(somme_liveBirths/AUTOTAL_LB,3)
  )

# - New module
df_merged_filtered %>% 
  filter(module=="New") %>%
  distinct(id_unik, nb_total_gross_NewModule, nb_live_births_onlynew, methode_conception_perID_cat_simple) %>%
  group_by(methode_conception_perID_cat_simple) %>% 
  summarise(
    somme_total_gross = sum(nb_total_gross),  # total number of pregnancies
    somme_liveBirths = sum(nb_live_births_all),  # total number of Live birth
    LBR_per_ConceptionMethod = round(somme_liveBirths / somme_total_gross, 2)  # LBR
  ) %>% 
  mutate(
    AUTOTAL_Gross = sum(somme_total_gross),  
    prop_grossesse_par_methode = round(somme_total_gross / AUTOTAL_Gross, 3)  
  )

grossesse_mod2 %>% filter(id_unik=="01D8AB06EB368B00A11F")

```
TEST
```{r}
sansgrossesseencours_final %>%
  distinct(
    id_unik,
    IdPregnancy,
    status_terme_all,
    # module,
    live_birth_bin_atleast_once_all,
    live_birth_bin_atleast_once_onlynew,
    # live_birth_bin_all,
    # live_birth_bin_onlynew,
    # type_preg_all,
    # type_preg_onlynew,
    # nb_live_births_all,
    # nb_live_births_onlynew,
    # nb_total_gross,
    # nb_total_gross_NewModule,
    # nb_total_gross_OldModule,
    # taux_LB_ALL_all,
    # taux_LB_ALL_onlynew,
    # taux_LB_NEWMODULE,
    # age_maternal,
    # first_preg_age,
    # last_preg_age_mod,
    # Methode_conception_bin_perID,
    # Methode_conception_bin_perPreg,
    # methode_conception_perID_cat,
    # methode_conception_perID_cat_simple,
    # methode_conception_perPreg_cat,
    # methode_conception_perPreg_cat_simple,
    # Mode_accouchement,
    # mode_Césarienne,
    # Mode_accouchement,
    # Travail_accouchement ,
    # Travail_accouchement_cat,
    # allaitement_YN,
    # PriseEnCharge_bin,
    # PriseenCharge_long,
    # total_complications_bin ,
    # total_complications,
    # total_complications_cat
  ) %>%  #Complication_value,
  select(-id_unik, -IdPregnancy) %>%
  tbl_summary(
    by = NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    label = list()
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption(glue::glue(
    "**Pregnancies characteristics (n={nrow(df_merged_filtered)})**"
  )) %>%
  bold_labels()
```


## Table 2 : Characteristics related to the period and diseases (Endometriosis and Adenomyosis)

```{r}
# ALL POPULATION
tbl2_all <-df_merged_filtered%>% 
  distinct(
    id_unik,
    #module,
    
    # Règles 
    age_regles_ini,
    age_regle_cat,
    first_rapport,
    
    # Endo/Adeno
    age_symptome, 
    age_symptome_cat,
    delai_diag_cat,
    signe_symptome_premiere_consult,
    type_endo,
    family_atcd_EndoAdeno, 
    family_atcd_doul_pelv,
    Nb_specialites_consult_cat,
    stade_dvpt_maladie, 
    stade_dvpt_legere_severe,
    motifs_operation1_cat, 
    motifs_operation2_cat, 
    indic_ovaire_touche,
    num_localisation_endo,
    localisation_final, 
    taille_kyste_IRM_max,
    tercile_taille_kyste_IRM,trompe_touche2,
    nb_operation_endo, 
    nb_operation_endo_cat
  )%>%
  mutate(stade_dvpt_maladie=
           factor(stade_dvpt_maladie,
                  levels= c("Non", "Je ne sais pas", 
                           "Stade I","Stade II",
                           "Stade III","Stade IV"),
                  ordered = TRUE))%>%
  select(-id_unik)%>%
  tbl_summary(
    by =NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    label = list()
  )%>%
  #add_overall() %>% 
  modify_header(label = "**Variable**") %>%
  modify_caption(
    "**Tableau 2. Endometriosis/Adenomyosis chracteristics (n=1107)**"
  ) %>%
  bold_labels()  ; tbl2_all

# TWO DIFFERENTS POPULATIONS by Modules
tbl2_modules <-df_merged_filtered%>% 
  distinct(
    id_unik,
    module,
    
    # Règles 
    age_regles_ini,
    age_regle_cat,
    first_rapport,
    
    # Endo/Adeno
    age_symptome, 
    age_symptome_cat,
    delai_diag_cat,
    signe_symptome_premiere_consult,
    type_endo,
    family_atcd_EndoAdeno, 
    family_atcd_doul_pelv,
    Nb_specialites_consult_cat,
    stade_dvpt_maladie, 
    stade_dvpt_legere_severe,
    motifs_operation1_cat, 
    motifs_operation2_cat, 
    indic_ovaire_touche,
    num_localisation_endo,
    localisation_final, 
    taille_kyste_IRM_max,
    tercile_taille_kyste_IRM,trompe_touche2,
    nb_operation_endo, 
    nb_operation_endo_cat
  )%>%
  mutate(stade_dvpt_maladie=
           factor(stade_dvpt_maladie,
                  levels= c("Non", "Je ne sais pas", 
                           "Stade I","Stade II",
                           "Stade III","Stade IV"),
                  ordered = TRUE))%>%
  select(-id_unik)%>%
  tbl_summary(
    by =module,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    label = list()
  )%>%
  modify_header(label = "**Variable**") %>%
  modify_caption(
    "**Tableau 2. Endometriosis/Adenomyosis chracteristics (n=1107)**"
  ) %>%
  bold_labels()  ; tbl2_modules
```

## Table 3 : All types of pain related to diseases 

```{r}
df_merged_filtered %>% 
  select(
    id_unik, 
    # Dysmenorrhea
    doul_pelvienne_regle,
    # Dysurenia
    doul_rapport_sexuel,
    douleur_abdo, 
    douleur_miction,
    trouble_digestif
    )%>%
  distinct()%>%
  select(-id_unik)%>%
  tbl_summary(
    by = NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    label = list()
  )%>%
  modify_header(label = "**Variable**") %>%
  modify_caption(
    "**Tableau 3. Pain related to Endometriosis/Adenomyosis (n=1390)**"
  ) %>%
  bold_labels()

```


# 여기서부터!!!!
한 환자 당 하나의 값을 갖는 변수들만 골라서 table 만들어야 함. 
Population : sans grossesses en cours et New/old module 
- grossesse à terme : Oui/Non 
- Taux de grossesse à terme 

```{r}
df_merged_filtered %>% 
  select(
    # Grossesse 
    id_unik, #IdPregnancy,
    #module,
    live_birth_bin_atleast_once,
    #live_birth_bin_all, live_birth_bin_onlynew,
    #status_terme_all, status_terme_onlynew,
    #type_preg_all, type_preg_onlynew,
    nb_live_births_all, nb_live_births_onlynew,
    nb_incomplete_preg_pt_all, nb_incomplete_preg_pt_onlynew,
    nb_total_gross,
    taux_LB_all, taux_LB_onlynew,
    first_preg_age, last_preg_age_mod,
    Methode_conception_bin, #starts_with("methode_"),
    starts_with("mode_"),-Mode_accouchement,
    starts_with("travail_"),-Travail_accouchement,
    nb_nevivants_parpt,
    allaitement_YN, nb_allaitement,
    PriseEnCharge_bin,starts_with("any_"), #PriseenCharge_long,
    starts_with("total_complicatioins"),
    starts_with("complic_"),-Complication_value
    ) %>% 
  distinct() %>% 
  select(-id_unik)%>%
  tbl_summary(
    by = NULL,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    label = list()
  )%>%
  #add_overall() %>% 
  modify_header(label = "**Variable**") %>%
  modify_caption(
    "**Tableau 4. Pregnancies information (n=1107)**"
  ) %>%
  bold_labels()

```




```{r}
df_merged_filtered %>% 
    distinct(id_unik, taux_LB_ALL_all) %>% 
    select(taux_LB_ALL_all) %>% 
    ggplot(aes(x = taux_LB_ALL_all)) +
    geom_histogram(aes(y = ..density..), bins = 30, 
                   fill = "lightblue", color = "black", 
                   binwidth = 0.2) +
    geom_density(color = "red", size = 0.8) +
    theme_minimal() +
    labs(
        title = "Distribution of LBR per patient",
        x = "LBR per patient",
        y = "Density"
    )

df_merged_filtered %>% 
  distinct(id_unik, taux_LB_ALL_all) %>% 
  group_by(taux_LB_ALL_all) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = taux_LB_ALL_all, y = n)) +
  geom_bar(stat = "identity", width = 0.15, fill = "lightblue") +
  geom_smooth(se = FALSE, colour = "red", linewidth = 0.8) +
  geom_text(aes(label = n), vjust = -0.8, size = 3) + 
  coord_cartesian(ylim = c(0, 800)) +  # 여기!
  theme_bw() +
  labs(
    title = "Distribution of LBR",
    y = "Frequency", 
    x = "LBR per patient"
  )

```


```{r}

```

