---
title: "03_O2_Data_EDA"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 03_O2_Data_EDA
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Obejctive 2, Data preparation and EDA 
*Date of creation* : 2025-04-09
*Date of update* : 2025-05-13
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

Informations : 
- 02 Mai 2025 : décidé à enlever l'outcome du taux de grossesse à terme par Marina et sélectionné les variables pour l'analyse liée à "Live birth" 

$Libraries$
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(purrr) # reduce()
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gt)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
library(chromote)
library(car)
library(mediation)
library(shiny)
library(shinydashboard)
library(data.table)
library(DT)
library(mice)
library(dagitty)
library(ggdag)
```

# 0. Import the dataset 

```{r}
# df_final_LB_1 <- read_csv("df_final_LB_1.csv")%>% 
#   dplyr::select(-`...1`)
df_final_LB_1 <- read_csv("df_final_LB_2.csv")%>% 
  dplyr::select(-`...1`)

# Import the final merged dataset : df_merged_filtered
df_final_LB_1 %>% colnames
#df_final_LB_2%>% colnames
```

# I. Data check and create new variables 

```{r}
df_final_LB_1_mod <- df_final_LB_1

#--------------------------------------------------
#  First pregnancy age in numeric to categorical
#--------------------------------------------------
quantiles <- df_final_LB_1_mod$first_preg_age %>% 
  quantile(probs = c(1/3,2/3), na.rm = TRUE) %>% 
  round()

df_final_LB_1_mod <- df_final_LB_1 %>%
  mutate(
    first_preg_age= round(first_preg_age),
    first_preg_age_cat= 
      case_when(
        first_preg_age< quantiles[[1]] ~ paste0("<", quantiles[1]),
        first_preg_age>= quantiles[[1]] &
          first_preg_age< quantiles[[2]] ~ paste0(quantiles[1], "-", quantiles[2]),
        first_preg_age>= quantiles[[2]] ~ paste0(">=", quantiles[2])
      )) %>% 
  dplyr::select(-id_unik)

rm(quantiles)

#-----------------------------------
#     Character to factor
#-----------------------------------
df_final_LB_1_mod <- df_final_LB_1_mod %>%
  mutate(across(where(is.character), ~ as.factor(.)))

#--------------------------------------
# Remove the 'Not_answered' modality
#--------------------------------------
df_final_LB_1_mod <- df_final_LB_1_mod %>%
  mutate(across(everything(), ~ {
    if (is.character(.) || is.factor(.)) {
      x <- as.character(.)
      x[x == "Not_answered"] <- NA
      return(factor(x))
    } else {
      return(.)
    }
  }))

```


# II. EDA 
**1. Data strucutre**
install.packages("prettyR")
library(prettyR)
describe(df_final_LB_1_mod)

**2. Data type**
- sapply(df_final_LB_1_mod, class)

**3. Check basic statistics and distribution**
- Summary of numeric data
- Count and proportion
- Histogram
- Boxplot

**4. Correlation analysis**
- Correlation Matrix and Visualisation

## 1) gtsummary
-  2 groups : live birth Yes /No 

```{r}
library(gtsummary)

df_final_LB_1_mod %>%
  tbl_summary(
    by = live_birth_bin_atleast_once_all,
    statistic = list(all_continuous() ~ "{mean} ({sd})",  
                     all_categorical() ~ "{n} ({p}%)"),    
    missing = "ifany" 
  ) %>%
  # add_n() %>%                
  # add_p() %>%               
  bold_labels()              
```


## 2) ShinyR : Distribution (Numeric, Categoric), Correlatoin, NA values
- R Shiny App for Data Distribution Visualization

### * UI
```{r}
#--------
#   UI
#--------
ui <- dashboardPage(
  dashboardHeader(title = "O2 : Data Distribution Viewer"),
  dashboardSidebar(sidebarMenu(
    menuItem(
      "Numeric Variables",
      tabName = "numeric",
      icon = icon("chart-line")
    ),
    menuItem(
      "Categorical Variables",
      tabName = "categorical",
      icon = icon("bar-chart")
    ),
    menuItem(
      "Missing Values Analysis",
      tabName = "missing",
      icon = icon("table")
    ),
    # New Menu Item for Correlation
    menuItem(
      "Variable Correlation",
      tabName = "correlation",
      icon = icon("link")
    ),
    menuItem(
      "Data Table",
      tabName = "datatable",
      icon = icon("table"))
  )),
  dashboardBody(tabItems(
    #-----------------------
    # Numeric variables tab
    #-----------------------
    tabItem(tabName = "numeric", fluidRow(
      box(
        title = "Select Numeric Variable",
        status = "primary",
        solidHeader = TRUE,
        selectInput("numericVar", "Choose variable:",
                    choices = NULL)
      ),
      box(
        title = "Summary Statistics",
        status = "primary",
        solidHeader = TRUE,
        verbatimTextOutput("numericSummary")
      )
    ), fluidRow(
      box(
        title = "Box Plot",
        status = "primary",
        solidHeader = TRUE,
        plotOutput("boxPlot")
      ),
      box(
        title = "Histogram",
        status = "primary",
        solidHeader = TRUE,
        plotOutput("histPlot")
      )
    )),
    #----------------------------
    # Categorical variables tab
    #----------------------------
    tabItem(tabName = "categorical", fluidRow(
      box(
        title = "Select Categorical Variable",
        status = "success",
        solidHeader = TRUE,
        selectInput("categoricalVar", "Choose variable:",
                    choices = NULL)
      ),
      box(
        title = "Bar Plot",
        status = "success",
        solidHeader = TRUE,
        plotOutput("barPlot")
      )
    ), fluidRow(
      box(
        title = "Frequency Table",
        status = "success",
        solidHeader = TRUE,
        width = 12,
        dataTableOutput("freqTable")
      )
    )),
    #-------------
    # NA analysis
    #-------------
    tabItem(tabName = "missing", fluidRow(
      box(
        title = "Missing Values Summary",
        status = "warning",
        solidHeader = TRUE,
        width = 12,
        dataTableOutput("missingTable")
      )
    ), fluidRow(
      box(
        title = "Missing Values Visualization",
        status = "warning",
        solidHeader = TRUE,
        width = 12,
        plotOutput("missingPlot", height = "600px")
      )
    )),
    #-------------------------
    # Correlation analysis tab
    #-------------------------
    tabItem(tabName = "correlation",
            fluidRow(
              box(
                title = "Select Analysis Type",
                status = "info",
                solidHeader = TRUE,
                width = 12,
                selectInput("correlationType", "Choose Analysis:",
                            choices = c("Numeric vs Numeric",
                                        "Categorical vs Categorical",
                                        "Numeric vs Categorical"))
              )
            ),
            # Conditional UI based on analysis type
            uiOutput("correlationUI")
    ),
    #----------------
    # Data table tab
    #----------------
    tabItem(
      tabName = "datatable",
      box(
        title = "Data Table",
        status = "warning",
        solidHeader = TRUE,
        width = 12,
        dataTableOutput("fullDataTable")
      )
    )
  ))
)
```

### * Server
```{r}
data <- df_final_LB_1_mod
server <- function(input, output, session) {

  # Reactive values to store column names
  col_types <- reactiveValues(numeric = NULL, categorical = NULL)

  #------------------------------------------
  # Identify numeric and categorical columns
  #------------------------------------------
  observe({
    req(data) # Ensure data is loaded

    is_numeric_col <- sapply(data,
                             function(x) is.numeric(x) && !is.factor(x))

    col_types$numeric <- names(data)[is_numeric_col]
    col_types$categorical <- names(data)[!is_numeric_col]

    # Update select inputs for existing tabs
    updateSelectInput(session, "numericVar",
                      choices = col_types$numeric)
    updateSelectInput(session, "categoricalVar",
                      choices = col_types$categorical)
  })

  #-------------------------------------
  # Original Tab Outputs (Existing code)
  #-------------------------------------

  # Histogram for numeric variables
  output$histPlot <- renderPlot({
    req(input$numericVar, data)
    ggplot(data, aes_string(x = input$numericVar)) +
      geom_histogram(fill = "steelblue", color = "white", bins = 30) +
      labs(title = paste("Distribution of", input$numericVar),
           x = input$numericVar, y = "Frequency") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5))
  })

  # Box plot for numeric variables
  output$boxPlot <- renderPlot({
    req(input$numericVar, data)
    ggplot(data, aes_string(y = input$numericVar)) +
      geom_boxplot(fill = "steelblue") +
      labs(title = paste("Box Plot of", input$numericVar),
           x = "", y = input$numericVar) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5))
  })

  # Summary statistics for numeric variables
  output$numericSummary <- renderPrint({
    req(input$numericVar, data)
    summary(data[[input$numericVar]])
  })

  # Bar plot for categorical variables
  output$barPlot <- renderPlot({
    req(input$categoricalVar, data)
    var_sym <- sym(input$categoricalVar)

    freq_data <- data %>%
      filter(!is.na(!!var_sym)) %>% # Exclude NA for counting
      count(!!var_sym) %>%
      arrange(desc(n))

    # Limit to top 20 + "Others"
    if (nrow(freq_data) > 20) {
        top_20 <- freq_data[1:20, ]
        others_n <- sum(freq_data[21:nrow(freq_data), ]$n)
        # Ensure correct structure for rbind
        others_row <- tibble(!!var_sym := factor("Others", levels=c(levels(top_20[[1]]),"Others")), n = others_n)
        freq_data <- rbind(top_20, others_row)
        freq_data[[1]] <- factor(freq_data[[1]], levels = c(as.character(top_20[[1]]), "Others")) # Set factor levels for plotting order
         freq_data$n <- as.numeric(freq_data$n)
    } else {
       freq_data[[1]] <- factor(freq_data[[1]], levels = as.character(freq_data[[1]])) # Ensure factor for consistent plotting
    }


    ggplot(freq_data, aes(x = !!var_sym, y = n)) + # Use original factor levels for order
      geom_bar(stat = "identity", fill = "darkgreen") +
      labs(
        title = paste("Distribution of", input$categoricalVar),
        x = input$categoricalVar,
        y = "Count"
      ) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            axis.text.x = element_text(angle = 45, hjust = 1))
  })

  # Frequency table for categorical variables
  output$freqTable <- renderDataTable({
    req(input$categoricalVar, data)
    data %>%
      count(!!sym(input$categoricalVar)) %>%
      mutate(Percentage = round(n / sum(n) * 100, 2)) %>%
      arrange(desc(n)) %>%
      rename(Category = 1, Count = n) %>%
      datatable(options = list(pageLength = 10))
  })

  # NA analysis table
  output$missingTable <- renderDataTable({
    req(data)
    na_summary <- data.frame(
      Variable = names(data),
      Type = sapply(data, function(x) ifelse(is.numeric(x) && !is.factor(x), "Numeric", "Categorical")),
      Missing_Count = sapply(data, function(x) sum(is.na(x))),
      Missing_Percent = sapply(data, function(x) round(sum(is.na(x)) / length(x) * 100, 2))
    )
    na_summary <- na_summary[order(-na_summary$Missing_Count), ]
    datatable(na_summary, options = list(pageLength = 20), rownames = FALSE)
  })

  # NA value visualisation
  output$missingPlot <- renderPlot({
    req(data)
    na_summary <- data.frame(
      Variable = names(data),
      Missing_Percent =
        sapply(data,
               function(x) sum(is.na(x)) / length(x) * 100)
    ) %>% filter(Missing_Percent > 0) # Only show vars with missing values

    if(nrow(na_summary) == 0) return(NULL) # Don't plot if no missing values

    na_summary <- na_summary[order(-na_summary$Missing_Percent), ]

    if(nrow(na_summary) > 30) {
      na_summary <- na_summary[1:30, ]
    }

    na_summary$color <- cut(
      na_summary$Missing_Percent,
      breaks = c(-1, 0.00001, 10, 30, 50, 101), # Adjusted breaks for >0
      labels = c("darkgreen", "green", "orange", "red", "darkred"),
      right = FALSE
    )

    ggplot(na_summary, aes(
      x = reorder(Variable, Missing_Percent),
      y = Missing_Percent,
      fill = color
    )) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=sprintf("%.1f%%", Missing_Percent)), hjust=-0.1, size=3.5) + # Add percentage labels
      scale_fill_identity() +
      coord_flip() + # Flip coordinates for better readability
      labs(title = "Missing Values by Variable (Top 30 with Missing Data)",
           x = "Variable", y = "Missing Values (%)") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5))
  }, height = 600) # Adjust height dynamically maybe?


  # Full data table
  output$fullDataTable <- renderDataTable({
    req(data)
    datatable(data, options = list(
      pageLength = 10,
      scrollX = TRUE
    ))
  })

  #-------------------------------
  # Correlation Tab Server Logic
  #-------------------------------
  # Dynamic UI for correlation analysis based on type selection
  output$correlationUI <- renderUI({
    req(input$correlationType, col_types$numeric, col_types$categorical)

    if (input$correlationType == "Numeric vs Numeric") {
      tagList(
        fluidRow(
          box(
            title = "Select Variables", status = "info", solidHeader = TRUE, width = 6,
            selectInput("numVar1", "Numeric Variable 1:", choices = col_types$numeric),
            selectInput("numVar2", "Numeric Variable 2:", choices = col_types$numeric, selected = if(length(col_types$numeric)>1) col_types$numeric[2] else NULL)
          ),
          box(
            title = "Correlation Coefficient", status = "info", solidHeader = TRUE, width = 6,
            verbatimTextOutput("numNumCorr")
          )
        ),
        fluidRow(
          box(
            title = "Scatter Plot", status = "info", solidHeader = TRUE, width = 12,
            plotOutput("numNumPlot")
          )
        )
      )
    } else if (input$correlationType == "Categorical vs Categorical") {
      tagList(
          fluidRow(
            box(
                title = "Select Variables", status = "info", solidHeader = TRUE, width = 12,
                selectInput("catVar1", "Categorical Variable 1:", choices = col_types$categorical),
                selectInput("catVar2", "Categorical Variable 2:", choices = col_types$categorical, selected = if(length(col_types$categorical)>1) col_types$categorical[2] else NULL)
            )
         ),
        fluidRow(
          box(
            title = "Contingency Table", status = "info", solidHeader = TRUE, width = 6,
             DT::dataTableOutput("catCatTable") # Use DT for consistency
          ),
          box(
            title = "Chi-Squared Test", status = "info", solidHeader = TRUE, width = 6,
            verbatimTextOutput("catCatTest")
          )
        ),
         fluidRow(
            box(
                title = "Association Plot", status = "info", solidHeader = TRUE, width = 12,
                plotOutput("catCatPlot")
            )
         )
      )
    } else if (input$correlationType == "Numeric vs Categorical") {
      tagList(
        fluidRow(
            box(
                title = "Select Variables", status = "info", solidHeader = TRUE, width = 12,
                selectInput("numVarCorr", "Numeric Variable:", choices = col_types$numeric),
                selectInput("catVarCorr", "Categorical Variable:", choices = col_types$categorical)
            )
        ),
        fluidRow(
          box(
            title = "Grouped Box Plot", status = "info", solidHeader = TRUE, width = 6,
            plotOutput("numCatPlot")
          ),
          box(
            title = "ANOVA Summary", status = "info", solidHeader = TRUE, width = 6,
            verbatimTextOutput("numCatTest")
          )
        )
      )
    }
  })

  # --- Numeric vs Numeric ---
  output$numNumPlot <- renderPlot({
    req(input$numVar1, input$numVar2, data)
    # Ensure different variables are selected
    if (input$numVar1 == input$numVar2) {
        plot(0, type='n', axes=FALSE, ann=FALSE)
        text(1, 0, "Please select two different numeric variables.", cex = 1.2)
        return()
    }

    ggplot(data, aes_string(x = input$numVar1, y = input$numVar2)) +
      geom_point(alpha = 0.6) +
      geom_smooth(method = "lm", se = FALSE, color = "red") + # Add linear regression line
      labs(title = paste("Scatter Plot:", input$numVar1, "vs", input$numVar2),
           x = input$numVar1, y = input$numVar2) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5))
  })

  output$numNumCorr <- renderPrint({
      req(input$numVar1, input$numVar2, data)
      # Ensure different variables are selected
      if (input$numVar1 == input$numVar2) {
          cat("Please select two different numeric variables.")
          return()
      }

      # Calculate Pearson correlation, handling potential NAs
      cor_test_result <- tryCatch({
          # Ensure we use pairwise complete observations for robustness
          cor.test(data[[input$numVar1]], data[[input$numVar2]], method = "pearson", use = "pairwise.complete.obs")
      }, error = function(e) {
          # Return a structure indicating error
          list(estimate = NA, p.value = NA, error_message = conditionMessage(e))
      })

      cat("Pearson Correlation Analysis\n")
      cat("----------------------------\n")
      cat("Variable 1:", input$numVar1, "\n")
      cat("Variable 2:", input$numVar2, "\n\n")

      # Check for errors during calculation
      if (!is.null(cor_test_result$error_message)) {
          cat("Error calculating correlation:", cor_test_result$error_message, "\n")
          cat("Note: Correlation could not be computed (check for NAs or lack of variance).\n")
      } else if (is.na(cor_test_result$estimate) || is.na(cor_test_result$p.value)) {
          cat("Correlation Coefficient (r): NA\n")
          cat("P-value: NA\n")
          cat("Note: Correlation could not be computed (likely due to extensive NA values or lack of variance).\n")
      } else {
          r_value <- cor_test_result$estimate
          p_value <- cor_test_result$p.value
          abs_r <- abs(r_value)

          # Determine correlation strength
          strength <- cut(abs_r,
                          breaks = c(-0.01, 0.10, 0.30, 0.50, 0.70, 1.00), 
                          labels = c("Very weak / none", "Weak", "Moderate", "Strong", "Very strong"),
                          right = TRUE) # Intervals are (lower, upper]

          cat("Correlation Coefficient (r):", round(r_value, 3), "\n")
          cat("Strength of Correlation:", as.character(strength), "\n") # Display strength
          cat("P-value:", format.pval(p_value, digits = 3, eps = 0.001), "\n")

          # Interpret significance
          if (p_value < 0.05) {
              cat("Significance: Statistically significant correlation (p < 0.05)\n")
          } else {
              cat("Significance: No statistically significant correlation (p >= 0.05)\n")
          }
      }
  })

  # --- Categorical vs Categorical ---
   cat_cat_data <- reactive({
       req(input$catVar1, input$catVar2, data)
       # Ensure different variables are selected
       if (input$catVar1 == input$catVar2) return(NULL)

       # Create contingency table, handling NAs by default (table removes them)
        cont_table <- table(data[[input$catVar1]], data[[input$catVar2]])

       # Check if table has dimensions before proceeding
        if (nrow(cont_table) < 2 || ncol(cont_table) < 2) {
            return(list(table = NULL, test = NULL, plot_data = NULL, message="Requires at least 2 levels in each variable after removing NA"))
        }

        # Perform Chi-squared test
        chi_test <- tryCatch({
            chisq.test(cont_table)
        }, warning = function(w) {
            # Capture warning (e.g., expected frequencies < 5)
            list(test = chisq.test(cont_table, simulate.p.value = TRUE), # Try simulation if needed
                 warning_message = paste("Chi-squared test warning:", conditionMessage(w)))
        }, error = function(e) {
            list(test = NULL, error_message = paste("Chi-squared test error:", conditionMessage(e)))
        })

        # Prepare data for ggplot (stacked bar chart)
        plot_df <- as.data.frame(cont_table)
        names(plot_df) <- c("Var1", "Var2", "Freq")


       list(table = cont_table, test_result = chi_test, plot_data = plot_df, message=NULL)
   })


  output$catCatTable <- DT::renderDataTable({
      result <- cat_cat_data()
      req(result)
      if(!is.null(result$message)){
           # Display message instead of table if invalid
           return(datatable(data.frame(Message = result$message)))
      }
      req(result$table)
      # Convert table to data frame for DT
      df_table <- as.data.frame.matrix(result$table)
      # Add row names as a column
      df_table <- cbind(Category1 = rownames(df_table), df_table)
      rownames(df_table) <- NULL # Remove row names for cleaner look

      datatable(df_table, options = list(pageLength = 5, scrollX = TRUE), caption = paste("Contingency Table:", input$catVar1, "vs", input$catVar2))
  })

  output$catCatTest <- renderPrint({
      result <- cat_cat_data()
      req(result)

      if(!is.null(result$message)){
          cat(result$message)
          return()
      }

      if (input$catVar1 == input$catVar2) {
        cat("Please select two different categorical variables.")
        return()
      }

      req(result$test_result)

      cat("Chi-Squared Test of Independence\n")
      cat("--------------------------------\n")
      cat("Variable 1:", input$catVar1, "\n")
      cat("Variable 2:", input$catVar2, "\n\n")

      if (!is.null(result$test_result$error_message)) {
          cat(result$test_result$error_message, "\n")
      } else {
           test_output <- result$test_result$test %||% result$test_result # Handle potential list structure from warning
           print(test_output)
           if (!is.null(result$test_result$warning_message)) {
                cat("\nWarning:", result$test_result$warning_message, "\n")
                if(grepl("simulate.p.value", result$test_result$warning_message)){
                     cat("P-value was simulated due to low expected frequencies.\n")
                }
           }
           if (test_output$p.value < 0.05) {
             cat("\nResult: Significant association (p < 0.05)\n")
           } else {
             cat("\nResult: No significant association (p >= 0.05)\n")
           }
      }
      # Helper function to handle NULL
      `%||%` <- function(a, b) if (!is.null(a)) a else b


  })

   output$catCatPlot <- renderPlot({
        result <- cat_cat_data()
        req(result)

        if (input$catVar1 == input$catVar2) {
            plot(0, type='n', axes=FALSE, ann=FALSE)
            text(1, 0, "Please select two different categorical variables.", cex = 1.2)
            return()
        }

        if(!is.null(result$message)){
            plot(0, type='n', axes=FALSE, ann=FALSE)
            text(1, 0, result$message, cex = 1.2)
           return()
       }

        req(result$plot_data)
        df_plot <- result$plot_data

       # Use ggplot for a stacked bar chart or heatmap
        ggplot(df_plot, aes(x = Var1, y = Freq, fill = Var2)) +
            geom_bar(stat = "identity", position = "fill") + # position="fill" shows proportions
            scale_y_continuous(labels = scales::percent_format()) +
            labs(title = paste("Proportional Stacked Bar Chart:", input$catVar1, "vs", input$catVar2),
                 x = input$catVar1,
                 y = "Proportion",
                 fill = input$catVar2) +
            theme_minimal() +
            theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0.5))

   })


  # --- Numeric vs Categorical ---
   num_cat_data <- reactive({
        req(input$numVarCorr, input$catVarCorr, data)

        num_var <- input$numVarCorr
        cat_var <- input$catVarCorr

        # Check number of levels in categorical variable after removing NAs for both vars
        valid_data <- data %>%
            filter(!is.na(.data[[num_var]]) & !is.na(.data[[cat_var]]))
        
        n_levels <- length(unique(valid_data[[cat_var]]))

        if (n_levels < 2) {
            return(list(plot_data=NULL, test=NULL, message="Categorical variable must have at least 2 levels with non-NA numeric data."))
        }

        # Perform ANOVA
        formula_str <- paste("`",num_var, "` ~ `", cat_var,"`", sep="") #Use backticks for non-standard names
        anova_test <- tryCatch({
            summary(aov(as.formula(formula_str), data = data))
        }, error = function(e) {
            list(error_message = paste("ANOVA error:", conditionMessage(e)))
        })

        list(plot_data = valid_data, test_result = anova_test, message=NULL)
   })


  output$numCatPlot <- renderPlot({
      result <- num_cat_data()
       req(result)

      if(!is.null(result$message)){
           plot(0, type='n', axes=FALSE, ann=FALSE)
           text(1, 0, result$message, cex = 1.2)
           return()
       }
       req(result$plot_data)
       df_plot <- result$plot_data
       num_var <- input$numVarCorr
       cat_var <- input$catVarCorr

       ggplot(df_plot, aes_string(x = cat_var, y = num_var, fill = cat_var)) +
         geom_boxplot(show.legend = FALSE) + # Fill by category for distinction
         labs(title = paste("Box Plot:", num_var, "by", cat_var),
              x = cat_var, y = num_var) +
         theme_minimal() +
         theme(axis.text.x = element_text(angle = 45, hjust = 1),
               plot.title = element_text(hjust = 0.5))
  })

  output$numCatTest <- renderPrint({
      result <- num_cat_data()
      req(result)

       if(!is.null(result$message)){
           cat(result$message)
           return()
       }
      req(result$test_result)


      cat("ANOVA (Analysis of Variance)\n")
      cat("----------------------------\n")
      cat("Dependent Variable (Numeric):", input$numVarCorr, "\n")
      cat("Independent Variable (Categorical):", input$catVarCorr, "\n\n")

      if (!is.null(result$test_result$error_message)) {
          cat(result$test_result$error_message, "\n")
      } else {
          anova_summary <- result$test_result
          print(anova_summary)

          # Extract p-value
          p_value <- tryCatch(anova_summary[[1]]$`Pr(>F)`[1], error = function(e) NA)

          if (is.na(p_value)) {
              cat("\nCould not extract p-value from ANOVA summary.\n")
          } else if (p_value < 0.05) {
            cat("\nResult: Significant difference in means across groups (p < 0.05)\n")
          } else {
            cat("\nResult: No significant difference in means across groups (p >= 0.05)\n")
          }
          cat("\nNote: ANOVA assumes normality of residuals and homogeneity of variances.\n")
          cat("Consider checking assumptions or using non-parametric alternatives (e.g., Kruskal-Wallis) if assumptions are violated.\n")
      }
  })

}

# Run the application
shinyApp(ui = ui, server = server)
```


## 3) Correlation between all numeric data
- Removing correlated variables
### - Correlation check
#### * Numeric data 
- There is a correlation between the age at the inclusion of the study and the maternal age at the first pregnancy (0.6014242), so it would be better to select one numeric variable.
- However, I'd like to use the maternal age variable as well, so I transformed this numeric variable into a categorical one. Then, I will check for multicollinearity to see if I can include this variable.

```{r}
#--------------------------------
#  1.  Identify numeric columns
#--------------------------------
is_numeric_col <- sapply(df_final_LB_1_mod,
                         function(x) is.numeric(x) &&
                           !is.factor(x))
numeric_columns <- names(df_final_LB_1_mod)[is_numeric_col]

#--------------------------
#  2. Correlation Matrix
#---------------------------
cor_matrix <- cor(df_final_LB_1_mod[, numeric_columns],
                  use="complete.obs")

#--------------------
# 3. Visualisation
#--------------------
library(corrplot)
corrplot(cor_matrix, method="circle")

# Interpretation 
# - age_firstOP_endoadeno ~ Age_premier_diag_endo_ade : STRONG CORRELATION
# - All types of age data tend to be correlated : select only age variable
# - All mean pain score  are correlated : select one representative variable
```

=> Eliminating Correlated Variables
Q) enlever first pregnancy age ? (250507 by Paul)
Answer : Remove this variable for the multivariate logistic regression 

### => Eliminate correlated variable

```{r}
df_final_LB_1_mod2 <- df_final_LB_1_mod %>% 
  dplyr::select(- first_preg_age)
```


## 3) Outlier 

- Already done in preprocessing data
- In the final dataframe, all numeric data don't have any outliers to treat

## 4) Treatment of Missing Data

- Check missing data : library(naniar);library(VIM);library(visdat); library(smdi)
- smdi_rf (MNAR, MCAR, MAR) -> visualisation (md.pattern) -> Imputation (mice or MissForest) -> Analysis modeling -> pooling 

    
```{r}
# Final model : outcome + covariables  
all_var <- c("live_birth_bin_atleast_once_all", "age","statut_marital_ini_cat","situation_financiere" ,"nb_mal_sansEndoAde_cat", "first_preg_age_cat","type_endo", "operation_endo_bin")

# Complete cases proportion = 66.2 % (737/1113)
sum(complete.cases(df_final_LB_1_mod2[all_var]))/ nrow(df_final_LB_1_mod2[all_var]) *100
```


### - Missing Data Check
#### - Proportion
```{r}
# 1 : Summary of missing data
miss_summary <- tibble(
  variable = names(df_final_LB_1_mod2),
  count = colSums(is.na(df_final_LB_1_mod2)),
  percent = round(colMeans(is.na(df_final_LB_1_mod2)) * 100,1)
) %>%
  arrange(desc(percent)) ; miss_summary

# 1:  Interactive plot : Plotly
library(plotly)
plot_ly(miss_summary %>% slice_max(order_by = percent, n = 25), 
        x = ~variable, y = ~ percent, 
        type = "bar", text= ~ percent,
        marker = list(color = "steelblue")) %>%
  layout(title = "Top 25 : Proportion of missing data per variable",
         xaxis = list(title = ""),
         yaxis = list(title = "Missing data prop (%)"))

# 1: Visualisation 
library(naniar)
gg_miss_var(df_final_LB_1_mod2) +
  labs(y = "NA counts", x = "Variable")

# # Missing data pattern correlation heatmap between variables
# gg_miss_upset(df_final_LB_1_mod2)

# 1: Visualisation 
na_summary_df <- data.frame(
      Variable = names(df_final_LB_1_mod2),
      Missing_Percent =
        sapply(df_final_LB_1_mod2,
               function(x) round(sum(is.na(x)) / length(x) * 100,1))
    ) %>% 
  filter(Missing_Percent > 0) %>% 
  arrange(desc(Missing_Percent)) %>% 
  slice_max(Missing_Percent, n = 35) %>%  # Top 35
  mutate(
    color= cut(
      Missing_Percent,
      breaks = c(-1, 0.00001, 10, 30, 50, 101),
      labels = c("darkgreen", "green", "orange", "red", "darkred"),
      right = FALSE
    )
  )


na_summary_df<-
  na_summary_df[!str_detect(rownames(na_summary_df), "_participated"), ]

ggplot(na_summary_df, aes(
  x = reorder(Variable, Missing_Percent), 
  y = Missing_Percent,
  fill = color)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label=Missing_Percent), hjust=-0.1, size=3.5) +
  scale_fill_identity() +
  coord_flip() +
  labs(title = "Missing Values by Variable (Top 35)",
       x = "Variable", y = "Missing Values (%)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

#### - Pattern and mechanism check

```{r}
#---------------------------------------
# Check MCAR for only continuous data
#---------------------------------------
#install.packages("MissMech")
library(MissMech)
# 1. Only numeric data
df_numeric <- df_final_LB_1_mod2 %>% 
  dplyr::select(where(is.numeric))

# 2. TestMCARNormality 
#TestMCARNormality(df_numeric)
# - interpretation : Probablement MCAR (p>0.05)

# naniar : mcar_test
library(naniar)
mcar_test(df_numeric)  

#--------------------------------
#   Missing data pattern check  
#--------------------------------
# library(smdi)
# res <- smdi_rf(data = df_final_LB_1_mod2,
#                set_seed = 2025,
#                ntree=1000,
#                n_cores=1)
# Interpretation : AUC ~> 0.5 = MCAR, AUC ~> 1 = MNAR 


# library(purrr)
# # Results' dataframe
# rf_results_df <- map_dfr(res, ~ .x$rf_table) #map_dfr() :  row bind
# rf_results_df <- rf_results_df %>%
#   mutate(rf_auc = as.numeric(rf_auc)) %>%
#   arrange(rf_auc)  
# 
# # Visualisation of results
# ggplot(rf_results_df, aes(x = reorder(covariate, rf_auc), 
#                           y = rf_auc)) +
#   geom_bar(stat = "identity", fill = "skyblue") +
#   coord_flip() + 
#   labs(
#     x = "Covariates",
#     y = "AUC (Random Forest)",
#     title = "Missing data mechanism test"
#   ) +
#   theme_minimal()

#------------------------------------
# Missing data pattern visualisation 
#-------------------------------------
# 1. Monotonicity : monotone vs non monotone
md.pattern(df_final_LB_1_mod2[,10:15], plot = TRUE,
           rotate.names = TRUE) # missing data pattern 
# - each row corresponds to a missing pattern(1= observed, 0= missing)
# - last column and row containes row and column counts, respectively 

# 2. influx and outflux of multivariate missing data patterns 
mice::flux(df_final_LB_1_mod2)
# - statistics of the missing data pattern : influx and outflux -> selecting predictors that sould go into the imputation model
# - influx : number of variables pairs (missing/total number of observed data cells, observation/total number of observed data cells), depending on the proportion of missing data of the variable => completely observed influx = 0 <->completely missing 1
# - outflux : variable with higher outflux is better connected to the missing data, and thus potentially more useful for imputing other variables 
# - FICO :outbound statistic defined by the fraction of incomplete cases among cases with observed 

# 2-1. Visualisation : Fluxplot of the missing data pattern 
mice::fluxplot(
  df_final_LB_1_mod2,
  local = names(df_final_LB_1_mod2),
  plot = TRUE,
  labels = TRUE,
  xlim = c(0, 1),
  ylim = c(0, 1),
  las = 1,
  xlab = "Influx",
  ylab = "Outflux",
  main = paste("Influx-outflux pattern for", deparse(substitute(df_final_LB_1_mod2))),
  eqscplot = TRUE)

#---------------------------------
#    Imputation of missing data 
#---------------------------------
library(missForest) # Non parametric Missing Value Imputation using Random Forest
library(mice)

# data <- df_final_LB_1_mod2 %>% 
#   dplyr::select()
# imp_mar <- mice(df_clean, m = 5, seed = 2025, printFlag = FALSE)
```

# III. Imputation of missing data 
## - MissForest (Simple imputation)
-                              
!
```{r}
#--------------------------------
# missForest : Simple imputation
#--------------------------------
# 1. Prepare data types
# 1-1) Check 
sapply(df_final_LB_1_mod2, class)
# 1-2) Transform all ordinal factor to factor 
df_clean <- data.frame(lapply(df_final_LB_1_mod2, function(x) {
  if (is.factor(x)) {
    return(as.factor(as.character(x)))
  } else if (is.numeric(x) || is.integer(x)) {
    return(x)
  } else {
    return(as.factor(as.character(x)))
  }
}))
# 1-3) Recheck 
sapply(df_clean, class)

# 2. Find a optimal number of trees : Optimal 300
ntree_vals <- c(50, 100, 200, 300,400)
results <- data.frame(ntree = ntree_vals, NRMSE = NA, PFC = NA)
set.seed(2025)
iris.mis <- prodNA(df_clean, noNA = 0.1)
for (i in 1:length(ntree_vals)) {
  res <- missForest(iris.mis, ntree = ntree_vals[i])
  results$NRMSE[i] <- res$OOBerror[1]
  results$PFC[i] <- res$OOBerror[2]
}
print(results)
#데이터에 따라 NRMSE(Normalized Root Mean Squared Error (NRMSE))와 PFC 중 더 중요한 지표가 달라질 수 있어. 예를 들어, 대부분 범주형이면 Proportion of falsely classified entries (PFC)를 중점적으로 보는 게 좋아.

# 3. Apply MissForest to the dataframe with optimal number of trees
library(missForest)
set.seed(2025)
imputed_result <- missForest(df_clean, ntree = 300)
imputed_data_O2 <- imputed_result$ximp

# 4. Check
sapply(imputed_data_O2, function(x) sum(is.na(x)))

# 5. Compare the original data vs imputed data
for (var in names(imputed_data_O2)) {
  cat("\n=== Variable:", var, "===\n")
  cat("Original distribution:\n")
  print(table(df_clean[[var]], useNA = "always"))
  cat("After imputation:\n")
  print(table(imputed_data_O2[[var]], useNA = "always"))
}

# 6. Remove 'df_clean' 
rm(df_clean)
```

# IV. Save the imputed final dataset
updata : 25-07-11

```{r}
# Original dataset : df_final_LB_1_mod2
# write.csv(imputed_data_O2,"imputed_data_O2_V2.csv")
```


# V. Dataset for the random effect model

```{r}
# df_final_LB_1 <- read_csv("Stage_M2_Complic_Gross/df_final_LB_1.csv")
id_unik<-df_final_LB_1$id_unik
new_dataframe<- cbind(id_unik,imputed_data_O2)

# Add the random effects informations 
df_final_LB_perPreg <- new_dataframe %>%
  left_join(
    df_merged_filtered %>%
      distinct(
        id_unik,
        IdPregnancy,
        age_maternal,
        live_birth_bin_all,
        gravida,
        para
      ),
    by="id_unik"
  )%>%
  dplyr::select(-live_birth_bin_atleast_once_all,
                -IdPregnancy)%>%
  mutate(live_birth_bin_all=
           ifelse(live_birth_bin_all == "Oui", 1, 0),
         id_unik = id_unik%>%as.factor())

#sapply(df_final_LB_perPreg, function(x) sum(is.na(x)))


library(geepack)
gee_model <- geeglm(
  live_birth_bin_all ~
    delai_diag_cat_simple +
    
    age +
    age_maternal+
    
    statut_marital_ini_cat +
    diplome_ini2 +
    
    gynecologique_diseases+
    
    operation_endo_bin +
    endo_ade_categorie +
    DIE,
  id = df_final_LB_perPreg$id_unik,
  data = df_final_LB_perPreg,
  family = binomial,
  corstr = "exchangeable"
)  
summary(gee_model)
QIC(gee_model)



glm_model <- glm(
  live_birth_bin_all ~
    age +
    age_maternal+
    statut_marital_ini_cat +
    diplome_ini2 +
    gynecologique_diseases+
    nb_mal_sansEndoAde_cat+
    gravida +
    age_firstOP_endoadeno_cat +
    endo_ade_categorie +
    DIE,
  data = df_final_LB_perPreg,
  family = binomial)

vif(glm_model)
```

