---
title: "04_O3_Statistical_Analysis_GLM"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : **3 Objective**, Statistical Analysis for the participant population
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Statistical Analysis
*Date of creation* : 2025-05-13
*Date of update* : 2025-06-02
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

# *O3 per ID*

$Libraries$
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(purrr) # reduce()
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gt)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
library(chromote)
library(car)
library(mediation)
library(shiny)
library(shinydashboard)
library(data.table)
library(DT)
library(mice)
library(dagitty)
library(ggdag)
library(gridExtra)
library(grid)
library(dotwhisker)
```

# 0. Import the imputed data 

```{r}
imputed_data_O3_perID <- 
  read.csv("C:/Stage_Master2/Stage_M2_Complic_Gross/imputed_data_O3_perID.csv") %>%
  dplyr::select(-X)

#----------------------
#  Imputed Dataframe
#----------------------
imputed_data_O3_perID <- imputed_data_O3_perID %>%
  mutate(
    diplome_ini2 = factor(
      diplome_ini2,
      levels = c("<=Baccalauréat/Aucun/Autre", "Bac+2", ">=Bac+3"),
      ordered = FALSE
    ),
    
    situation_financiere   = factor(
      situation_financiere,
      levels = c(
        "Balanced",
        "Difficult/In debt",
        "Comfortable/Very comfortable"
      ),
      ordered = FALSE
    ),
    
    family_atcd_EndoAdeno_cat =
      factor(
        family_atcd_EndoAdeno_cat,
        levels = c("Pas d'antécédents", "Antécédents")
      ),
    
    family_atcd_doul_pelv =
      factor(
        family_atcd_EndoAdeno_cat,
        levels = c("Pas d'antécédents", "Antécédents")
      ),
    
    endo_ade_categorie = factor(
      endo_ade_categorie,
      levels = c("Only endometriosis", "Only adenomyosis", "Both")
    ),
    
    type_endo =
      factor(
        type_endo,
        levels = c(
          "DIE only",
          "SPE only",
          "OMA only",
          "ADENO only",
          "Multiple types",
          "Aucun mentionné/nsp"
        )
      ),
    
    stade_dvpt_legere_severe = factor(
      stade_dvpt_legere_severe,
      levels = c(
        "I-II",
        "III-IV",
        "No surgery",
        "Non et nsp"
      )
    ), 
    
    op_avant_1_preg = factor(
      op_avant_1_preg,
      levels = c("Non","Oui")
    ),
    
    complication_operation = factor(
      complication_operation,
      levels = c("Non", "Oui", "No surgery")
    ),
    
    age_symptome_cat = factor(
      age_symptome_cat,
      levels = c(
        "<= 13",
        "(13;15]",
        "(15;22]",
        "22+",
        "No symptoms"
      ),
      ordered = FALSE
    ),

    age_firstOP_endoadeno_cat = factor(
      age_firstOP_endoadeno_cat,
      levels = c("No operation","<27", "27-31", "31>="),
      ordered = FALSE
    ),
    
    nb_mal_sansEndoAde_cat = factor(
      nb_mal_sansEndoAde_cat,
      levels = c("0", "1", "2 ou plus"),
      ordered = FALSE
    ),
    
    comorbidities_bin = case_when(
      nb_mal_sansEndoAde_cat =="0" ~ "Non",
      TRUE~ "Oui"
    ),
    
    delai_diag_cat = factor(
      delai_diag_cat,
      levels = c("<= 5", "(5;10]", "(10;15]", "15+", "No symptoms"),
      ordered = FALSE
    ),
    
    Travail_accouchement_perID = factor(
      Travail_accouchement_perID,
      levels = c(
        "Seul travail provoqué",
        "Seul travail spontané",
        "Travail mixte",
        "Pas de travail",
        "No_delivery"
      )
    ),
    
    Mode_accouchement_perID = factor(
      Mode_accouchement_perID,
      levels = c(
        "Seule voie Basse", 
        "Seule Césarienne", 
        "Both", 
        "No_delivery")
    ),
    
    first_preg_age_cat = factor(
      first_preg_age_cat,
      levels = c("<29", "29-34", ">=34"),
      ordered = FALSE
    ),
    
    total_complications_bin_perID = 
      as.factor(total_complications_bin_perID )
  ) 
  
```

#------------------------------------
# I. Binary outcome 
#-------------------------------------
## 1. Univariable analysis adjusted for age 
### 1) Results Table 


```{r}
#----------------------
#  Define the data
#----------------------
data <- imputed_data_O3_perID %>% 
  dplyr::select(-total_complications_cat_perID)

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables 
outcome_variable <- "total_complications_bin_perID"
adjustment_variable <- "age"
# adjustment_variable <- c("age",
#                          "first_preg_age_cat",
#                          "statut_marital_ini_cat",
#                          "situation_financiere")

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure outcome variable is suitable for logistic regression (factor or 0/1 numeric)
if (!is.factor(imputed_data_O3_perID[[outcome_variable]]) && !all(imputed_data_O3_perID[[outcome_variable]] %in% c(0, 1, NA))) {
  stop(paste("Outcome variable '", outcome_variable, "' must be a factor or numeric with values 0/1."))
  }

# Convert outcome to factor if it's 0/1 numeric, as tbl_summary needs factor for 'by'
if (is.numeric(data[[outcome_variable]])) {
    data <- data %>%
        mutate(!!sym(outcome_variable) := factor(!!sym(outcome_variable)))
    print(paste("INFO: Converted numeric outcome variable '",
                outcome_variable, "' to factor."))
    }

# Ensure adjustment variable exists
if (!all(adjustment_variable %in% names(data))) {
  stop(paste("Some adjustment variables not found in the dataframe:", 
             paste(setdiff(adjustment_variable, names(data)), collapse = ", ")))
}
# if (!adjustment_variable %in% names(data)) {
#   stop(paste("Adjustment variable '", adjustment_variable, "' not found in the dataframe."))
#   }

missing_vars <- setdiff(adjustment_variable, names(data))
if (length(missing_vars) > 0) {
  stop(paste("Adjustment variable(s) not found in the dataframe:", paste(missing_vars, collapse = ", ")))
  }

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(data), 
                               c(outcome_variable, adjustment_variable))

if (length(predictor_variables) == 0) {
    stop("No predictor variables found after excluding outcome and adjustment variables.")
  }

print(paste("Outcome:", outcome_variable))
print(paste("Adjustment:", paste(adjustment_variable, collapse=", ")))
print(paste("Predictors:", paste(predictor_variables, collapse=", ")))


#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>% 
  tbl_summary(
    by = all_of(outcome_variable),           # Stratify by the outcome
    statistic = list(                         # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",                           
    digits = list(all_continuous() ~ 1, 
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%     # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>% 
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted Multivariate Logistic Regression Table
#-------------------------------------------------------------
# This function fits glm(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_glm <- function(predictor, outcome_var, adjust_var, data) {
  # Use backticks for safety with non-standard names
  adjust_str <- paste(sprintf("`%s`", adjustment_variable), collapse = " + ")
    # adjust_vars는 벡터
  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)
  
  # Fit the model
  model <- tryCatch(
    glm(as.formula(formula_str), data = data, family = binomial(link = "logit")),
    error = function(e) {
      warning(paste("Could not fit model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )
  
  # If model failed, return NULL
  if (is.null(model)) return(NULL)
  
  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int=TRUE) %>%
    # Keep rows where 'variable' matches the predictor name
    # This correctly handles factor variables which have multiple rows
    modify_table_body(~ dplyr::filter(.x, variable == predictor)) 
  }

# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_glm(.x, outcome_variable,
                                            adjustment_variable,
                                            data))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables, 
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
            )
          )
      ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted OR (95% CI)**"
    )
  print("Step 2: Adjusted Odds Ratios Table Created.")
} else {
  warning("No successful models were fitted for adjusted ORs.", 
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
adjust_str <- paste(adjustment_variable, collapse = " + ")

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      "**Age-Adjusted Model**")
    ) %>%
      modify_caption(
        paste0(
          "**Multivariate Logistic Regression Analysis for '",
          outcome_variable, 
          "' (Adjusted for '", adjust_str, "')**"))

    print("Step 3: Merged Table Created.")
    
    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '", 
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results 
#------------------------------------------------------------
table_res2_final <- as_tibble(tbl_final) %>% 
  dplyr::select(-"**OR**",-"**95% CI**",-"**p-value**")

# save the table : 2025-05-30
#table_univariable_O3_bin<- table_res2_final
table_res2_final
```

### 2) Pre-selection of variables for the multivariate regression model
 - P-value <= 0.2
```{r}
ls_var_uni_results <- tbl_final$table_body %>% filter(p.value_2<=0.2) %>% distinct(variable) %>% pull()
ls_var_uni_results
```


**Results** 250602
- diplome_ini2 : >=Bac+3 (0.45 (0.24, 0.82)) 
- gynecologique_diseases : Oui (2.89 (1.10, 9.05)) -> Large CI
- infertilite_diag : Oui (2.66 (1.00, 8.39))
- age_regle_cat_simple : >=12 (0.38 (0.20, 0.67))
- stade_dvpt_III_IV : Oui (1.94 (1.01, 3.89))
- age_symptome_cat : (15;22] (0.49 (0.24, 0.98)), 22+ (0.53 (0.28, 0.97))
- age_firstOP_endoadeno_cat :  31>= (3.55 (1.43, 9.09)) -> Large CI
- history_FC_preterm_IVGIMG_perID : Oui (2.49 (1.41, 4.55))
- Mode_accouchement_perID : Seule Césarienne (2.57 (1.48, 4.62))
- VD_perID : Oui (0.40 (0.22, 0.69))
- Csection_perID : Oui (2.50 (1.46, 4.40))
- labour_natural_perID : Oui (0.57 (0.36, 0.91))
- doul_abdo_avant_1_preg : Oui (2.19 (1.36, 3.56))

## 2. Multivariate logistic regression

### 1) DAG
```{r}
#---------------------
#  Creation  of DAG 
#---------------------
library(dagitty)
library(ggdag)
dag_code <- "dag {
  age -> total_complications_bin_perID
  age -> statut_marital_ini_cat
  age -> diplome_ini2
  diplome_ini2 -> situation_financiere
  age -> situation_financiere
  age -> situation_emploi_ini2
  age -> operation_endo_bin
  age -> nb_mal_sansEndoAde_cat
  age -> metabolic_syndrome
  age -> nb_total_gross
  
  age_regle_cat_simple -> metabolic_syndrome
  age_regle_cat_simple -> total_complications_bin_perID
  
  metabolic_syndrome -> nb_mal_sansEndoAde_cat
  metabolic_syndrome -> total_complications_bin_perID
  metabolic_syndrome -> history_FC_preterm_IVG_IMG
  
  gynecologique_diseases -> total_complications_bin_perID
  gynecologique_diseases -> nb_mal_sansEndoAde_cat
  gynecologique_diseases -> Csection_perID
  gynecologique_diseases -> nb_total_gross
  gynecologique_diseases -> infertilite_diag
  gynecologique_diseases -> history_FC_preterm_IVG_IMG
  
  autoimmune_diseases -> history_FC_preterm_IVG_IMG
  autoimmune_diseases -> nb_mal_sansEndoAde_cat
  autoimmune_diseases -> total_complications_bin_perID
  
  nb_mal_sansEndoAde_cat -> total_complications_bin_perID
  nb_mal_sansEndoAde_cat -> Csection_perID

  Csection_perID -> total_complications_bin_perID
  total_complications_bin_perID -> Csection_perID
  
  type_endo -> operation_endo_bin
  type_endo -> stade_dvpt_III_IV
  type_endo -> stade_dvpt_legere_severe
  type_endo -> Endo_ade_categorie
  type_endo -> indic_ovaire_touche
  
  DIE -> stade_dvpt_III_IV
  stade_dvpt_III_IV -> operation_endo_bin
  operation_endo_bin -> indic_ovaire_touche
  operation_endo_bin -> trompe_anomalie 
  operation_endo_bin -> nb_total_gross
  operation_endo_bin -> complication_operation

  indic_ovaire_touche -> infertilite_diag
  trompe_anomalie -> infertilite_diag
  trompe_anomalie -> nb_total_gross
  
  op_avant_1_preg -> operation_endo_bin
  op_avant_1_preg -> infertilite_diag
  
  infertilite_diag -> operation_endo_bin
  infertilite_diag -> conception_PMA_perID
  infertilite_diag -> conception_operation_perID
  
  conception_operation_perID -> operation_endo_bin
  conception_operation_perID -> nb_total_gross
  
  maternal_age -> infertilite_diag
  maternal_age -> conception_PMA_perID
  maternal_age -> total_complications_bin_perID
  maternal_age -> nb_total_gross
  maternal_age -> history_FC_preterm_IVG_IMG
  
  first_preg_age_cat -> infertilite_diag
  first_preg_age_cat ->  conception_PMA_perID
  first_preg_age_cat -> total_complications_bin_perID
  first_preg_age_cat -> nb_total_gross
  first_preg_age_cat -> history_FC_preterm_IVG_IMG
  
  conception_PMA_perID -> total_complications_bin_perID
  conception_PMA_perID -> nb_total_gross
  
  nb_total_gross -> nb_live_births_all
  nb_live_births_all -> total_complications_bin_perID
  nb_live_births_all -> live_birth_bin_atleast_once_all
  
  history_FC_preterm_IVGIMG_perID -> nb_live_births_all
  history_FC_preterm_IVGIMG_perID -> nb_total_gross
  
  first_preg_age_cat -> op_avant_1_preg
  age_firstOP_endoadeno_cat -> op_avant_1_preg
  
  conception_operation_perID -> methode_conception_perID_cat_simple_mod
  conception_PMA_perID -> methode_conception_perID_cat_simple_mod

  methode_conception_perID_cat_simple_mod -> live_birth_bin_atleast_once_all
  methode_conception_perID_cat_simple_mod -> operation_endo_bin 
  methode_conception_perID_cat_simple_mod -> total_complications_bin_perID
}"

dag_model <- dagitty(dag_code)

#------------------------
# Visualisation of DAG
#------------------------
library(ggplot2)
library(ggrepel)
library(ggdag)

# tidy DAG
dag_tidy <- tidy_dagitty(dag_model)
nodes_df<- dag_tidy$data
edges <- edges(dag_model)
ggplot() +
  # Edges
  geom_segment(data = nodes_df, 
               aes(x = x, y = y, xend = xend, yend = yend),
               arrow = arrow(length = unit(0.01, "npc"), 
                             angle = 40, type="closed"), 
               color = "gray50", linewidth = 0.6) +  # size 대신 linewidth
  
  # Node
  geom_point(data = nodes_df, 
             aes(x = x, y = y), 
             shape = 21, fill = "lightcyan", 
             color = "steelblue", size = 10, stroke = 0.4,
             alpha = 0.2) +
  
  # label text 
  geom_text_repel(data = distinct(nodes_df, name, .keep_all = TRUE), 
                  aes(x = x, y = y, label = name), 
                  color = "brown", size = 2.5) +
  
  
  # theme
  theme_dag(base_size = 10, base_family = "Arial") +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "white"),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_blank()
  )


library(dplyr)
# geom_point size = 10 →  10*0.035 = 0.35 
r <- 0.35  

edges_adjusted <- nodes_df %>%
  mutate(
    dx = xend - x,
    dy = yend - y,
    dist = sqrt(dx^2 + dy^2),
    xend_adj = xend - r * (dx / dist),
    yend_adj = yend - r * (dy / dist)
  )

ggplot() +
  geom_segment(
    data = edges_adjusted,
    aes(
      x = x,
      y = y,
      xend = xend_adj,
      yend = yend_adj
    ),
    arrow = arrow(
      length = unit(0.015, "npc"),
      angle = 30,
      type = "open"
    ),
    color = "gray50",
    linewidth = 0.6
  ) +
  geom_point(
    data = nodes_df,
    aes(x = x, y = y),
    shape = 21,
    fill = "lightcyan",
    color = "steelblue",
    size = 10,
    stroke = 0.4,
    alpha = 0.2
  ) +
  geom_text_repel(
    data = distinct(nodes_df, name, .keep_all = TRUE),
    aes(x = x, y = y, label = name),
    color = "brown",
    size = 2.5
  ) +
  theme_dag(base_size = 10, base_family = "Arial") +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "white"),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_blank()
  )

#-------------------------------------
# Finding the adjustements variables
#-------------------------------------
adjustmentSets(dag_model, 
               exposure ="age_regle_cat_simple", 
               outcome = "total_complications_bin_perID",
               type= "canonical")
```

### 2) Final Model and Results

#### * Model_version1
```{r}
#-------------------------------
#   logsitic regression model
#-------------------------------
mod_glm_multivariate <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        type_endo +
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        indic_ovaire_touche+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID,
    family = binomial
  )

#------------
#  Results 
#------------
# 1) summary()
summary(mod_glm_multivariate)

# 2) tidy(exponentiate=TRUE) 
library(tibble)
res_multivar_glm <- tidy(mod_glm_multivariate, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_mod <- res_multivar_glm[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_mod <- res_multivar_glm_mod %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_mod
```


##### - Assumptions Check and Sensitivity analysis
###### > Sensitivity test : Outlier
```{r}
#--------------------------------
#  Multicollinearity Check : <2
#--------------------------------
vif(mod_glm_multivariate)

#----------------------------------
#  Outliers and Influences Check : 17 and 169
#----------------------------------
library(car)
influencePlot(mod_glm_multivariate)
# StudRes(Studentized Residuals) : <-2 ou >2 -> outlier
# Hat (Leverage, Hat value) : high leverage point (p/n, p = number of variables, n= number of samples)
# Cook's distance : Residus + Leverage
# CookD > 4/n (n= number of samples)
# - 4/378 = 0.016

plot(mod_glm_multivariate, which = 4)  # Cook’s distance
#-> 17 and 169 : Big residus, high leverage -> Outlier and big influence

#--------------------------
#   Sensitivity analysis  
#--------------------------
# 1. Create a new model without 17, 169 individuals
mod_glm_multivariate_sensitivity <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        type_endo +
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        #operation_endo_bin+
        indic_ovaire_touche+
        #conception_PMA_perID+
        #conception_operation_perID+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        #op_avant_1_preg+
        doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID[-c(17,169),],
    family = binomial
  )
summary(mod_glm_multivariate_sensitivity)

# 2. Compare the AIC of two differents model for the sensitivity analysis 
AIC(mod_glm_multivariate, mod_glm_multivariate_sensitivity)

# 3. Compare the AUC of two models for the sensitivity analysis 
library(pROC)
# 3-1) Prepare the FULL DATASET model for ROC and AUC
model_data <- na.omit(imputed_data_O3_perID)
mod_glm_multivariate <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        type_endo +
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        #operation_endo_bin+
        indic_ovaire_touche+
        #conception_PMA_perID+
        #conception_operation_perID+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        #op_avant_1_preg+
        doul_abdo_avant_1_preg,
    data = model_data,
    family = binomial
  )
prob <- predict(mod_glm_multivariate, type = "response")
roc_full <- roc(response = model_data$total_complications_bin_perID, predictor = prob)

# 3-2) Prepare the MODIFIED DATASET model for ROC and AUC
mod_glm_multivariate_sensitivity <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        type_endo +
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        #operation_endo_bin+
        indic_ovaire_touche+
        #conception_PMA_perID+
        #conception_operation_perID+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        #op_avant_1_preg+
        doul_abdo_avant_1_preg,
    data = model_data[-c(17,169),],
    family = binomial
  )
prob_sensitivity <- predict(mod_glm_multivariate_sensitivity, type = "response")
roc_no_outlier <- roc(response = model_data[-c(17,169),]$total_complications_bin_perID, predictor = prob_sensitivity)

# 4. Compare 
plot(roc_full, col="blue")
plot(roc_no_outlier, add=TRUE, col="red")
auc(roc_full) # 0.7986
auc(roc_no_outlier) # 0.7998


# 5. Results 
# There is no big differences between two model, so I'll keep the two individus with big leverage and big influence 
```


###### > Sensitivity test : Inclusion of variable

- model with VS without Macrophenotype of endometriosis : No big differences 
- model with DIE vs with Macrophenotype of endometriosis : No big differences 

```{r}
##################################################################################
# model with VS without Macrophenotype of endometriosis : no big differences 
#--------------------------
#   Sensitivity analysis  
#--------------------------
# 0. Original model with type_endo
mod_glm_multivariate <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        type_endo + # HERE
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        indic_ovaire_touche+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        doul_abdo_avant_1_preg,
    data = model_data,
    family = binomial
  )

# 1. Create a new model without type_endo
mod_glm_multivariate_sensitivity <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        indic_ovaire_touche+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID,
    family = binomial
  )
summary(mod_glm_multivariate_sensitivity)

# 2. Compare the AIC of two differents model for the sensitivity analysis 
AIC(mod_glm_multivariate, mod_glm_multivariate_sensitivity)

# 3. Compare the AUC of two models for the sensitivity analysis 
library(pROC)
# 3-1) Prepare the FULL DATASET model for ROC and AUC
model_data <- na.omit(imputed_data_O3_perID)
prob <- predict(mod_glm_multivariate, type = "response")
roc_type_endo <- roc(response = model_data$total_complications_bin_perID, predictor = prob)

# 3-2) ROC and AUC
prob_sensitivity <- predict(mod_glm_multivariate_sensitivity, type = "response")
roc_without <- roc(response = model_data$total_complications_bin_perID, predictor = prob_sensitivity)

# 3-3) Visualisation and the AUC results
plot(roc_type_endo, col="blue")
plot(roc_without, add=TRUE, col="red")
auc(roc_type_endo) # 0.7956
auc(roc_without) # 0.7816

# 4. Results 
# There is no big differences between two model, so I'll keep the two individus with big leverag


##################################################################################
# Model with DIE vs with Macrophenotype of endometriosis : no big differences 
#--------------------------
#   Sensitivity analysis  
#--------------------------
# 0. Original model with type_endo
mod_glm_multivariate <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        type_endo + # HERE
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        indic_ovaire_touche+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        doul_abdo_avant_1_preg,
    data = model_data,
    family = binomial
  )

# 1. Create a new model with DIE insetead of type_endo
mod_glm_multivariate_sensitivity <-
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        DIE+ # HERE
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        indic_ovaire_touche+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID,
    family = binomial
  )
summary(mod_glm_multivariate_sensitivity)

# 2. Compare the AIC of two differents model for the sensitivity analysis 
AIC(mod_glm_multivariate, mod_glm_multivariate_sensitivity)

# 3. Compare the AUC of two models for the sensitivity analysis 
library(pROC)
# 3-1) Prepare the FULL DATASET model for ROC and AUC
model_data <- na.omit(imputed_data_O3_perID)
prob <- predict(mod_glm_multivariate, type = "response")
roc_type_endo <- roc(response = model_data$total_complications_bin_perID, predictor = prob)

# 3-2) ROC and AUC
prob_sensitivity <- predict(mod_glm_multivariate_sensitivity, type = "response")
roc_DIE <- roc(response = model_data$total_complications_bin_perID, predictor = prob_sensitivity)

# 3-3) Visualisation and the AUC results
plot(roc_type_endo, col="blue")
plot(roc_DIE, add=TRUE, col="red")

auc(roc_type_endo) # 0.7956
auc(roc_without) # 0.7816
auc(roc_DIE) # 0.786

# 4. Results 
# There is no big differences between two model, so I'll keep the two individus with big leverage
```


#### * Model_version2_250620
After the meeting with Marina in 19th of june
 1) Considering the results of univariable analysis with *p<=0.2*
 2) Considering correlation between :
  - operation_endo_bin and age_firstOP_endoadeno_cat 
  - > age_firstOP_endoadeno_cat
 3) Considering the correlation : 
  - type_endo -> endo_ade_categorie(endo, adeno, both)
  - type_endo -> DIE
  - Remove stade information because of potential correlation with DIE
  
```{r}
# list of significative variables with p-value<=0.2
ls_var_uni_results

#-------------------------------
#   logsitic regression model
#-------------------------------
mod_glm_multivariate_1 <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      metabolic_syndrome +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg +
      dysuria_avant_1_preg,
    data = imputed_data_O3_perID,
    family = binomial
  )

summary(mod_glm_multivariate_1)

#------------
#  Results 
#------------
# 1) summary()
summary(mod_glm_multivariate_1)

# 2) tidy(exponentiate=TRUE) 
library(tibble)
res_multivar_glm_1 <- tidy(mod_glm_multivariate_1, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_mod_1 <- res_multivar_glm_1[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_mod_1 <- res_multivar_glm_mod_1 %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_mod_1

```

##### - Assumptions Check and Sensitivity analysis
###### > Sensitivity test : Outlier and influence check
```{r}
#--------------------------------
#  Multicollinearity Check : <2
#--------------------------------
vif(mod_glm_multivariate_1)

#----------------------------------
#  Outliers and Influences Check    : 17 and 316
#----------------------------------
library(car)
influencePlot(mod_glm_multivariate_1)
# StudRes(Studentized Residuals) : <-2 ou >2 -> outlier
# Hat (Leverage, Hat value) : high leverage point (p/n, p = number of variables, n= number of samples)
# Cook's distance : Residus + Leverage
# CookD > 4/n (n= number of samples)
# - 4/378 = 0.016

plot(mod_glm_multivariate_1, which = 4)  # Cook’s distance
#-> 17 and 316 : Big residus, high leverage -> Outlier and big influence

#--------------------------
#   Sensitivity analysis  
#--------------------------
# 1. Create a new model without 17 and 316 individuals
mod_glm_multivariate_sensitivity <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      metabolic_syndrome +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg +
      dysuria_avant_1_preg,
    data = imputed_data_O3_perID[-c(17,316),],
    family = binomial
  )
summary(mod_glm_multivariate_sensitivity)

# 2. Compare the AIC of two differents model for the sensitivity analysis 
AIC(mod_glm_multivariate_1, mod_glm_multivariate_sensitivity)
# Sensitivity : 389.6878
# Original model : 401.5950

# 3. Compare the AUC of two models for the sensitivity analysis 
library(pROC)
# 3-1) Prepare the FULL DATASET model for ROC and AUC
model_data <- na.omit(imputed_data_O3_perID)
mod_glm_multivariate <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      metabolic_syndrome +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg +
      dysuria_avant_1_preg,
    data = model_data,
    family = binomial
  )
prob <- predict(mod_glm_multivariate, type = "response")
roc_full <- roc(response = model_data$total_complications_bin_perID, predictor = prob)

# 3-2) Prepare the MODIFIED DATASET model for ROC and AUC
mod_glm_multivariate_sensitivity <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      metabolic_syndrome +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg +
      dysuria_avant_1_preg,
    data = model_data[-c(17,316),],
    family = binomial
  )

prob_sensitivity <- predict(mod_glm_multivariate_sensitivity, type = "response")
roc_no_outlier <- roc(response = model_data[-c(17,316),]$total_complications_bin_perID, 
                      predictor = prob_sensitivity)

# 4. Compare 
plot(roc_full, col="blue")
plot(roc_no_outlier, add=TRUE, col="red")
auc(roc_full) # 0.7908
auc(roc_no_outlier) # 0.7894
```


#### * Final model 

```{r}
mod_glm_multivariate_final <- 
  glm(total_complications_bin_perID ~
        diplome_ini2+
        age +
        gynecologique_diseases +
        infertilite_diag+
        autoimmune_diseases+
        metabolic_syndrome+
        nb_mal_sansEndoAde_cat+
        nb_live_births_all+
        history_FC_preterm_IVGIMG_perID+
        age_regle_cat_simple+
        type_endo +
        stade_dvpt_III_IV+
        age_symptome_cat +
        age_firstOP_endoadeno_cat+
        Csection_perID+
        indic_ovaire_touche+
        methode_conception_perID_cat_simple_mod+
        first_preg_age_cat+
        doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID,
    family = binomial
  )

#------------
#  Results 
#------------
# 1) summary()
summary(mod_glm_multivariate_final)

# 2) tidy(exponentiate=TRUE) 
library(tibble)
res_multivar_glm <- tidy(mod_glm_multivariate_final, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_final <- res_multivar_glm[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_final <- res_multivar_glm_final %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_mod
```


### 3) Visualization of results

```{r}
#------------
#  gt table
#------------
library(gt)
res_multivar_glm_final %>%
  gt() %>%
  tab_header(
    title = md("**Multivariate Logistic Regression: **  
    *Factors Associated with Live Birth*")
  ) %>%
  cols_label(
    term = "Variables",
    estimate = "Adjusted OR",
    IC95 = "95% CI",
    p.value = "p-value"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = p.value,
      rows = as.numeric(gsub("<", "", p.value)) < 0.05
    )
  ) %>%
  cols_align(
    align = "left",
    columns = term
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95, p.value)
  )

#---------------
#  Forest plot 
#---------------
res_multivar_glm_final$term <- 
  factor(res_multivar_glm_final$term,
         levels = rev(res_multivar_glm_final$term))

ggplot(res_multivar_glm_final, aes(x = term, y = estimate)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray") +
  coord_flip() +  
  labs(
    title = "Résultats du modèle de régression logistique multivariée",
    x = NULL,
    y = "OR (Odds Ratio avec IC 95%)"
  ) +
  theme_minimal()

#------------------------------
#  Forest plot with the ORs
#------------------------------
library(dotwhisker)
library(grid)

# 1. Limite the maximum conf.high for the forest plot  
res_multivar_glm_final <- res_multivar_glm_final %>% 
  mutate(conf.high_mod = 
           ifelse(conf.high>9, 9.1, conf.high)) %>% 
  dplyr:: select(-conf.high) %>% 
  mutate(conf.high= conf.high_mod) %>% 
  dplyr:: select(-conf.high_mod)

# 2. Redefine the terms
res_multivar_glm_final$term <- c(
  "Intercept",
  "High school diploma or less/Other vs Bac+2",
  "High school diploma or less/Other vs >=Bac+3",
  "Age",
  "Gynecologic diseases: No vs Yes",
  "Infertility diagnostic/treatement: No vs Yes",
  "Autoimmune_diseases: No vs Yes",
  "Metabolic syndrome: No vs Yes",
  "Nb_comorbidity: 0 vs 1",
  "0 vs 2+",
  "Parity",
  "History of misscarriage/preterm/IVGIMG: No vs Yes",
  "age at menarche : <12 vs >=12",
  "<12 vs unknown",
  "type_endo: SPE only vs DIE only",
  "SPE only vs OMA only",
  "SPE only vs ADENO only",
  "SPE only vs Multiple",
  "SPE only vs Unknown",
  "Endo III-IV stage: No vs Yes",
  "Age at symptom onset : <=13 vs (13;15]",
  "<=13 vs (15;22]",
  "<=13 vs 22+",
  "<=13 vs No symptom",
  "Age at 1stOP: <27 vs 27-31",
  "<27 vs 31>=",
  "<27 vs No OP",
  "C-section: No vs Yes",
  "impacted/removed ovaries: No vs Yes",
  "Conception Method: Natural vs Operation",
  "Natural vs Operation+ART",
  "Natural vs ART/Medicaments",
  "Age_1stPregnancy: <30 vs 30-34",
  "<30 vs >=34",
  "Abdominal pain befor 1st preg : No vs Yes"
)

# 3. Create the plot 
O3_1_forestplot <- dwplot(res_multivar_glm_final, show_intercept = TRUE) + 
  geom_text(aes(x = 9.1, y = term, 
                label = paste(estimate, "(", IC95, ")")),
            hjust = -0.1, size = 3) +
  theme_minimal() +
  labs(title = "Results of Multivariate Logistic Regression",
       x= "OR with 95% CI") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.2, size = 12, face = "bold"),
    axis.title.x = element_text(hjust = 0.3, vjust = 0),
    panel.grid = element_blank(),                
    panel.background = element_blank(),           
    axis.line.x =element_blank(), 
    axis.line.y = element_blank(),                
    axis.ticks = element_line(color = "black"),   
    axis.text = element_text(color = "black")     
  ) +
  geom_linerange(aes(xmin = conf.low, xmax = conf.high, y = term),
                 color = "red", size = 0.5, linetype=1)+
  scale_x_continuous(breaks = seq(0, 9, 1)) +
  coord_cartesian(xlim = c(0, 11), ylim = c(1, 37)) +
  geom_segment(aes(x = 1, xend = 1, y = 0.8, yend = 37),
               color = "black", size = 0.5)+
  annotation_custom(
    grob = textGrob("OR (95%CI)", 
                    gp = gpar(fontsize = 10, col = "black",fontface = "bold")),
    xmin = 8.5, ymin = 37) +
  geom_segment(aes(x = -0.3, xend = 9, y = 0.5, yend = 0.5),  
               color = "black", size = 0.5);O3_1_forestplot

O3_1_forestplot

# Save 
#ggsave("C:/Stage_Master2/Stage_M2_Complic_Gross/O3_1_forestplot.jpg", O3_1_forestplot, width = 10, units = "in", dpi = 300)
```


#---------------------------------------------------------
# II. Ordinal outcome 
#---------------------------------------------------------

## 1. Univariable analysis adjusted for age 
### 1) Results Table 

```{r}
# Load necessary libraries
library(dplyr)
library(gtsummary)
library(purrr)
library(MASS) # For polr
library(brant) # Test 

#----------------------
# Define the data
#----------------------
data <- imputed_data_O3_perID %>%
  dplyr::select(-total_complications_bin_perID) %>% 
  mutate(across(where(~ is.factor(.) | is.character(.)), as.factor))

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure the outcome variable is an ordered factor
data$total_complications_cat_perID <-
  factor(data$total_complications_cat_perID,
         levels = c("0", "1", "2+"), 
         ordered = TRUE)

data <- data %>% mutate(
  nb_live_births_all_cat= ifelse(nb_live_births_all>=2,
                                 "2+",nb_live_births_all) %>% as.factor()) 

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables
outcome_variable <- "total_complications_cat_perID"
adjustment_variable <- "age" 

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(data), 
                               c(outcome_variable, adjustment_variable))

#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>%
  tbl_summary(
    by = all_of(outcome_variable),         # Stratify by the outcome
    statistic = list(                     # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%   # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>%
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted Multivariate Ordinal Regression Table
#-------------------------------------------------------------
# This function fits polr(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_polr <- function(predictor, outcome_var, adjust_var, data) {
  # Use backticks for safety with non-standard names and collapse adjustment variables
  # adjust_str <- paste(sprintf("`%s`", adjustment_variable), collapse = " + ") # This was for a vector
  # For a single adjustment variable, it's simpler:
  adjust_str <- sprintf("`%s`", adjust_var) # If adjust_var is a single string

  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)

  # Fit the model
  model <- tryCatch(
    MASS::polr(as.formula(formula_str), data = data, Hess = TRUE), # Use polr from MASS, Hess=TRUE is crucial
    error = function(e) {
      warning(paste("Could not fit ordinal regression model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )

  # If model failed, return NULL
  if (is.null(model)) return(NULL)

  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int = TRUE) %>% # exponentiate = TRUE for Proportional OR
    # Keep rows where 'variable' matches the predictor name
    # This correctly handles factor variables which have multiple rows
    modify_table_body(~ dplyr::filter(.x, variable == predictor))
}

# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_polr(.x, outcome_variable,
                                             adjustment_variable,
                                             data))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables,
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
          )
        )
    ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted Proportional OR (95% CI)**" # Clarify as Proportional OR
    )
  print("Step 2: Adjusted Proportional Odds Ratios Table Created.")
} else {
  warning("No successful ordinal regression models were fitted for adjusted ORs.",
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
# The 'adjustment_variable' is a single string in this setup, so no paste collapse needed
adjust_str <- adjustment_variable

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      paste0("**Age-Adjusted Proportional Odds Model**")) # Update spanner for clarity
    ) %>%
      modify_caption(
        paste0(
          "**Multivariate Ordinal Logistic Regression Analysis for '",
          outcome_variable,
          "' (Adjusted for '", adjust_str, "')**")) # Update caption for clarity

    print("Step 3: Merged Table Created.")

    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '",
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results
#------------------------------------------------------------
O3_tab_2_univariable_final <- as_tibble(tbl_final) %>%
  dplyr::select(-"**OR**",-"**95% CI**",-"**p-value**") # These columns might not exist or be named differently, adjust as needed based on tbl_final structure
# Note: The column names like "**OR**", "**95% CI**", "**p-value**" might not be exact after using polr and gtsummary's custom formatting.
# You might need to inspect `names(as_tibble(tbl_final))` to get the precise column names to exclude.


O3_tab_2_univariable_final

```

*Results* 2025-06-02
diplome_ini2 >=Bac+3 : 0.58 (0.35, 0.97)
infertilite_diag : 	1.58 (1.04, 2.41)
age_regle_cat_simple : >=12 (0.50 (0.31, 0.81)), nsp (0.39 (0.15, 0.97))
stade_dvpt_III_IV : Oui 1.92 (1.10, 3.36)
age_firstOP_endoadeno_cat 31>= : 2.51 (1.09, 5.88)
nb_live_births_all: 1.64 (1.06, 2.53)
history_FC_preterm_IVGIMG_perID : 1.98 (1.23, 3.20)

Mode_accouchement_perID : only c-section 	2.13 (1.33, 3.42)
VD_perID : 	0.48 (0.30, 0.76)
Csection_perID : 	2.08 (1.32, 3.29)

labour_natural_perID : 	0.65 (0.43, 0.99)
doul_abdo_avant_1_preg : 1.90 (1.22, 2.98)


#### - Porportional Odds Assumption (POA) Test
Results : 
	                                                     pval
situation_financiereComfortable/Very comfortable 0.02654584
age_regle_cat_simple>=12                         0.04952934
type_endoOMA only                                0.03962746
age_firstOP_endoadeno_cat31>=                    0.02924595
age_firstOP_endoadeno_catNo operation            0.02520263
nb_live_births_all                               0.01167016
methode_conception_perID_cat_simple_modOperation 0.01624656


```{r}
library(purrr)
library(MASS)
library(brant)
#-----------------------------------------
#   PO TEST for each predictor variable
#-----------------------------------------
brant_results <- map(predictor_variables, function(pred) {
  formula_str <- as.formula(paste0("total_complications_cat_perID ~ ", pred, " + age"))
  model <- tryCatch(
    polr(formula_str, data = data, Hess = TRUE),
    error = function(e) return(NULL)
  )
  if (!is.null(model)) {
    tryCatch(
      brant(model),
      error = function(e) return(NULL)
    )
  } else {
    NULL
  }
})

#
result_list <- list()
for (i in seq_along(brant_results)) {
  pvals <- brant_results[[i]][, 3]
  
  # p < 0.05
  significant_vars <- names(pvals)[pvals < 0.05 & names(pvals) != "Omnibus"]
  significant_pvals <- pvals[pvals < 0.05 & names(pvals) != "Omnibus"]
  
  if (length(significant_vars) > 0) {
    # save in the result_list
    result_list[[i]] <- data.frame(pval = significant_pvals)
  }
}

significant_vars_vector <- do.call(rbind, result_list); significant_vars_vector
significant_vars_vector %>% View
```

## 2. Multivariable  Model and Results
### 1) Model

```{r}
#-----------------------------
#  Ordinal regression model
#-----------------------------
mod_polr_multivariate <-
  polr(
    total_complications_cat_perID  ~
    diplome_ini2 + 
    age + 
    age_regle_cat_simple +
    delai_diag_cat_simple + 
    gynecologique_diseases +
    infertilite_diag + 
    autoimmune_diseases +
    metabolic_syndrome + 
    nb_mal_sansEndoAde_cat +
    age_firstOP_endoadeno_cat + 
    endo_ade_categorie+
    DIE +
    stade_dvpt_III_IV +
    indic_ovaire_touche + 
    trompe_anomalie +
    nb_live_births_all_cat + 
    history_FC_preterm_IVGIMG_perID +
    first_preg_age_cat + 
    conception_PMA_perID +
    conception_operation_perID + 
    Csection_perID +
    labour_natural_perID +
    doul_abdo_avant_1_preg,
    
    data = data,
    Hess = TRUE
  )

#------------
#  Results 
#------------
# 1) summary()
summary(mod_polr_multivariate)
tidy(mod_polr_multivariate, exponentiate = TRUE)

res <- tidy(mod_polr_multivariate)
ci <- confint.default(mod_polr_multivariate)
ci_df <- as.data.frame(ci)
colnames(ci_df) <- c("conf.low", "conf.high")
ci_df$term <- rownames(ci_df)
res2 <- res %>%
  left_join(ci_df, by = "term")
res_multivar_polr <- res2 %>%
  mutate(
    estimate = exp(estimate) %>% round(2),
    conf.low = exp(conf.low) %>% round(2),
    conf.high = exp(conf.high) %>% round(2)
  ) %>% 
  dplyr::select(-std.error,-statistic, -coef.type)

#res_multivar_polr

# 2) Result Table
res_multivar_polr <- res_multivar_polr %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95,conf.low,conf.high) 
res_multivar_polr

# Significants variables : 
res_multivar_polr %>% filter(conf.low<1 & conf.high<1 | conf.high>1 & conf.low>1) 
```

#### - Assumptions check 
- VIF
- POA 
Question : D'autres hypothèses à vérifier ?

```{r}
#---------------------------------------
# Linearity and Independence of Errors
#---------------------------------------
# ???

#------------------------------------
#     Multicollinearity Check : <2
#------------------------------------
library(car)
glm_proxy <- glm(
  as.numeric(total_complications_cat_perID) ~
    diplome_ini2 + 
    age + 
    age_regle_cat_simple +
    delai_diag_cat_simple + 
    
    gynecologique_diseases +
    infertilite_diag + 
    autoimmune_diseases +
    metabolic_syndrome + 
    nb_mal_sansEndoAde_cat +
    
    age_firstOP_endoadeno_cat + 
    endo_ade_categorie+
    DIE +
    stade_dvpt_III_IV +
    indic_ovaire_touche + 
    trompe_anomalie +
    
    nb_live_births_all_cat + 
    history_FC_preterm_IVGIMG_perID +
    first_preg_age_cat + 
    
    conception_PMA_perID +
    conception_operation_perID + 
    
    Csection_perID +
    labour_natural_perID +
    
    doul_abdo_avant_1_preg,
  data = data
)
vif(glm_proxy) # all < 2



#--------------------------------------
# Proportional Odds Aussumption Test
#--------------------------------------
brant(mod_polr_multivariate)
# Omnibus p-value >0.05
# Question 
# age_regle_cat_simple>=12			p-value=0.02 <0.05


```

### 2) Final model

Question 

```{r}
library(VGAM) 
library(purrr)

#-------------------------------------------
#  VGAM : Partial Proportional Odds Model  
# - cumulative logit model
#-------------------------------------------
# In general, the proportional odds assumption is required in ordinal logistic regression. However, for some variables that do not satisfy this assumption, you can set `parallel = FALSE` for those specific variables, using `parallel = FALSE ~ variable_name`, to treat them as exceptions.
# parallel = FALSE ~ age_regle_cat_simple => light proportional odds assumption, meaning that different coefficients are allowed for each cutoff point

mod_vglm_multivariate <- vglm(
  total_complications_cat_perID ~
    diplome_ini2 +
    age +
    age_regle_cat_simple+ 
    delai_diag_cat_simple +
    gynecologique_diseases +
    infertilite_diag +
    autoimmune_diseases +
    metabolic_syndrome +
    nb_mal_sansEndoAde_cat +
    age_firstOP_endoadeno_cat +
    endo_ade_categorie +
    DIE +
    stade_dvpt_III_IV +
    indic_ovaire_touche +
    trompe_anomalie +
    nb_live_births_all_cat +
    history_FC_preterm_IVGIMG_perID +
    first_preg_age_cat +
    conception_PMA_perID +
    conception_operation_perID +
    Csection_perID +
    labour_natural_perID +
    doul_abdo_avant_1_preg,
  family = cumulative(link = "logitlink",
                      parallel = FALSE ~ age_regle_cat_simple), 
  data = data,
  crit = "coefs"
)

summary(mod_vglm_multivariate)

##################
#  모델 해석법
##################
# (1) Coefficients (계수) 해석
# Estimate는 로짓 (log odds) 변화량을 의미해요.
# - 양수면 그 변수 값이 커질수록 (0<1<+2) 높은 카테고리에 속할 확률이 높아지고, 음수면 그 반대입니다.
# Ex ) 
# diplome_ini2>=Bac+3의 Estimate는 0.639로, exp(0.639)=1.89이니까,
# Bac+3 이상 학력이면 복합 합병증 카테고리가 높아질 (더 나쁠) 확률이 약 1.89배 증가한다고 해석 가능해요.
# doul_abdo_avant_1_pregOui (임신 전 복통 있음) 의 Estimate는 -0.78, exp(-0.78) ≈ 0.46 복통이 있으면 더 높은 합병증 카테고리에 속할 확률이 낮아진다고 해석 가능하네요.
# 
# (2) 병렬성 가정을 깬 변수 해석
# age_regle_cat_simple 변수는 병렬성 가정을 깼기 때문에 각 카테고리 구간(1, 2)마다 다른 효과가 있음.
# age_regle_cat_simple>=12:1과 age_regle_cat_simple>=12:2를 따로 해석해야 해요.

#------------
#  Results 
#------------
# 1) summary()
summary(mod_vglm_multivariate)

# 2) Result Table
library(tibble)

coef_mat <- coef(summary(mod_vglm_multivariate))
res_multivar_vglm <- data.frame(
  variable = rownames(coef_mat),
  estimate = coef_mat[, "Estimate"],
  se = coef_mat[, "Std. Error"]
)
res_multivar_vglm <- res_multivar_vglm %>%
  mutate(
    OR = round(exp(estimate),1),
    conf.low = round(exp(estimate - 1.96 * se),2),
    conf.high = round(exp(estimate + 1.96 * se),2)
  )
res_multivar_vglm <- res_multivar_vglm %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>%
  dplyr::select(variable, OR, IC95, conf.low, conf.high)
```

### 3) Visualization of results

```{r}
#------------
#  gt table
#------------
library(gt)
res_multivar_vglm %>%
  gt() %>%
  tab_header(
    title = md("**Multivariate Ordinal Regression:**\n
    - Factors Associated with number of obstetric complications")
  ) %>%
  cols_label(
    variable   = "Variables",
    OR = "Adjusted OR",
    IC95 = "95% CI"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = OR,
      rows = (conf.low > 1 & conf.high > 1) | (conf.low < 1 & conf.high < 1)
    )
  ) %>%
  cols_align(
    align = "left",
    columns = variable
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95)
  ) %>% 
 cols_hide(columns = c(conf.low, conf.high))

#---------------------------
#   Simple Visualization 
#---------------------------
# 1) OR (IC95)
res_multivar_vglm <- res_multivar_vglm %>%
  filter(!grepl("\\(Intercept\\)", variable))

# 2) Forest plot
ggplot(res_multivar_vglm, aes(x = reorder(variable, OR), 
                              y = ifelse(OR >4,NA,OR))) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = ifelse(conf.high>4,4,conf.high)), width = 0.2) +
  coord_flip() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(x = "Var", y = "OR(IC95)", title = "Influence per variable") +
  theme_minimal()
```

```{r}
#------------------------------
#  Forest plot with the ORs
#------------------------------
library(dotwhisker)
library(grid)
library(ggplot)

# 1. Limite the maximum conf.high for the forest plot  
res_multivar_vglm <- res_multivar_vglm %>% 
  mutate(conf.high_mod = 
           ifelse(conf.high>11, 11.1, conf.high)) %>% 
  dplyr:: select(-conf.high) %>% 
  mutate(conf.high= conf.high_mod,
         term=variable) %>% 
  dplyr:: select(-conf.high_mod, -variable) 

# 2. Remove intercept informations 
res_multivar_vglm_mod <- res_multivar_vglm[-c(1,2),]

# 3. Set the order of term values 
res_multivar_vglm_mod$term <- factor(res_multivar_vglm_mod$term, 
                                 levels = rev(res_multivar_vglm_mod$term))

# 4. Create the forest plot 
O3_2_forestplot <- ggplot(res_multivar_vglm_mod, aes(y = term, x = OR)) +
  geom_point(color = "blue") +  # OR 점
  geom_linerange(aes(xmin = conf.low, xmax = conf.high), 
                 color = "red", size = 0.8) +  
  geom_text(aes(x = 12.1, label = paste0(round(OR, 2), " (", IC95, ")")),
            hjust = 0, size = 3) +
  geom_vline(xintercept = 1, color = "black") +  # 기준선 OR=1
  scale_x_continuous(breaks = seq(0, 11, 1), limits = c(0, 13)) +
  coord_cartesian(ylim = c(0.5, length(levels(res_multivar_vglm_mod$term)) + 1.5),
                  xlim= c(0,15)) +
  theme_minimal() +
  labs(title = "Results of Multivariate Logistic Regression",
       x = "Odds Ratio (95% CI)",
       y = NULL) +
  theme(
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  ) + 
  annotation_custom(
    grob = textGrob("OR (IC95)", 
                    gp = gpar(fontsize = 10, fontface = "bold")),
    xmin = 13, xmax = 13, 
    ymin = length(levels(res_multivar_vglm_mod$term)) + 1.3,  
    ymax = length(levels(res_multivar_vglm_mod$term)) + 1.3
  )
O3_2_forestplot



# 2. Redefine the terms
# res_multivar_glm_mod$term <- c(
#   "High school diploma or less/Other vs Bac+2",
#   "High school diploma or less/Other vs >=Bac+3",
#   "Age",
#   "age at menarche : <12 vs >=12 : 0 vs 1",
#   "<12 vs >=12 : 1 vs 2+",
#   "<12 vs nsp : 0 vs 1",
#   "<12 vs nsp : 1 vs 2+",
#   "Delay diagnostic years :<8 vs 13+",
#   "<8 vs 8-13",
#   "<8 vs No symptom",
#   "Gynecologic diseases: No vs Yes",
#   "Infertility diagnostic/treatement: No vs Yes",
#   "Autoimmune_diseases: No vs Yes",
#   "Metabolic syndrome: No vs Yes",
#   "Nb_comorbidity: 0 vs 1",
#   "0 vs 2+",
#   "Age at 1stOP: <27 vs 27-31",
#   "<27 vs 31>=",
#   "<27 vs No OP",
#   "Category : only Endo vs only adeno",
#   "only Endo vs Both",
#   "DIE : No vs Yes",
#   "Endo III-IV stage: No vs Yes",
#   "impacted/removed ovaries: No vs Yes",
#   "Anormal fallopian tube: No vs Yes",


#   "History of misscarriage/preterm/IVGIMG: No vs Yes",
#   "age at menarche : <12 vs >=12",
#   "<12 vs unknown",
#   "type_endo: SPE only vs DIE only",
#   "SPE only vs OMA only",
#   "SPE only vs ADENO only",
#   "SPE only vs Multiple",
#   "SPE only vs Unknown",

#   "Age at symptom onset : <=13 vs (13;15]",
#   "<=13 vs (15;22]",
#   "<=13 vs 22+",
#   "<=13 vs No symptom",

#   "C-section: No vs Yes",


#   "Conception Method: Natural vs Operation",
#   "Natural vs Operation+ART",
#   "Natural vs ART/Medicaments",
#   "Age_1stPregnancy: <30 vs 30-34",
#   "<30 vs >=34",
#   "Abdominal pain befor 1st preg : No vs Yes"
# )

# 3. Create the plot 
O3_2_forestplot <- dwplot(res_multivar_vglm, show_intercept = TRUE) + 
  geom_text(aes(x = 9.1, y = term, 
                label = paste(OR, "(", IC95, ")")),
            hjust = -0.1, size = 3) +
  theme_minimal() +
  labs(title = "Results of Multivariate Logistic Regression",
       x= "OR with 95% CI") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.2, size = 12, face = "bold"),
    axis.title.x = element_text(hjust = 0.3, vjust = 0),
    panel.grid = element_blank(),                
    panel.background = element_blank(),           
    axis.line.x =element_blank(), 
    axis.line.y = element_blank(),                
    axis.ticks = element_line(color = "black"),   
    axis.text = element_text(color = "black")     
  ) +
  geom_linerange(aes(xmin = conf.low, xmax = conf.high, y = variable),
                 color = "red", size = 0.5, linetype=1)+
  scale_x_continuous(breaks = seq(0, 9, 1)) +
  coord_cartesian(xlim = c(0, 11), ylim = c(1, 37)) +
  geom_segment(aes(x = 1, xend = 1, y = 0.8, yend = 37),
               color = "black", size = 0.5)+
  annotation_custom(
    grob = textGrob("OR (95%CI)", 
                    gp = gpar(fontsize = 10, col = "black",fontface = "bold")),
    xmin = 8.5, ymin = 37) +
  geom_segment(aes(x = -0.3, xend = 9, y = 0.5, yend = 0.5),  
               color = "black", size = 0.5);O3_2_forestplot

O3_2_forestplot
# Save 
#ggsave("Stage_M2_Complic_Gross/O3_2_forestplot.jpg", O3_2_forestplot,height = plot_height, width = 7, units = "in", dpi = 300)
```



