---
title: "04_O3_Statistical_Analysis_V2"
author: "Minji NAM"
date: "2025-05-13"
output:
  pdf_document: default
  html_document: default
---

*Name* : **3 Objective**, Statistical Analysis for the participant population 
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Statistical Analysis
*Date of creation* : 2025-06-20
*Date of update* : 2025-06-20
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

# * V2 : O3 per ID*

$Libraries$
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(purrr) # reduce()
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gt)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
library(chromote)
library(car)
library(mediation)
library(shiny)
library(shinydashboard)
library(data.table)
library(DT)
library(mice)
library(dagitty)
library(ggdag)
library(gridExtra)
library(grid)
library(dotwhisker)
#install.packages("writexl")
library(writexl)
```

# 0. Import the imputed data 
 Population : 310 participants without nulliparous women(n=57) and having NA value in outcome variable(n=11)
 
```{r}
imputed_data_O3_perID_V2<- 
  read.csv("imputed_data_O3_perID_V2.csv") %>%
  dplyr::select(-X)

#----------------------
#  Imputed Dataframe
#----------------------
imputed_data_O3_perID_V2 <- imputed_data_O3_perID_V2 %>%
  # Filtering nulliparous participants and NA outcome
  filter(!is.na(total_complications_bin_perID)) %>% 
   mutate(across(where(~ is.factor(.) | is.character(.)),
                 as.factor)) %>% 
  mutate(
    diplome_ini2 = factor(
      diplome_ini2,
      levels = c("<=Baccalauréat/Aucun/Autre", "Bac+2", ">=Bac+3"),
      ordered = FALSE
    ),
    
    situation_financiere   = factor(
      situation_financiere,
      levels = c(
        "Balanced",
        "Difficult/In debt",
        "Comfortable/Very comfortable"
      ),
      ordered = FALSE
    ),
    
    family_atcd_EndoAdeno_cat =
      factor(
        family_atcd_EndoAdeno_cat,
        levels = c("Pas d'antécédents", "Antécédents")
      ),
    
    family_atcd_doul_pelv =
      factor(
        family_atcd_EndoAdeno_cat,
        levels = c("Pas d'antécédents", "Antécédents")
      ),
    
    endo_ade_categorie = factor(
      endo_ade_categorie,
      levels = c("Only endometriosis", "Only adenomyosis", "Both")
    ),
    
    type_endo =
      factor(
        type_endo,
        levels = c(
          "DIE only",
          "SPE only",
          "OMA only",
          "ADENO only",
          "Multiple types",
          "Aucun mentionné/nsp"
        )
      ),
    
    stade_dvpt_legere_severe = factor(
      stade_dvpt_legere_severe,
      levels = c(
        "No surgery",
        "I-II",
        "III-IV",
        "Non et nsp"
      )
    ), 
    
    op_avant_1_preg = factor(
      op_avant_1_preg,
      levels = c("Non","Oui")
    ),
    
    complication_operation = factor(
      complication_operation,
      levels = c("Non", "Oui", "No surgery")
    ),
    
    age_symptome_cat = factor(
      age_symptome_cat,
      levels = c(
        "<= 13",
        "(13;15]",
        "(15;22]",
        "22+",
        "No symptoms"
      ),
      ordered = FALSE
    ),

    age_firstOP_endoadeno_cat = factor(
      age_firstOP_endoadeno_cat,
      levels = c("No operation","<28", "28-31", "31>="),
      ordered = FALSE
    ),
    
    nb_mal_sansEndoAde_cat = factor(
      nb_mal_sansEndoAde_cat,
      levels = c("0", "1", "2 ou plus"),
      ordered = FALSE
    ),
    
    comorbidities_bin = case_when(
      nb_mal_sansEndoAde_cat =="0" ~ "Non",
      TRUE~ "Oui"
    ),
    
    delai_diag_cat = factor(
      delai_diag_cat,
      levels = c("<= 5", "(5;10]", "(10;15]", "15+", "No symptoms"),
      ordered = FALSE
    ),
    
    nb_live_births_all_cat= 
      ifelse(as.numeric(as.character(nb_live_births_all)) >=2, 
             "2+",nb_live_births_all) %>% as.factor(),
    
    nb_total_gross_cat = 
      ifelse(as.numeric(as.character(nb_total_gross)) >=2, 
             "2+",nb_total_gross) %>% as.factor(),
    
    Travail_accouchement_perID = factor(
      Travail_accouchement_perID,
      levels = c(
        "Seul travail spontané",
        "Seul travail provoqué",
        "Travail mixte",
        "Pas de travail"
      )
    ),
    
    Mode_accouchement_perID = factor(
      Mode_accouchement_perID,
      levels = c(
        "Seule voie Basse", 
        "Seule Césarienne", 
        "Both")
    ),
    
    first_preg_age_cat = factor(
      first_preg_age_cat,
      levels = c("<29", "29-34", ">=34"),
      ordered = FALSE
    ),
    
    total_complications_bin_perID = 
      as.factor(total_complications_bin_perID),
    
    total_complications_cat_perID= factor(
      total_complications_cat_perID, 
      levels= c("0","1","2+"),
      ordered = TRUE
    )
  )
```

#------------------------------------
# I. Binary outcome 
#-------------------------------------

## 0. Definition of data
```{r}
data  <-  imputed_data_O3_perID_V2 %>%
  dplyr::select(-total_complications_cat_perID,
                -Complications_cat_perID_grouped,
                -Age_premier_diag_endo_ade_cat,
                -stade_dvpt_I_II,
                -stade_dvpt_III_IV,
                -conception_natural_perID,
                -conception_natural_perID,
                -conception_operation_perID,
                -conception_PMA_perID,
                -labour_natural_perID,
                -labour_induced_perID,
                -delai_diag, 
                -delai_diag_cat,
                -situation_emploi_ini2_cat,
                -Csection_perID,
                -VD_perID,
                -age_regles_ini,
                -age_regle_cat,
                -nb_live_births_all_cat,
                -nb_total_gross_cat,
                -nb_mal_sansEndoAde_cat,
                -nb_total_gross,
                -first_preg_age_cat) %>% 
  dplyr::select(37,1,2,3,4,26,27,28,29,30,31,
                38,5,6,7,8,9,19,10,11,12,13,14,15,16,17,18,
                21,22,23,24,25,
                20,32,33,34,35,36)

```


## 1. Function for the results

```{r}
create_results_table_nested_binary <- 
  function(data, outcome_variable, adjust_sets) {
  
  # 0. Predictor variables
  all_vars <- colnames(data)
  all_adjust_vars <- unique(unlist(adjust_sets))
  predictor_variables <- setdiff(all_vars, c(outcome_variable, all_adjust_vars))
  
  # 1. Descriptive Table
  tbl_descriptive <- data %>%
    dplyr::select(all_of(outcome_variable),
                  all_of(predictor_variables)) %>%
    tbl_summary(
      by = all_of(outcome_variable),
      statistic = list(
        all_continuous() ~ "{mean} ({sd})",
        all_categorical() ~ "{n} ({p}%)"
      ),
      missing = "no"
    ) %>%
    modify_header(label ~ "**Characteristic**") %>%
    bold_labels()
  
  # 2. Function for one model run
  fit_model <- function(predictor, outcome, adjust_vars, data) {
    adjust_str <- if (is.null(adjust_vars)) "" else paste(sprintf("`%s`", adjust_vars), collapse = " + ")
    formula_str <- if (adjust_str == "") {
      paste0("`", outcome, "` ~ `", predictor, "`")
    } else {
      paste0("`", outcome, "` ~ `", predictor, "` + ", adjust_str)
    }
    
    model <- tryCatch(
      glm(as.formula(formula_str), data = data, family = binomial),
      error = function(e) NULL
    )
    
    if (is.null(model)) return(NULL)
    
    tbl_regression(model, exponentiate = TRUE, conf.int = TRUE) %>%
      modify_table_body(~ dplyr::filter(.x, variable == predictor))
  }
  
  # 3. Run models for each predictor and adjustment set
  all_models <- list()
  
  for (adjust_name in names(adjust_sets)) {
    adjust_vars <- adjust_sets[[adjust_name]]
    tables <- map(predictor_variables, ~ fit_model(.x, outcome_variable, adjust_vars, data))
    tables <- tables[!sapply(tables, is.null)]
    
    if (length(tables) > 0) {
      all_models[[adjust_name]] <- tbl_stack(tables, group_header = NULL) %>%
        modify_table_body(
          ~ .x %>%
            mutate(
              est_ci = if_else(
                !is.na(estimate),
                sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
                NA_character_
              )
            )
        ) %>%
        modify_header(
          label ~ "**Characteristic**",
          est_ci ~ paste0("**", adjust_name, " OR (95% CI)**")
        )
    }
  }
  
  # 4. Merge all model tables side-by-side
  if (length(all_models) > 0) {
    tbl_adjusted_all <- tbl_merge(
      tbls = c(list(tbl_descriptive), all_models),
      tab_spanner = c("**Descriptive Statistics**", names(all_models))
    ) %>%
      modify_caption(
        paste0("**Logistic Regression Models for '", outcome_variable, "'**")
      )
    
    return(tbl_adjusted_all)
  } else {
    warning("No models fitted.")
    return(tbl_descriptive)
  }
}
```


## 2. Tables of results 

```{r}
#--------------------------
# Definition of variables
#--------------------------
outcome_variable <- "total_complications_bin_perID"
adjust_sets <- list(
  
  # # CRUDE without adjustment
  # "Crude" = NULL,
  
  # Adjustment : Age
  "Age_adjusted" = c("age"),
  
  # Adjustment : Age + sociodemographic
  "Age_Socio_adjusted" = c(
    "age",
    "diplome_ini2",
    "situation_financiere",
    "statut_marital_ini_cat"
  ),
  
  # Adjustment : Age + sociodemographic + pregnancy related variables
  "Fully_adjusted" = c(
    "age",
    "diplome_ini2",
    "situation_financiere",
    "statut_marital_ini_cat",
    "nb_live_births_all",
    "history_FC_preterm_IVGIMG_perID",
    "first_preg_age",
    "Mode_accouchement_perID",
    "Travail_accouchement_perID",
    "methode_conception_perID_cat_simple_mod"
  )
)

#----------------------------
#  Application of function 
#----------------------------
res_table_adjustment_binary <- 
  create_results_table_nested_binary(data = data , 
                            outcome_variable = outcome_variable, 
                            adjust_sets = adjust_sets) %>% as.data.frame()

#------------------------------
#  Final Results table
#------------------------------
res_table_adjustment_binary_final<- res_table_adjustment_binary %>%
  dplyr::select(1,2,3, 7,11,15) 

res_table_adjustment_binary_final %>% View

#------------------
#  Save the table
#------------------
write_xlsx(res_table_adjustment_binary_final, 
           "C:/Stage_Master2/Stage_M2_Complic_Gross/O3_bin_res_table_adjusted.xlsx")
```


## 3. Minji_Multivariable analysis
### 3-1. univariable analysis adjusted for age 
#### 1) Results Table 

```{r}
library(pmsampsize)
# data %>% group_by(total_complications_bin_perID) %>% summarise(n=n()) %>% mutate(somme= sum(n), prop= n/somme)

pmsampsize(type = "b",     # 'b' for binary outcome
           rsquared = 0.05, 
           parameters = 13, 
           prevalence = 0.59, 
           shrinkage = 0.9,
           cstatistic = 0.75)
```



```{r}
#----------------------
#  Define the data
#----------------------
data <- imputed_data_O3_perID_V2 %>% 
  dplyr::select(-total_complications_cat_perID,
                -Complications_cat_perID_grouped,
                -Age_premier_diag_endo_ade_cat,
                -stade_dvpt_I_II,
                -stade_dvpt_III_IV)

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables 
outcome_variable <- "total_complications_bin_perID"
adjustment_variable <- "age"

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure outcome variable is suitable for logistic regression (factor or 0/1 numeric)
if (!is.factor(data[[outcome_variable]]) && !all(data[[outcome_variable]] %in% c(0, 1, NA))) {
  stop(paste("Outcome variable '", outcome_variable, "' must be a factor or numeric with values 0/1."))
  }

# Convert outcome to factor if it's 0/1 numeric, as tbl_summary needs factor for 'by'
if (is.numeric(data[[outcome_variable]])) {
    data <- data %>%
        mutate(!!sym(outcome_variable) := factor(!!sym(outcome_variable)))
    print(paste("INFO: Converted numeric outcome variable '",
                outcome_variable, "' to factor."))
    }

# Ensure adjustment variable exists
if (!all(adjustment_variable %in% names(data))) {
  stop(paste("Some adjustment variables not found in the dataframe:",
             paste(setdiff(adjustment_variable, names(data)), collapse = ", ")))
}
# if (!adjustment_variable %in% names(data)) {
#   stop(paste("Adjustment variable '", adjustment_variable, "' not found in the dataframe."))
#   }

missing_vars <- setdiff(adjustment_variable, names(data))
if (length(missing_vars) > 0) {
  stop(paste("Adjustment variable(s) not found in the dataframe:", paste(missing_vars, collapse = ", ")))
  }

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(data),
                               c(outcome_variable, adjustment_variable))

if (length(predictor_variables) == 0) {
    stop("No predictor variables found after excluding outcome and adjustment variables.")
  }

print(paste("Outcome:", outcome_variable))
print(paste("Adjustment:", paste(adjustment_variable, collapse=", ")))
print(paste("Predictors:", paste(predictor_variables, collapse=", ")))

#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>% 
  tbl_summary(
    by = all_of(outcome_variable),           # Stratify by the outcome
    statistic = list(                         # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",                           
    digits = list(all_continuous() ~ 1, 
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%     # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>% 
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted multivariable Logistic Regression Table
#-------------------------------------------------------------
# This function fits glm(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_glm <- function(predictor, outcome_var, adjust_var, data) {
  # Use backticks for safety with non-standard names
  adjust_str <- paste(sprintf("`%s`", adjustment_variable), collapse = " + ")
    # adjust_vars는 벡터
  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)
  
  # Fit the model
  model <- tryCatch(
    glm(as.formula(formula_str), data = data, family = binomial(link = "logit")),
    error = function(e) {
      warning(paste("Could not fit model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )
  
  # If model failed, return NULL
  if (is.null(model)) return(NULL)
  
  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int=TRUE) %>%
    # Keep rows where 'variable' matches the predictor name
    # This correctly handles factor variables which have multiple rows
    modify_table_body(~ dplyr::filter(.x, variable == predictor)) 
  }

# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_glm(.x, outcome_variable,
                                            adjustment_variable,
                                            data))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables, 
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
            )
          )
      ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted OR (95% CI)**"
    )
  print("Step 2: Adjusted Odds Ratios Table Created.")
} else {
  warning("No successful models were fitted for adjusted ORs.", 
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
adjust_str <- paste(adjustment_variable, collapse = " + ")

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      "**Age-Adjusted Model**")
    ) %>%
      modify_caption(
        paste0(
          "**multivariable Logistic Regression Analysis for '",
          outcome_variable, 
          "' (Adjusted for '", adjust_str, "')**"))

    print("Step 3: Merged Table Created.")
    
    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '", 
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results 
#------------------------------------------------------------
table_res2_final <- as_tibble(tbl_final) %>% 
  dplyr::select(-"**OR**",-"**95% CI**",-"**p-value**")

# save the table : 2025-05-30
#table_univariable_O3_bin<- table_res2_final
table_res2_final
```

#### 2) Pre-selection of variables for the multivariable regression model

 - P-value <= 0.2
```{r}
ls_var_uni_results <- tbl_final$table_body %>% 
  filter(p.value_2<=0.2) %>%
  distinct(variable) %>% pull()
ls_var_uni_results
```


### 3-2. multivariable logistic regression

#### 1) DAG
```{r}
#---------------------
#  Creation  of DAG 
#---------------------
library(dagitty)
library(ggdag)
dag_code <- "dag {
  age -> total_complications_bin_perID
  age -> statut_marital_ini_cat
  age -> diplome_ini2
  diplome_ini2 -> situation_financiere
  age -> situation_financiere
  age -> situation_emploi_ini2
  age -> operation_endo_bin
  age -> nb_mal_sansEndoAde_cat
  age -> metabolic_syndrome
  age -> nb_total_gross
  
  age_regle_cat_simple -> metabolic_syndrome
  age_regle_cat_simple -> total_complications_bin_perID
  
  metabolic_syndrome -> nb_mal_sansEndoAde_cat
  metabolic_syndrome -> total_complications_bin_perID
  metabolic_syndrome -> history_FC_preterm_IVG_IMG
  
  gynecologique_diseases -> total_complications_bin_perID
  gynecologique_diseases -> nb_mal_sansEndoAde_cat
  gynecologique_diseases -> Csection_perID
  gynecologique_diseases -> nb_total_gross
  gynecologique_diseases -> infertilite_diag
  gynecologique_diseases -> history_FC_preterm_IVG_IMG
  
  autoimmune_diseases -> history_FC_preterm_IVG_IMG
  autoimmune_diseases -> nb_mal_sansEndoAde_cat
  autoimmune_diseases -> total_complications_bin_perID
  
  nb_mal_sansEndoAde_cat -> total_complications_bin_perID
  nb_mal_sansEndoAde_cat -> Csection_perID

  Csection_perID -> total_complications_bin_perID
  total_complications_bin_perID -> Csection_perID
  
  type_endo -> operation_endo_bin
  type_endo -> stade_dvpt_III_IV
  type_endo -> stade_dvpt_legere_severe
  type_endo -> Endo_ade_categorie
  type_endo -> indic_ovaire_touche
  
  DIE -> stade_dvpt_III_IV
  stade_dvpt_III_IV -> operation_endo_bin
  operation_endo_bin -> indic_ovaire_touche
  operation_endo_bin -> trompe_anomalie 
  operation_endo_bin -> nb_total_gross
  operation_endo_bin -> complication_operation

  indic_ovaire_touche -> infertilite_diag
  trompe_anomalie -> infertilite_diag
  trompe_anomalie -> nb_total_gross
  
  op_avant_1_preg -> operation_endo_bin
  op_avant_1_preg -> infertilite_diag
  
  infertilite_diag -> operation_endo_bin
  infertilite_diag -> conception_PMA_perID
  infertilite_diag -> conception_operation_perID
  
  conception_operation_perID -> operation_endo_bin
  conception_operation_perID -> nb_total_gross
  
  maternal_age -> infertilite_diag
  maternal_age -> conception_PMA_perID
  maternal_age -> total_complications_bin_perID
  maternal_age -> nb_total_gross
  maternal_age -> history_FC_preterm_IVG_IMG
  
  first_preg_age_cat -> infertilite_diag
  first_preg_age_cat ->  conception_PMA_perID
  first_preg_age_cat -> total_complications_bin_perID
  first_preg_age_cat -> nb_total_gross
  first_preg_age_cat -> history_FC_preterm_IVG_IMG
  
  conception_PMA_perID -> total_complications_bin_perID
  conception_PMA_perID -> nb_total_gross
  
  nb_total_gross -> nb_live_births_all
  nb_live_births_all -> total_complications_bin_perID
  nb_live_births_all -> live_birth_bin_atleast_once_all
  
  history_FC_preterm_IVGIMG_perID -> nb_live_births_all
  history_FC_preterm_IVGIMG_perID -> nb_total_gross
  
  first_preg_age_cat -> op_avant_1_preg
  age_firstOP_endoadeno_cat -> op_avant_1_preg
  
  conception_operation_perID -> methode_conception_perID_cat_simple_mod
  conception_PMA_perID -> methode_conception_perID_cat_simple_mod

  methode_conception_perID_cat_simple_mod -> live_birth_bin_atleast_once_all
  methode_conception_perID_cat_simple_mod -> operation_endo_bin 
  methode_conception_perID_cat_simple_mod -> total_complications_bin_perID
}"

dag_model <- dagitty(dag_code)

#------------------------
# Visualisation of DAG
#------------------------
library(ggplot2)
library(ggrepel)
library(ggdag)

# tidy DAG
dag_tidy <- tidy_dagitty(dag_model)
nodes_df<- dag_tidy$data
edges <- edges(dag_model)
ggplot() +
  # Edges
  geom_segment(data = nodes_df, 
               aes(x = x, y = y, xend = xend, yend = yend),
               arrow = arrow(length = unit(0.01, "npc"), 
                             angle = 40, type="closed"), 
               color = "gray50", linewidth = 0.6) +  # size 대신 linewidth
  
  # Node
  geom_point(data = nodes_df, 
             aes(x = x, y = y), 
             shape = 21, fill = "lightcyan", 
             color = "steelblue", size = 10, stroke = 0.4,
             alpha = 0.2) +
  
  # label text 
  geom_text_repel(data = distinct(nodes_df, name, .keep_all = TRUE), 
                  aes(x = x, y = y, label = name), 
                  color = "brown", size = 2.5) +
  
  
  # theme
  theme_dag(base_size = 10, base_family = "Arial") +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "white"),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_blank()
  )


library(dplyr)
# geom_point size = 10 →  10*0.035 = 0.35 
r <- 0.35  

edges_adjusted <- nodes_df %>%
  mutate(
    dx = xend - x,
    dy = yend - y,
    dist = sqrt(dx^2 + dy^2),
    xend_adj = xend - r * (dx / dist),
    yend_adj = yend - r * (dy / dist)
  )

ggplot() +
  geom_segment(
    data = edges_adjusted,
    aes(
      x = x,
      y = y,
      xend = xend_adj,
      yend = yend_adj
    ),
    arrow = arrow(
      length = unit(0.015, "npc"),
      angle = 30,
      type = "open"
    ),
    color = "gray50",
    linewidth = 0.6
  ) +
  geom_point(
    data = nodes_df,
    aes(x = x, y = y),
    shape = 21,
    fill = "lightcyan",
    color = "steelblue",
    size = 10,
    stroke = 0.4,
    alpha = 0.2
  ) +
  geom_text_repel(
    data = distinct(nodes_df, name, .keep_all = TRUE),
    aes(x = x, y = y, label = name),
    color = "brown",
    size = 2.5
  ) +
  theme_dag(base_size = 10, base_family = "Arial") +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "white"),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_blank()
  )

#-------------------------------------
# Finding the adjustements variables
#-------------------------------------
adjustmentSets(dag_model, 
               exposure ="age_regle_cat_simple", 
               outcome = "total_complications_bin_perID",
               type= "canonical")
```

#### 2) Final Model and Results
##### (1) Reference Model
After the meeting with Marina in 19th of june
 1) Considering the results of univariable analysis with *p<=0.2*
 2) Considering correlation between :
  - operation_endo_bin and age_firstOP_endoadeno_cat 
  - > age_firstOP_endoadeno_cat
 3) Considering the correlation : 
  - type_endo -> endo_ade_categorie(endo, adeno, both)
  - type_endo -> DIE
  - Remove stade information because of potential correlation with DIE
  4) age at menarche and age at 1st pregnancy in continuous 
  
  
```{r}
# list of significative variables with p-value<=0.2
ls_var_uni_results

#-----------------------------------------
#   Reference logsitic regression model
#-----------------------------------------
mod_glm_multivariable_0 <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      # situation_financiere+
      # statut_marital_ini_cat+
    
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +               #ref
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +                 # ref
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID_V2,
    family = binomial
  )

# summary(mod_glm_multivariable_0)
# AIC(mod_glm_multivariable_0)
#------------
#  Results 
#------------
# 1) summary()
summary(mod_glm_multivariable_0)

# 2) tidy(exponentiate=TRUE) 
library(tibble)
library(broom)
res_multivar_glm_0 <- tidy(mod_glm_multivariable_0, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_mod_0 <- res_multivar_glm_0[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_mod_0 <- res_multivar_glm_mod_0 %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_mod_0


# Resultats : variables significatives
res_multivar_glm_mod_0 %>% filter(p.value<0.05) %>% distinct(term, IC95)
```

##### (2) Log-linearity test : OK

```{r}
# In R
df <- imputed_data_O3_perID_V2
df$log_x <- log(df$age)
df$x_logx <- df$age * log(df$age)

model <- glm(total_complications_bin_perID ~ age + x_logx, family = binomial, data = df)
summary(model)  # Check p-value for x_logx

```

##### (3) Test : AIC + Pseudo-R² + AUC + Hosmer–Lemeshow test

mod1: first_preg_age_cat + age_regles_ini
mod2: first_preg_age + age_regles_ini
*mod3: first_preg_age + age_regle_cat_simple*
  - AIC : le plus bas
  - Pseudo-R² (Explication: >0.2 bon, >=0.3 Très fort) : le plus élevé
  - AUC : haut
  - Hosmer–Lemeshow test (goodness-of-fit) : p > 0.05
mod4: first_preg_age_cat + age_regle_cat_simple

```{r}
library(pROC)
library(pscl)
library(ResourceSelection)
library(car)
#-----------------------------
# Compare differents models  
#-----------------------------
compare_model_metrics <- function(model, data, outcome_var) {
  # Without NA
  data <- data[!is.na(data[[outcome_var]]), ]

  # AUC
  pred <- predict(model, type = "response")
  roc_obj <- roc(data[[outcome_var]], pred)
  auc_val <- auc(roc_obj)

  # Pseudo-R2 (McFadden)
  pseudo_r2 <- pR2(model)["McFadden"]

  # Hosmer-Lemeshow
  y_obs <- ifelse(df$total_complications_bin_perID == "Oui", 1, 0)
  hoslem <- hoslem.test(y_obs, fitted(model), g = 10)
  hoslem_p <- hoslem$p.value

  return(list(
    AIC = AIC(model),
    Pseudo_R2 = pseudo_r2,
    AUC = auc_val,
    Hosmer_Lemeshow_p = hoslem_p
  ))
}

# Remove NA in outcome variable
df <- imputed_data_O3_perID_V2[!is.na(imputed_data_O3_perID_V2$total_complications_bin_perID), ]

# Common variables
common_formula <- paste(
  "diplome_ini2 + age +",
  
  "gynecologique_diseases + infertilite_diag + autoimmune_diseases +   comorbidities_bin +",
  
  "endo_ade_categorie + DIE + age_firstOP_endoadeno_cat + indic_ovaire_touche +",
  
  "age_symptome_cat + nb_live_births_all + history_FC_preterm_IVGIMG_perID +",
  "Mode_accouchement_perID + Travail_accouchement_perID + methode_conception_perID_cat_simple_mod +",
  
  "doul_abdo_avant_1_preg"
)

# Model formula 
f1 <- as.formula(paste("total_complications_bin_perID ~", common_formula, "+ first_preg_age_cat + age_regles_ini"))
f2 <- as.formula(paste("total_complications_bin_perID ~", common_formula, "+ first_preg_age + age_regles_ini"))
f3 <- as.formula(paste("total_complications_bin_perID ~", common_formula, "+ first_preg_age + age_regle_cat_simple"))
f4 <- as.formula(paste("total_complications_bin_perID ~", common_formula, "+ first_preg_age_cat + age_regle_cat_simple"))

# Modeling
mod1 <- glm(f1, data = df, family = binomial)
mod2 <- glm(f2, data = df, family = binomial)
mod3 <- glm(f3, data = df, family = binomial)
mod4 <- glm(f4, data = df, family = binomial)

# Results
res1 <- compare_model_metrics(mod1, df, "total_complications_bin_perID")
res2 <- compare_model_metrics(mod2, df, "total_complications_bin_perID")
res3 <- compare_model_metrics(mod3, df, "total_complications_bin_perID")
res4 <- compare_model_metrics(mod4, df, "total_complications_bin_perID")

# Results dataframe
results_df <- data.frame(
  Model = c("mod1: first_preg_age_cat + age_regles_ini",
            "mod2: first_preg_age + age_regles_ini",
            "mod3: first_preg_age + age_regle_cat_simple",
            "mod4: first_preg_age_cat + age_regle_cat_simple"),
  AIC = c(res1$AIC, res2$AIC, res3$AIC, res4$AIC),
  Pseudo_R2 = c(res1$Pseudo_R2, res2$Pseudo_R2, res3$Pseudo_R2, res4$Pseudo_R2),
  AUC = c(res1$AUC, res2$AUC, res3$AUC, res4$AUC),
  Hosmer_Lemeshow_p = c(res1$Hosmer_Lemeshow_p, res2$Hosmer_Lemeshow_p, res3$Hosmer_Lemeshow_p, res4$Hosmer_Lemeshow_p)
)

print(results_df)

```

##### (4) V2 Model with *first_preg_age + age_regle_cat_simple*
- Marina a voulu tenter cette variable en continue au lieu de categoriel

```{r}
#-----------------------------------------
#   Reference logsitic regression model
#-----------------------------------------
mod_glm_multivariable_1 <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +           # in categorical
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +                 # in continuous
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID_V2,
    family = binomial
  )

summary(mod_glm_multivariable_1)

#------------
#  Results 
#------------
# 1) summary()
summary(mod_glm_multivariable_1)

# 2) tidy(exponentiate=TRUE) 
library(tibble)
res_multivar_glm_1 <- tidy(mod_glm_multivariable_1, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_mod_1 <- res_multivar_glm_1[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_mod_1 <- res_multivar_glm_mod_1 %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_mod_1
```


##### (5) Assumptions Check and Sensitivity analysis
###### - VIF
```{r}
#--------------------------------
#  Multicollinearity Check : <5
#--------------------------------
vif(mod_glm_multivariable_1)
```

###### - Outlier and influence check : Sensitivity test

There is no big difference without and with 262,172
 -> ALL participantes 
 
```{r}
#----------------------------------
#  Outliers and Influences Check    : 262,172
#----------------------------------
library(car)
influencePlot(mod_glm_multivariable_1)
# StudRes(Studentized Residuals) : <-2 ou >2 -> outlier
# Hat (Leverage, Hat value) : high leverage point (p/n, p = number of variables, n= number of samples)
# Cook's distance : Residus + Leverage
# CookD > 4/n (n= number of samples)
# - 4/378 = 0.016

plot(mod_glm_multivariable_1, which = 4)  # Cook’s distance
#-> 17 and 316 : Big residus, high leverage -> Outlier and big influence

#--------------------------
#   Sensitivity analysis  
#--------------------------
# 1. Create a new model without 262,172 individus
mod_glm_multivariable_sensitivity <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID_V2[-c(262,172),],
    family = binomial
  )
summary(mod_glm_multivariable_sensitivity)

# 2. Compare the AIC of two differents model for the sensitivity analysis 
AIC(mod_glm_multivariable_1, mod_glm_multivariable_sensitivity)
# Sensitivity : 398.2349
# Original model : 399.3754

# 3. Compare the AUC of two models for the sensitivity analysis 
library(pROC)
# 3-1) Prepare the FULL DATASET model for ROC and AUC
model_data <- na.omit(imputed_data_O3_perID_V2)
mod_glm_multivariable <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    data = model_data,
    family = binomial
  )
prob <- predict(mod_glm_multivariable, type = "response")
roc_full <- roc(response = model_data$total_complications_bin_perID, predictor = prob)

# 3-2) Prepare the MODIFIED DATASET model for ROC and AUC
mod_glm_multivariable_sensitivity <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age_cat +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    data = model_data[-c(262,172),],
    family = binomial
  )

prob_sensitivity <- predict(mod_glm_multivariable_sensitivity, type = "response")
roc_no_outlier <- roc(response = model_data[-c(262,172),]$total_complications_bin_perID, 
                      predictor = prob_sensitivity)

# 4. Compare 
plot(roc_full, col="blue")
plot(roc_no_outlier, add=TRUE, col="red")
auc(roc_full) # 0.7948
auc(roc_no_outlier) # 0.7936
```


##### (6) Comparison of models for the final model 

- Comparison of models 
```{r}
# Model for the comparison
mod_test <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases + 
      comorbidities_bin +
      
      # endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat+
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      #age_symptome_cat +
      
      nb_live_births_all + 
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      
      Csection_perID + 
      conception_PMA_perID +
      
      doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID_V2,
    family = binomial
  )

mod_test2 <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      #infertilite_diag +
      #autoimmune_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +           
      #age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +            
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID_V2,
    family = binomial
  )

mod_glm_multivariable_1 <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +           # in categorical
      # age_symptome_cat +
      
      nb_live_births_all +
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +                 # in continuous
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    data = imputed_data_O3_perID_V2,
    family = binomial
  )

# TEST
res1 <- compare_model_metrics(mod_test, imputed_data_O3_perID_V2,
                      "total_complications_bin_perID")
res2 <- compare_model_metrics(mod_test2, imputed_data_O3_perID_V2,
                      "total_complications_bin_perID")
res3 <- compare_model_metrics(mod_glm_multivariable_1,
                    imputed_data_O3_perID_V2,"total_complications_bin_perID")
res4 <- compare_model_metrics(mod_glm_multivariable_0,
                    imputed_data_O3_perID_V2,"total_complications_bin_perID")


# Results dataframe
results_df <- data.frame(
  Model = c("mod1:",
            "mod2:",
            "mod3:",
            "mod4:"),
  AIC = c(res1$AIC, res2$AIC, res3$AIC,res4$AIC),
  Pseudo_R2 = c(res1$Pseudo_R2, res2$Pseudo_R2, res3$Pseudo_R2,res4$Pseudo_R2),
  AUC = c(res1$AUC, res2$AUC, res3$AUC,res4$AUC),
  Hosmer_Lemeshow_p = 
    c(res1$Hosmer_Lemeshow_p, res2$Hosmer_Lemeshow_p, res3$Hosmer_Lemeshow_p,res4$Hosmer_Lemeshow_p)
)

print(results_df)
```


##### (7) Final model 
I think that, in general, model 2 is the most appropriate model

Results :  250630
  term                                    estimate IC95     
  <chr>                                      <dbl> <chr>    
1 diplome_ini2>=Bac+3                         0.33 0.15–0.68
2 DIEOui                                      1.85 1.05–3.31
3 age_firstOP_endoadeno_cat31>=               3.27 1.3–8.73 
4 indic_ovaire_toucheOui                      0.36 0.13–0.94
5 age_regle_cat_simple>=12                    0.25 0.11–0.52
6 age_regle_cat_simplensp                     0.26 0.07–0.99
7 history_FC_preterm_IVGIMG_perIDOui          2.59 1.32–5.27
8 Mode_accouchement_perIDSeule Césarienne     2.58 1.16–5.98
9 doul_abdo_avant_1_pregOui                   2.46 1.37–4.5 

```{r}
#mod_glm_multivariable_final_V2 <-  mod_test2
mod_glm_multivariable_final_V2 <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +

      gynecologique_diseases +
      comorbidities_bin +
      endo_ade_categorie +
      age_firstOP_endoadeno_cat + 
      
      # indic_ovaire_touche + # pas sure
      # age_regle_cat_simple + # pas sure
      # doul_abdo_avant_1_preg + # pas sure

      nb_live_births_all + 
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod,
    data = imputed_data_O3_perID_V2,
    family = binomial
  )

summary(mod_glm_multivariable_final_V2)

#------------
#  Results 
#------------
# 1) summary()
summary(mod_glm_multivariable_final_V2)

# 2) tidy(exponentiate=TRUE) 
library(tibble)
res_multivar_glm_V2 <- tidy(mod_glm_multivariable_final_V2, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_final_V2 <- res_multivar_glm_V2[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_final_V2 <- res_multivar_glm_final_V2 %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_final_V2

#-------------------------
# Significant variables
#-------------------------
res_multivar_glm_final_V2 %>% filter(p.value<=0.05) %>%
  distinct(term, estimate, IC95) %>%  View

#---------------------
# Save : update 250722
#---------------------
write_xlsx(
  res_multivar_glm_final_V2 %>% distinct(term, estimate, IC95), 
           "C:/Stage_Master2/Stage_M2_Complic_Gross/O3_bin_res_table_multivariable.xlsx")
```

```{r}
library(gtsummary)
library(dplyr)
library(broom)

# 1. 기술통계 테이블 res_description 정리
desc_table <- data %>%
  tbl_summary(
    by = all_of(outcome_variable),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Characteristic**") %>%
  bold_labels() %>%
  as.data.frame()
desc_table

# write_xlsx(desc_table, 
#            "C:/Stage_Master2/Stage_M2_Complic_Gross/O3_bin_res_Descriptive_table_multivariable.xlsx")
```


###### - Analyse de sensibilité : Imputed dataset vs Complete case dataset
- 33.88 % of participants are excluded for the sensitivity analysis

Results :  250625
  term                               estimate IC95      
1 diplome_ini2>=Bac+3                    0.19 0.06–0.55 
2 diplome_ini2Bac+2                      0.19 0.04–0.72 
3 age_firstOP_endoadeno_cat28-31         4.39 1.16–18.04
4 age_firstOP_endoadeno_cat31>=          4.84 1.37–18.45
5 age_regle_cat_simple>=12               0.16 0.06–0.39 
6 history_FC_preterm_IVGIMG_perIDOui     3.65 1.46–9.74 
7 doul_abdo_avant_1_pregOui              2.83 1.28–6.49 

```{r}
data_sensibility <- 
  df_final_complic_perID_V2_mod2 %>% 
  mutate(comorbidities_bin =
           ifelse(nb_mal_sansEndoAde_cat=="0","Non","Oui"))

mod_glm_sensibility_V2 <-
  glm(
    total_complications_bin_perID ~
      diplome_ini2 +
      age +

      gynecologique_diseases +
      comorbidities_bin +
      endo_ade_categorie +
      age_firstOP_endoadeno_cat + 
      
      # indic_ovaire_touche + # pas sure
      # age_regle_cat_simple + # pas sure
      # doul_abdo_avant_1_preg + # pas sure
    

      nb_live_births_all + 
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Mode_accouchement_perID +
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod,
    data = data_sensibility %>% na.omit() , 
    family = binomial
  )

summary(mod_glm_sensibility_V2)


# results
res_multivar_glm_sensibility <- tidy(mod_glm_sensibility_V2, 
                         exponentiate = TRUE, conf.int = 0.95) %>%
  as.tibble()

res_multivar_glm_sensibility <- 
  res_multivar_glm_sensibility[, c("term", "estimate", "conf.low", "conf.high", "p.value")] %>%
  mutate(across(c(estimate,conf.low, conf.high), ~ round(.x, 2)),
         p.value= p.value %>% round(4))

res_multivar_glm_sensibility <- res_multivar_glm_sensibility %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95 , p.value,conf.low,conf.high)
res_multivar_glm_sensibility

#-------------------------
# Significants variables
#-------------------------
res_multivar_glm_sensibility %>% filter(p.value<=0.05) %>%
  distinct(term, estimate, IC95) %>%  View

sig_var_bin <- res_multivar_glm_sensibility %>% filter(p.value<=0.05) %>%
  distinct(term) %>% pull()

```


#### 3) Visualization of results
##### - Table
```{r}
data <- imputed_data_O3_perID_V2
outcome_variable <- "total_complications_bin_perID"
covariables <- c(
  "diplome_ini2",
  "age",
  "gynecologique_diseases",
  "comorbidities_bin",
  "endo_ade_categorie",
  "age_firstOP_endoadeno_cat" ,
  "nb_live_births_all",
  "history_FC_preterm_IVGIMG_perID",
  "first_preg_age",
  "Mode_accouchement_perID",
  "Travail_accouchement_perID",
  "methode_conception_perID_cat_simple_mod"
)

# 2. Descriptive table
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(covariables)) %>%
  tbl_summary(
    by = all_of(outcome_variable),         
    statistic = list(                     
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) 
  ) %>%
  modify_header(label ~ "**Characteristic**") %>%
  bold_labels()

# 3. Multivariable analysis results table
tbl_glm_multivariable <- tbl_regression(mod_glm_multivariable_final_V2, exponentiate = TRUE, conf.int = TRUE)

# 4. Combined table
tbl_combined <- tbl_merge(
  tbls = list(tbl_descriptive, tbl_glm_multivariable),
  tab_spanner = c("**Descriptive Statistics**", "**Multivariable Logistic Regression**")
);tbl_combined

# 5. tbl_merge to dataframe
df_combined <- as_tibble(tbl_combined)

# 6. Save as Excel : update: 250722
write_xlsx(df_combined, path = "O3_glm_multivar_Complications.xlsx")
```


```{r}
#------------
#  gt table
#------------
library(gt)
result <- res_multivar_glm_final_V2
result <-  result[-1,]

res_table_O3_bin_multivariable <- result %>%
  gt() %>%
  tab_header(
    title = md("**multivariable Logistic Regression: **  
    *Factors Associated with Obstetric Complications experience at least once *")
  ) %>%
  cols_label(
    term = "Variables",
    estimate = "Adjusted OR",
    IC95 = "95% CI",
    p.value = "p-value"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = p.value,
      rows = as.numeric(gsub("<", "", p.value)) < 0.05
    )
  ) %>%
  cols_align(
    align = "left",
    columns = term
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95, p.value)
  )


#---------------
#  Forest plot 
#---------------
result$term <- 
  factor(result$term,levels = rev(result$term))

ggplot(result, aes(x = term, y = estimate)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray") +
  coord_flip() +  
  labs(
    title = "Résultats du modèle de régression logistique multivariée",
    x = NULL,
    y = "OR (Odds Ratio avec IC 95%)"
  ) +
  theme_minimal()

#------------------------------
#  Forest plot with the ORs
#------------------------------
library(dotwhisker)
library(grid)

# 1. Limite the maximum conf.high for the forest plot  
result_mod <- result %>% 
  mutate(conf.high_mod = 
           ifelse(conf.high>9, 9.1, conf.high)) %>% 
  dplyr:: select(-conf.high) %>% 
  mutate(conf.high= conf.high_mod) %>% 
  dplyr:: select(-conf.high_mod)

# 2. Redefine the terms
result_mod$term <- c(
  # "Intercept",
  "High school diploma or less/Other vs Bac+2",
  "High school diploma or less/Other vs >=Bac+3",
  "Age",
  
  "Gynecologic diseases: No vs Yes",
  # "Infertility diagnostic/treatement: No vs Yes",
  # "Autoimmune_diseases: No vs Yes",
  # "Metabolic syndrome: No vs Yes",
  "Comorbidity: No vs Yes",
  "Category of disease : Only endo vs only Adeno",
  "Only endo vs Both",
  #"Deep Inflating endometriosis : No vs Yes",
  "Age at 1stOP: No OP vs <28",
  "No OP vs 28-31",
  "<No OP vs 31>=",
  
  # "impacted/removed ovaries: No vs Yes",       # Removed by Marina
  # "age at menarche : <12 vs >=12",             # Removed by Marina
  # "<12 vs unknown",                            # Removed by Marina
  # "Abdominal pain before 1st preg: No vs Yes", # Removed by Marina
  
  # "Age at symptom onset : <=13 vs (13;15]",
  # "<=13 vs (15;22]",
  # "<=13 vs 22+",
  # "<=13 vs No symptoms",

  "Parity",
  "History of misscarriage/preterm/IVGIMG: No vs Yes",
  "Age_1stPregnancy",
  "Delivery Mode : only NSVD vs only C-section",
  "only NSVD vs Both",
  "Labour : only Natural vs only Induced",
  "Natural vs Mixed",
  "Natural vs No labour",
  "Conception Method: Natural vs Operation",
  "Natural vs Operation+ART",
  "Natural vs ART/Medicaments"
  )

# 3. Create the plot 
O3_1_forestplot <- dwplot(result_mod, show_intercept = TRUE) + 
  geom_text(aes(x = 9.1, y = term, 
                label = paste(estimate, "(", IC95, ")")),
            hjust = -0.1, size = 3) +
  theme_minimal() +
  labs(title = "Results of Multivariable Logistic Regression",
       subtitle= "Experience of at least one obstetric complication : Yes/No",
       x= "OR with 95% CI") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.2, size = 12, face = "bold"),
    plot.subtitle = element_text(hjust= 0.15, size= 10,face="bold"),
    axis.title.x = element_text(hjust = 0.3, vjust = 0),
    panel.grid = element_blank(),                
    panel.background = element_blank(),           
    axis.line.x =element_blank(), 
    axis.line.y = element_blank(),                
    axis.ticks = element_line(color = "black"),   
    axis.text = element_text(color = "black")     
  ) +
  geom_linerange(aes(xmin = conf.low, xmax = conf.high, y = term),
                 color = "red", size = 0.5, linetype=1)+
  scale_x_continuous(breaks = seq(0, 9, 1)) +
  coord_cartesian(xlim = c(0, 11), ylim = c(1, result_mod$term %>% length() +1)) +
  geom_segment(aes(x = 1, xend = 1, y = 0.8, yend = result_mod$term %>% length +1),
               color = "black", size = 0.5)+
  annotation_custom(
    grob = textGrob("OR (95%CI)", 
                    gp = gpar(fontsize = 10, 
                              col = "black",fontface = "bold")),
    xmin = 8.7, ymin = result_mod$term %>% length +1) +
  geom_segment(aes(x = -0.3, xend = 9, y = 0.5, yend = 0.5),  
               color = "black", size = 0.5);O3_1_forestplot

O3_1_forestplot

# Save  : Update 250722
ggsave("C:/Stage_Master2/Stage_M2_Complic_Gross/O3_1_forestplot.jpg", O3_1_forestplot, width = 10, units = "in", dpi = 300)
```


#---------------------------------------------------------
# II. Ordinal outcome : 0 vs 1 vs 2+ 
#---------------------------------------------------------

## 1. Univariable analysis adjusted for age 
### 1) Results Table

```{r}
# Load necessary libraries
library(dplyr)
library(gtsummary)
library(purrr)
library(MASS) # For polr
library(brant) # Test 
library(broom)

#----------------------
# Define the data
#----------------------
data <- imputed_data_O3_perID_V2 %>%
  dplyr::select(-total_complications_bin_perID,
                -Complications_cat_perID_grouped) %>% 
  mutate(across(where(~ is.factor(.) | is.character(.)), as.factor))

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure the outcome variable is an ordered factor
data$total_complications_cat_perID <-
  factor(data$total_complications_cat_perID,
         levels = c("0", "1", "2+"), 
         ordered = TRUE)

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables
outcome_variable <- "total_complications_cat_perID"
adjustment_variable <- "age" 

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(data), 
                               c(outcome_variable, adjustment_variable))

#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>%
  tbl_summary(
    by = all_of(outcome_variable),         # Stratify by the outcome
    statistic = list(                     # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%   # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>%
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted multivariable Ordinal Regression Table
#-------------------------------------------------------------
# This function fits polr(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_polr <- function(predictor, outcome_var, adjust_var, data) {
  # Use backticks for safety with non-standard names and collapse adjustment variables
  # adjust_str <- paste(sprintf("`%s`", adjustment_variable), collapse = " + ") # This was for a vector
  # For a single adjustment variable, it's simpler:
  adjust_str <- sprintf("`%s`", adjust_var) 

  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)

  # Fit the model
  model <- tryCatch(
    MASS::polr(as.formula(formula_str), data = data, Hess = TRUE), 
    error = function(e) {
      warning(paste("Could not fit ordinal regression model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL)
    }
  )

  # If model failed, return NULL
  if (is.null(model)) return(NULL)

  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int = TRUE) %>% # exponentiate = TRUE for Proportional OR
    # Keep rows where 'variable' matches the predictor name
    # This correctly handles factor variables which have multiple rows
    modify_table_body(~ dplyr::filter(.x, variable == predictor))
}

# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_polr(.x, outcome_variable,
                                             adjustment_variable,
                                             data))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables,
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
          )
        )
    ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted Proportional OR (95% CI)**" # Clarify as Proportional OR
    )
  print("Step 2: Adjusted Proportional Odds Ratios Table Created.")
} else {
  warning("No successful ordinal regression models were fitted for adjusted ORs.",
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
# The 'adjustment_variable' is a single string in this setup, so no paste collapse needed
adjust_str <- adjustment_variable

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      paste0("**Age-Adjusted Proportional Odds Model**")) # Update spanner for clarity
    ) %>%
      modify_caption(
        paste0(
          "**multivariable Ordinal Logistic Regression Analysis for '",
          outcome_variable,
          "' (Adjusted for '", adjust_str, "')**")) # Update caption for clarity

    print("Step 3: Merged Table Created.")

    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '",
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results
#------------------------------------------------------------
O3_tab_2_univariable_final <- as_tibble(tbl_final) %>%
  dplyr::select(-"**OR**",-"**95% CI**")  

O3_tab_2_univariable_final
```

### 2) Pre-selection of variables for the multivariable regression model
 - P-value <= 0.2 : 
diplome_ini2
situation_financiere
nb_mal_sansEndoAde_cat
comorbidities_bin
infertilite_diagOui
age_regles_ini
age_regle_cat
age_regle_cat_simple
type_endo
DIE
stade_dvpt_III_IV
age_symptome_cat
operation_endo_bin
age_firstOP_endoadeno_cat
nb_total_gross
nb_live_births_all
nb_live_births_all_cat
history_FC_preterm_IVGIMG_perID
methode_conception_perID_cat_simple_mod
conception_PMA_perID
conception_natural_perID
conception_operation_perID
Mode_accouchement_perID
VD_perID
Csection_perID
Travail_accouchement_perID
labour_natural_perID
doul_abdo_avant_1_preg

```{r}
# Load necessary libraries
library(dplyr)
library(gtsummary)
library(purrr)
library(MASS) # For polr
library(brant) # Test 
library(broom)

#----------------------
# Define the data
#----------------------
data <- imputed_data_O3_perID_V2 %>%
  dplyr::select(-total_complications_bin_perID) %>% 
  mutate(across(where(~ is.factor(.) | is.character(.)), as.factor))

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure the outcome variable is an ordered factor
data$total_complications_cat_perID <-
  factor(data$total_complications_cat_perID,
         levels = c("0", "1", "2+"), 
         ordered = TRUE)

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables
outcome_variable <- "total_complications_cat_perID"
adjustment_variable <- "age" 

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(data), 
                               c(outcome_variable, adjustment_variable))

#-----------------------------------------------------
# Modeling function with the calculation of p-value
#-----------------------------------------------------
fit_adjusted_polr_with_pval <- function(predictor, outcome_var, adjust_var, data) {
  adjust_str <- sprintf("`%s`", adjust_var)
  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)

  model <- tryCatch(
    MASS::polr(as.formula(formula_str), data = data, Hess = TRUE),
    error = function(e) {
      warning(paste("Could not fit ordinal regression model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL)
    }
  )

  if (is.null(model)) return(NULL)

  # Extract model summary and coefficient table
  model_summary <- summary(model)
  coef_table <- coef(model_summary)
  
  # Calculate p-values using Wald z-test
  z_values <- coef_table[, "Value"] / coef_table[, "Std. Error"]
  p_values <- 2 * (1 - pnorm(abs(z_values)))

  results_df <- as.data.frame(coef_table)
  results_df$p.value <- p_values
  results_df$variable <- rownames(results_df)

  # Filter to keep only the predictor (and its levels if factor)
  results_df <- results_df %>%
    dplyr::filter(grepl(paste0("^", predictor), variable)) %>%
    dplyr::mutate(
      OR = exp(Value),
      CI_low = exp(Value - 1.96 * `Std. Error`),
      CI_high = exp(Value + 1.96 * `Std. Error`),
      est_ci = sprintf("%.2f (%.2f, %.2f)", OR, CI_low, CI_high),
      p.value = signif(p.value, 3)
    ) %>%
    dplyr::select(variable, est_ci, p.value)

  return(results_df)
}

#------------------------------
# Application of this function 
#------------------------------
list_of_adj_tables <- map(
  predictor_variables,
  ~ fit_adjusted_polr_with_pval(.x, outcome_variable, adjustment_variable, data)
)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables, is.null)]

#-----------------
# list to tibble
#-----------------
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- bind_rows(list_of_adj_tables) %>%
    dplyr::rename(
      `**Characteristic**` = variable,
      `**Adjusted Proportional OR (95% CI)**` = est_ci,
      `**p-value**` = p.value
    )
  print("Step 2: Adjusted ORs with p-values Table Created.")
} else {
  warning("No successful ordinal regression models were fitted for adjusted ORs with p-values.",
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

tbl_adjusted_ors <- bind_rows(list_of_adj_tables)

#----------------------------
#  Filtering p-value <= 0.2 
#----------------------------
tbl_adjusted_ors %>% filter(p.value<=0.2) %>% distinct(variable) %>% pull()
```

### 3) Porportional Odds Assumption (POA) Test
Results : 250625
                                                        pval
situation_financiereComfortable/Very comfortable 0.029443330
age_regle_cat>13                                 0.048911840
age_firstOP_endoadeno_cat<28                     0.006269611
nb_live_births_all                               0.011670157
methode_conception_perID_cat_simple_modOperation 0.016246565
nb_live_births_all_cat2+                         0.048886173

=> Violation du POA
Les hypothèses de base du modèle de régression logistique ordonnée ne sont pas respectées. Mais ces résultats seraient différents dans l’analyse multivariée

```{r}
library(purrr)
library(MASS)
library(brant)
#-----------------------------------------
#   PO TEST for each predictor variable
#-----------------------------------------
brant_results <- map(predictor_variables, function(pred) {
  formula_str <- as.formula(paste0("total_complications_cat_perID ~ ", pred, " + age"))
  model <- tryCatch(
    polr(formula_str, data = data, Hess = TRUE),
    error = function(e) return(NULL)
  )
  if (!is.null(model)) {
    tryCatch(
      brant(model),
      error = function(e) return(NULL)
    )
  } else {
    NULL
  }
})

#
result_list <- list()
for (i in seq_along(brant_results)) {
  pvals <- brant_results[[i]][, 3]
  
  # p < 0.05
  significant_vars <- names(pvals)[pvals < 0.05 & names(pvals) != "Omnibus"]
  significant_pvals <- pvals[pvals < 0.05 & names(pvals) != "Omnibus"]
  
  if (length(significant_vars) > 0) {
    # save in the result_list
    result_list[[i]] <- data.frame(pval = significant_pvals)
  }
}

significant_vars_vector <- do.call(rbind, result_list); significant_vars_vector

```

## 2. Multivariable  Model and Results
### 1) Model
#### - polr()
```{r}
#-----------------------------
#  Ordinal regression model
#-----------------------------
mod_polr_multivariable <-
  polr(
    total_complications_cat_perID  ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      infertilite_diag +
      autoimmune_diseases +
      comorbidities_bin + 
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      nb_live_births_all_cat +

      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Csection_perID +

      conception_PMA_perID +
      conception_operation_perID+
      
      doul_abdo_avant_1_preg,
    
    data = data,
    Hess = TRUE # Calculate standard error
  )

#------------
#  Results 
#------------
# 1) summary(), tidy()
summary(mod_polr_multivariable) # no p-value
tidy(mod_polr_multivariable, exponentiate = TRUE)

# 2) Anova()
# Instead of considering each coefficient individually, check the impact across the entire factor level
# Type=2 : main effects when there is no interactions
Anova(mod_polr_multivariable,type=2) # with p-value

# 3) Result Table
res <- tidy(mod_polr_multivariable)
ci <- confint.default(mod_polr_multivariable)
ci_df <- as.data.frame(ci)
colnames(ci_df) <- c("conf.low", "conf.high")
ci_df$term <- rownames(ci_df)
res2 <- res %>%
  left_join(ci_df, by = "term")
res_multivar_polr <- res2 %>%
  mutate(
    estimate = exp(estimate) %>% round(2),
    conf.low = exp(conf.low) %>% round(2),
    conf.high = exp(conf.high) %>% round(2)
  ) %>% 
  dplyr::select(-std.error,-statistic, -coef.type)
#res_multivar_polr

res_multivar_polr <- res_multivar_polr %>%
  mutate(IC95 = paste0(conf.low, "–", conf.high)) %>% 
  dplyr::select(term, estimate, IC95,conf.low,conf.high) 
res_multivar_polr

##################################################################
# Significants variables : 
res_multivar_polr %>% filter(conf.low<1 & conf.high<1 | conf.high>1 & conf.low>1)
```

#### - clm() : cumulative link model

```{r}
library(ordinal)
mod_clm <- clm(
  total_complications_cat_perID  ~
    diplome_ini2 +
    age +
    
    gynecologique_diseases +
    comorbidities_bin +
    
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    
    age_regle_cat_simple +
    nb_live_births_all_cat +
    
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +
    Travail_accouchement_perID +
    methode_conception_perID_cat_simple_mod +
    
    doul_abdo_avant_1_preg,
  data = data
)
summary(mod_clm)
```

### 2) Assumptions check 
#### (1) VIF 
##### - polr() : vif<5
```{r}
#------------------------------------
#     Multicollinearity Check : <5
#------------------------------------
library(car)

#################
## polr() model
#################
glm_proxy <- glm(
  as.numeric(total_complications_cat_perID) ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      #infertilite_diag +
      #autoimmune_diseases +
      comorbidities_bin + 
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      nb_live_births_all_cat+

      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Csection_perID +

      conception_PMA_perID +
      conception_operation_perID+
      
      doul_abdo_avant_1_preg,
    
    data = data
)
vif(glm_proxy) 
```

##### - clm() : vif <5

```{r}
glm_proxy <- glm(
  as.numeric(total_complications_cat_perID) ~
    diplome_ini2 +
    age +
    
    gynecologique_diseases +
    comorbidities_bin +
    
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    
    age_regle_cat_simple +
    nb_live_births_all_cat +
    
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +
    Travail_accouchement_perID +
    methode_conception_perID_cat_simple_mod +
    
    doul_abdo_avant_1_preg,
  data= data
)
vif(glm_proxy) 
```

#### (2) Proportional Odds Assumption 
- brant() for polr()
- nominal_test for clm()
- vglm() for lrtest()

##### - polr() + brant()    
- age_regle_cat_simple 

```{r}
# brant 
brant(mod_polr_multivariable)
# Omnibus p-value < 0.03
# age_regle_cat_simple>=12			p-value=0.05 
```

##### - clm()+ nominal_test()
- nb_live_births_all_cat, age_regle_cat_simple, age_firstOP_endoadeno_cat

```{r}
nominal_test(mod_clm,type = "II")
```

##### - vglm() for lrtest()
- Comparison : 
  age_regle_cat_simple 
  age_regle_cat_simple + nb_live_births_all_cat
  age_regle_cat_simple + nb_live_births_all_cat + age_firstOP_endoadeno_cat
  
```{r}
library(VGAM)
# vglm and lrtest
# POA : proportional odds vs non-proportional odds
model_PO <- vglm(total_complications_cat_perID ~
    diplome_ini2 +
    age +
    gynecologique_diseases +
    # infertilite_diag +
    # autoimmune_diseases +
    comorbidities_bin + 
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    age_regle_cat_simple +
    nb_live_births_all_cat +
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +
    conception_PMA_perID +
    conception_operation_perID +
    doul_abdo_avant_1_preg, 
    family = cumulative(parallel = TRUE), 
    data = data)

model_nonPO_1 <- vglm(
  total_complications_cat_perID ~
    diplome_ini2 +
    age +
    gynecologique_diseases +
    infertilite_diag +
    autoimmune_diseases +
    comorbidities_bin + 
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    age_regle_cat_simple +
    nb_live_births_all_cat +
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +
    conception_PMA_perID +
    conception_operation_perID +
    doul_abdo_avant_1_preg,
  
  family = cumulative(
    link = "logitlink",
    parallel = FALSE ~ age_regle_cat_simple),
  data = data
)

model_nonPO_2 <- vglm(
  total_complications_cat_perID ~
    diplome_ini2 +
    age +
    gynecologique_diseases +
    infertilite_diag +
    autoimmune_diseases +
    comorbidities_bin + 
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    age_regle_cat_simple +
    nb_live_births_all_cat +
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +
    conception_PMA_perID +
    conception_operation_perID +
    doul_abdo_avant_1_preg,
  
  family = cumulative(
    link = "logitlink",
    parallel = FALSE ~ age_regle_cat_simple + nb_live_births_all_cat),
  data = data
)

model_nonPO_3 <- vglm(
  total_complications_cat_perID ~
    diplome_ini2 +
    age +
    gynecologique_diseases +
    infertilite_diag +
    autoimmune_diseases +
    comorbidities_bin + 
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    age_regle_cat_simple +
    nb_live_births_all_cat +
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +
    conception_PMA_perID +
    conception_operation_perID +
    doul_abdo_avant_1_preg,
  
  family = cumulative(
    link = "logitlink",
    parallel = FALSE ~ age_regle_cat_simple+nb_live_births_all_cat+age_firstOP_endoadeno_cat),
  data = data
)

lrtest(model_PO, model_nonPO_1)
lrtest(model_PO, model_nonPO_2)
lrtest(model_PO, model_nonPO_3)
```


comparison of different models
- parallel = FALSE ~ age_regle_cat_simple+nb_live_births_all_cat+age_firstOP_endoadeno_cat
Residual deviance: 583.5399 on 588 degrees of freedom
Log-likelihood: -291.7699 on 588 degrees of freedom

- parallel = FALSE ~ age_regle_cat_simple + nb_live_births_all_cat 
Residual deviance: 590.8341 on 591 degrees of freedom
Log-likelihood: -295.4171 on 591 degrees of freedom

- parallel = FALSE ~ age_regle_cat_simple
Residual deviance: 594.8087 on 592 degrees of freedom
Log-likelihood: -297.4044 on 592 degrees of freedom


### 3) Final model
#### - clm() 

```{r}
mod_clm_adjusted_final <-
  clm(
    total_complications_cat_perID  ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      comorbidities_bin +
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      nb_live_births_all_cat +
      
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Csection_perID +
      
      Travail_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      
      doul_abdo_avant_1_preg,
    nominal = ~ age_regle_cat_simple + nb_live_births_all_cat + age_firstOP_endoadeno_cat,
    data = data
  )

summary(mod_clm_adjusted_final)

#---------------------------
#  Model goodneeds of fit 
#---------------------------
anova(mod_clm_adjusted_final)

AIC(mod_clm_adjusted_final)
BIC(mod_clm_adjusted_final)
logLik(mod_clm_adjusted_final)

#------------------------------------
# Nominal Effects Significance Tests
#------------------------------------
anova(mod_clm, mod_clm_adjusted_final)
# - mod_clm_adjusted_final better model 
```

#### - vglm() : polr() + brant()
```{r}
library(VGAM) 
library(purrr)
#-------------------------------------------
#  VGAM : Partial Proportional Odds Model  
# - cumulative logit model
#-------------------------------------------
# In general, the proportional odds assumption is required in ordinal logistic regression. However, for some variables that do not satisfy this assumption, you can set `parallel = FALSE` for those specific variables, using `parallel = FALSE ~ variable_name`, to treat them as exceptions.
# parallel = FALSE ~ age_regle_cat_simple => light proportional odds assumption, meaning that different coefficients are allowed for each cutoff point

mod_vglm_multivariable_final <- vglm(
  total_complications_cat_perID ~
    diplome_ini2 +
    age +
    gynecologique_diseases +
    infertilite_diag +
    autoimmune_diseases +
    comorbidities_bin +
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    age_regle_cat_simple +
    nb_live_births_all_cat +
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +
    conception_PMA_perID +
    conception_operation_perID +
    doul_abdo_avant_1_preg,
  
  family = cumulative(
    link = "logitlink",
    parallel = FALSE ~ age_regle_cat_simple + nb_live_births_all_cat + age_firstOP_endoadeno_cat
  ),
  data = data
)

Anova(mod_vglm_multivariable_final)
# RESULTATS: DIE, age_firstOP_endoadeno_cat, age_regle_cat_simple,nb_live_births_all_cat,history_FC_preterm_IVGIMG_perID,Csection_perID,doul_abdo_avant_1_preg

summary(mod_vglm_multivariable_final)
```

250625
vglm Exponentiated coefficients: 

diplome_ini2>=Bac+3  = 1.9448096
DIEOui  : 0.5964814
age_firstOP_endoadeno_cat31>=:1 = 0.4059979
age_regle_cat_simple>=12:1 = 3.6951970
age_regle_cat_simplensp:1 = 4.4880889
age_regle_cat_simplensp:2 = 12.0490893
nb_live_births_all_cat2+:2 = 0.4398247
history_FC_preterm_IVGIMG_perIDOui = 0.5496407
Csection_perIDOui = 0.5585332
doul_abdo_avant_1_pregOui = 0.5549930


### 4) Visualization of results
#### - clm() 
##### Table
                                term   OR        IC95 
                   1|2+.(Intercept) 15.1 1.78–128.21   
       0|1.age_regle_cat_simple>=12  3.7   1.87–7.41     
        0|1.age_regle_cat_simplensp  3.9  1.12–13.22     
       1|2+.age_regle_cat_simplensp 10.3  1.15–92.77    
  0|1.age_firstOP_endoadeno_cat31>=  0.4   0.17–0.88     
                diplome_ini2>=Bac+3  0.5   0.28–0.92    
                             DIEOui  1.6   1.02–2.65     
 history_FC_preterm_IVGIMG_perIDOui  1.9   1.09–3.15     
          doul_abdo_avant_1_pregOui  1.8   1.07–2.92    


```{r}
#------------
#  Results 
#------------
# 1) summary()
summary(mod_clm_adjusted_final)

# 2) Result Table
library(tibble)

res_multivar_clm <-  tidy(mod_clm_adjusted_final) %>% 
  as.data.frame() %>% na.omit()

res_multivar_clm <- res_multivar_clm %>%
  mutate(
    OR = round(exp(estimate),1),
    se =std.error,
    conf.low = round(exp(estimate - 1.96 * se),2),
    conf.high = round(exp(estimate + 1.96 * se),2),
    IC95 = paste0(conf.low, "–", conf.high)) %>%
  dplyr::select(term, OR, IC95, conf.low, conf.high, p.value)

# Check only significant variables
res_multivar_clm %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% View 


sig_var_clm <- res_multivar_clm %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(term) %>% pull()

#
res_multivar_clm_final <- res_multivar_clm %>% dplyr::select(term,OR, IC95)
res_multivar_clm_final %>% View

#------------
#  gt table
#------------
library(gt)
res_multivar_clm %>%
  mutate(p.value= round(p.value, 2)) %>% 
  gt() %>%
  tab_header(
    title = md("**multivariable Ordinal logistic Regression:**\n
    - Factors Associated with the number of obstetric complications")
  ) %>%
  cols_label(
    term   = "Variables",
    OR = "Adjusted OR",
    IC95 = "95% CI"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = OR,
      rows = (conf.low > 1 & conf.high > 1) | (conf.low < 1 & conf.high < 1)
    )
  ) %>%
  cols_align(
    align = "left",
    columns = term
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95)
  ) %>% 
 cols_hide(columns = c(conf.low, conf.high))
```

# -----------REDO IT---------
##### Forest Plot

```{r}
#------------------------------
#  Forest plot with the ORs
#------------------------------
library(dotwhisker)
library(grid)
library(ggplot2)

# 1. Limite the maximum conf.high for the forest plot  
res_multivar_vglm <- res_multivar_clm %>% 
  mutate(conf.high_mod = 
           ifelse(conf.high>11, 11.1, conf.high)) %>% 
  dplyr:: select(-conf.high) %>% 
  mutate(conf.high= conf.high_mod,
         term=variable) %>% 
  dplyr:: select(-conf.high_mod, -variable) 

# 2. Remove intercept informations 
res_multivar_vglm_mod <- res_multivar_vglm[-c(1,2),]

# 3. Set the order of term values 
res_multivar_vglm_mod$term <- factor(res_multivar_vglm_mod$term, 
                                 levels = rev(res_multivar_vglm_mod$term))

# 4. Create the forest plot 
O3_2_forestplot <- ggplot(res_multivar_vglm_mod, aes(y = term, x = OR)) +
  geom_point(color = "blue") +  # OR 점
  geom_linerange(aes(xmin = conf.low, xmax = conf.high), 
                 color = "red", size = 0.8) +  
  geom_text(aes(x = 12.1, label = paste0(round(OR, 2), " (", IC95, ")")),
            hjust = 0, size = 3) +
  geom_vline(xintercept = 1, color = "black") +  # 기준선 OR=1
  scale_x_continuous(breaks = seq(0, 11, 1), limits = c(0, 13)) +
  coord_cartesian(ylim = c(0.5, length(levels(res_multivar_vglm_mod$term)) + 1.5),
                  xlim= c(0,15)) +
  theme_minimal() +
  labs(title = "Results of multivariable Logistic Regression",
       x = "Odds Ratio (95% CI)",
       y = NULL) +
  theme(
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  ) + 
  annotation_custom(
    grob = textGrob("OR (IC95)", 
                    gp = gpar(fontsize = 10, fontface = "bold")),
    xmin = 13, xmax = 13, 
    ymin = length(levels(res_multivar_vglm_mod$term)) + 1.3,  
    ymax = length(levels(res_multivar_vglm_mod$term)) + 1.3
  )
O3_2_forestplot



# 2. Redefine the terms
# res_multivar_glm_mod$term <- c(
#   "High school diploma or less/Other vs Bac+2",
#   "High school diploma or less/Other vs >=Bac+3",
#   "Age",
#   "age at menarche : <12 vs >=12 : 0 vs 1",
#   "<12 vs >=12 : 1 vs 2+",
#   "<12 vs nsp : 0 vs 1",
#   "<12 vs nsp : 1 vs 2+",
#   "Delay diagnostic years :<8 vs 13+",
#   "<8 vs 8-13",
#   "<8 vs No symptom",
#   "Gynecologic diseases: No vs Yes",
#   "Infertility diagnostic/treatement: No vs Yes",
#   "Autoimmune_diseases: No vs Yes",
#   "Metabolic syndrome: No vs Yes",
#   "Nb_comorbidity: 0 vs 1",
#   "0 vs 2+",
#   "Age at 1stOP: <27 vs 27-31",
#   "<27 vs 31>=",
#   "<27 vs No OP",
#   "Category : only Endo vs only adeno",
#   "only Endo vs Both",
#   "DIE : No vs Yes",
#   "Endo III-IV stage: No vs Yes",
#   "impacted/removed ovaries: No vs Yes",
#   "Anormal fallopian tube: No vs Yes",


#   "History of misscarriage/preterm/IVGIMG: No vs Yes",
#   "age at menarche : <12 vs >=12",
#   "<12 vs unknown",
#   "type_endo: SPE only vs DIE only",
#   "SPE only vs OMA only",
#   "SPE only vs ADENO only",
#   "SPE only vs Multiple",
#   "SPE only vs Unknown",

#   "Age at symptom onset : <=13 vs (13;15]",
#   "<=13 vs (15;22]",
#   "<=13 vs 22+",
#   "<=13 vs No symptom",

#   "C-section: No vs Yes",


#   "Conception Method: Natural vs Operation",
#   "Natural vs Operation+ART",
#   "Natural vs ART/Medicaments",
#   "Age_1stPregnancy: <30 vs 30-34",
#   "<30 vs >=34",
#   "Abdominal pain befor 1st preg : No vs Yes"
# )

# 3. Create the plot 
O3_2_forestplot <- dwplot(res_multivar_vglm, show_intercept = TRUE) + 
  geom_text(aes(x = 9.1, y = term, 
                label = paste(OR, "(", IC95, ")")),
            hjust = -0.1, size = 3) +
  theme_minimal() +
  labs(title = "Results of multivariable Logistic Regression",
       x= "OR with 95% CI") +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.2, size = 12, face = "bold"),
    axis.title.x = element_text(hjust = 0.3, vjust = 0),
    panel.grid = element_blank(),                
    panel.background = element_blank(),           
    axis.line.x =element_blank(), 
    axis.line.y = element_blank(),                
    axis.ticks = element_line(color = "black"),   
    axis.text = element_text(color = "black")     
  ) +
  geom_linerange(aes(xmin = conf.low, xmax = conf.high, y = variable),
                 color = "red", size = 0.5, linetype=1)+
  scale_x_continuous(breaks = seq(0, 9, 1)) +
  coord_cartesian(xlim = c(0, 11), ylim = c(1, 37)) +
  geom_segment(aes(x = 1, xend = 1, y = 0.8, yend = 37),
               color = "black", size = 0.5)+
  annotation_custom(
    grob = textGrob("OR (95%CI)", 
                    gp = gpar(fontsize = 10, col = "black",fontface = "bold")),
    xmin = 8.5, ymin = 37) +
  geom_segment(aes(x = -0.3, xend = 9, y = 0.5, yend = 0.5),  
               color = "black", size = 0.5);O3_2_forestplot

O3_2_forestplot
# Save 
#ggsave("Stage_M2_Complic_Gross/O3_2_forestplot.jpg", O3_2_forestplot,height = plot_height, width = 7, units = "in", dpi = 300)
```



#### - vglm() : Table

```{r}
#------------
#  Results 
#------------
# 1) summary()
summary(mod_vglm_multivariable_final)

# 2) Result Table
library(tibble)

res_multivar_vglm <- summary(mod_vglm_multivariable_final) %>% 
  coef() %>% as.data.frame() %>%  na.omit() 
res_multivar_vglm$term <- rownames(res_multivar_vglm)
res_multivar_vglm <- res_multivar_vglm %>%
  mutate(
    OR = round(exp(Estimate),1),
    se =`Std. Error`,
    p.value =`Pr(>|z|)`,
    conf.low = round(exp(Estimate - 1.96 * se),2),
    conf.high = round(exp(Estimate + 1.96 * se),2),
    IC95 = paste0(conf.low, "–", conf.high)) %>%
  dplyr::select(term, OR, IC95, conf.low, conf.high, p.value)

rownames(res_multivar_vglm) <- NULL

# Check only significant variables
res_multivar_vglm %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% View 

sig_var_vglm <- res_multivar_vglm %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(term) %>% pull()

#
res_multivar_vglm %>% dplyr::select(term,OR, IC95)

#------------
#  gt table
#------------
library(gt)
res_multivar_vglm %>%
  mutate(p.value= round(p.value, 2)) %>% 
  gt() %>%
  tab_header(
    title = md("**multivariable Ordinal logistic Regression:**\n
    - Factors Associated with the number of obstetric complications")
  ) %>%
  cols_label(
    term   = "Variables",
    OR = "Adjusted OR",
    IC95 = "95% CI"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = OR,
      rows = (conf.low > 1 & conf.high > 1) | (conf.low < 1 & conf.high < 1)
    )
  ) %>%
  cols_align(
    align = "left",
    columns = term
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95)
  ) %>% 
 cols_hide(columns = c(conf.low, conf.high))
```


```{r}
#------------
#  gt table
#------------
library(gt)
res_multivar_vglm %>%
  mutate(p.value= round(p.value, 2)) %>% 
  gt() %>%
  tab_header(
    title = md("**multivariable Ordinal logistic Regression:**\n
    - Factors Associated with the number of obstetric complications")
  ) %>%
  cols_label(
    term   = "Variables",
    OR = "Adjusted OR",
    IC95 = "95% CI"
  ) %>% 
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.align = "left",
    column_labels.font.weight = "bold",
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "#f9f9f9"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = OR,
      rows = (conf.low > 1 & conf.high > 1) | (conf.low < 1 & conf.high < 1)
    )
  ) %>%
  cols_align(
    align = "left",
    columns = term
  ) %>%
  cols_align(
    align = "center",
    columns = c(IC95)
  ) %>% 
 cols_hide(columns = c(conf.low, conf.high))

#---------------------------
#   Simple Visualization 
#---------------------------
# 1) OR (IC95)
res_multivar_vglm <- res_multivar_vglm %>%
  filter(!grepl("\\(Intercept\\)", term))

# 2) Forest plot
ggplot(res_multivar_vglm, aes(x = reorder(term, OR), 
                              y = ifelse(OR >4,NA,OR))) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, 
                    ymax = ifelse(conf.high>4,4,conf.high)), 
                width = 0.2) +
  coord_flip() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(x = "Var", y = "OR(IC95)", title = "Influence per variable") +
  theme_minimal()
```

#---------------------------------------------------------
# III. (Nb) Multinomial outcome without considering order  
#---------------------------------------------------------

## 0.Define Data

```{r}
#----------------------
# Define the data
#----------------------
data <- imputed_data_O3_perID_V2 %>%
  dplyr::select(-total_complications_bin_perID,
                -Complications_cat_perID_grouped) %>% 
  mutate(across(where(~ is.factor(.) | is.character(.)), as.factor))

#-----------------------------
#  Data Preparation & Checks
#-----------------------------
# Ensure the outcome variable is an ordered factor
data$total_complications_cat_perID <-
  factor(data$total_complications_cat_perID,
         levels = c("0", "1", "2+"), 
         ordered = FALSE)
```


## 1. univariable analysis adjusted for age 
### 1) Results Table

```{r}
O3_tab_multinom_univariable_final %>% View
```

```{r}
library(nnet)

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables
outcome_variable <- "total_complications_cat_perID"
adjustment_variable <- "age" 

# Identify predictor variables (excluding outcome and adjustment variable)
predictor_variables <- setdiff(names(data), 
                               c(outcome_variable, adjustment_variable))

#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>%
  tbl_summary(
    by = all_of(outcome_variable),         # Stratify by the outcome
    statistic = list(                     # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%   # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>%
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted multivariable multinomial Regression Table
#-------------------------------------------------------------
# This function fits multinom(outcome ~ predictor + ajusted variables ) and returns a gtsummary table row
fit_adjusted_multinom <- function(predictor, outcome_var, adjust_var, data) {
  adjust_str <- sprintf("`%s`", adjust_var) # If adjust_var is a single string

  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)

  # Fit the model
  model <- tryCatch(
    nnet::multinom(as.formula(formula_str), data = data), 
    error = function(e) {
      warning(paste("Could not fit ordinal regression model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )

  # If model failed, return NULL
  if (is.null(model)) return(NULL)

  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int = TRUE) %>% 
    modify_table_body(~ dplyr::filter(.x, variable == predictor))
}

# Apply the function to each predictor variable
list_of_adj_tables <- map(predictor_variables,
                          ~fit_adjusted_multinom(.x, outcome_variable,
                                             adjustment_variable,
                                             data))

# Remove any NULL results (from failed models)
list_of_adj_tables <- list_of_adj_tables[!sapply(list_of_adj_tables,
                                                 is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables,
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
          )
        )
    ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted Proportional OR (95% CI)**" # Clarify as Proportional OR
    )
  print("Step 2: Adjusted Proportional Odds Ratios Table Created.")
} else {
  warning("No successful ordinal regression models were fitted for adjusted ORs.",
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
# The 'adjustment_variable' is a single string in this setup, so no paste collapse needed
adjust_str <- adjustment_variable

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      paste0("**Age-Adjusted Proportional Odds Model**")) # Update spanner for clarity
    ) %>%
      modify_caption(
        paste0(
          "**multivariable Ordinal Logistic Regression Analysis for '",
          outcome_variable,
          "' (Adjusted for '", adjust_str, "')**")) # Update caption for clarity

    print("Step 3: Merged Table Created.")

    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '",
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results
#------------------------------------------------------------
O3_tab_multinom_univariable_final <- as_tibble(tbl_final) %>%
  dplyr::select(-"**OR**",-"**95% CI**")  
```

### 2) Pre-selection of variables for the multivariable regression model
 - P-value <= 0.2 : 

diplome_ini2
situation_emploi_ini2_cat
situation_financiere

nb_mal_sansEndoAde_cat
comorbidities_bin

gynecologique_diseases
infertilite_diag

age_regles_ini
age_regle_cat_simple

endo_ade_categorie
type_endo
ADENO
DIE
stade_dvpt_III_IV

delai_diag
delai_diag_cat
delai_diag_cat_simple

age_symptome_cat

operation_endo_bin
age_firstOP_endoadeno_cat

nb_live_births_all_cat
nb_total_gross_cat

history_FC_preterm_IVGIMG_perID

methode_conception_perID_cat_simple_mod
conception_PMA_perID
conception_natural_perID
conception_operation_perID

Mode_accouchement_perID
VD_perID
Csection_perID

Travail_accouchement_perID
labour_natural_perID
labour_induced_perID

doul_abdo_avant_1_preg

first_preg_age_cat

```{r}
O3_tab_multinom_univariable_final <- 
  O3_tab_multinom_univariable_final %>% 
  mutate(ind= rep(1:nrow(O3_tab_multinom_univariable_final),1)) %>% 
  dplyr::select(ind, everything()) 


ls <- O3_tab_multinom_univariable_final %>% filter(`**p-value**`<0.2) %>% distinct(ind) %>% pull()

O3_tab_multinom_univariable_final %>% View
```


## 2. Multivariable  Model and Results
### 1) Model

```{r}
mod_multinom_multivariable <-
  nnet::multinom(
    total_complications_cat_perID  ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      # infertilite_diag +
      # autoimmune_diseases +
      comorbidities_bin + 
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      nb_live_births_all_cat +

      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Csection_perID +

      methode_conception_perID_cat_simple_mod+
      Travail_accouchement_perID+
      
      doul_abdo_avant_1_preg,
    
    data = data)

summary(mod_multinom_multivariable)
tidy(mod_multinom_multivariable, exponentiate = TRUE, conf.int = TRUE) %>% View

```

### 2) Final model

```{r}
mod_multinom_multivariable_final <-
  nnet::multinom(
    total_complications_cat_perID  ~
      diplome_ini2 +
      age +
      
      gynecologique_diseases +
      # infertilite_diag +
      # autoimmune_diseases +
      comorbidities_bin + 
      
      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +
      indic_ovaire_touche +
      
      age_regle_cat_simple +
      nb_live_births_all_cat +

      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Csection_perID +

      methode_conception_perID_cat_simple_mod+
      Travail_accouchement_perID+
      
      doul_abdo_avant_1_preg,
    
    data = data)
```

### 3) Assumptions check 
#### - VIF  
- Use 'glm' model 

```{r}
glm_proxy <- glm(
  as.numeric(total_complications_cat_perID) ~
    diplome_ini2 +
    age +
    
    gynecologique_diseases +
    # infertilite_diag +
    # autoimmune_diseases +
    comorbidities_bin +
    
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    indic_ovaire_touche +
    
    age_regle_cat_simple +
    nb_live_births_all_cat +
    
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Csection_perID +

    methode_conception_perID_cat_simple_mod +
    Travail_accouchement_perID +
    
    doul_abdo_avant_1_preg,
  data= data
)
vif(glm_proxy) 

AIC(glm_proxy)
```


### 4) Visualization of results
#### - Table

   y.level                               term   OR       IC95
1        1                diplome_ini2>=Bac+3 0.33  0.14-0.73
2        1      age_firstOP_endoadeno_cat31>= 2.95  1.05-8.28
3        1           age_regle_cat_simple>=12 0.26   0.12-0.6
4        1 history_FC_preterm_IVGIMG_perIDOui 2.65  1.24-5.66
5        1          doul_abdo_avant_1_pregOui 2.65   1.35-5.2
6       2+                diplome_ini2>=Bac+3 0.33  0.13-0.81
7       2+                             DIEOui 2.05  1.01-4.17
8       2+      age_firstOP_endoadeno_cat31>= 3.42 1.13-10.39
9       2+             indic_ovaire_toucheOui 0.28  0.09-0.86
10      2+           age_regle_cat_simple>=12 0.23   0.1-0.54
11      2+            age_regle_cat_simplensp 0.05      0-0.6
12      2+ history_FC_preterm_IVGIMG_perIDOui 2.80  1.26-6.26
13      2+          doul_abdo_avant_1_pregOui 2.32  1.11-4.83

```{r}
res_multivar_multinom <- tidy(mod_multinom_multivariable_final, 
     conf.int = TRUE, exponentiate = TRUE) %>% as.data.frame() %>% 
  mutate(
    OR= round(estimate,2),
    p.value= round(p.value,2),
    conf.low = round(conf.low, 2),
    conf.high = round(conf.high, 2),
    IC95 = paste0(conf.low,"-",conf.high)
  )

# Check only significant variables
res_multivar_multinom %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(y.level,term,OR,IC95) %>% View 


sig_var_multinom_number <- res_multivar_multinom %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(term) %>% pull()

# Visualisation of table
library(kableExtra)
tbl_regression(mod_multinom_multivariable_final, exp = TRUE)

tidy(mod_multinom_multivariable_final, 
     conf.int = TRUE, exponentiate = TRUE) %>%
  kable() %>%
  kable_styling("basic", full_width = FALSE) 
```

#### - Forest plot 
```{r}
ggplot(res_multivar_multinom, 
       aes(x = term, y = OR, ymin = conf.low, ymax = conf.high, 
           color = y.level)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  labs(title = "Odds Ratios with 95% CI from Multinomial Logistic Regression",
       y = "Odds Ratio (OR)", x = "Variable") +
  theme_minimal()
```

#### - Effect plot  

```{r}
#Effect Plot
library(effects)
library(ggeffects)

# Prediction probability
#age_firstOP_endoadeno_cat
pred <- ggpredict(mod_multinom_multivariable_final, 
                  terms = "history_FC_preterm_IVGIMG_perID")
plot(pred)+theme_bw()

```


#---------------------------------------------------------
# IV. (Types) Multinominal outcome 
#---------------------------------------------------------
## 0.Define Data

```{r}
#----------------------
# Define the data
#----------------------
data <- imputed_data_O3_perID_V2 %>%
  dplyr::select(-total_complications_bin_perID,
                -total_complications_cat_perID) %>% 
  mutate(across(where(~ is.factor(.) | is.character(.)), as.factor))

data <- data %>% 
  mutate(
    Complications_cat_perID_grouped_mod = case_when(
      Complications_cat_perID_grouped == "Hemorr_postpartum_single" ~ 
        "Other_single",
      TRUE ~ Complications_cat_perID_grouped
    ) %>% factor(levels= c("No_complication","Metabolic_Hypertensive_single",
                           "Preterm_related_single","Other_single",
                           "Multiple_complications")),
    Complications_cat_perID_grouped = 
      factor(Complications_cat_perID_grouped, 
             levels= c("No_complication","Metabolic_Hypertensive_single",
                       "Preterm_related_single","Hemorr_postpartum_single",
                       "Other_single","Multiple_complications"))
  )

#check
data$Complications_cat_perID_grouped %>% table(useNA = "a")

# 
data <-  data %>%
  dplyr::select(-Age_premier_diag_endo_ade_cat,
                -stade_dvpt_I_II,
                -stade_dvpt_III_IV,
                
                -conception_natural_perID,
                -conception_natural_perID,
                -conception_operation_perID,
                -conception_PMA_perID,
                -labour_induced_perID,
                -delai_diag, 
                -delai_diag_cat,
                -situation_emploi_ini2_cat,
                -Csection_perID,
                -VD_perID,
                -age_regles_ini,
                -age_regle_cat,
                -nb_total_gross_cat,
                -nb_mal_sansEndoAde_cat,
                -nb_total_gross,
                -first_preg_age_cat)
```

Regroupement of response variable 
```{r}
data %>%
  group_by(Complications_cat_perID_grouped) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = Complications_cat_perID_grouped, y = n)) + 
  geom_col(fill = "lightgreen", color = "darkgreen") +
  geom_text(aes(label = n), vjust = -0.3, size = 3.5) +
  labs(
    title = "Distribution of types of pregnancy complications",
    x = "Types of pregnancy complication",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

data %>%
  group_by(Complications_cat_perID_grouped_mod) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = Complications_cat_perID_grouped_mod, y = n)) + 
  geom_col(fill ="lightblue", color = "darkblue") +
  geom_text(aes(label = n), vjust = -0.3, size = 3.5) +  
  labs(
    title = "Distribution of types of pregnancy complications",
    x = "Types of pregnancy complication",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



## 1. Function 
- Function to create a result table from a multivariable multinomial logistic regression model


```{r}
library(broom)
library(nnet)
#----------------------------------
# Final dataset for the analysis
#----------------------------------
data  <- data %>%
  dplyr::select(-Complications_cat_perID_grouped)

#--------------------------------------------------------
# Multinomial Univariable adjusted result table Function
#--------------------------------------------------------
multinomial_univariable_table <- function(variables, outcome, adjust_var, data) {
  table_dfs <- list()
  variables <- predictor_variables
  adjust_var <- adjust_sets$Age_adjusted
  outcome <- outcome_variable

  for (var in variables) {
    # Dynamically create the adjustment part of the formula
    # If adjust_var is NULL or empty, it should be an empty string
    if (is.null(adjust_var) || length(adjust_var) == 0) {
      adjust_part <- ""
    } else {
      # Add backticks around each adjustment variable name and collapse with " + "
      adjust_part <- paste0(" + `", paste(adjust_var, collapse = "` + `"), "`")
    }

    # Create the complete model formula string
    formula_str <- paste0("`", outcome, "` ~ `", var, "`", adjust_part)
    formula <- as.formula(formula_str)

    # Application of multinom()
    model <- nnet::multinom(formula, data = data, trace = FALSE)

    # Results with tidy()
    model_tidy <- broom::tidy(model,
                             exponentiate = TRUE,
                             conf.int = TRUE)

    # Create a new variable : OR (95% CI)
    model_tidy <- model_tidy %>%
      mutate(
        OR = round(estimate, 1),
        CI_low = round(conf.low, 1),
        CI_high = round(conf.high, 1),
        OR_IC = paste0(OR, " (", CI_low, ";", CI_high, ")"),
        Outcome = y.level
      ) %>%
      dplyr::select(Outcome, term, OR_IC)

    # Arrange variables' name
    model_tidy <- model_tidy %>%
      filter(stringr::str_detect(term, paste0("^", var))) %>% # Using stringr::str_detect for clarity
      mutate(
        Characteristic = stringr::str_remove(term, paste0("^", var))
      ) %>%
      dplyr::select(Outcome, Characteristic, OR_IC)

    # wide format
    joint_table <- model_tidy %>%
      pivot_wider(names_from = Outcome,
                  values_from = OR_IC,
                  names_prefix = "OR_")

    # Descriptive analysis without adjust_var
    # Note: CreateTableOne is for descriptive stats, not directly affected by adjustment vars in the model.
    # It seems you want a table of the predictor variable distribution by outcome.
    table_one <- CreateTableOne(data = data,
                                vars = var,
                                strata = outcome)

    table_one <- as.data.frame(print(table_one, showAllLevels = TRUE))
    table_one$var_col <- rownames(table_one) # Use a new name to avoid conflict with 'var' loop variable
    rownames(table_one) <- NULL

    # Clean up the 'level' column: fill in missing levels("") and remove empty or NA rows
    table_one <- table_one %>%
      mutate(level = ifelse(level == "" & var_col != "n", NA, level), # Use var_col here
             level = ifelse(is.na(level), var_col, level)) %>% # Use var_col here
      filter(var_col == var_col) %>%
      filter(level != "" & !is.na(level))


    # Join the descriptive table with the OR/CI table by the 'level' column
    final_table <- left_join(table_one, joint_table,
                             by = c("level" = "Characteristic"))
    final_table <- final_table %>%
      rename(Characteristic = level) %>%
      dplyr::select(-p, -test, -var_col) # Select out the new var_col

    # Save the result to the list of data frames, one per variable
    table_dfs[[var]] <- final_table
  }

  # After the loop: combine all tables into one data frame, keeping the original variable name
  df_final <- bind_rows(table_dfs, .id = "Variable")
  return(df_final)
}


#--------------------------------------------------------------
# Definition of outcome, adjustement set, predictor variables
#--------------------------------------------------------------
outcome_variable <- "Complications_cat_perID_grouped_mod" # Modify
all_vars <- colnames(data) # all column's name

# Progressive adjustment variables set
adjust_sets <- list(
  
  # # CRUDE without adjustment
  # "Crude" = NULL,
  
  # Adjustment : Age
  "Age_adjusted" = c("age"),
  
  # Adjustment : Age + sociodemographic
  "Age_Socio_adjusted" = c(
    "age",
    "diplome_ini2",
    "situation_financiere",
    "statut_marital_ini_cat"
  ),
  
  # Adjustment : Age + sociodemographic + pregnancy related variables
  "Fully_adjusted" = c(
    "age",
    
    "diplome_ini2",
    "situation_financiere",
    "statut_marital_ini_cat",
    
    "nb_live_births_all",
    "history_FC_preterm_IVGIMG_perID",
    "first_preg_age",
    "Mode_accouchement_perID",
    "labour_natural_perID",
    "methode_conception_perID_cat_simple_mod"
  )
)

predictor_variables <- 
  setdiff(all_vars, 
          c(outcome_variable, unique(unlist(adjust_sets))))
```

## 2. Results Table
### 1) Age adjusted
```{r}
table_dfs <- list()
variables <- predictor_variables
adjust_var <- adjust_sets$Age_adjusted
outcome <- outcome_variable

for (var in variables) {
  # Dynamically create the adjustment part of the formula
  # If adjust_var is NULL or empty, it should be an empty string
  if (is.null(adjust_var) || length(adjust_var) == 0) {
    adjust_part <- ""
  } else {
    # Add backticks around each adjustment variable name and collapse with " + "
    adjust_part <- paste0(" + `", paste(adjust_var, collapse = "` + `"), "`")
  }
  
  # Create the complete model formula string
  formula_str <- paste0("`", outcome, "` ~ `", var, "`", adjust_part)
  formula <- as.formula(formula_str)
  
  # Application of multinom()
  model <- nnet::multinom(formula, data = data, trace = FALSE)
  
  # Results with tidy()
  model_tidy <- broom::tidy(model, exponentiate = TRUE, conf.int = TRUE)
  
  # Create a new variable : OR (95% CI)
  model_tidy <- model_tidy %>%
    mutate(
      OR = round(estimate, 1),
      CI_low = round(conf.low, 1),
      CI_high = round(conf.high, 1),
      OR_IC = paste0(OR, " (", CI_low, ";", CI_high, ")"),
      Outcome = y.level
    ) %>%
    dplyr::select(Outcome, term, OR_IC)
  
  # Arrange variables' name
  model_tidy <- model_tidy %>%
    filter(stringr::str_detect(term, paste0("^", var))) %>% # Using stringr::str_detect for clarity
    mutate(Characteristic = stringr::str_remove(term, paste0("^", var))) %>%
    dplyr::select(Outcome, Characteristic, OR_IC)
  
  # wide format
  joint_table <- model_tidy %>%
    pivot_wider(
      names_from = Outcome,
      values_from = OR_IC,
      names_prefix = "OR_"
    )
  
  # Descriptive analysis without adjust_var
  # Note: CreateTableOne is for descriptive stats, not directly affected by adjustment vars in the model.
  # It seems you want a table of the predictor variable distribution by outcome.
  table_one <- CreateTableOne(data = data,
                              vars = var,
                              strata = outcome)
  
  table_one <- as.data.frame(print(table_one, showAllLevels = TRUE))
  table_one$var_col <- rownames(table_one) # Use a new name to avoid conflict with 'var' loop variable
  rownames(table_one) <- NULL
  
  # Clean up the 'level' column: fill in missing levels("") and remove empty or NA rows
  table_one <- table_one %>%
    mutate(
      level = ifelse(level == "" &
                       var_col != "n", NA, level),
      # Use var_col here
      level = ifelse(is.na(level), var_col, level)
    ) %>% # Use var_col here
    filter(var_col == var_col) %>%
    filter(level != "" & !is.na(level))
  
  
  # Join the descriptive table with the OR/CI table by the 'level' column
  final_table <- left_join(table_one, joint_table, by = c("level" = "Characteristic"))
  final_table <- final_table %>%
    rename(Characteristic = level) %>%
    dplyr::select(-p, -test, -var_col) # Select out the new var_col
  
  # Save the result to the list of data frames, one per variable
  table_dfs[[var]] <- final_table
}

# After the loop: combine all tables into one data frame, keeping the original variable name
df_final_age <- bind_rows(table_dfs, .id = "Variable")
df_final_age
```



```{r}
# 1) Age adjusted 
final_table_age <- multinomial_univariable_table(
  predictor_variables, # each predictor/exposition variable
  outcome_variable,
  adjust_sets$Age_adjusted, 
  data = data) %>%
  mutate(
    across(starts_with("OR_"),~ case_when(
      Characteristic!="" & is.na(.) ~ "ref",
      TRUE ~ .))
  )%>%
  group_by(Variable) %>%
  mutate(Variable = ifelse(row_number() == 1, Variable, "")) %>%
  ungroup()

final_table_age %>% dplyr::select(OR_Metabolic_Hypertensive_single:OR_Multiple_complications) %>% View


# Save as Excel file
write_xlsx(df_final_age, 
           "C:/Stage_Master2/Stage_M2_Complic_Gross/multinom_univar_table_age.xlsx")
```

### 2) Age+sociodemographic adjusted
```{r}
table_dfs <- list()
variables <- predictor_variables
adjust_var <- adjust_sets$Age_Socio_adjusted
outcome <- outcome_variable

for (var in variables) {
  # Dynamically create the adjustment part of the formula
  # If adjust_var is NULL or empty, it should be an empty string
  if (is.null(adjust_var) || length(adjust_var) == 0) {
    adjust_part <- ""
  } else {
    # Add backticks around each adjustment variable name and collapse with " + "
    adjust_part <- paste0(" + `", paste(adjust_var, collapse = "` + `"), "`")
  }
  
  # Create the complete model formula string
  formula_str <- paste0("`", outcome, "` ~ `", var, "`", adjust_part)
  formula <- as.formula(formula_str)
  
  # Application of multinom()
  model <- nnet::multinom(formula, data = data, trace = FALSE)
  
  # Results with tidy()
  model_tidy <- broom::tidy(model, exponentiate = TRUE, conf.int = TRUE)
  
  # Create a new variable : OR (95% CI)
  model_tidy <- model_tidy %>%
    mutate(
      OR = round(estimate, 1),
      CI_low = round(conf.low, 1),
      CI_high = round(conf.high, 1),
      OR_IC = paste0(OR, " (", CI_low, ";", CI_high, ")"),
      Outcome = y.level
    ) %>%
    dplyr::select(Outcome, term, OR_IC)
  
  # Arrange variables' name
  model_tidy <- model_tidy %>%
    filter(stringr::str_detect(term, paste0("^", var))) %>% # Using stringr::str_detect for clarity
    mutate(Characteristic = stringr::str_remove(term, paste0("^", var))) %>%
    dplyr::select(Outcome, Characteristic, OR_IC)
  
  # wide format
  joint_table <- model_tidy %>%
    pivot_wider(
      names_from = Outcome,
      values_from = OR_IC,
      names_prefix = "OR_"
    )
  
  # Descriptive analysis without adjust_var
  # Note: CreateTableOne is for descriptive stats, not directly affected by adjustment vars in the model.
  # It seems you want a table of the predictor variable distribution by outcome.
  table_one <- CreateTableOne(data = data,
                              vars = var,
                              strata = outcome)
  
  table_one <- as.data.frame(print(table_one, showAllLevels = TRUE))
  table_one$var_col <- rownames(table_one) # Use a new name to avoid conflict with 'var' loop variable
  rownames(table_one) <- NULL
  
  # Clean up the 'level' column: fill in missing levels("") and remove empty or NA rows
  table_one <- table_one %>%
    mutate(
      level = ifelse(level == "" &
                       var_col != "n", NA, level),
      # Use var_col here
      level = ifelse(is.na(level), var_col, level)
    ) %>% # Use var_col here
    filter(var_col == var_col) %>%
    filter(level != "" & !is.na(level))
  
  
  # Join the descriptive table with the OR/CI table by the 'level' column
  final_table <- left_join(table_one, joint_table, by = c("level" = "Characteristic"))
  final_table <- final_table %>%
    rename(Characteristic = level) %>%
    dplyr::select(-p, -test, -var_col) # Select out the new var_col
  
  # Save the result to the list of data frames, one per variable
  table_dfs[[var]] <- final_table
}

# After the loop: combine all tables into one data frame, keeping the original variable name
df_final_age_socio <- bind_rows(table_dfs, .id = "Variable")
df_final_age_socio
```


```{r}
# Age + socio demographic adjusted
final_table_age_socio <- multinomial_univariable_table(
  predictor_variables, # each predictor/exposition variable
  outcome_variable,
  adjust_sets$Age_Socio_adjusted, 
  data = data) %>%
  mutate(
    across(starts_with("OR_"),~ case_when(
      Characteristic!="" & is.na(.) ~ "ref",
      TRUE ~ .))
  )%>%
  group_by(Variable) %>%
  mutate(Variable = ifelse(row_number() == 1, Variable, "")) %>%
  ungroup()

final_table_age_socio %>% dplyr::select(OR_Metabolic_Hypertensive_single:OR_Multiple_complications) %>% View

# Save as Excel file
write_xlsx(df_final_age_socio,
           "C:/Stage_Master2/Stage_M2_Complic_Gross/multinom_univar_table_age_socio.xlsx")
```

### 3) Age+socio+preg adjusted
```{r}
table_dfs <- list()
variables <- predictor_variables
adjust_var <- adjust_sets$Fully_adjusted
outcome <- outcome_variable

for (var in variables) {
  # Dynamically create the adjustment part of the formula
  # If adjust_var is NULL or empty, it should be an empty string
  if (is.null(adjust_var) || length(adjust_var) == 0) {
    adjust_part <- ""
  } else {
    # Add backticks around each adjustment variable name and collapse with " + "
    adjust_part <- paste0(" + `", paste(adjust_var, collapse = "` + `"), "`")
  }
  
  # Create the complete model formula string
  formula_str <- paste0("`", outcome, "` ~ `", var, "`", adjust_part)
  formula <- as.formula(formula_str)
  
  # Application of multinom()
  model <- nnet::multinom(formula, data = data, trace = FALSE)
  
  # Results with tidy()
  model_tidy <- broom::tidy(model, exponentiate = TRUE, conf.int = TRUE)
  
  # Create a new variable : OR (95% CI)
  model_tidy <- model_tidy %>%
    mutate(
      OR = round(estimate, 1),
      CI_low = round(conf.low, 1),
      CI_high = round(conf.high, 1),
      OR_IC = paste0(OR, " (", CI_low, ";", CI_high, ")"),
      Outcome = y.level
    ) %>%
    dplyr::select(Outcome, term, OR_IC)
  
  # Arrange variables' name
  model_tidy <- model_tidy %>%
    filter(stringr::str_detect(term, paste0("^", var))) %>% # Using stringr::str_detect for clarity
    mutate(Characteristic = stringr::str_remove(term, paste0("^", var))) %>%
    dplyr::select(Outcome, Characteristic, OR_IC)
  
  # wide format
  joint_table <- model_tidy %>%
    pivot_wider(
      names_from = Outcome,
      values_from = OR_IC,
      names_prefix = "OR_"
    )
  
  # Descriptive analysis without adjust_var
  # Note: CreateTableOne is for descriptive stats, not directly affected by adjustment vars in the model.
  # It seems you want a table of the predictor variable distribution by outcome.
  table_one <- CreateTableOne(data = data,
                              vars = var,
                              strata = outcome)
  
  table_one <- as.data.frame(print(table_one, showAllLevels = TRUE))
  table_one$var_col <- rownames(table_one) # Use a new name to avoid conflict with 'var' loop variable
  rownames(table_one) <- NULL
  
  # Clean up the 'level' column: fill in missing levels("") and remove empty or NA rows
  table_one <- table_one %>%
    mutate(
      level = ifelse(level == "" &
                       var_col != "n", NA, level),
      # Use var_col here
      level = ifelse(is.na(level), var_col, level)
    ) %>% # Use var_col here
    filter(var_col == var_col) %>%
    filter(level != "" & !is.na(level))
  
  
  # Join the descriptive table with the OR/CI table by the 'level' column
  final_table <- left_join(table_one, joint_table, by = c("level" = "Characteristic"))
  final_table <- final_table %>%
    rename(Characteristic = level) %>%
    dplyr::select(-p, -test, -var_col) # Select out the new var_col
  
  # Save the result to the list of data frames, one per variable
  table_dfs[[var]] <- final_table
}

# After the loop: combine all tables into one data frame, keeping the original variable name
df_final_fully <- bind_rows(table_dfs, .id = "Variable")
df_final_fully
```

```{r}
# Age + socio demographic + pregnancy related variables adjusted
final_table_fully <- multinomial_univariable_table(
  predictor_variables, # each predictor/exposition variable
  outcome_variable,
  adjust_sets$Fully_adjusted, 
  data = data) %>%
  mutate(
    across(starts_with("OR_"),~ case_when(
      Characteristic!="" & is.na(.) ~ "ref",
      TRUE ~ .))
  )%>%
  group_by(Variable) %>%
  mutate(Variable = ifelse(row_number() == 1, Variable, "")) %>%
  ungroup()

write_xlsx(df_final_fully, 
           "C:/Stage_Master2/Stage_M2_Complic_Gross/multinom_univar_table_fully.xlsx")
```


## 3. Multivarible analysis

```{r}
data <- imputed_data_O3_perID_V2 %>%
  dplyr::select(-total_complications_bin_perID,
                -total_complications_cat_perID) %>% 
  mutate(across(where(~ is.factor(.) | is.character(.)), as.factor))

data <- data %>% 
  mutate(
    Complications_cat_perID_grouped_mod = case_when(
      Complications_cat_perID_grouped == "Hemorr_postpartum_single" ~ 
        "Other_single",
      TRUE ~ Complications_cat_perID_grouped
    ) %>% factor(levels= c("No_complication","Metabolic_Hypertensive_single",
                           "Preterm_related_single","Other_single",
                           "Multiple_complications")),
    Complications_cat_perID_grouped = 
      factor(Complications_cat_perID_grouped, 
             levels= c("No_complication","Metabolic_Hypertensive_single",
                       "Preterm_related_single","Hemorr_postpartum_single",
                       "Other_single","Multiple_complications"))
  )

data  <- data %>%
  dplyr::select(-Complications_cat_perID_grouped,
                -Age_premier_diag_endo_ade_cat,
                -stade_dvpt_I_II,
                -stade_dvpt_III_IV)
```


### 3-1. univariable analysis adjusted for age 
#### 1) Results Table
```{r}
final_table_age
```


```{r}
library(nnet)
# data  <- data %>%
#   dplyr::select(-Complications_cat_perID_grouped)

#---------------
# Configuration
#---------------
# 1. Define your outcome and adjustment variables
outcome_variable <- "Complications_cat_perID_grouped_mod"
adjustment_variable <- "age"
predictor_variables <-
  setdiff(names(data), c(outcome_variable,adjustment_variable))

#-----------------------------------------------------------------
# 1. Create Descriptive Statistics Table (Stratified by Outcome)
#-----------------------------------------------------------------
# This table shows N(%) or Mean (SD) for each predictor within each outcome group
tbl_descriptive <- data %>%
  dplyr::select(all_of(outcome_variable), all_of(predictor_variables)) %>%
  tbl_summary(
    by = all_of(outcome_variable),         # Stratify by the outcome
    statistic = list(                     # Define stats for variable types
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  # add_p() %>%   # Optional: Add p-value for difference between groups
  modify_header(label ~ "**Characteristic**") %>%
  # Remove the overall column if you only want stratified stats (optional)
  # remove_overall() %>%
  bold_labels()

print("Step 1: Descriptive Statistics Table Created.")
# tbl_descriptive # Uncomment to view intermediate table

#-------------------------------------------------------------
# 2. Create Adjusted univariable multinomial Regression Table
#-------------------------------------------------------------
# This function fits multinom(outcome ~ predictor + ajusted variables) and returns a gtsummary table row
fit_adjusted_multinom <- function(predictor, outcome_var, adjust_var, data) {

  adjust_str <- sprintf("`%s`", adjust_var)
  formula_str <- paste0("`", outcome_var, "` ~ `", predictor, "` + ", adjust_str)

  # Fit the model
  model <- tryCatch(
    nnet::multinom(as.formula(formula_str), data = data),
    error = function(e) {
      warning(paste("Could not fit ordinal regression model for predictor:", predictor, "- Error:", conditionMessage(e)), call. = FALSE)
      return(NULL) # Return NULL if model fails
    }
  )

  # If model failed, return NULL
  if (is.null(model)) return(NULL)

  # Create a regression table, keeping only the row(s) for the main predictor
  tbl_regression(model, exponentiate = TRUE, conf.int = TRUE) %>%
    modify_table_body(~ dplyr::filter(.x, variable == predictor))
}

# Apply the function to each predictor variable
list_of_adj_tables <- map(
  predictor_variables,
  ~fit_adjusted_multinom(.x, outcome_variable,
                         adjustment_variable,
                         data))

# Remove any NULL results (from failed models)
list_of_adj_tables <-
  list_of_adj_tables[!sapply(list_of_adj_tables,is.null)]

# Stack the individual predictor tables into one
if (length(list_of_adj_tables) > 0) {
  tbl_adjusted_ors <- tbl_stack(list_of_adj_tables,
                                group_header = NULL) %>%
    modify_table_body(
      ~ .x %>%
        mutate(
          est_ci = if_else(
            !is.na(estimate),
            sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
            NA_character_
          )
        )
    ) %>%
    modify_header(
      label ~"**Characteristic**",
      est_ci ~"**Adjusted Proportional OR (95% CI)**" # Clarify as Proportional OR
    )
  print("Step 2: Adjusted Odds Ratios Table Created.")
} else {
  warning("No successful ordinal regression models were fitted for adjusted ORs.",
          call. = FALSE)
  tbl_adjusted_ors <- NULL
}

#----------------------------------------------
# 3. Merge Descriptive and Adjusted OR Tables
#----------------------------------------------
# The 'adjustment_variable' is a single string in this setup, so no paste collapse needed
adjust_str <- adjustment_variable

if (!is.null(tbl_adjusted_ors)) {
    # Merge the two tables. gtsummary aligns rows based on the 'label' column.
    tbl_final <- tbl_merge(
      tbls = list(tbl_descriptive, tbl_adjusted_ors),
      # Add spanners to differentiate the table sections
      tab_spanner = c("**Descriptive Statistics by Outcome**",
                      paste0("**Age-Adjusted Proportional Odds Model**")) # Update spanner for clarity
    ) %>%
      modify_caption(
        paste0(
          "**multivariable Ordinal Logistic Regression Analysis for '",
          outcome_variable,
          "' (Adjusted for '", adjust_str, "')**")) # Update caption for clarity

    print("Step 3: Merged Table Created.")

    # --- Display the final table ---
    tbl_final
    } else {
    print("Final merged table could not be created as the adjusted OR table is empty.")
    print("Displaying only the descriptive statistics table:")
    tbl_descriptive %>%
        modify_caption(
          paste0(
            "**Descriptive Statistics for Predictors of '",
            outcome_variable,"'**"))
    }
#tbl_final

#------------------------------------------------------------
# MERGED TABLE as tibble for the facility of coping results
#------------------------------------------------------------
O3_tab_multinom_univariable_Complic_final <- as_tibble(tbl_final) %>%
  dplyr::select(-"**OR**",-"**95% CI**")

O3_tab_multinom_univariable_Complic_final <- 
  O3_tab_multinom_univariable_Complic_final %>%
    dplyr::select(1, 7, 8, 9) %>% 
    dplyr::rename(
      Characteristic = 1,
      Group = 2,
      p_value = 3,
      OR_95CI = 4
    ) %>% 
  mutate(p_value= as.numeric(as.character(p_value)))

# O3_tab_multinom_univariable_Complic_final %>% glimpse

# write.csv(O3_tab_multinom_univariable_Complic_final,"O3_tab_multinom_age_socio_preg_adjusted_Complic_final.csv")
```



#### 2) Pre-selection of variables for the multivariable regression model
 - Results of univariable analysis adjusted for age : P-value <= 0.2 

diplome_ini2
situation_financiere
gynecologique_diseases

age_regles_ini
age_regle_cat
age_regle_cat_simple

type_endo
DIE

age_symptome_cat
Age_premier_diag_endo_ade_cat
operation_endo_bin
complication_operation
age_firstOP_endoadeno_cat

nb_total_gross

nb_live_births_all
nb_live_births_all_cat

first_preg_age
first_preg_age_cat

history_FC_preterm_IVGIMG_perID
methode_conception_perID_cat_simple_mod

Mode_accouchement_perID
VD_perID
Csection_perID

Travail_accouchement_perID

doul_abdo_avant_1_preg

```{r}
O3_tab_multinom_univariable_Complic_final <- 
  O3_tab_multinom_univariable_Complic_final %>% 
  mutate(ind= rep(1:nrow(O3_tab_multinom_univariable_Complic_final),1)) %>% 
  dplyr::select(ind, everything())
# O3_tab_multinom_univariable_Complic_final %>% glimpse

# List of variables with p < 0.2
ls <- O3_tab_multinom_univariable_Complic_final %>% filter(p_value<0.2) %>% distinct(ind) %>% pull()
#O3_tab_multinom_univariable_Complic_final %>% View
```


### 3-2. Multivariable  Model and Results
#### 1) Reference Model
 - Without DIE 949.3822 vs With DIE 947.4213   => With DIE
```{r}
mod_multinom_multivariable_Complic <-
  nnet::multinom(
    Complications_cat_perID_grouped_mod ~
      age +
      diplome_ini2 +
      
      gynecologique_diseases +
      comorbidities_bin +
      
      DIE +
      endo_ade_categorie +
      age_firstOP_endoadeno_cat +
      
      age_regle_cat_simple +
      nb_live_births_all_cat +
      
      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Mode_accouchement_perID +
      methode_conception_perID_cat_simple_mod + # sensitivity analysis
      Travail_accouchement_perID +              # sensitivity analysis
      
      doul_abdo_avant_1_preg,
    
    data = data
  )

summary(mod_multinom_multivariable_Complic)
options(scipen = 999)  
tidy(mod_multinom_multivariable_Complic, 
     exponentiate = TRUE, conf.int = TRUE)
```

#### 2) Log linearity test
- ALL OK
```{r}
library(dplyr)
library(ggplot2)
library(rlang) # Required for `!!sym()` if you don't already have it

plot_linearity_check_flexible <- function(data, category_to_test, variable_to_check) {

  # Ensure the variable to check is numeric
  if (!is.numeric(data[[variable_to_check]])) {
    stop(paste0("Error: '", variable_to_check, "' is not a numeric variable in the provided data. Please check your dataframe column names."))
  }

  # Create the binary outcome variable
  df_b <- data %>%
    mutate(y_binary = ifelse(Complications_cat_perID_grouped_mod == category_to_test, 1, 0))

  # Dynamically build the formula for the simple glm model
  formula_str <- paste0("y_binary ~ ", variable_to_check)
  model_formula <- as.formula(formula_str)

  # Run a simple logistic regression with the specified numeric predictor
  model_simple <- glm(model_formula, family = binomial, data = df_b)

  # Calculate the predicted log-odds (logit)
  df_b$logit_prob <- predict(model_simple, type = "link")

  # Generate the plot using tidy evaluation with aes()
  # Use `!!sym(variable_to_check)` to unquote the string `variable_to_check`
  # and treat it as a variable name within `aes()`
  p <- ggplot(df_b, aes(x = !!sym(variable_to_check), y = logit_prob)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "loess", se = FALSE, color = "blue", linetype = "solid", size = 1) +
    geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed", size = 0.8) +
    labs(title = paste0("Linearity Check for ", variable_to_check, ": ", category_to_test, " vs Others"),
         x = variable_to_check,
         y = paste0("Log-Odds of (", category_to_test, " vs Others)")) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))

  print(p)
}

# Define your lists of categories and variables 
# Multinomial complication categories
categories <- c(
  "Preterm_related_single",
  "Other_single",
  "Metabolic_Hypertensive_single",
  "No_complication",
  "Multiple_complications"
)

# The numeric variables you want to check for linearity
numeric_variables_to_check <- c("age", "first_preg_ageli", "nb_live_births_all")

# Loop through each combination and generate plots
for (category in categories) {
  for (variable in numeric_variables_to_check) {
    cat(paste0("\nGenerating plot for Category: '", category, "' and Variable: '", variable, "'\n"))
    tryCatch({
      plot_linearity_check_flexible(df, category, variable)
    }, error = function(e) {
      message("An error occurred while plotting: ", e$message)
      message(paste0("Skipping combination: Category='", category, "', Variable='", variable, "'"))
    })
  }
}

# Interpretation 
# 파란색 LOESS 스무딩 라인이 빨간색 점선(선형 적합)에 가깝게 따라가면 선형성 가정이 합리적이라고 볼 수 있습니다.
# 만약 LOESS 라인이 S자 형태, U자 형태 또는 다른 뚜렷한 곡선 형태를 보이면 비선형성이 존재할 가능성이 높습니다.
```



```{r}
df <- data

# In R
df$log_x <- log(df$nb_live_births_all)
df$x_logx <- df$nb_live_births_all * log(df$nb_live_births_all)

# test 1 : Preterm_related_single vs others
df_b <- df %>% mutate(y_binary = 
                        ifelse(Complications_cat_perID_grouped_mod  ==
                                 "Preterm_related_single", 1, 0))
df_b$x_logx <- df_b$nb_live_births_all * log(df_b$nb_live_births_all)
model_b <- glm(y_binary ~ nb_live_births_all + x_logx, family = binomial, data = df_b)
summary(model_b)  

# test 2 : Other_single vs others
df_b <- df %>% mutate(y_binary = ifelse(Complications_cat_perID_grouped_mod  == "Other_single", 1, 0))
df_b$x_logx <- df_b$nb_live_births_all * log(df_b$nb_live_births_all)
model_b <- glm(y_binary ~ nb_live_births_all + x_logx, family = binomial, data = df_b)
summary(model_b)  

# test 3 : Metabolic_Hypertensive_single vs others
df_b <- df %>% mutate(y_binary = ifelse(Complications_cat_perID_grouped_mod  == "Metabolic_Hypertensive_single", 1, 0))
df_b$x_logx <- df_b$nb_live_births_all * log(df_b$nb_live_births_all)
model_b <- glm(y_binary ~ nb_live_births_all + x_logx, family = binomial, data = df_b)
summary(model_b)  

# test 4 : No_complication vs others
df_b <- df %>% mutate(y_binary = ifelse(Complications_cat_perID_grouped_mod  == "No_complication", 1, 0))
df_b$x_logx <- df_b$nb_live_births_all * log(df_b$nb_live_births_all)
model_b <- glm(y_binary ~ nb_live_births_all + x_logx, family = binomial, data = df_b)
summary(model_b)  

# test 5 : Multiple_complications vs others
df_b <- df %>% mutate(y_binary = ifelse(Complications_cat_perID_grouped_mod  == "Multiple_complications", 1, 0))
df_b$x_logx <- df_b$nb_live_births_all * log(df_b$nb_live_births_all)
model_b <- glm(y_binary ~ nb_live_births_all + x_logx, family = binomial, data = df_b)
summary(model_b)  
```


#### 2) Comparison of models

Sensitivity analysis for the two variables with lots of categories
- methode_conception_perID_cat_simple_mod (4 levels)
- Travail_accouchement_perID (4 levels)

Model	   AIC	  LogLik	McFadden R² AUC
mod1	947.4213	-369.71 	0.1787	    0.7522 ====> Selected 
mod2	929.1704	-376.59 	0.1634	    0.7393
mod3	932.9020	-374.45	  0.1682 	    0.743
mod4	935.7774	-371.89	  0.174 	    0.7471
      small     big       big         big
      
- Questions 

```{r}
# Common variables
common_formula <- paste(
  "age +
  diplome_ini2 +
  gynecologique_diseases +
  comorbidities_bin +
  endo_ade_categorie +
  DIE +
  age_firstOP_endoadeno_cat +
  age_regle_cat_simple +
  nb_live_births_all_cat +
  history_FC_preterm_IVGIMG_perID +
  first_preg_age +
  Mode_accouchement_perID +
  doul_abdo_avant_1_preg"
)

# Model formula
f1 <- as.formula(
  paste(
    "Complications_cat_perID_grouped_mod ~",
    common_formula,
    "+ methode_conception_perID_cat_simple_mod + 
    Travail_accouchement_perID"
  )
)

f2 <- as.formula(
  paste(
    "Complications_cat_perID_grouped_mod ~",
    common_formula,
    "+ conception_PMA_perID + 
    labour_natural_perID"
  )
)

f3 <- as.formula(
  paste(
    "Complications_cat_perID_grouped_mod ~",
    common_formula,
    "+ conception_PMA_perID + 
    conception_operation_perID+ 
    labour_natural_perID"
  )
)

f4 <- as.formula(
  paste(
    "Complications_cat_perID_grouped_mod ~",
    common_formula,
    "+ methode_conception_perID_cat_simple_mod+ 
    labour_natural_perID"
  )
)

# Modeling
mod1 <- nnet::multinom(f1, data = data)
mod2 <- nnet::multinom(f2, data = data)
mod3 <- nnet::multinom(f3, data = data)
mod4 <- nnet::multinom(f4, data = data)

# Results
## AIC
AIC(mod1, mod2, mod3,mod4)
# mod1 = 944.5843
# mod2 = 928.0619
# mod3 = 931.1610
# mod4 = 933.6931

BIC(mod1, mod2, mod3, mod4)

# Log-Likelihood : fitness 
logLik(mod1) #-369.7106  
logLik(mod2) #-376.5852
logLik(mod3) #-374.451 
logLik(mod4) #-371.8887 

## Pseudo-R² 
library(pscl)
pR2(mod1)["McFadden"] # 0.1787132 
pR2(mod2)["McFadden"] # 0.1634418  
pR2(mod3)["McFadden"] # 0.1681828  
pR2(mod4)["McFadden"] # 0.1738748  


library(pROC)
multiclass.roc(data$Complications_cat_perID_grouped_mod, 
               predict(mod1, type = "prob")) #0.7522
multiclass.roc(data$Complications_cat_perID_grouped_mod, 
               predict(mod2, type = "prob")) #0.7393
multiclass.roc(data$Complications_cat_perID_grouped_mod, 
               predict(mod3, type = "prob")) #0.743
multiclass.roc(data$Complications_cat_perID_grouped_mod, 
               predict(mod4, type = "prob")) #0.7471
```

- LASSO for the selection of variables 

```{r}
library(glmnet)

# 1) matrix : X
X <- model.matrix(
  Complications_cat_perID_grouped_mod ~
    age +
    diplome_ini2 +
    
    gynecologique_diseases +
    comorbidities_bin +
    
    endo_ade_categorie +
    DIE +
    age_firstOP_endoadeno_cat +
    
    age_regle_cat_simple +
    nb_live_births_all_cat +
    
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    
    Mode_accouchement_perID +
    # conception_PMA_perID +
    # labour_natural_perID +
    methode_conception_perID_cat_simple_mod +
    Travail_accouchement_perID +
    
    doul_abdo_avant_1_preg,
  data = data
)

y <- data$Complications_cat_perID_grouped_mod

# 2) glmnet 
mod_lasso <- cv.glmnet(
  x = X,
  y = y,
  family = "multinomial",
  alpha = 1,            # LASSO
  type.measure = "class" 
)

# 3) Optimal lambda
best_lambda <- mod_lasso$lambda.min

# 4) Final model with the optimal lambda
final_model <- glmnet(
  x = X,
  y = y,
  family = "multinomial",
  alpha = 1,
  lambda = best_lambda
)

# 5) Coef
coef(final_model)
```


#### 3) Final Model

```{r}
#mod_multinom_multivariable_Complic_final <- mod1

mod_multinom_multivariable_Complic_final <-
  nnet::multinom(
    Complications_cat_perID_grouped_mod  ~
      age +
      diplome_ini2 +

      gynecologique_diseases +
      comorbidities_bin +

      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +

      age_regle_cat_simple +
      nb_live_births_all_cat +

      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Mode_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      Travail_accouchement_perID+

      doul_abdo_avant_1_preg,

    data = data
  )

summary(mod_multinom_multivariable_Complic_final)
```

##### - Sensibility analysis : Imputed dataset vs Complete dataset

```{r}
data_sensibility <-
  df_final_complic_perID_V2_mod2 %>%
  mutate(
    comorbidities_bin =
      ifelse(nb_mal_sansEndoAde_cat == "0", "Non", "Oui"),
    Complications_cat_perID_grouped_mod = case_when(
      Complications_cat_perID_grouped == "Hemorr_postpartum_single" ~
        "Other_single",
      TRUE ~ Complications_cat_perID_grouped
    ) %>% factor(
      levels = c(
        "No_complication",
        "Metabolic_Hypertensive_single",
        "Preterm_related_single",
        "Other_single",
        "Multiple_complications"
      )
    ),
    Complications_cat_perID_grouped =
      factor(
        Complications_cat_perID_grouped,
        levels = c(
          "No_complication",
          "Metabolic_Hypertensive_single",
          "Preterm_related_single",
          "Hemorr_postpartum_single",
          "Other_single",
          "Multiple_complications"
        )
      ),
    nb_live_births_all_cat= 
      ifelse(as.numeric(as.character(nb_live_births_all)) >=2, 
             "2+",nb_live_births_all) %>% as.factor()
  )

# Model with complete dataset
mod_multinom_sensibility <-
  nnet::multinom(
    Complications_cat_perID_grouped_mod  ~
      age +
      diplome_ini2 +

      gynecologique_diseases +
      comorbidities_bin +

      endo_ade_categorie +
      DIE +
      age_firstOP_endoadeno_cat +

      age_regle_cat_simple +
      nb_live_births_all_cat+

      history_FC_preterm_IVGIMG_perID +
      first_preg_age +
      Mode_accouchement_perID +
      methode_conception_perID_cat_simple_mod +
      Travail_accouchement_perID+

      doul_abdo_avant_1_preg,

    data = data_sensibility #%>% na.omit()
  )

summary(mod_multinom_sensibility)

res_sensibility <- 
  tidy(mod_multinom_sensibility, exponentiate = TRUE, conf.int = TRUE) %>%
  as.data.frame() %>% 
  mutate(
    OR= round(estimate,2),
    p.value= round(p.value,2),
    conf.low = round(conf.low, 2),
    conf.high = round(conf.high, 2),
    IC95 = paste0(conf.low,"-",conf.high),
    OR_IC95= paste0(OR, "(",IC95,")")
  )
 
sig_var_multinom_type_sensibility<- res_sensibility%>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(term) %>% pull() 

sig_var_multinom_type <- res_multivar_multinom_Complic %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(term) %>% pull() 

intersect(sig_var_multinom_type_sensibility,sig_var_multinom_type)
```


#### 4) Assumptions check for the final model
##### - VIF  
- Use 'glm' model 

```{r}
# Based on mod1
glm_proxy <- glm(
  as.numeric(Complications_cat_perID_grouped_mod) ~
    age +
    diplome_ini2 +
    
    gynecologique_diseases +
    comorbidities_bin +
    
    DIE +
    endo_ade_categorie +
    age_firstOP_endoadeno_cat +
    
    age_regle_cat_simple +
    nb_live_births_all_cat +
    
    history_FC_preterm_IVGIMG_perID +
    first_preg_age +
    Mode_accouchement_perID +
    # mod1
    methode_conception_perID_cat_simple_mod +
    Travail_accouchement_perID +
    
    #mod2    
    # conception_PMA_perID +
    # labour_natural_perID +
    
    doul_abdo_avant_1_preg,
  data= data
)
vif(glm_proxy) 
```




#### 5) Visualization of results

```{r}
O3_multinom_multivar_descriptive_res <- data %>%
  dplyr::select(
    Complications_cat_perID_grouped_mod,
    age,
    diplome_ini2,
    gynecologique_diseases,
    comorbidities_bin,
    endo_ade_categorie,
    DIE,
    age_firstOP_endoadeno_cat,
    age_regle_cat_simple,
    nb_live_births_all_cat,
    history_FC_preterm_IVGIMG_perID,
    first_preg_age,
    Mode_accouchement_perID,
    methode_conception_perID_cat_simple_mod,
    Travail_accouchement_perID,
    doul_abdo_avant_1_preg) %>% 
  tbl_summary(
    by=Complications_cat_perID_grouped_mod,
    statistic = list(                   
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no",
    digits = list(all_continuous() ~ 1,
                  all_categorical() ~ c(0, 1)) # Control digits
  ) %>%
  modify_header(label ~ "**Characteristic**") %>%
  bold_labels() %>% as.data.frame()
# O3_multinom_multivar_descriptive_res

# write_xlsx(O3_multinom_multivar_descriptive_res,
#            "C:/Stage_Master2/Stage_M2_Complic_Gross/O3_multinom_multivar_descriptive_res.xlsx")
```


##### - Table
```{r}
res_multivar_multinom_Complic <- 
  tidy(mod_multinom_multivariable_Complic_final, 
     conf.int = TRUE, exponentiate = TRUE) %>% as.data.frame() %>% 
  mutate(
    OR= round(estimate,2),
    p.value= round(p.value,2),
    conf.low = round(conf.low, 2),
    conf.high = round(conf.high, 2),
    IC95 = paste0(conf.low,"-",conf.high),
    OR_IC95= paste0(OR, " (",IC95,")")
  ) %>% 
  filter(term != "(Intercept)")

res_multivar_multinom_Complic_final<- res_multivar_multinom_Complic %>%
  dplyr::select(term, y.level, OR_IC95) %>%
  pivot_wider(
    names_from = y.level,   
    values_from = OR_IC95   
  )

# write_xlsx(res_multivar_multinom_Complic_final,"C:/Stage_Master2/Stage_M2_Complic_Gross/O3_multinomial_multivariable_res.xlsx")

######################################
# Check only significant variables
res_multivar_multinom_Complic %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(y.level,term,OR,IC95) %>% View 

sig_var_multinom_type <- res_multivar_multinom_Complic %>% filter((conf.low<1&conf.high<1) |(conf.low>1&conf.high>1)|p.value<0.05) %>% distinct(term) %>% pull() 
######################################

#-------------------------
# Visualisation of table
#-------------------------
library(kableExtra)
tidy(mod_multinom_multivariable_Complic_final, 
     conf.int = TRUE, exponentiate = TRUE) %>%
  kable() %>%
  kable_styling("basic", full_width = TRUE) 
```

```{r}
res_multivar_multinom_Complic %>% dplyr::select(y.level, term, OR_IC95) %>% View

library(dplyr)
library(tidyr)

res_multivar_multinom_Complic %>%
  dplyr::select(term, y.level, OR_IC95) %>%
  pivot_wider(
    names_from = y.level,   
    values_from = OR_IC95   
  ) %>%
  View()

```

##### - Forest plot 

```{r}
library(ggplot2)

clean_res_tab <- res_multivar_multinom_Complic %>% 
  filter(term != "(Intercept)")

clean_res_tab$conf.high_mod <- pmin(clean_res_tab$conf.high, 10.5)

clean_res_tab$facet_title <- dplyr::recode(clean_res_tab$y.level,
  "Metabolic_Hypertensive_single" = "Only Metabolic&hypertensive",
  "Preterm_related_single" = "Only preterm related",
  "Other_single" = "Only other complications",
  "Multiple_complications" = "Multiple"
)

ls_term <-  c(
  "Age",
  "Niveau d'étude: Bac<= vs Bac+2",
  "Bac<= vs >=Bac+3",
  "Maladie gynécologique: No vs Yes",
  "Comorbidités: No vs Yes",
  "Diagnostic: Only endo vs only adeno",
  "Only endo vs Both",
  "Deep Inflating Endometriosis: No vs Yes",
  "Age at the 1st operation: No op vs <28",
  "No op vs 28-31",
  "No op vs >=31",
  "Age at the menarche: <12 vs >=12",
  "<12 vs don't know",
  "Parity: 1 vs 2+",
  "History of incomplete preg: No vs Yes",
  "Maternal age at the 1st preg",
  "Only NSVD vs only Cesarean",
  "Only NSVD vs Both",
  "Conception method: Natural vs Operation",
  "Natural vs Operation+ART",
  "Natural vs ART/Medication",
  "Labor: Only natural vs only induced",
  "Only natural vs Both",
  "Only natural vs No labor",
  "abdominal pain before 1st preg: No vs Yes"
)
clean_res_tab$term_new <- rep(ls_term, 4)


# visualisation
ggplot(
  clean_res_tab,
  aes(
    x = estimate,
    y = term_new,
    xmin = conf.low,
    xmax = conf.high_mod,
    colour = y.level
  )
) +
  geom_point() +
  geom_errorbarh(height = 0.2, show.legend = FALSE) +
  geom_vline(xintercept = 1,
             linetype = "dashed",
             color = "black") +
  geom_vline(
    xintercept = seq(0, 10, by = 2),
    linetype = "dotted",
    color = "gray80",
    size = 0.3
  ) +
  facet_wrap( ~ facet_title) +
  scale_x_continuous(limits = c(0, 10.5), breaks = seq(0, 10.5, 2)) +
  scale_color_manual(
    values = c(
      "Metabolic_Hypertensive_single" = "darkblue",
      "Preterm_related_single" = "darkgreen",
      "Other_single" = "purple",
      "Multiple_complications" = "darkred"
    )
  ) +
  labs(x = "Odds Ratio with CI95%", y = "Variables", 
       title = "Forest Plot by Regrouped complication types") +
  theme_light() +
  theme(
    legend.position = "none",
    strip.text = element_text(
      color = "black",
      size = 9,
      face = "bold"
    )
  )
```


```{r}
library(dplyr)
library(ggplot2)
library(purrr)

# 데이터 facet_title 별로 나누기
plot_list <- clean_res_tab %>%
  group_split(facet_title) %>%
  map(~ {
    ggplot(
      .x,
      aes(
        x = estimate,
        y = term_new,
        xmin = conf.low,
        xmax = conf.high_mod,
        colour = y.level
      )
    ) +
      geom_point() +
      geom_errorbarh(height = 0.2, show.legend = FALSE) +
      geom_vline(xintercept = 1,
                 linetype = "dashed",
                 color = "black") +
      geom_vline(
        xintercept = seq(0, 10, by = 2),
        linetype = "dotted",
        color = "gray80",
        size = 0.3
      ) +
      scale_x_continuous(limits = c(0, 10.5), breaks = seq(0, 10.5, 2)) +
      scale_color_manual(
        values = c(
          "Metabolic_Hypertensive_single" = "darkblue",
          "Preterm_related_single" = "darkgreen",
          "Other_single" = "purple",
          "Multiple_complications" = "darkred"
        )
      ) +
      labs(x = "Odds Ratio with CI95%", y = "Variables", 
           title = paste("Forest Plot -", unique(.x$facet_title))) +
      theme_light() +
      theme(
        legend.position = "none",
        strip.text = element_text(
          color = "black",
          size = 9,
          face = "bold"
        )
      )
  })

```


##### - Effect plot  

```{r}
#Effect Plot
library(effects)
library(ggeffects)
library(ggplot2)

# Prediction probability
#age_firstOP_endoadeno_cat
pred <- ggpredict(mod_multinom_multivariable_Complic_final, 
                  terms = "history_FC_preterm_IVGIMG_perID")
plot(pred)+theme_bw()

```





