---
title: "01_clean_Endo_adeno_initial"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Endo_adeno_initial
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data for the initial information of endometriosis/adenomyosis 
*Date of creation* : 2025-05-13
*Date of update* : 2025-06-02
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0.Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
# Endo_Adeno_initial <- read.csv("~/Projet/0.Data/complication_pop_QST_0418_20250226_endo_ade_init.csv",fileEncoding = "latin1",colClasses = c(Sexe = "character"))

Endo_Adeno_initial <-
  read.csv("Complic_pop_QST_0418_20250407_endo_adeno_init.csv",
           fileEncoding = "latin1",
           colClasses = c(Sexe = "character"))

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(across(everything(), ~na_if(trimws(.), "")))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- Endo_Adeno_initial %>% sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count==nrow(Endo_Adeno_initial)) %>% as.vector() # index of columns having only NA
Endo_Adeno_initial <- Endo_Adeno_initial[,-index_col_NA]
```


# I. Preprocessing Data

## - Outlier 

```{r}
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  mutate(across(starts_with(c("Age_","age_")), as.numeric))

sapply(Endo_Adeno_initial %>% dplyr::select(starts_with("age_")), summary)

# id_unik =="AFA16A96910CB83E7EC6"
# - outlier : Age_endo = 4 years old

# id_unik=="BE7A323B5380CE130080"
# - outlier : age_diag_endo = 4 years old (age_symptome 16 years old))
# - outlier : age_diag_adeno = 4 years old (age_symptome 16 years old))

# id_unik=="C4F27F1D79FF7B37BF8A"
# - outlier : age_symtome_adeno = 0 (id_unik=="C4F27F1D79FF7B37BF8A")
# - outlier : age_symptom = 0 (id_unik == "4619847D44E1E3C0E021")
```

## - Create tnew variables 

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  mutate(Endo_Adeno_initial_participated=
           ifelse(is.na(IdAction),"Non","Oui"))

#---------------------------------
# Family history of endometriosis : family_atcd_endo
#---------------------------------
# 1.dplyr::selecting and Cleaning Columns
new_dataframe <- Endo_Adeno_initial %>%
  dplyr::select(id_unik, starts_with("atcd_endo__")) %>%
  mutate(across(starts_with("atcd_endo__"), str_trim))

# 2.Creating a New Variable Based on Conditions
# Non, Non applicable : pas d'atcd 
new_dataframe$family_atcd_endo <-
  apply(new_dataframe[, 1:ncol(new_dataframe)], 1, function(row) {
    ifelse("Oui" %in% row,"Antécédents",
           ifelse("Non" %in% row,"Pas d'antécédents",
                  ifelse("Non applicable" %in% row, "Pas d'antécédents", NA)
           ))
  }) %>% as.factor()

# 3.Merging the New Variable with the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, family_atcd_endo),
            by = "id_unik")

# 4.Removing unnecessary Columns 
Endo_Adeno_initial <- Endo_Adeno_initial %>%  
  dplyr::select(-starts_with("atcd_endo__"))

# 5.Removing the Temporary new_dataframe
rm(new_dataframe)

#---------------------------------
# Family history of adenomyosis
#---------------------------------
# 1.dplyr::selecting and Cleaning Columns
new_dataframe <- Endo_Adeno_initial %>%
  dplyr::select(id_unik, starts_with("atcd_adeno__")) %>%
  mutate(across(starts_with("atcd_adeno__"), str_trim))

# 2.Creating a New Variable Based on Conditions
new_dataframe$family_atcd_adeno <- apply(new_dataframe[, 1:ncol(new_dataframe)], 1, function(row) {
  ifelse("Oui" %in% row, "Antécédents", 
         ifelse("Non" %in% row, "Pas d'antécédents",
                ifelse("Non applicable"  %in% row, "Pas d'antécédents",NA)))
  }) %>% as.factor()

# 3.Merging the new variable with the original dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, family_atcd_adeno),
            by = "id_unik")

# 4.Removing unnecessary columns 
Endo_Adeno_initial <- Endo_Adeno_initial %>%  
  dplyr::select(-starts_with("atcd_adeno__"))

# 5.Removing the Temporary new_dataframe
rm(new_dataframe)


#-----------------------------------
# Family history of Endo or Adeno
#-----------------------------------
# 1. Create a temporary dataframe
new_dataframe <- Endo_Adeno_initial %>% 
  dplyr::select(id_unik,starts_with("family_atcd"))

# 2. Make a new variable 
new_dataframe <- new_dataframe %>% 
  mutate(family_atcd_EndoAdeno=
           case_when(
             family_atcd_endo=="Antécédents"& family_atcd_adeno=="Antécédents" ~ 
               "Both",
             family_atcd_endo=="Antécédents"& family_atcd_adeno!="Antécédents"~ 
               "Only endo",
             family_atcd_endo!="Antécédents"& family_atcd_adeno=="Antécédents"~ 
               "Only adeno",
             family_atcd_endo=="Pas d'antécédents"& family_atcd_adeno=="Pas d'antécédents"~ 
               "No family history",
             !is.na(family_atcd_endo) & is.na(family_atcd_adeno) ~"Only endo",
             is.na(family_atcd_endo) & !is.na(family_atcd_adeno) ~"Only adeno",
             TRUE ~ NA_character_
           ))

# 3. Add this new variable to the main dataframe 
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik,family_atcd_EndoAdeno))

#4. Delete the temporary dataframe 
rm(new_dataframe)

#--------------------------------
# Family history of pelvic pain
#--------------------------------
new_dataframe <- Endo_Adeno_initial %>%
  dplyr::select(id_unik, starts_with("douleur_pelvienne__")) %>%
  mutate(across(starts_with("douleur_pelvienne__"), str_trim))

new_dataframe$family_atcd_doul_pelv <-apply(new_dataframe[, 1:ncol(new_dataframe)], 1, function(row) {
  ifelse("Oui" %in% row, "Antécédents", 
         ifelse("Non" %in% row, "Pas d'antécédents",
                ifelse("Non applicable"  %in% row, "Pas d'antécédents",NA)))
  }) %>%  as.factor()

Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, family_atcd_doul_pelv),
            by = "id_unik")

Endo_Adeno_initial <- Endo_Adeno_initial %>%  
  dplyr::select(-starts_with("douleur_pelvienne__"))
rm(new_dataframe)


#----------------------------------------------------------------
#  Symptoms that led to consultation before the first diagnosis (symptome_consult_)
#----------------------------------------------------------------
# 1. Convert All Columns to Character Format
Endo_Adeno_initial[] <- lapply(Endo_Adeno_initial, as.character)

# 2. Handle non standard encodings
library(stringi) # advanced string manipulation functions
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(across(starts_with("symptome_consult_"), 
                ~ stri_replace_all_fixed(., c("\u0092", "\u0085"), c("'", "…"), vectorize_all = FALSE))) 
# \u0092 : a non-standard apostrophe, \u0085: an unusual line break character

# 3. Replace "Autre" with "Autre ..."
cols_modif_autre <- c("symptome_consult_1", "symptome_consult_2", "symptome_consult_3", "symptome_consult_4")
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(across(all_of(cols_modif_autre), ~ ifelse(. == "Autre", "Autre ...", .)))
#Alternative approach
# for (col in cols_modif_autre) {
#   Endo_Adeno_initial[[col]][Endo_Adeno_initial[[col]] == "Autre"] <- "Autre ..."
# }

# 4. Define Symptom Categories
#Here, I added a '\' before '(' due to an encoding issue; it might not work in your code, so you should #check
infertil = "Infertilité \\(tentative de concevoir pendant au moins 12 mois consécutifs sans obtenir de grossesse, même incomplète\\)"
douleur = "Douleurs \\(liées aux règles, lors des rapports sexuels, de l’évacuation d’urine ou de selles…\\)"
autre_dig =  "Autres symptômes \\(symptômes digestifs…\\)"
autre = "Autre ..."

# 5. Create a New Data Frame with Logical Indicators
new_dataframe <- Endo_Adeno_initial %>%
  dplyr::select(id_unik, symptome_consult_1:symptome_consult_4) %>%
  mutate(
    infertil_logique = rowSums(across(starts_with("symptome_consult_"), ~grepl(infertil, .))) > 0,
    douleur_logique = rowSums(across(starts_with("symptome_consult_"), ~ grepl(douleur, .))) > 0,
    autre_dig_logique = rowSums(across(starts_with("symptome_consult_"), ~ grepl(autre_dig, .))) > 0,
    autre_logique = rowSums(across(starts_with("symptome_consult_"), ~ grepl(autre, .))) > 0
    ) %>% 
# 6. Classify Symptoms 
  mutate(
    signe_symptome_premiere_consult = case_when(
      douleur_logique & !(infertil_logique | autre_dig_logique | autre_logique) ~ "Pain",
      autre_dig_logique & !(infertil_logique | douleur_logique | autre_logique) ~ "Other symptoms (digestive…)",
      infertil_logique & !(autre_dig_logique | douleur_logique | autre_logique) ~ "Infertility",
      autre_logique & !(autre_dig_logique | douleur_logique | infertil_logique) ~ "Other",
      douleur_logique | infertil_logique | autre_dig_logique | autre_logique ~ "Multiple signs and symptoms",
      TRUE ~ NA_character_)
  )

# 7. Merge the Classification Back into the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, signe_symptome_premiere_consult),
            by = "id_unik")


# 8. Remove new dataframe and variables symptome_conlsult_ from the original dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>%  
  dplyr::select(-starts_with("symptome_consult_"))
rm(new_dataframe)

#------------------------------------------------------------------
#   Create a variable that gives the age at the first diagnosis
#------------------------------------------------------------------
# 1. Convert all columns related to 'age' to numeric format 
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(across(starts_with("age_"),
                ~as.numeric(gsub("[^0-9.-]", "",
                                 as.character(.)))))

# 2. p-min :  find the row-wise minimum across multiple vectors or columns
Endo_Adeno_initial$Age_premier_diag_endo_ade <- 
  do.call(pmin, #  row-wise minimum
          c(Endo_Adeno_initial[, c("Age_endo", "age_diag_endo", "age_diag_adeno", "age_adeno", "age_diag_les2")], 
            na.rm = TRUE))

# 3. Create a categorical variable 
quantiles <- quantile(Endo_Adeno_initial$Age_premier_diag_endo_ade, probs=c(1/3,2/3), na.rm=TRUE)

Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  mutate(
    Age_premier_diag_endo_ade_cat= 
      case_when(
        Age_premier_diag_endo_ade< quantiles[[1]] ~ 
          paste0("<", quantiles[1]),
        Age_premier_diag_endo_ade>= quantiles[[1]] &
          Age_premier_diag_endo_ade< quantiles[[2]] ~
          paste0(quantiles[1], "-", quantiles[2]),
        Age_premier_diag_endo_ade>= quantiles[[2]] ~ 
          paste0(">=", quantiles[2])
      )
  )

# 4. Remove all variables related to the age of diagnosis sauf Age_premier_diag_endo_ade
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  dplyr::select(-age_diag_les2,-Age_endo,-age_diag_endo,-age_diag_adeno,-age_adeno)

# 5. Delete the quantile information
rm(quantiles)

#-----------------------
#   Diagnostic Delay    : numeric format and category format
#-----------------------
# Endo_Adeno_initial$age_symptome <- Endo_Adeno_initial$age_symptome %>% as.numeric()
# Endo_Adeno_initial$Age_premier_diag_endo_ade <- Endo_Adeno_initial$Age_premier_diag_endo_ade %>% as.numeric()
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  mutate(delai_diag = 
           Age_premier_diag_endo_ade - age_symptome)

quantiles <- quantile(Endo_Adeno_initial$delai_diag, probs=c(1/3,2/3), na.rm=TRUE)

#Endo_Adeno_initial$delai_diag=as.numeric(Endo_Adeno_initial$delai_diag)
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(delai_diag_cat = case_when(
    delai_diag <= 5  ~ "<= 5",
    delai_diag >5 & delai_diag <=10  ~ "(5;10]",
    delai_diag >10 & delai_diag <=15 ~ "(10;15]",
    delai_diag >=15 ~ "15+",
    symptomes == "Non"  ~ "No symptoms",
    TRUE ~ NA_character_) %>% 
      factor(
        levels=c("No symptoms","<= 5","(5;10]","(10;15]","15+"),
        ordered = TRUE),
    
    delai_diag_cat_simple = case_when(
      delai_diag <= 7  ~ "<= 7",
      delai_diag > 7 & delai_diag <= 13  ~ "8-13",
      delai_diag > 13 ~ "13+",
      symptomes == "Non" ~ "No symptoms",
      TRUE~ NA_character_
      ) %>% 
      factor(
        levels=c("No symptoms","<= 7","8-13","13+"),
        ordered = TRUE)
  )

#---------------------------
# Age at the first symptom  : category format
#---------------------------
#quantile(Endo_Adeno_initial$age_symptome,probs = c(1/4,2/4,3/4),na.rm = TRUE)
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(age_symptome_cat = case_when(
    age_symptome <= 13  ~ "<= 13",
    age_symptome >13 & age_symptome <=15  ~ "(13;15]",
    age_symptome >15 & age_symptome <=22 ~ "(15;22]",
    age_symptome >22 ~ "22+",
    symptomes == "Non"  ~ "No symptoms",
    TRUE ~ NA_character_  )%>% 
      factor(
        levels=c("No symptoms","<= 13","(13;15]","(15;22]","22+"),
        ordered = TRUE)
  )

#-------------------------
#  Type of endometriosis
#-------------------------
# 1. Extracting Relevant Columns
new_dataframe <-  Endo_Adeno_initial %>% 
  dplyr::select(id_unik, starts_with("type_endo_"))

# 2. Defining Patterns for Classification
patterns <- list(
  ADENO = "Une adénomyose",
  OMA = "Une endométriose ovarienne \\(autres termes : endométriome, kyste endométriosique\\)",
  SPE = "Une endométriose superficielle \\(autre terme : péritonéale\\)",
  DIE  = "Une endométriose profonde",
  NOP = "Aucun de ces termes ne m'a été mentionné",
  NSP = "Je ne m'en souviens plus"
)

# 3. Creating Indicator Variables (Binary Encoding= TRUE 1/FALSE 0)
for (var in names(patterns)) {
  new_dataframe[[var]] <- rowSums(sapply(new_dataframe[, 2:5], grepl, pattern = patterns[[var]], ignore.case = TRUE), na.rm = TRUE) > 0
}

# 4. Classifying Endometriosis Types 
new_dataframe <- new_dataframe %>%
  mutate(
    type_endo = case_when(
      NOP ~ "Aucun mentionné/nsp",
      NSP & !ADENO & !OMA & !DIE & !SPE ~ "Aucun mentionné/nsp", 
      ADENO & !OMA & !DIE & !SPE ~ "ADENO only",
      OMA & !DIE & !SPE ~ "OMA only",
      DIE & !OMA & !SPE ~ "DIE only",
      SPE & !OMA & !DIE ~ "SPE only",
      (SPE + OMA + DIE + ADENO) > 1 ~ "Multiple types",  
      TRUE ~ NA_character_
    )
  )

# 5. Convert Indicator Variables to "Oui"/"Non" AFTER type_endo is created
for (var in names(patterns)) {
  new_dataframe[[var]] <- ifelse(new_dataframe[[var]], "Oui", "Non")
}

# 5. Merging Back to the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  left_join(new_dataframe %>% dplyr::select(-starts_with("type_endo_"),-NOP,-NSP), 
            by = "id_unik")

# 6. Removing all variables related to the type of endometriosis and the new dataframe
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  dplyr::select(-starts_with("type_endo_"))
rm(new_dataframe)

#-------------------------------------------------------
# Rename columns : The number of professional consultants and the number of consultations before the diagnosis of endometriosis/adenomyosis
categories <-
  c(
    "gyn",
    "radio",
    "GP",
    "Gastro_E",
    "chirDigest",
    "uro",
    "urgence",
    "Sagefemmes",
    "PMA",
    "Autre"
  )
cols_to_rename <- grep("^nb_consult__", colnames(Endo_Adeno_initial), value = TRUE)

if (length(cols_to_rename) == length(categories)) {
  names(Endo_Adeno_initial)[names(Endo_Adeno_initial) %in% cols_to_rename] <-
    paste0("nb_consult__", categories)
} else {
  stop("Le nombre de colonnes nb_consult__ ne correspond pas au nombre de catégories !")
}
cols_to_rename1 <- grep("^Nb_pro_consult__", colnames(Endo_Adeno_initial), value = TRUE)
if (length(cols_to_rename) == length(categories)) {
  names(Endo_Adeno_initial)[names(Endo_Adeno_initial) %in% cols_to_rename1] <-
    paste0("Nb_pro_consult__", categories)
} else {
  stop("Le nombre de colonnes nb_consult__ ne correspond pas au nombre de catégories !")
}

# Modifie all columns starting "Nb_pro_" to numeric format
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(across(starts_with("Nb_pro_"), ~ as.numeric(gsub("[^0-9]", "", .))))

#------------------------------------------------
# Number of specialities consulted at least once
#------------------------------------------------
# 1. Extracting Relevant Columns
new_dataframe <- Endo_Adeno_initial %>% 
  dplyr::select(id_unik, starts_with("Nb_pro_consult__"))

# 2. Number of specialities consulted at least once
new_dataframe$Nb_specialites_consultees <- rowSums(new_dataframe[, -1] >= 1, na.rm = TRUE)

# 3. Merging Back to the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, Nb_specialites_consultees),
            by = "id_unik")

#-------
# CALCUL
#-------
# - Total number of professional consultants by specialty  
# Convert all columns to numeric (handling potential character encoding issues)  
new_dataframe[] <- lapply(new_dataframe, function(x) as.numeric(as.character(x)))

# - Count the number of patients who consulted at least one professional in each specialty  
nb_patients_par_specialite <- colSums(new_dataframe[, -1] >= 1, na.rm = TRUE); nb_patients_par_specialite

# - Total number of consultations for each specialty  
total_consultations_par_specialite <- colSums(new_dataframe[, -1], na.rm = TRUE); total_consultations_par_specialite

# - Compute the average number of consultations per specialty, considering only patients who had at least one consultation  
moyenne_consultations <- total_consultations_par_specialite / nb_patients_par_specialite ; moyenne_consultations

total_patients <- nrow(new_dataframe); total_patients
# - Percentage of patients with at least one visit in each specialty
prop_patients_par_specialite <- nb_patients_par_specialite / total_patients;prop_patients_par_specialite

# - Percentage of all patients have at least one visit in any specialty
total_patients_with_any_consult <- sum(rowSums(new_dataframe[, -1], na.rm = TRUE) > 0)
total_patients_with_any_consult / total_patients

# 5. Removing the new dataframe
rm(new_dataframe)

#------------------------------------------------------------------
# Catégories : the number of specialities consulted at least once
#------------------------------------------------------------------
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  mutate(Nb_specialites_consult_cat = factor(
    cut(Nb_specialites_consultees, 
        breaks = quantile(Nb_specialites_consultees, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
        include.lowest = TRUE,
        labels = c("0-2", "2-3", "3-4", "4-10")
    ),
    ordered = TRUE
  ))

#----------------------------------------------------------------
# Type de professionnels de santé consultés avant du diagnostics
#----------------------------------------------------------------
# 1.  Creating the new dataframe with variables of interest : Nb_pro_consult__
new_dataframe <- Endo_Adeno_initial %>%
  dplyr::select(id_unik, starts_with('Nb_pro_consult__')) %>%
  mutate(
    # Création d'une variable binaire pour chaque type de consultation
    pro_gyn         = ifelse(Nb_pro_consult__gyn >= 1, 1, 0),
    pro_radio       = ifelse(Nb_pro_consult__radio >= 1, 1, 0),
    pro_GP          = ifelse(Nb_pro_consult__GP >= 1, 1, 0),
    pro_Gastro_E    = ifelse(Nb_pro_consult__Gastro_E >= 1, 1, 0),
    pro_chirDigest  = ifelse(Nb_pro_consult__chirDigest >= 1, 1, 0),
    pro_uro         = ifelse(Nb_pro_consult__uro >= 1, 1, 0),
    pro_urgence     = ifelse(Nb_pro_consult__urgence >= 1, 1, 0),
    pro_Sagefemmes  = ifelse(Nb_pro_consult__Sagefemmes >= 1, 1, 0),
    pro_PMA         = ifelse(Nb_pro_consult__PMA >= 1, 1, 0),
    pro_Autre       = ifelse(Nb_pro_consult__Autre >= 1, 1, 0),
    
    # Nombre total de professionnels consultés
    nb_pro_consultes = pro_gyn + pro_radio + pro_GP + pro_Gastro_E + pro_chirDigest +
      pro_uro + pro_urgence + pro_Sagefemmes + pro_PMA + pro_Autre,
    
    type_pro_consult_avant_diag = case_when(
      nb_pro_consultes == 0 ~ "Aucun",
      nb_pro_consultes == 1 & pro_gyn == 1 ~ "Seulement Gyn",
      nb_pro_consultes == 1 & pro_radio == 1 ~ "Seulement radio",
      nb_pro_consultes == 1 & pro_GP == 1 ~ "Seulement GP",
      nb_pro_consultes == 1 &
        pro_Gastro_E == 1 ~ "Seulement Gastro",
      nb_pro_consultes == 1 &
        pro_chirDigest == 1 ~ "Seulement Chirurgien Digestif",
      nb_pro_consultes == 1 & pro_uro == 1 ~ "Seulement Urologue",
      nb_pro_consultes == 1 &
        pro_urgence == 1 ~ "Seulement Urgence",
      nb_pro_consultes == 1 &
        pro_Sagefemmes == 1 ~ "Seulement Sage-femme",
      nb_pro_consultes == 1 & pro_PMA == 1 ~ "Seulement PMA",
      nb_pro_consultes == 1 & pro_Autre == 1 ~ "Seulement Autre",
      nb_pro_consultes > 1 ~ "Multiples professionnels",
      TRUE ~ NA_character_
    )
  )

#new_dataframe$type_pro_consult_avant_diag %>% table(useNA = "a") 

# 2. Merging Back to the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial%>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, type_pro_consult_avant_diag), 
            by = "id_unik")

# 3. Removing the new dataframe
rm(new_dataframe)

#---------------------------------
#  Type of information sources    (info_suspect_)
#---------------------------------
# 1.  Create a new dataframe with columns of interest
new_dataframe <- Endo_Adeno_initial %>% 
  dplyr::select(id_unik, starts_with("info_suspect_"))

# 2. Define information categories
categories <- c("Dans les médias", "Sur internet", "Sur les réseaux sociaux", "Autre")

# 3. Create a vector containing the names of the indicator variables
indic_names <- c("info_media", "info_internet", "info_sns", "info_autre")

# 4. Create indicator variables (0/1) for each type of information
for (i in seq_along(categories)) {
  new_dataframe[[indic_names[i]]] <- rowSums(sapply(
    new_dataframe[, 2:5],
    grepl,
    pattern = categories[i],
    ignore.case = TRUE
  ),
  na.rm = TRUE) > 0
}

# 5. Create a new categorical variable based on the number of sources used
new_dataframe <- new_dataframe %>%
  mutate(
    type_info_source = case_when(
      info_media    == 1 &
        rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_media",
      info_internet == 1 &
        rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_internet",
      info_sns      == 1 &
        rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_sns",
      info_autre    == 1 &
        rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_autre",
      rowSums(dplyr::select(., all_of(indic_names))) > 1 ~ "multiples_infos",
      TRUE ~ NA_character_
    )
  )

# 6. Merging Back to the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  left_join(new_dataframe %>% dplyr::select(id_unik, type_info_source),
            by = "id_unik")

# 7. Removing the new dataframe
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  dplyr::select(-starts_with("info_suspect_"))
rm(new_dataframe)

#---------------------------
# elements of diagnostics   (element_diag_)               
#---------------------------
# 1. Sélectionner les colonnes des éléments de diagnostic
new_dataframe <- Endo_Adeno_initial %>%
  dplyr::select(id_unik, starts_with("element_diag_"))

# 2. Define information categories
categories <-
  c("Autre",
    "Echographie pelvienne",
    "IRM",
    "Opération chirurgicale",
    "Symptômes")

# 3. Create a vector containing the names of the indicator variables
indic_names <-
  c("diag_autre",
    "diag_echo",
    "diag_irm",
    "diag_chirurgie",
    "diag_symptomes")

# 4. Create indicator variables (0/1) for each type of information
for (i in seq_along(categories)) {
  new_dataframe[[indic_names[i]]] <- rowSums(sapply(
    new_dataframe[, 2:6],
    grepl,
    pattern = categories[i],
    ignore.case = TRUE
  ),
  na.rm = TRUE) > 0
}

# 5. Create a new categorical variable
new_dataframe <- new_dataframe %>%
  mutate(
    type_element_diag = case_when(
      diag_echo== 1 & rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_echo",
      diag_irm == 1 & rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_IRM",
      diag_chirurgie == 1 & rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_chirurgie",
      diag_symptomes== 1 & rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_symptom",
      diag_autre == 1 & rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ "only_autre",
      rowSums(dplyr::select(., all_of(indic_names))) > 1 ~ "multiples_ele_diag",
      TRUE ~ NA_character_
    )
  )

# 6. Merge Back to the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  left_join(new_dataframe %>% dplyr::select(id_unik, type_element_diag),
            by = "id_unik")

# 7. Removing the new dataframe and the orginal variables 
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  dplyr::select(-starts_with("element_diag_"))
rm(new_dataframe)

#---------------------------------------------------------------------
# Evocation des maladies par un professionnel de santé avant le diagnostic (pro_parle_mal)
#---------------------------------------------------------------------
# 1. Sélectionner les colonnes des éléments de diagnostic
new_dataframe <- Endo_Adeno_initial %>% 
  dplyr::select(id_unik, starts_with("pro_parle_mal_"))

# 2. Define information categories
categories <-
  c(
    "Non",
    "Autre",
    "Chir_digestif",
    "Gastro_e",
    "Gyn",
    "medecin_PMA",
    "GP",
    "Radio",
    "Sage_femme",
    "Uro",
    "Urgences"
  )

# 3. Create a vector containing the names of the indicator variables
indic_names <-
  c(
    "pro_non",
    "pro_autre",
    "pro_chir_digestif",
    "pro_gastro",
    "pro_gyn",
    "pro_pma",
    "pro_gp",
    "pro_radio",
    "pro_sagefemme",
    "pro_uro",
    "pro_urgence"
  )

# 4. Create indicator variables (0/1) for each type of information
for (i in seq_along(categories)) {
  new_dataframe[[indic_names[i]]] <- rowSums(sapply(
    new_dataframe[, 2:ncol(new_dataframe)],
    grepl,
    pattern = categories[i],
    ignore.case = TRUE
  ),
  na.rm = TRUE) > 0
}

# 5. Create a new categorical variable
new_dataframe <- new_dataframe %>%
  mutate(
    type_pro_parle_mal = case_when(
      rowSums(dplyr::select(., all_of(indic_names))) == 1 ~ paste0("only_", indic_names[max.col(dplyr::select(., all_of(indic_names)), ties.method = "first")]),
      rowSums(dplyr::select(., all_of(indic_names))) > 1 ~ "multiples_pro",
      TRUE ~ NA_character_
    )
  )

# 6. Merge Back to the Original Dataset
Endo_Adeno_initial <- Endo_Adeno_initial %>%
  left_join(new_dataframe %>% dplyr::select(id_unik, type_pro_parle_mal),
            by = "id_unik")

# 7. Removing the new dataframe and the orginal variables 
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  dplyr::select(-starts_with("pro_parle_mal_"))
rm(new_dataframe)

#---------------------------------------------------------------------
# Evocation d’endométriose ou d’adénomyose par un professionnel de santé sans diagnostic (evok_maladie_ssdiag_1)
#---------------------------------------------------------------------
# strage...... but I ignore this information that I don't need for the analysis
#Endo_Adeno_initial$evok_maladie_ssdiag_1 %>% table


#-----------------------------------
#  Family history of endometriosis
#-----------------------------------
Endo_Adeno_initial <- Endo_Adeno_initial %>% 
  mutate(
    family_atcd_EndoAdeno_cat =
      case_when(
        family_atcd_EndoAdeno %in% c("Both", "Only adeno", "Only endo") ~
          "Antécédents",
        family_atcd_EndoAdeno == "No family history" ~ "Pas d'antécédents",
        TRUE~ NA_character_
      )
    )
```


# II. Final dataset

```{r}
Endo_Adeno_initial_final <- Endo_Adeno_initial %>%
  dplyr::select(
    id_unik,
    IdAction,
    
    family_atcd_EndoAdeno_cat,
    family_atcd_doul_pelv,

    age_symptome_cat,
    signe_symptome_premiere_consult,
    Age_premier_diag_endo_ade,
    Age_premier_diag_endo_ade_cat,
    
    delai_diag,
    delai_diag_cat,
    delai_diag_cat_simple,
    
    type_endo,
    ADENO,OMA,SPE,DIE,
    #Nb_specialites_consultees,
    Nb_specialites_consult_cat,
    type_pro_consult_avant_diag,
    type_info_source,
    type_element_diag,
    type_pro_parle_mal,
    
    Endo_Adeno_initial_participated)%>%
  distinct() %>% 
  
  # Add the new modality 'Not_answered'
  mutate(across(-c(id_unik, IdAction, delai_diag,Age_premier_diag_endo_ade),
                ~ case_when(
                  is.na(IdAction) ~ "Not_answered",
                  TRUE ~ .
    )
  )) %>% 
  dplyr::select(-IdAction)
```

# III. Save the final dataframe 

```{r}
# update : 2025-06-02
#write.csv(Endo_Adeno_initial_final,"Stage_M2_Complic_Gross/Endo_Adeno_initial_final.csv")
```

