---
title: "01_clean_Douleur_pelvienne_abdominale"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Douleur_pelvienne_abdominale
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```


# 0. Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
# Endo_douleur_pelvic_abd <- read.csv("~/Projet/0.Data/complication_pop_QST_0463_20250226_doul_pel_abdo.csv",encoding = "latin1",colClasses = c(Sexe = "character"))

Endo_douleur_pelvic_abd <-
  read.csv("Stage_M2_Complic_Gross/Complic_pop_QST_0463_20250407_doul_pel_abdo.csv",
           encoding = "latin1",
           colClasses = c(Sexe = "character"))

#---------------------------------
#  Modify bizarre name of column 
#---------------------------------
# Colname 29 index is "VAR31" : I checked it manually 
names(Endo_douleur_pelvic_abd)[29] <- "inter_sport_16"

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
Endo_douleur_pelvic_abd <- Endo_douleur_pelvic_abd %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- Endo_douleur_pelvic_abd %>% 
  sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count==nrow(Endo_douleur_pelvic_abd))%>% as.vector(); index_col_NA # index of columns having only NA

#--------------------------
#  Pain score in numeric
#--------------------------
Endo_douleur_pelvic_abd <- Endo_douleur_pelvic_abd %>% 
  mutate(across(starts_with(c("niv_severite",
                              "niv_intensite")), 
                ~ as.numeric(as.character(.))),
         across(starts_with("Date"),
                ~ as.Date(.,format="%d/%m/%Y")))
```


# I. Preprocessing Data

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
Endo_douleur_pelvic_abd <- Endo_douleur_pelvic_abd %>% 
  mutate(Endo_dysmenorrhea_abd_participated=ifelse(is.na(IdAction),
                                                "Non","Oui"))

#-------------------------------------------------
# Mean dysmernorrhea score along the entire life
#-------------------------------------------------
library(dplyr)
# 1. Create a temporary dataframe 
# new_dataframe <- Endo_douleur_pelvic_abd %>%
#   dplyr::select(id_unik, starts_with("niv_severite_doul_pelv_")) %>%
#   dplyr::mutate(
#     total_score = rowSums(select(
#       ., starts_with("niv_severite_doul_pelv_")
#     ), na.rm = TRUE),
#     na_count = 5 - rowSums(is.na(select(
#       ., starts_with("niv_severite_doul_pelv_")
#     ))),
#     mean_niv_severite_dysmernorrhea = total_score / na_count
#   ) %>%
#   dplyr::select(id_unik, mean_niv_severite_dysmernorrhea) 

new_dataframe <- Endo_douleur_pelvic_abd %>%
  dplyr::select(id_unik, starts_with("niv_severite_doul_pelv_")) %>%
  dplyr::mutate(
    total_score = rowSums(across(starts_with("niv_severite_doul_pelv_")), na.rm = TRUE),
    na_count = rowSums(!is.na(across(starts_with("niv_severite_doul_pelv_")))),
    mean_niv_severite_dysmernorrhea = total_score / na_count
  ) %>%
  dplyr::select(id_unik, mean_niv_severite_dysmernorrhea)


# 2. Add a new categorical variables 
quantiles <- new_dataframe$mean_niv_severite_dysmernorrhea %>% quantile(probs=c(1/3,2/3), na.rm=TRUE) %>% round()

new_dataframe <- new_dataframe %>%
  mutate(
    mean_niv_severite_dysmernorrhea_cat =
      case_when(
        mean_niv_severite_dysmernorrhea < quantiles[[1]] ~
          paste0("<", quantiles[1]),
        mean_niv_severite_dysmernorrhea >= quantiles[[1]] &
          mean_niv_severite_dysmernorrhea < quantiles[[2]] ~
          paste0(quantiles[1], "-", quantiles[2]),
        mean_niv_severite_dysmernorrhea >= quantiles[[2]] ~
          paste0(">=", quantiles[2])
      )
  )

# 3. Add a new variable to the main dataframe
Endo_douleur_pelvic_abd <- Endo_douleur_pelvic_abd %>% 
  left_join(new_dataframe, by="id_unik") 

# 4. Modify the new variable by adding a 'No pain' modality 
Endo_douleur_pelvic_abd <- Endo_douleur_pelvic_abd %>% 
  mutate(
    mean_niv_severite_dysmernorrhea_cat= 
      case_when(
        doul_pelvienne_regle=="Non" ~ "No pain",
        TRUE ~ mean_niv_severite_dysmernorrhea_cat
      )
    )

# 5. Delete the temporary dataframe and quantile information
rm(new_dataframe);rm(quantiles)

#-------------------------------------------------
# Mean of the worst severity level of abdominal pain over the entire lifespan
#-------------------------------------------------
# 1. Create a temporary dataframe 
new_dataframe <- Endo_douleur_pelvic_abd %>%
  select(id_unik, starts_with("niv_intensite_doul_abdo_")) %>%
  mutate(
    total_score = rowSums(select(
      ., starts_with("niv_intensite_doul_abdo_")
    ), na.rm = TRUE),
    na_count = 5 - rowSums(is.na(select(
      ., starts_with("niv_intensite_doul_abdo_")
    ))),
    mean_niv_intensite_doul_abdo = total_score / na_count
  ) %>%
  select(id_unik, mean_niv_intensite_doul_abdo) 

# 2. Add a new categorical variables 
quantiles <- new_dataframe$mean_niv_intensite_doul_abdo %>% quantile(probs=c(1/3,2/3), na.rm=TRUE) %>% round()

new_dataframe <- new_dataframe %>%
  mutate(
    mean_niv_intensite_doul_abdo_cat =
      case_when(
        mean_niv_intensite_doul_abdo < quantiles[[1]] ~
          paste0("<", quantiles[1]),
        mean_niv_intensite_doul_abdo >= quantiles[[1]] &
          mean_niv_intensite_doul_abdo < quantiles[[2]] ~
          paste0(quantiles[1], "-", quantiles[2]),
        mean_niv_intensite_doul_abdo >= quantiles[[2]] ~
          paste0(">=", quantiles[2])
      )
  )

# 3. Add a new variable to the main dataframe
Endo_douleur_pelvic_abd <- Endo_douleur_pelvic_abd %>% 
  left_join(new_dataframe, by="id_unik")

# 4. Modify the new variable by adding a 'No pain' modality 
Endo_douleur_pelvic_abd <- Endo_douleur_pelvic_abd %>% 
  mutate(
    mean_niv_intensite_doul_abdo_cat= 
      case_when(
        douleur_abdo=="Non" ~ "No pain",
        TRUE ~ mean_niv_intensite_doul_abdo_cat
      )
    )

# 5. Delete the temporary dataframe 
rm(new_dataframe); rm(quantiles)
```

# II. Final dataset
## 1) Selection of variables
```{r}
Endo_douleur_pelvic_abd_final <-
  Endo_douleur_pelvic_abd %>%
  dplyr::select(
    id_unik,
    IdAction,
    doul_pelvienne_regle,
    regle_avant_15a,
    periode2:periode5,
    douleur_abdo,
    doul_abdo_15a,
    periode2_abdo:periode5_abdo,
    
    starts_with(c("niv_severite","niv_intensite")),
    mean_niv_severite_dysmernorrhea_cat,
    mean_niv_intensite_doul_abdo_cat,
    DateValidation,
    Endo_dysmenorrhea_abd_participated
  ) %>% 
  distinct() %>% 
  mutate(across(
    - c(id_unik, IdAction, DateValidation, starts_with("niv_")),
    ~ case_when(is.na(IdAction) ~ "Not_answered", TRUE  ~ .)
    )) %>%
  dplyr::select(-IdAction)
```

## 2) Create a new variable
```{r}
#------------------------------------------------------------------
# Experience of abdominal and dysmenorrhea before 1st pregnancy
#------------------------------------------------------------------
# 1. Create a temporary dataframe with certaine information of pregnancy module
new_dataframe <- Endo_douleur_pelvic_abd_final %>% 
  left_join(grossesse_mod_final %>% 
              select(id_unik, merged_DateDebut,
                     first_preg_age,grossesse_encours,
                     Date_naissance),
            by="id_unik")


# 2. Create two new variables : 
# - Experience of abdominal pain before first pregnancy
# - Experience of dysmenorrhea before first pregnancy
new_dataframe <- new_dataframe %>%
  mutate(
    first_preg_age = first_preg_age %>% round(),
    age_inclusion = as.numeric(as.character((
      DateValidation - Date_naissance
    ) / 365)) %>%
      round(),
    #-----------------------------------------------------
    # Experience of abdominal pain before first pregnancy
    #-----------------------------------------------------
    doul_abdo_avant_1_preg = case_when(
      # No answer provided in the questionnaire
      douleur_abdo == "Not_answered" ~ "Not_answered",
      
      # First pregnancy occurred after study inclusion
      first_preg_age > age_inclusion &
        douleur_abdo == "Oui" ~ "Oui",
      first_preg_age > age_inclusion &
        douleur_abdo == "Non" ~ "Non et nsp",
      
      # First pregnancy occurred before or at inclusion and no pain reported
      first_preg_age <= age_inclusion &
        douleur_abdo == "Non" ~ "Non et nsp",
      
      # First pregnancy before or at inclusion and before age 15
      first_preg_age <= age_inclusion &
        first_preg_age <= 15 ~ case_when(
          doul_abdo_15a == "Oui" ~ "Oui",
          doul_abdo_15a == "Non" ~ "Non et nsp",
          doul_abdo_15a == "Je ne m'en souviens plus" ~ "Non et nsp"
        ),
      
      # First pregnancy between 16–20 years old
      between(first_preg_age, 16, 20) ~ case_when(
        first_preg_age <= age_inclusion &
          (doul_abdo_15a == "Oui" | periode2_abdo == "Oui") ~ "Oui",
        first_preg_age < age_inclusion &
          doul_abdo_15a == "Non" & periode2_abdo == "Non" ~ "Non et nsp",
        first_preg_age <= age_inclusion &
          (
            doul_abdo_15a == "Je ne m'en souviens plus" |
              periode2_abdo == "Je ne m'en souviens plus"
          ) ~ "Non et nsp"
      ),
      
      # First pregnancy between 21–30 years old
      between(first_preg_age, 21, 30) ~ case_when(
        first_preg_age <= age_inclusion &
          (
            doul_abdo_15a == "Oui" |
              periode2_abdo == "Oui" | periode3_abdo == "Oui"
          ) ~ "Oui",
        first_preg_age <= age_inclusion &
          doul_abdo_15a == "Non" &
          periode2_abdo == "Non" & periode3_abdo == "Non" ~ "Non et nsp",
        first_preg_age <= age_inclusion &
          (
            doul_abdo_15a == "Je ne m'en souviens plus" |
              periode2_abdo == "Je ne m'en souviens plus" |
              periode3_abdo == "Je ne m'en souviens plus"
          ) ~ "Non et nsp"
      ),
      
      # First pregnancy between 31–40 years old
      between(first_preg_age, 31, 40) ~ case_when(
        first_preg_age <= age_inclusion &
          (
            doul_abdo_15a == "Oui" | periode2_abdo == "Oui" |
              periode3_abdo == "Oui" | periode4_abdo == "Oui"
          ) ~ "Oui",
        first_preg_age <= age_inclusion &
          doul_abdo_15a == "Non" & periode2_abdo == "Non" &
          periode3_abdo == "Non" & periode4_abdo == "Non" ~ "Non et nsp",
        first_preg_age <= age_inclusion &
          (
            doul_abdo_15a == "Je ne m'en souviens plus" |
              periode2_abdo == "Je ne m'en souviens plus" |
              periode3_abdo == "Je ne m'en souviens plus" |
              periode4_abdo == "Je ne m'en souviens plus"
          ) ~ "Non et nsp"
      ),
      
      # First pregnancy at or after age 41
      first_preg_age >= 41 ~ case_when(
        first_preg_age <= age_inclusion &
          (
            doul_abdo_15a == "Oui" | periode2_abdo == "Oui" |
              periode3_abdo == "Oui" | periode4_abdo == "Oui" |
              periode5_abdo == "Oui"
          ) ~ "Oui",
        first_preg_age <= age_inclusion &
          doul_abdo_15a == "Non" & periode2_abdo == "Non" &
          periode3_abdo == "Non" & periode4_abdo == "Non" &
          periode5_abdo == "Non" ~ "Non et nsp",
        first_preg_age <= age_inclusion &
          (
            doul_abdo_15a == "Je ne m'en souviens plus" |
              periode2_abdo == "Je ne m'en souviens plus" |
              periode3_abdo == "Je ne m'en souviens plus" |
              periode4_abdo == "Je ne m'en souviens plus" |
              periode5_abdo == "Je ne m'en souviens plus"
          ) ~ "Non et nsp"
      ),
      
      TRUE ~ NA_character_
    ),
    #------------------------------------------------------
    #  Experience of dysmenorrhea before first pregnancy
    #------------------------------------------------------
    dysmenorrhea_avant_1_preg = case_when(
      # No answer provided
      doul_pelvienne_regle == "Not_answered" ~ "Not_answered",
      
      # First pregnancy occurred after study inclusion
      first_preg_age > age_inclusion &
        doul_pelvienne_regle == "Oui" ~ "Oui",
      first_preg_age > age_inclusion &
        doul_pelvienne_regle == "Non" ~ "Non et nsp",
      
      # First pregnancy occurred before or at inclusion and no pain reported
      first_preg_age <= age_inclusion &
        doul_pelvienne_regle == "Non" ~ "Non et nsp",
      
      # First pregnancy before age 15
      first_preg_age <= age_inclusion &
        first_preg_age <= 15 ~ case_when(
          regle_avant_15a == "Oui" ~ "Oui",
          regle_avant_15a == "Non" ~ "Non et nsp",
          regle_avant_15a == "Je ne m'en souviens plus" ~ "Non et nsp"
        ),
      
      # First pregnancy between 16–20
      between(first_preg_age, 16, 20) ~ case_when(
        first_preg_age <= age_inclusion &
          (regle_avant_15a == "Oui" | periode2 == "Oui") ~ "Oui",
        first_preg_age < age_inclusion &
          regle_avant_15a == "Non" & periode2 == "Non" ~ "Non et nsp",
        first_preg_age <= age_inclusion &
          (
            regle_avant_15a == "Je ne m'en souviens plus" |
              periode2 == "Je ne m'en souviens plus"
          ) ~ "nsp"
      ),
      
      # First pregnancy between 21–30
      between(first_preg_age, 21, 30) ~ case_when(
        first_preg_age <= age_inclusion &
          (regle_avant_15a == "Oui" |
             periode2 == "Oui" | periode3 == "Oui") ~ "Oui",
        first_preg_age <= age_inclusion &
          regle_avant_15a == "Non" &
          periode2 == "Non" & periode3 == "Non" ~ "Non et nsp",
        first_preg_age <= age_inclusion &
          (
            regle_avant_15a == "Je ne m'en souviens plus" |
              periode2 == "Je ne m'en souviens plus" |
              periode3 == "Je ne m'en souviens plus"
          ) ~ "Non et nsp"
      ),
      
      # First pregnancy between 31–40
      between(first_preg_age, 31, 40) ~ case_when(
        first_preg_age <= age_inclusion &
          (
            regle_avant_15a == "Oui" | periode2 == "Oui" |
              periode3 == "Oui" | periode4 == "Oui"
          ) ~ "Oui",
        first_preg_age <= age_inclusion &
          regle_avant_15a == "Non" & periode2 == "Non" &
          periode3 == "Non" & periode4 == "Non" ~ "Non",
        first_preg_age <= age_inclusion &
          (
            regle_avant_15a == "Je ne m'en souviens plus" |
              periode2 == "Je ne m'en souviens plus" |
              periode3 == "Je ne m'en souviens plus" |
              periode4 == "Je ne m'en souviens plus"
          ) ~ "Non et nsp"
      ),
      
      # First pregnancy at or after 41
      first_preg_age >= 41 ~ case_when(
        first_preg_age <= age_inclusion &
          (
            regle_avant_15a == "Oui" | periode2 == "Oui" |
              periode3 == "Oui" | periode4 == "Oui" |
              periode5 == "Oui"
          ) ~ "Oui",
        first_preg_age <= age_inclusion &
          regle_avant_15a == "Non" & periode2 == "Non" &
          periode3 == "Non" & periode4 == "Non" &
          periode5 == "Non" ~ "Non et nsp",
        first_preg_age <= age_inclusion &
          (
            regle_avant_15a == "Je ne m'en souviens plus" |
              periode2 == "Je ne m'en souviens plus" |
              periode3 == "Je ne m'en souviens plus" |
              periode4 == "Je ne m'en souviens plus" |
              periode5 == "Je ne m'en souviens plus"
          ) ~ "Non et nsp"
      ),
      
      TRUE ~ NA_character_
    )
  )


##################
# TEST
new_dataframe <- new_dataframe %>% 
  mutate(
  first_preg_age_cat =
    case_when(
      first_preg_age < 16 ~ "<16",
      first_preg_age >=16 & first_preg_age <21 ~"16-20",
      first_preg_age >=21 & first_preg_age <31 ~ "21-30",
      first_preg_age >=31 & first_preg_age <41 ~ "31-40",
      first_preg_age >=41 ~ "41+",
      TRUE ~ NA_character_
    ),
  
  
  
  )
###################




# 3. Add the new variables to the main dataframe
Endo_douleur_pelvic_abd_final <- 
  Endo_douleur_pelvic_abd_final %>%
  left_join(
    new_dataframe %>%
      select(
        id_unik,
        doul_abdo_avant_1_preg,
        dysmenorrhea_avant_1_preg
      ),
    by = "id_unik"
  )

# 4. Delete unnecessary variables
Endo_douleur_pelvic_abd_final <- 
  Endo_douleur_pelvic_abd_final %>%
  select(-starts_with("periode"),
         -DateValidation,
         -regle_avant_15a,
         -doul_abdo_15a)

# 5. Delete the temporary dataframe 
rm(new_dataframe)
```

```{r}
Endo_douleur_pelvic_abd_final%>%glimpse
```



# III. Save the final dataframe 

```{r}
#write.csv(Endo_douleur_pelvic_abd_final,"Stage_M2_Complic_Gross/Endo_douleur_pelvic_abd_final.csv")
```

