---
title: "01_clean_comorbidités"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_comorbidités
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data for the comorbidity information 
*Date of creation* : 2025-05-13
*Date of update* : 2025-06-02
*Author* : Minji NAM, stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import raw data : pop_demo_inclusion 
- nb_maladie, IdDisease, DiseaseNotFound

```{r}
#-----------------------
# Import a main dataset
#-----------------------
# pop_comorbidite <-
#   read.csv("~/Projet/0.Data/complication_pop_comorbidite_20250226.csv",fileEncoding = "latin1",colClasses = c(Sexe = "character"))

pop_comorbidite <-
  read.csv("Complic_pop_comorbidite_20250407.csv",
           fileEncoding = "latin1",
           colClasses = c(Sexe = "character"))

#------------------------------------------------
# Replace cells with nothing or spaces with NA
#------------------------------------------------
pop_comorbidite <-
  pop_comorbidite %>% mutate(across(everything(), ~ na_if(trimws(.), ""))) %>% 
  mutate_all(~str_trim(.))

#---------------------
#  char -> numeric
#---------------------
pop_comorbidite$nb_maladie <-
  as.numeric(pop_comorbidite$nb_maladie)

#-----------------------------------------------------
# Replace non-standard apostrophes with standard ones 
#-----------------------------------------------------
pop_comorbidite <- pop_comorbidite %>%
    mutate(
    across(where(is.character), ~ gsub("\u009c", "œ", .)),
    across(where(is.character), ~ gsub("\u0092", "'", .)),
    across(where(is.character), ~ gsub("\u0085", "", .))
  )
# \u0092 : a non-standard apostrophe, \u0085: an unusual line break character, "\u009c" : control character(String Terminator(ST))
```

# I. Preprocessing Data
## 1) IdDisease
```{r}
# List of comobidities
list_com <- pop_comorbidite$IdDisease %>% unique

# Create different categories of comorbidities
# Endométriose/Adénomyose
endometriose_adenomyose <- c("Endométriose", "Adénomyose")

# Gynécologique
gynecologique <- c(
  "Fibrome utérin",
  "Kyste ovarien, syndrome des ovaires polykystiques (SOPK)",
  "Trouble de la fertilité"
)

# Endocrino-métabolique (Hormonal and metabolic disorders)
endocrine_metabolic <- c(
  "hypercholestérolémie / hypertriglycéridémie",
  "Hyperprolactinémie",
  "Autre maladie endocrinienne",
  "Syndrome de Cushing",
  "Hypoparathyroidie chronique",
  "Diabète insipide",
  "Hyperaldostéronisme primaire",
  "Maladie de Basedow",
  "Autre Hyperthyroïdie",
  "Maladie d'Addison (insuffisance surrénalienne)",
  "Adénome hypophysaire",
  "Maladie de Gilbert",
  "Hypothyroïdie",
  "Diabète",
  "Obésité (Indice de Masse Corporelle>30)",
  "Syndrome de Turner",
  "Anémie de Biermer (anémie mégaloblastique, anémie pernicieuse)",
  "Déficit en facteur de coagulation/troubles de la coagulation"
)

# Neuro_pain (Nervous system disorder and chronic pain conditions)
neuro_pain <- c(
  "Trouble du déficit de l'attention/hyperactivité (TDAH)",
  "Fibromyalgie",
  "Vertiges",
  "Algie vasculaire de la face",
  "Antécédent d'accident vasculaire cérébral (AVC)",
  "Autre maladie neurologique",
  "Syndrome des jambes sans repos",
  "Syndrome douloureux régional complexe (SDRC 1 ou algodystrophie et SDCR 2 ou algoneurodystrophie)",
  "Céphalées chroniques (dont migraine)",
  "Autres syndromes parkinsoniens",
  "Epilepsie",
  "Autres démences",
  "Narcolepsie / Maladie de Gelineau",
  "Algies du trijumeau (névralgies essentielles, douleurs neuropathiques du nerf trijumeau)",
  "Syndrome de la queue de cheval",
  "Syringomyélie",
  "Neuropathie périphérique",
  "Autisme",
  "Malformation d'Arnold-Chiari",
  "Trouble stomatologique ou maxillofacial (SADAM)",
  "Kystes de Tarlov, kyste péri-radiculaire, kyste méningé sacré",
  "Maladie de Charcot-Marie-Tooth",
  "Névralgie d'Arnold"
)

# Cardio_circulatory (Heart and blood vessel disorders)
cardio_circulatory <- c(
  "Hypertension artérielle",
  "Troubles du rythme cardiaque (y compris Maladie de Bouveret)",
  "Antécédent d'accident ischémique transitoire (AIT)",
  "Antécédent d'infarctus du myocarde",
  "Insuffisance veineuse chronique",
  "Artérite/sténose carotidienne/artériopathie oblitérante des membres inférieurs",
  "Hypertension artérielle pulmonaire",
  "Antécédent de phlébite ou d'embolie pulmonaire",
  "Arythmie cardiaque par fibrillation auriculaire : ACFA",
  "Angor/angine de poitrine/coronaropathie/maladie coronarienne (sans infarctus du myocarde)",
  "Insuffisance cardiaque",
  "Autre maladie cardiaque ou vasculaire",
  "Valvulopathie, Remplacement valvulaire (y compris Maladie de Barlow)",
  "Trouble de conduction (Pace Maker)",
  "Cardiopathie congénitale",
  "Endocardite",
  "Syndrome de Wolff-Parkinson-White"
)

# Respiratory (Lung and airway disorders)
respiratory <- c(
  "Asthme",
  "Sinusite chronique",
  "Polypose naso-sinusiennes",
  "Bronchite chronique, broncho pneumopathie chronique obstructive (BPCO), emphysème",
  "Rhinite chronique",
  "Dilatation des bronches",
  "Apnée du sommeil"
)

# Uro_nephro (Renal and urinary disorders)
uro_nephro <- c(
  "Lithiase urinaire chronique (calculs rénaux)",
  "Incontinence urinaire",
  "Autre maladie rénale, urinaire ou génitale",
  "Cystites redicivantes",
  "Insuffisance rénale chronique",
  "Polykystose rénale",
  "Maladie de Berger",
  "Maladie de Cacchi et Ricci"
)

# Gastrointestinal (Digestive system disorders)
gastrointestinal <-
  c(
    "Reflux gastro oesophagien, achalasie, hernie hiatale",
    "Colon irritable / Colopathie fonctionnelle",
    "Ulcère gastro duodénal, gastrite chronique",
    "Maladie de Crohn",
    "Rectocolite hémorragique (RCH)",
    "Maladie coeliaque",
    "Diverticules",
    "Autre maladie du système digestif",
    "Stéatose hépatique (NASH)",
    "Hépatite/Cirrhose auto-immune",
    "Insuffisance pancréatique (pancréatite chronique)",
    "Hépatite B",
    "Hépatite C",
    "Autre maladie du foie"
  )

# Musculoskeletal (Bone, Joint, and muscle disorders)
musculoskeletal <- c(
  "Scoliose",
  "Autre maladie rhumatologique",
  "Ostéoporose / Ostéopénie",
  "Arthrose",
  "Tendinites chroniques (périarthrite, épicondylite, capsulite...)",
  "Polyarthrite rhumatoïde",
  "Rhumatisme psoriasique",
  "Arthropathies microcristallines (goutte, chondrocalcinose)",
  "Infection ostéo-articulaire (arthrite/spondylodiscite)",
  "Épine Calcanéenne",
  "Syndromes canalaires (Syndrome du canal carpien, Syndrome de la loge de Guyon)",
  "Maladie de Dupuytren",
  "Névrome de Morton",
  "Lombalgie chronique (douleur du bas du dos, hernie discale lombaire, sciatique)",
  "Cervicalgie chronique, dorsalgie chronique (douleur du haut du dos, hernie discale cervicale, hernie discale dorsale)",
  "Epine Calcanéenne"
)

# Autoimmune_inflammatory (Inflammatory Disorders)
autoimmune_inflammatory <- c(
  "Maladie de Lyme",
  "Maladie de Raynaud",
  "Syndrome de Goujerot-Sjögren",
  "Allergie",
  "Thyroïdite d'Hashimoto",
  "Lupus",
  "Mastocytose, syndrome d'activation mastocytaire (SAMA)",
  "Déficit immunitaire (hors VIH)",
  "Anémie hémolytique auto-immune",
  "Purpura thrombopénique immunologique",
  "Syndrome des antiphospholipides",
  "Maladie de Behçet",
  "Vascularite cryoglobulinémique",
  "Maladie de Still de l'adulte",
  "Myélome (gammapathie monoclonale, MGUS, plasmocytome, maladie de Waldenström)",
  "Autre maladie systémique",
  "Syndrome de Cogan",
  "Syndrome de fatigue chronique (Encéphalomyélite myalgique)",
  "Spondyloarthrite axiale (anciennement spondylarthrite ankylosante)",
  "Sarcoïdose",
  "Lichen plan"
)

# Mental_health (Mental and behavioral disorders)
mental_health <-
  c(
    "Troubles obsessionnels compulsifs :  TOC",
    "Troubles anxieux",
    "Dépression",
    "Trouble du comportement alimentaire (anorexie, boulimie)",
    "Addictions (alcool, jeux, autres drogues...)",
    "Syndrome de stress post traumatique",
    "Syndrome de fatigue chronique",
    "Troubles du sommeil",
    "Trouble de la personnalité",
    "Psychose chronique (schizophrénie...)",
    "Autre maladie psychiatrique",
    "Trouble bipolaire, maladie maniaco-dépressive"
  )

# Dermatological (Skin disorders)
dermatological <- c(
  "Dyshidrose",
  "Vitiligo",
  "Dyshydrose",
  "Dermatose bulleuse chronique",
  "Zona",
  "Eczéma (dermatite atopique)",
  "Psoriasis",
  "Autre maladie de la peau",
  "Dermatite séborrhéique",
  "Urticaire",
  "Rosacée",
  "Pelade",
  "Acné",
  "Maladie de Verneuil / hidradénite suppurée"
)

# Ophtalmo (Ophthalmologic disorders)
ophtalmo <-
  c(
    "Trouble de la vision (myopie, presbytie, astigmatisme)",
    "Cataracte",
    "Dégénérescence maculaire (DMLA)",
    "Rétinopathie",
    "Glaucome chronique",
    "Autre maladie des yeux",
    "Décollement de rétine ou décollement du vitré"
  )

# ORL_auditif (ORL and auditory disorders)
ORL_auditif <- c("Troubles de l'audition / surdité/ acouphènes",
                 "Rhinite chronique",
                 "Otospongiose",
                 "Vertiges (Maladie de Menière, vertige paroxystique positionnel bénin)",
                 "Autre maladie ORL ou respiratoire")

# Hemato (Hematologic disorders)
hemato <- c(
  "Leucémie lymphoïde chronique",
  "Autres hémolyses constitutionnelles (thalassémies, sphérocytose)",
  "Drépanocytose",
  "Hémophilie",
  "Maladie de Willebrand",
  "Leucémie myéloïde chronique",
  "Autre maladie du sang",
  "Syndrome myéloprolifératif (polyglobulie, maladie de Vaquez, thrombocytémie essentielle, myélofibrose primitive)"
)

# Oncological (Oncologic disorders)
oncological <- c(
  "Cancer du col de l'utérus",
  "Cancer du rein",
  "Cancer du sein",
  "Cancer de la peau de type Carcinome baso-cellulaire",
  "Mélanome",
  "Autre cancer",
  "Cancer du rectum",
  "Cancer du rein",
  "Polypes coliques",
  "Cancer de la thyroïde"
)

# Infectious (Infectious Disorders)
infectious <- c(
  "Covid-19 (AVEC symptômes persistants)",
  "Covid-19 aigue (SANS symptôme persistant)",
  "Autre maladie infectieuse chronique",
  "Herpes",
  "Tuberculose"
)

# Other
other <- c(
  "Sclérose en plaque",
  "NA",
  "Spina bifida",
  "Hypersensibilité chimique multiple - MCS",
  "Lymphoedème",
  "Meningiome",
  "Sclérodermie",
  "Myasthénie",
  "Syndrome de Marfan",
  "Neurofibromatose de type 1 (NF1) ou maladie de Von Recklinghausen",
  "Hémochromatose",
  "Myopathie génétique",
  "Malformations artérioveineuses cérébrales",
    "Syndrome d'Ehlers Danlos"
)

# Create new dataframe from the created list of categories
all_comorbidities <-
  c(
    endometriose_adenomyose,
    gynecologique,
    endocrine_metabolic,
    neuro_pain,
    cardio_circulatory,
    respiratory,
    uro_nephro,
    gastrointestinal,
    musculoskeletal,
    autoimmune_inflammatory,
    mental_health,
    dermatological,
    ophtalmo,
    ORL_auditif,
    hemato,
    oncological,
    infectious,
    other
  )
df_comorbidites <- data.frame(comorbidite = unique(all_comorbidities))

# Create new vide variable with NA
df_comorbidites$categorie <- NA

# Fill the new variable (categorie) 
df_comorbidites$categorie[df_comorbidites$comorbidite %in% endometriose_adenomyose] <-
  "Endométriose/Adénomyose"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% gynecologique] <-
  "Gynécologique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% endocrine_metabolic] <-
  "Endocrino-métabolique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% neuro_pain] <-
  "Neurologique et Douleurs chroniques"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% cardio_circulatory] <-
  "Cardiovasculaire"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% respiratory] <-
  "Respiratoire"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% uro_nephro] <-
  "Uro-néphrologique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% gastrointestinal] <-
  "Gastro-intestinal"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% musculoskeletal] <-
  "Musculo-squelettique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% autoimmune_inflammatory] <-
  "Inflammatoire"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% mental_health] <-
  "Santé mentale"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% dermatological] <-
  "Dermatologique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% ophtalmo] <-
  "Ophtalmologique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% ORL_auditif] <-
  "ORL et Auditif"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% hemato] <-
  "Hématologique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% oncological] <-
  "Oncologique"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% infectious] <-
  "Infectieuse"
df_comorbidites$categorie[df_comorbidites$comorbidite %in% other] <-
  "Autre"

names(df_comorbidites)[1] <- "IdDisease"

pop_comorbidite <- pop_comorbidite %>% 
  left_join(df_comorbidites, by = "IdDisease")

pop_comorbidite <- pop_comorbidite %>%
  mutate(categorie = case_when(
    is.na(categorie)&str_detect(IdDisease, "^hypercholes") ~ "Endocrino-métabolique",  
    is.na(categorie)&str_detect(IdDisease, "^Troubles obsession") ~ "Santé mentale",
    TRUE ~ categorie  
  ))
```


## 2) DiseaseNotFound 
: 160 maladies différentes it's doable 

```{r}
list_dis <- pop_comorbidite$DiseaseNotFound %>% 
  tolower() %>% 
  unique() %>% 
  as.vector()

#1. Make some lists of diseases in terms of categories of diseases
# Endométriose/Adénomyose
endometriose_adenomyose <-
  c(
    "endométriose digestive",
    "endométriose broncho-pulmonaire",
    "endométriose rectale",
    "adénomyose",
    "endometrite chronique",
    "diagnostiquée sous une forme d'endométriose de 2014 à 2019 (mauvais diagnostic)",
    "l'endometriose",
    "endometriose"
  )

# Gynecologique
gynecologique <-
  c(
    "infertilité",
    "syndrome douloureux pelvien chronique",
    "insuffisance ovarienne",
    "insuffisance ovarienne précoce",
    "insuffisance ovarienne précose",
    "syndrome de congestion pelvienne",
    "syndrome vésical douloureux",
    "trouble dysphorique prémenstruel",
    "endosalpyngiose",
    "incompetence cervicale",
    "hypersensibilité pelvienne",
    "cogestions pelviennes - varices utérines",
    "prolapsus",
    "syndrome de congestion pelvienne / varices pelviennes",
    "hypersensibilité pelvienne et abdominale"
  )

# Neuro_pain (Nervous system disorder and chronic pain conditions)
neuro_pain <-
  c(
    "neuropathie",
    "douleur neuropathique séquellaire",
    "neuropathie à petites fibres",
    "neuropathie des petites fibres",
    "persistent postural-perceptual dizziness",
    "névrite vestibulaire",
    "dysphonie spasmodique",
    "migraine ophtalmique",
    "spasmophilie",
    "tsa",
    "htic idiopathique",
    "blépharospasme",
    "hypertension intracranienne idiopathique.",
    "hypertension intracrânienne idiopathique",
    "névralgie cervico brachiale",
    "névralgie pudendale",
    "névralgie inguinale",
    "douleurs neuropathiques et névralgies multiples",
    "leucopathie vasculaire fazekas 1",
    "névralgie cervico-bracciale",
    "neuropathie membres inférieurs",
    "leucopathie vasculaire",
    "syndrome de sensibilité centrale"
    
  )

# Cardio_circulatory (Heart and blood vessel disorders)
cardio_circulatory <-
  c(
    "arrêt cardiaque sans cause déterminée",
    "hypertension intracrânienne idiopathique",
    "varices pelviennes bilatérales",
    "varices pelviennes",
    "varices orbitaires",
    "syndrome des anticorps anti phospholipides",
    "mesotheliome peritoneal",
    "varices pelviennes bilatérales",
    "péricardite",
    "pericardite",
    "myocardite",
    "trouble de la conduction",
    "syndrome de tachycardie orthostatique posturale",
    "tachycardie sinusale inappropriée"
  )

# Respiratory (Lung and airway disorders)
respiratory <-
  c("trouble ventilatoire restrictif",
    "hypoxie sévère diurne et nocturne")

# Autoimmune_inflammatory (Inflammatory Disorders)
autoimmune_inflammatory <-
  c(
    "maladie auto immune inflammatoire non étiquetée",
    "lupus anticoagulant circulant",
    "maladie auto immune d hashimoto",
    "maladie auto immune d hashimoto"
  )

# Uro_nephro (Renal and urinary disorders)
uro_nephro <- c(
  "rétention urinaire",
  "hyperactivité de la vessie",
  "rein atrophié sans insuffisance rénale",
  "rétention urinaire chronique",
  "cystite interstitielle"
)

# Gastrointestinal (Digestive system disorders)
gastrointestinal <-
  c(
    "syndrome du côlon irritable",
    "syndrôme de l'intestin irritable",
    "colite lymphocytaire",
    "intolérance alimentaire",
    "œsophagite à éosinophile",
    "gastrite chronique",
    "iléite chronique",
    "malabsorption",
    "dyskinésie billiare",
    "hyperplasies nodulaires focales",
    "intolérance alimentaire",
    "syndrome de l'intestin irritable",
    "malformation anale",
    "constipation",
    "intolérance au lactose",
    "panniculite mésentérique",
    "œsophage en marteau piqueur",
    "maladie hémorroïdaire"
  )

# Musculoskeletal (Bone, Joint, and muscle disorders)
musculoskeletal <-
  c(
    "discopathie cervicales",
    "discopathie cervical",
    "tendino bursite du supra épineux",
    "hernie discale l5-s1",
    "syndrôme rotulien",
    "chondropathie fémoro patellaire",
    "enthésopathie calcifiante",
    "dysplasie congénitale des hanches",
    "synovite villo nodulaire",
    "arthrose (légère) des vertèbres dorsales",
    "bursite",
    "bruxisme",
    "trouble de la colone verterbrale",
    "dysplasie fibreuse du clivus",
    "periartrites calcifiantes épaules",
    "maladie de hoffa",
    "fibromyalgie",
    "maladie de la graisse de hoffa (genou)",
    "sacroilite",
    "syndrome du défilé thoraco-brachial",
    "polyrhumatisme inflammatoire"
  )

# Mental_health (Mental and behavioral disorders)
mental_health <-
  c(
    "trouble anxieux-dépressif",
    "depression post partum",
    "burn out",
    "trouble du comportement alimentaire",
    "tag",
    "hypersomnie"
  )

# Dermatological (Skin disorders)
dermatological <-
  c(
    "dermatite atopique",
    "lichen scléro-atrophique",
    "dermographisme",
    "morphée",
    "syndrôme de sweet"
  )


# Endocrino-métabolique (Hormonal and metabolic disorders)
endocrine_metabolic <-
  c(
    "hyperinsulinisme",
    "hyperthyroïdie",
    "hyperthyroïdie par autonomisation thyroïdienne",
    "hypothyroïdie",
    "insuffisance hypophysaire",
    "cortisol bas",
    "lipoedème",
    "lipœdème",
    "lipodoeme",
    "éctopie thyroïdienne",
    "diabète gestationnel",
    "hyperthyroidie"
  )

# Ophtalmo (Ophthalmologic disorders)
ophtalmo <- c("keratite", "photosensibilité suite kératite")

# ORL_auditif (ORL and auditory disorders)
ORL_auditif <-
  c("hyperaccousie",
    "hypertrophie des cornets",
    "laryngite chronique")

# Hemato (Hematologic disorders)
hemato <- c("méthémoglobinémie",
            "leiden facteur ii")

# Oncological (Oncologic disorders)
oncological <- c("tumeur gist",
                 "carcinome chromophobe du rein",
                 "cancer du sein métastatique",
                 "cancer du sein métastatique")


# Infectious (Infectious Disorders)
infectious <-
  c("mononucléose à cytomégalovirus", 
    "toxoplasmose congénitale")

# Other
other <-   c(
  "fibromes et kystes aux seins",
  "syndrome de gardner",
  "mutation génétique de type brca2",
  "nouvelle maladie",
  "syndrome de lynch",
  "cytolyse hepatique chronique",
  "parodontie",
  "syndrome d'ehlers-danlos",
  "fibroadénomatose bilatérale (prédominante au  niveau du sein gauche)",
  "congestation",
  "hydrocéphalie",
  "parondontite",
  "syndrome d'hypersensibilité chimique multiple",
  "hypersensibilité chimique multiple",
  "parodontite",
  "mih - hypominéralisation des molaires et des incisives",
  "polypose adénomateuse familiale",
  "syndrome d'elhers danlos hypermobile",
  "hallux  valgus"
)

# 2. Create new dataframe from the created list of categories
all_comorbidities_other <-
  c(
    endometriose_adenomyose,
    gynecologique,
    endocrine_metabolic,
    neuro_pain,
    cardio_circulatory,
    respiratory,
    uro_nephro,
    gastrointestinal,
    musculoskeletal,
    autoimmune_inflammatory,
    mental_health,
    dermatological,
    ophtalmo,
    ORL_auditif,
    hemato,
    oncological,
    infectious,
    other
  )

# 3. Create a dataframe for a list of diseases and their categories 
df_comorbidites_other <- data.frame(comorbidite = unique(all_comorbidities_other))

# 4. Create new vide variable with NA
df_comorbidites_other$categorie <- NA

# 5. Fill the new variable (categorie) 
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% endometriose_adenomyose] <-
  "Endométriose/Adénomyose"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% gynecologique] <-
  "Gynécologique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% endocrine_metabolic] <-
  "Endocrino-métabolique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% neuro_pain] <-
  "Neurologique et Douleurs chroniques"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% cardio_circulatory] <-
  "Cardiovasculaire"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% respiratory] <-
  "Respiratoire"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% uro_nephro] <-
  "Uro-néphrologique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% gastrointestinal] <-
  "Gastro-intestinal"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% musculoskeletal] <-
  "Musculo-squelettique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% autoimmune_inflammatory] <-
  "Inflammatoire"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% mental_health] <-
  "Santé mentale"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% dermatological] <-
  "Dermatologique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% ophtalmo] <-
  "Ophtalmologique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% ORL_auditif] <-
  "ORL et Auditif"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% hemato] <-
  "Hématologique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% oncological] <-
  "Oncologique"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% infectious] <-
  "Infectieuse"
df_comorbidites_other$categorie[df_comorbidites_other$comorbidite %in% other] <-
  "Autre"

# 6. Modifie a name of variable 'comorbidite' like "DiseaseNotFound"
names(df_comorbidites_other)[1] <- "DiseaseNotFound"

# 7. Put categories created to the main dataset "pop_comorbidite"
pop_comorbidite <- pop_comorbidite %>%
  mutate(DiseaseNotFound= tolower(DiseaseNotFound)) %>% 
  left_join(df_comorbidites_other, by = "DiseaseNotFound") %>%
  mutate(categorie = 
           if_else(!is.na(categorie.y), 
                   categorie.y, categorie.x)) %>% 
  dplyr::select(-categorie.x, -categorie.y)
```


## 3) Creating new variables

```{r}
#----------------------------------------------------------------
# Number of comorbidities without endometriosis and adenomyosis
#----------------------------------------------------------------
pop_comorbidite_mod <- pop_comorbidite %>% 
  group_by(id_unik) %>% 
  mutate(
    nb_mal_sansEndoAde = case_when(
      IdDisease == "Endométriose" & !("Adénomyose" %in% IdDisease) ~ 
        nb_maladie - 1,
      IdDisease == "Adénomyose" & !("Endométriose" %in% IdDisease) ~ 
        nb_maladie - 1,
      "Endométriose" %in% IdDisease & "Adénomyose" %in% IdDisease ~ 
        nb_maladie - 2,
      TRUE ~ nb_maladie
    ),
 
    # New variable : Endometriosis, Adenomyosis, Two
    endo_ade_categorie = case_when(
      "Endométriose" %in% IdDisease & "Adénomyose" %in% IdDisease ~ 
        "Both",
      "Endométriose" %in% IdDisease & !("Adénomyose" %in% IdDisease) ~ 
        "Only endometriosis",
      "Adénomyose" %in% IdDisease & !("Endométriose" %in% IdDisease) ~ 
        "Only adenomyosis",
      TRUE ~ "Aucun"
    )
  ) %>% 
  slice_tail(n = 1) %>% 
  ungroup() %>% 
  dplyr::select(id_unik, nb_mal_sansEndoAde, endo_ade_categorie) %>% 
  right_join(pop_comorbidite, by="id_unik")

#-------------------------------------------
#   Comorbidity in binary and categorical
#-------------------------------------------
pop_comorbidite_mod <-  pop_comorbidite_mod %>% 
  mutate(
    comorbidite_bin=ifelse(nb_mal_sansEndoAde==0,"Non","Oui"),
    
    nb_mal_sansEndoAde_cat= cut(
      nb_mal_sansEndoAde,
      breaks= c(-Inf, 0, 1, Inf),
      labels = c("0", "1", "2 ou plus"),
      right=TRUE, 
      ordered_result = TRUE)
    )

#--------------------------------------
# Comorbidity : Obesity , Diabete, HTA
#---------------------------------------
#US National Cholesterol Education Program (NCEP):  Adult Treatment Panel (ATP) III report 
pop_comorbidite_mod <- pop_comorbidite_mod %>%
  group_by(id_unik) %>%
  mutate(
    metabolic_syndrome = if_else(
      any(IdDisease %in% c("Diabète", 
                           "Hypertension", 
                           "Obésité (Indice de Masse Corporelle>30)",
                           "hypercholestérolémie / hypertriglycéridémie")
          ),
      "Oui",
      "Non"
    )
  ) %>%
  ungroup()

# #---------------------
# # Comorbidity : SOPK
# #---------------------
# new_dataframe <- pop_comorbidite_mod %>%
#   group_by(id_unik) %>%
#   mutate(SOPK = case_when(
#     any(IdDisease == "Kyste ovarien, syndrome des ovaires polykystiques (SOPK)") ~"Oui",
#     TRUE ~  "Non")) %>%
#   ungroup()
# 
# pop_comorbidite_mod <- pop_comorbidite_mod %>% 
#   left_join(new_dataframe %>% distinct(id_unik, SOPK),
#             by="id_unik")
# rm(new_dataframe)

#----------------------------------------
# D'autres gynécologiques comorbidités 
#----------------------------------------
pop_comorbidite_mod <- pop_comorbidite_mod %>%
  group_by(id_unik) %>%
  mutate(gynecologique_diseases = ifelse(
    IdDisease %in% c(
      "Fibrome utérin",
      "Kyste ovarien, syndrome des ovaires polykystiques (SOPK)"
    ) |
      DiseaseNotFound %in% c(
        "syndrome douloureux pelvien chronique",
        "insuffisance ovarienne",
        "insuffisance ovarienne précoce",
        "insuffisance ovarienne précose",
        "syndrome de congestion pelvienne",
        "syndrome vésical douloureux",
        "trouble dysphorique prémenstruel",
        "endosalpyngiose",
        "incompetence cervicale",
        "hypersensibilité pelvienne",
        "cogestions pelviennes - varices utérines",
        "prolapsus",
        "syndrome de congestion pelvienne / varices pelviennes",
        "hypersensibilité pelvienne et abdominale"
      ),
    "Oui",
    "Non"
    ),
  gynecologique_diseases =
    case_when(any(gynecologique_diseases=="Oui") ~ "Oui",
              all(gynecologique_diseases=="Non") ~ "Non",
              TRUE~ NA_character_)
  ) %>%
  ungroup() %>% 
  distinct()

#-------------------------------
#    infertility Comorbidity 
#-------------------------------
#pop_comorbidite_mod <- pop_comorbidite_mod %>%
#  group_by(id_unik) %>%
#  mutate(infertility = case_when(
#    any(IdDisease=="Trouble de la fertilité")|
#      any(DiseaseNotFound%in%c("infertilité")) ~
#      "Oui",
#    TRUE ~ "Non"
#  )) %>%
#  ungroup()

pop_comorbidite_mod <- pop_comorbidite_mod %>% 
   group_by(id_unik) %>% 
   mutate(infertility_comorbidite = case_when(
     any(IdDisease=="Trouble de la fertilité")| 
      any(DiseaseNotFound == "infertilité") ~
       "Oui",
     TRUE ~ "Non"
   )) %>% 
   ungroup()



#-----------------------------
#     Auto-immune diseases
#-----------------------------
autoimmune_list <- unique(
  c(
    "Anémie de Biermer (anémie mégaloblastique, anémie pernicieuse)",
    "Thyroïdite d'Hashimoto",
    "Syndrome de Goujerot-Sjögren",
    "Sclérodermie",
    "Maladie de Basedow",
    "Lupus",
    "Maladie de Crohn",
    "Polyarthrite rhumatoïde",
    "Psoriasis",
    "Rhumatisme psoriasique",
    "Rectocolite hémorragique (RCH)",
    "Vitiligo",
    "Syndrome des antiphospholipides",
    "Maladie de Behçet",
    "Maladie de Still de l'adulte",
    "Sclérose en plaque",
    "Hépatite/Cirrhose auto-immune",
    "Lichen plan",
    "Syndrome de Goujerot-Sjögren",
    "Thyroïdite d'Hashimoto",
    "Lupus",
    "Anémie hémolytique auto-immune",
    "Purpura thrombopénique immunologique",
    "Syndrome des antiphospholipides",
    "Maladie de Behçet",
    "Vascularite cryoglobulinémique",
    "Syndrome de Cogan",
    "Sarcoïdose",
    "maladie auto immune inflammatoire non étiquetée",
    "lupus anticoagulant circulant",
    "maladie auto immune d hashimoto"
  )
)
pop_comorbidite_mod <- pop_comorbidite_mod %>%
  mutate(autoimmune_diseases = 
           case_when(
             IdDisease %in% autoimmune_list |
               DiseaseNotFound %in% autoimmune_list ~ "Oui",
             TRUE ~ "Non"
           ))%>%
  group_by(id_unik) %>%
  mutate(autoimmune_diseases = 
           case_when(
             any(autoimmune_diseases=="Oui") ~"Oui",
             all(autoimmune_diseases=="Non") ~ "Non",
             TRUE ~ NA_character_
             )
         )%>% ungroup()
```


# II. Final Dataset

```{r}
pop_comorbidite_final <- pop_comorbidite_mod %>%
  distinct(id_unik,
           nb_mal_sansEndoAde,
           nb_mal_sansEndoAde_cat,
           endo_ade_categorie,
           comorbidite_bin,
           metabolic_syndrome,
           #SOPK,
           gynecologique_diseases,
           infertility_comorbidite,
           autoimmune_diseases)
```

# III. Save the final dataframe 

```{r}
# update : 2025-06-02
write.csv(pop_comorbidite_final,"Stage_M2_Complic_Gross/pop_comorbidite_final.csv")
```

