---
title: "01_clean_PHQ9_GAD7"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_PHQ9_GAD7
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 
 PHQ9 = Patient Health Questionnaire-9 (Depression)
 GAD7 = Generalized Anxiety Disorder-7
```{r}
#-----------------------
# Import a main dataset
#-----------------------
PHQ9_GAD7 <- read.csv(
  "Complic_pop_QST_0335_20250407_PHQ9_GAD7.csv",
  encoding = "latin1",
  colClasses = c(Sexe = "character")
  )

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
PHQ9_GAD7 <- PHQ9_GAD7 %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- PHQ9_GAD7 %>% sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count==nrow(PHQ9_GAD7)) %>% as.vector() # index of columns having only NA
```

# I. Preprocessing Data

```{r}
PHQ9_GAD7 <- PHQ9_GAD7%>%
  mutate(across(starts_with("score_"),~as.numeric(.)))

#---------------------------
#  Participation : Yes/No
#---------------------------
PHQ9_GAD7 <- PHQ9_GAD7 %>% 
  mutate(PHQ9_GAD7_participated=ifelse(is.na(IdAction),
                                       "Non","Oui"))

#----------------------------------
# Score : numeric -> categorical
#----------------------------------
# 1. Create new categorical variables 
PHQ9_GAD7 <- PHQ9_GAD7 %>% 
  mutate(
    score_GAD7_cat= 
      case_when(
        score_GAD7<5 ~ "Minimal",
        score_GAD7>= 5 & score_GAD7< 10 ~ "Mild",
        score_GAD7>= 10 & score_GAD7< 15 ~ "Moderate",
        score_GAD7>= 15 & score_GAD7< 22 ~ "Severe",
        TRUE~NA_character_
          ) %>% 
      factor(
        levels=c("Minimal", "Mild","Moderate","Severe"),
        ordered = TRUE),
    
    score_PHQ9_cat= 
      case_when(
        score_PHQ9<5 ~ "Minimal",
        score_PHQ9>= 5 & score_PHQ9< 10 ~ "Mild",
        score_PHQ9>= 10& score_PHQ9< 15 ~ "Moderate",
        score_PHQ9>= 15& score_PHQ9< 20 ~ "Moderately severe",
        score_PHQ9>= 20& score_PHQ9< 28 ~ "Severe",
        TRUE~NA_character_
        ) %>% 
      factor(
        levels=c("Minimal", "Mild","Moderate",
                 "Moderately severe","Severe"), 
        ordered = TRUE)
  )
```

# II. Final dataset

```{r}
PHQ9_GAD7_final <- PHQ9_GAD7 %>% 
  dplyr::select(id_unik,
                IdAction,
                score_GAD7_cat,
                score_PHQ9_cat,
                PHQ9_GAD7_participated
                ) %>% 
  distinct() %>% 
  mutate(across(-c(id_unik, IdAction), ~
                  case_when(
                    is.na(IdAction) ~ "Not_answered",
                    TRUE  ~ .
                  ))) %>%
  select(-IdAction)
```

# III. Save the final dataframe 

```{r}
#write.csv(PHQ9_GAD7_final,"PHQ9_GAD7_final.csv")
```

