---
title: "01_clean_Situation_financière"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Situation_financière
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
situation_finan <- read.csv(
  "Complic_pop_QST_0315_20250407_situation_financiere.csv",
  fileEncoding = "latin1",
  colClasses = c(Sexe = "character")
)

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
situation_finan <- situation_finan %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim))
```

# I. Preprocessing Data

```{r}
situation_finan <- situation_finan %>%
  mutate(
    situation_financiere =
      situation_financiere %>%
      factor(
        levels = c(
          "Je suis obligé(e) de m'endetter pour m'en sortir",
          "C'est difficile",
          "Un peu juste, je dois faire attention à mes finances",
          "A l'équilibre",
          "Plûtot à l'aise",
          "Très à l'aise"
        ),
        labels = c(
          "In debt",
          "Difficult",
          "A little tight",
          "Balanced",
          "Comfortable",
          "Very comfortable"
        )
      ),
    
    situation_financiere = 
      case_when(
        situation_financiere %in% c("In debt", "Difficult") ~
          "Difficult/In debt",
        situation_financiere %in% c("A little tight","Balanced") ~
          "Balanced",
        situation_financiere %in% c("Comfortable", "Very comfortable") ~
          "Comfortable/Very comfortable"
      ) %>% 
      factor(ordered = FALSE),

    # Particpation : Yes or No
    situation_financiere_participated =
      ifelse(is.na(situation_financiere), "Non", "Oui")
  )
```

# II. Final dataframe

```{r}
situation_finan_final <- situation_finan
```

# III. Save the final dataframe 

```{r}
#write.csv(situation_finan_final,"situation_finan_final.csv")
```

