---
title: "01_clean_module_grossesse"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_module_grossesse without ongoning pregnancies
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-04-09
*Date of update* : 2025-06-16
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

Information after selecting just 'terminated pregnancies'
 - 1107 patients : New module = 
 - 1619 pregnancies : New  module= 515 , Old module= 1104 
 
$Libraries$ 
```{r}
library(dplyr)
library(readr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```


# 0. Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
# Original : 1955 Idpreg, 1402 id_unik
grossesse <- read.csv(
  "Complic_pop_Module_grossesse_20250407.csv",
  fileEncoding = "latin1",
  colClasses = c(Sexe = "character"))

#-----------------------------------------------------
# Replace cells with nothing or spaces with NA and
# Replace non-standard apostrophes with standard ones 
#-----------------------------------------------------
grossesse <- grossesse %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), ~na_if(trimws(.), " ")),
         across(everything(), ~ gsub("\u0092", "'", .)))%>% 
  mutate_all(~str_trim(.))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- grossesse %>% sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count==nrow(grossesse)) %>% as.vector() # index of columns having only NA
grossesse <- grossesse[,-index_col_NA]

#---------------------------------
# Outlier : DateDebut et DateFin  : A LOT 
#---------------------------------
# 1. Character to 'date' data
grossesse <-  grossesse %>%
  mutate(across(starts_with("Date"), 
                ~ as.Date(.,format="%d/%m/%Y")))

# 2. Correct some evident outliers
grossesse <- grossesse %>%
  mutate(
    DateDebut = case_when(
      # outlier : 8015-03-30-> 2015-03-30
      id_unik == "3B44783A6F2E5375F12F" ~ as.Date("2015-03-30"),
      # outlier : DateDebut== 2002-01-18 -> 2001-05-01
      id_unik == "77746B417C56728035D3" ~ as.Date("2001-05-01"),
      # outlier : DateDebut=="2024-09-03" -> "2024-08-14"
      IdPregnancy=="PRG_5522" ~ as.Date("2024-08-14"),
      # outlier : DateDebut 2019-08-08-> 2018-08-08
      IdPregnancy=="PRG_1401" ~ as.Date("2018-08-08"),
      # outlier : DateDebut 2023-05-11->2022-05-11
      IdPregnancy=="PRG_4358" ~ as.Date("2022-05-11"),
      # outlier  : DateDebut 2023-06-15 -> 2022-06-15
      IdPregnancy=="PRG_4056" ~ as.Date("2022-06-15"),
      #outlier  : DateDebut 2019-08-08 -> 2018-08-08
      IdPregnancy=="PRG_1401" ~ as.Date("2018-08-08"),
      # outlier : DateDebut 2016-07-01 -> 2015-07-01
      IdPregnancy =="PRG_4081" ~as.Date("2015-07-01"),
      # outlier : DateDebut 2022-12-14 -> 2021-12-14
      IdPregnancy=="PRG_4204"~ as.Date("2021-12-14"),
      # outlier : DateDebut 2007-11-10 -> 2006-11-10
      IdPregnancy=="PRG_3973"~ as.Date("2006-11-10"),
      # outlier : DateDebut 2019-08-08 -> 2018-08-08
      IdPregnancy=="PRG_1401" ~ as.Date("2018-08-08"),
      # outlier : DateDebut 2019-08-04 -> 2018-08-04
      IdPregnancy=="PRG_2385" ~ as.Date("2018-08-04"),
      # outlier : DateDebut 2018-05-25 -> 2017-05-25
      IdPregnancy=="PRG_400" ~ as.Date("2017-05-25"),
      # outlier : DateDebut 2018-05-27 -> 2017-05-27
      IdPregnancy=="PRG_453" ~ as.Date("2017-05-27"),
      # outlier : DateDebut 2018-05-20 -> 2017-05-20
      IdPregnancy=="PRG_1549" ~ as.Date("2017-05-20"),
      # outlier : DateDebut 2017-05-02 -> 2016-05-02
      IdPregnancy=="PRG_345" ~ as.Date("2016-05-02"),
      # outlier : DateDebut 2021-06-14 -> 2020-06-14
      IdPregnancy=="PRG_2894" ~ as.Date("2020-06-14"),
      # outlier : DateDebut 2020-07-07 -> 2019-07-07
      IdPregnancy=="PRG_1907" ~ as.Date("2019-07-07"),
      # outlier : DateDebut 2016-04-15 -> 2015-04-15
      IdPregnancy=="PRG_2740" ~ as.Date("2015-04-15"),
      IdPregnancy =="PRG_1471" ~ DateFin,
      # outlier : DateDebut 2019-10-13 -> 2020-10-13
      IdPregnancy=="PRG_2101" ~ as.Date("2020-10-13"),
      # outlier : DateDebut 2018-10-17 -> 2020-10-17
      IdPregnancy=="PRG_2288" ~ as.Date("2020-10-17"),
      # outlier : DateDebut 2021-11-05 -> 2022-11-05
      IdPregnancy=="PRG_2935" ~ as.Date("2022-11-05"),
      # outlier : 2012-04-15 -> 2014-04-15
      IdPregnancy=="PRG_586" ~ as.Date("2014-04-15"),
      # outlier : 2019-01-23 -> 2020-01-23
      IdPregnancy=="PRG_1097" ~ as.Date("2020-01-23"),
      
      IdPregnancy== "PRG_2360" ~ as.Date("2018-12-01"),
      # outlier : 2016-09-07 -> 2017-09-07
      IdPregnancy=="PRG_1118" ~ as.Date("2017-09-07"),
      # outlier : 2015-02-26 -> 2016-02-26
      IdPregnancy=="PRG_2861" ~ as.Date("2016-02-26"),
      # outlier : 2019-01-25 -> 2020-01-25
      IdPregnancy=="PRG_1075" ~ as.Date("2020-01-25"),
      TRUE ~ DateDebut
    ),
    DateFin = case_when(
      # outlier : 3020-01-15 -> 2020-01-15
      id_unik == "4EBAFCACC71D96FF1C77" ~ as.Date("2020-01-15"),
      # outlier : 2013-10-20 -> 2023-10-20
      IdPregnancy=="PRG_4099" ~ as.Date("2023-10-20"),
      # outlier : DateFin == 2001-05-01 -> 2002-01-18
      id_unik == "77746B417C56728035D3" ~ as.Date("2002-01-18"),
      # outlier : 2016-01-08 -> 2018-01-08
      IdPregnancy=="PRG_544" ~ as.Date("2018-01-08"),
      # outlier : 2023-01-07 -> 2024-01-07
      IdPregnancy=="PRG_5158" ~ as.Date("2024-01-07"),
      # outlier : DateFin=="2024-08-14" -> "2024-09-03" 
      IdPregnancy=="PRG_5522" ~ as.Date("2024-09-03"),
      
      IdPregnancy =="PRG_1471" ~ DateDebut,
      IdPregnancy== "PRG_2360" ~ as.Date("2019-01-08"),
      TRUE ~ DateFin
    )
  )

# # outlier : 8015-03-30-> 2015-03-30
# grossesse$DateDebut[grossesse$id_unik=="3B44783A6F2E5375F12F"]<- "2015-03-30"
# # outlier : 3020-01-15 -> 2020-01-15
# grossesse$DateFin[grossesse$id_unik=="4EBAFCACC71D96FF1C77"]<- "2020-01-15"
# # outlier : IdPregnancy=="PRG_4099" 2013-10-20 -> 2023-10-20
# grossesse$DateFin[grossesse$id_unik=="C7C8E9716FF7B28E1B5F"] <- "2023-10-20"
# # outlier : DateDebut== 2002-01-18 , DateFin ==2001-05-01 -> change each other
# grossesse$DateDebut[grossesse$id_unik=="77746B417C56728035D3"] <- "2001-05-01"
# grossesse$DateFin[grossesse$id_unik=="77746B417C56728035D3"] <- "2002-01-18"
# semaine_amenor : -71 -> PRG_544 DateFin== as.Date("2018-01-08") 
# semaine_amenor : -39 ->C47B5E8F7A3572590AC5 and PRG_5158 DateFin== as.Date("2024-01-07")
# semaine_amenor : -3 -> AFF39127F8E135C14FBF and PRG_5522 DateFin==as.Date("2024-09-03") and DateDebut=as.Date("2024-08-14")
# PRG_1401 : DateDebut 2019-08-08-> 2018-08-08
# PRG_4358 : DateDebut 2023-05-11->2022-05-11
# PRG_4056 : DateDebut 2023-06-15 -> 2022-06-15
#PRG_1401 : DateDebut 2019-08-08 -> 2018-08-08
# PRG_4081: DateDebut 2016-07-01 -> 2015-07-01
# PRG_4204 : DateDebut 2022-12-14 -> 2021-12-14
# PRG_3973 : DateDebut 2007-11-10 -> 2006-11-10

#-------------------------------------------
# Outlier : moyen duration of breastfeeding 
#--------------------------------------------
# - maximum 120
grossesse$mois_allaitement %>% as.numeric() %>% summary() 

# Correct the outlier : 120/30 jours -> 4 mois  (confirmé par Marina)
grossesse$mois_allaitement[grossesse$id_unik=="5747A857A42403B262AC"] <- 4

#--------------------------------------------------
#  Delete some repetead rows (중복된 임신 데이터를 제거)
# Same pregnancies with a little bit different date
#--------------------------------------------------
# 1. Check if there are some repeted data
grossesse %>%
  group_by(id_unik, DateFin) %>%
  filter(n() > 1) %>%
  ungroup()

# 2. Remove STRANGE DATA
# 9A38939EB57EAEF9B2EA (3) : same DateFin (-PRG_5825, - PRG_5800)
# D1E2C77AE3BC83F50613 (3) : same DateFin (-PRG_5742, - PRG_5744)
# 38C1703D0D31744BBF95 (5) : same DateFin but different DateDebut (- PRG_2392)
# Same Pregnancy but ongoing pregnancy at the moment of first response
# 161650AB86A4BF788555 (2) : DateDebut 09/03/2024  (-PRG_5252)
grossesse_mod2 <- grossesse %>%
  filter(!IdPregnancy %in% 
           c("PRG_5825", "PRG_5800", "PRG_5744", "PRG_5742", "PRG_5252"))

# # + Alternative method
# # Like some rows having the same information of pregnancies but different IdPregnancy 
# # - DateFin est un champ obligatoire et DateDebut n'est pas obligatoire
# grossesse_mod2 <- grossesse %>% 
#   distinct(id_unik,DateFin, .keep_all = TRUE) 

#-------------------------------
#      Ongoing pregnancy        
# : 369 femmes et 369 grossesses = 367 en cours + 2 (pas de réponse)
#-------------------------------
# In new module : 104 pregnancies + 2 Pas de réponse 
# In old module : 263 pregnancies

# 1. Create a temporary dataframe 
new_dataframe <- grossesse_mod2 %>% 
  # DateDebut exists but DateFin, terme2,fin2,issue2 don't 
  filter(is.na(DateFin) & is.na(terme2) & is.na(fin2) & is.na(issue2)) %>% 
  distinct(id_unik,IdPregnancy,DateDebut,DateFin,DateCreation,DateModification,Debut_1_2,Vivant,terme2) %>% 
  mutate(SA_estime = round(as.numeric(DateCreation - DateDebut) / 7 + 2),
         grossesse_encours= 
           case_when(
             as.numeric(SA_estime)<= 42 ~ "Oui",
             as.numeric(SA_estime) > 42 ~ "Pas de reponse",
             TRUE ~ NA_character_
           )
  )

# 2. Add new variable to the main dataset 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe %>% dplyr::select(IdPregnancy, grossesse_encours),
            by="IdPregnancy")

# 3. Add new class 'Non' to the new variable 'grossesse_encours'
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(grossesse_encours=
           case_when(
             !is.na(grossesse_encours) ~ grossesse_encours,
             is.na(grossesse_encours) ~ "Non",
             TRUE~ NA_character_
           ))

# 4. Delete the temporary dataframe 
rm(new_dataframe)

#-----------------------------------
#    Filter ongoing pregnancies
#-----------------------------------
grossesse_mod2 <- grossesse_mod2 %>%
  filter(grossesse_encours=="Non")%>% 
  dplyr::select(-grossesse_encours) %>% 
  distinct()  

test <-  grossesse_mod2
#write.csv(grossesse_mod2, "V2_grossesse_mod2.csv")
#test <- read.csv("V2_grossesse_mod2.csv")
```


*INFORMATION* after cleaning up the data
- 1113 patients
- 1575 pregnancies

```{r}
grossesse_mod2 %>% distinct(id_unik) %>% n_distinct() # 1113 patients
grossesse_mod2 %>% distinct(IdPregnancy) %>% n_distinct() #1575 preg
```


# I. Preprocessing Data

```{r}
#--------------------------------------------------------
# New 'DateDebut' with data of 'amenorrhee' and 'FinMois'
#--------------------------------------------------------
# Semaine de grossesse = Semaine d'amenorrhee -2
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(
    # DateFin - (number of amenorrhea - 2)*7 days = probable DateDebut(Date de debut de grossesse)
    DateDebut_grossesse = DateFin - (as.numeric(amenorrhee)-2) * 7,
    
    # DateFin - number of FinMois days = probable DateDebut (if DateDebut is missing)
    Finjours = as.numeric(FinMois) * 30,
    DateDebut_FinMois = DateFin - Finjours,
    
    # Create a new variable with original data, DateDebut_amenorrhee, and DateDebut_FinMois data
    merged_DateDebut = case_when(
      !is.na(DateDebut) ~ DateDebut,
      !is.na(DateDebut_grossesse)& is.na(DateDebut_FinMois) ~
        DateDebut_grossesse,
      is.na(DateDebut) & is.na(DateDebut_grossesse) & !is.na(DateDebut_FinMois) ~
        DateDebut_FinMois,
      TRUE ~ as.Date(NA)
    )
  ) %>% 
  dplyr::select(-DateDebut_grossesse, -Finjours, -DateDebut_FinMois)

#-----------------------------------------------
#  Weeks of amenorrhea = End Date - start Date  
#-----------------------------------------------
#  There are a lot of strange data
# 1. Calculate weeks of amenorrhea by using Date data
# SA = SG (=DateFin-DateDebut) + 2
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(preg_jours = as.numeric(difftime(DateFin, merged_DateDebut, 
                                          units = "days")),
         semaine_amenor = round(preg_jours / 7)+2) %>% 
  dplyr::select(-preg_jours)


# 2. Correct strange datas
# It's normal to have some strange data because of incomplete information from old module 
# So I don't take your time to all correct 

# 3. Merge this new variable to amenorrhee variable
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    amenorrhee = as.numeric(amenorrhee),
    amenorrhea_final =
      case_when(
        terme2 == "Oui" & is.na(semaine_amenor) ~
          amenorrhee,
        terme2 == "Oui" & !is.na(semaine_amenor) & semaine_amenor > 0 ~
          semaine_amenor,
        terme2 == "Oui" & !is.na(semaine_amenor) & semaine_amenor < 0 ~
          amenorrhee,
        (terme2 == "Non" | is.na(terme2)) & !is.na(semaine_amenor) ~
          semaine_amenor,
        TRUE ~ NA_real_
      )
  )

# 4. Correct some outliers 
#66902F57FCB43D60994B & PRG_4657: 4 -> 42
grossesse_mod2$amenorrhea_final[grossesse_mod2$id_unik=="66902F57FCB43D60994B"] <- 42
#99F6B8BA60CD0ED959EB & PRG_5326 : 2 -> 50 
grossesse_mod2$amenorrhea_final[grossesse_mod2$IdPregnancy=="PRG_5326"] <- 50

# 5. Just check just for the participants in the new module
# Because there are lots of bizarre data in the old module in terms of Date data
grossesse_mod2$amenorrhea_final[grossesse_mod2$terme2=="Oui"] %>% table()

#-------------------------------
#      Ongoing pregnancy        
# : 369 femmes et 369 grossesses = 367 en cours + 2 (pas de réponse)
#-------------------------------
# In new module : 104 pregnancies + 2 Pas de réponse 
# In old module : 263 pregnancies

# 1. Create a temporary dataframe 
new_dataframe <- grossesse_mod2 %>% 
  # DateDebut exists but DateFin, terme2,fin2,issue2 don't 
  filter(!is.na(DateDebut) & is.na(DateFin) & is.na(terme2) & is.na(fin2) & is.na(issue2)) %>% 
  dplyr::select(id_unik,IdPregnancy,DateDebut,DateCreation,DateModification,Debut_1_2,Vivant,terme2) %>% 
  mutate(SA_estime = round(as.numeric(DateCreation - DateDebut) / 7 + 2),
         grossesse_encours= 
           case_when(
             as.numeric(SA_estime)<= 42 ~ "Oui",
             as.numeric(SA_estime) > 42 ~ "Pas de reponse",
             TRUE ~ NA_character_
           )
  )

# 2. Add new variable to the main dataset 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe %>% dplyr::select(IdPregnancy, grossesse_encours), by="IdPregnancy")

# 3. Add new class 'Non' to the new variable 'grossesse_encours'
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(grossesse_encours=
           case_when(
             !is.na(grossesse_encours) ~ grossesse_encours,
             is.na(grossesse_encours) ~ "Non",
             TRUE~ NA_character_
           ))

# 4. Delete the temporary dataframe 
rm(new_dataframe)

#-------------------------------------
#  Correction of variable terme2
#-------------------------------------
# PRG_3699 : terme2 =="Non"
grossesse_mod2$terme2[grossesse_mod2$IdPregnancy=="PRG_3699"] <- "Non"

#----------------------------------------
#       Terme/incomplete pregnancy
#----------------------------------------
# status_terme_all 
# status_terme_onlynew
# live_birth_bin_all
# live_birth_bin_onlynew
# type_preg_all
# type_preg_all_simple 
# type_preg_onlynew
# type_preg_onlynew_simple
#----------------------------------------
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    # 1. Create a binary variable : Terme / Incomplete
    # 1-1) Consider 'fin2' with just two class (Terme, Incomplete)
    status_terme_all =
      case_when(
        terme2 == "Oui" ~ "Terme",
        terme2=="Non" ~ "Incomplete",
        is.na(terme2) & fin2 != "Naissance" ~ "Incomplete",
        is.na(terme2) & fin2=="Naissance" ~ "Terme",
        TRUE ~NA_character_
      ),
    # 1-2) Don't consider 'fin2' (just new module)
    status_terme_onlynew = 
      case_when(
        terme2 == "Oui" ~ "Terme",
        terme2=="Non" ~ "Incomplete",
        TRUE ~NA_character_
      ),
    # 1-3) Consider only old module
    status_terme_onlyold = 
      case_when(
        fin2=="Naissance"~ "Terme",
        fin2!="Naissance" ~ "Incomplete",
        TRUE ~NA_character_
      ),
    
    # 2. Create a binary variable : Live Birth Yes or No (both modules)
    # 2-1) both modules
    live_birth_bin_all = 
      case_when(
        status_terme_all=="Terme" ~ "Oui",
        status_terme_all=="Incomplete"~ "Non",
        TRUE ~ NA_character_
      ),
    # 2-2) only new module
    live_birth_bin_onlynew = 
      case_when(
        status_terme_onlynew=="Terme" ~ "Oui",
        status_terme_onlynew=="Incomplete"~ "Non",
        TRUE ~ NA_character_
      ),
    # 2-3) only old module
    live_birth_bin_onlyold= 
      case_when(
        status_terme_onlyold=="Terme" ~ "Oui",
        status_terme_onlyold== "Incomplete" ~ "Non",
        TRUE ~ NA_character_
      ),
    
    # 3. Create a categorical variable : fullterm/preterm and different types of incomplete pregnancies
    # 3-1) Consider 'fin2' (both modules)
    type_preg_all = 
      case_when(
        status_terme_all=="Terme"  & amenorrhea_final >= 37 ~ "Fullterm",
        status_terme_all=="Terme"  & amenorrhea_final < 37 ~ "Preterm",
        status_terme_all=="Incomplete" & is.na(terme2) ~ fin2,
        status_terme_all=="Incomplete" & terme2=="Non" & is.na(fin2) &
          issue2 %in%c("Fausse-couche", 
                       "Grossesse extra-utérine (par ex : grossesse dans la trompe de Fallope)",
                       "Grossesse môlaire",
                       "Autre")~ issue2,
        status_terme_all=="Incomplete" & terme2=="Non" & is.na(fin2)&
          issue2== "Interruption volontaire de grossesse (IVG)"~"IVG",
        status_terme_all=="Incomplete" & terme2=="Non" & is.na(fin2)&
          issue2== "Interruption médicale de grossesse (IMG)" ~ "IMG",
        TRUE ~NA_character_
      ),
    
    type_preg_all_simple= 
      case_when(
        type_preg_all %in% c("Grossesse extra-utérine (par ex : grossesse dans la trompe de Fallope)",
                            "Grossesse môlaire",
                            "Mort foetale intra-uterine",
                            "Autre") ~"Autre",
        type_preg_all %in% c("IMG","IVG") ~ "IVG/IMG",
        TRUE ~ type_preg_all
      ),
    
    # 3-2) Don't consider 'fin2' (just new module)
    type_preg_onlynew = 
      case_when(
        status_terme_onlynew=="Terme"  & amenorrhea_final >= 37 ~ "Fullterm",
        status_terme_onlynew=="Terme" & amenorrhea_final < 37 ~ "Preterm",
        status_terme_onlynew=="Incomplete" & 
          issue2 %in%
          c("Fausse-couche", 
            "Grossesse extra-utérine (par ex : grossesse dans la trompe de Fallope)",
            "Grossesse môlaire",
            "Autre")~ issue2,
        status_terme_onlynew=="Incomplete" &
          issue2== "Interruption volontaire de grossesse (IVG)" ~ "IVG",
        status_terme_onlynew=="Incomplete" &
          issue2== "Interruption médicale de grossesse (IMG)" ~ "IMG",
        TRUE ~NA_character_
      ),
    
    type_preg_onlynew_simple= 
      case_when(
        type_preg_onlynew %in% c("Grossesse extra-utérine (par ex : grossesse dans la trompe de Fallope)",
                            "Grossesse môlaire",
                            "Mort foetale intra-uterine",
                            "Autre") ~"Autre",
        type_preg_onlynew %in% c("IMG","IVG") ~ "IVG/IMG",
        TRUE ~ type_preg_onlynew
      ),
  ) 

#-------------------------------
#   Live birth at least once    : Binary
#-------------------------------
# 1. Create a temporary dataframe
new_dataframe <- grossesse_mod2 %>% 
  group_by(id_unik) %>% 
  mutate(live_birth_bin_atleast_once_all = 
           case_when(any(live_birth_bin_all=="Oui")~ "Oui",
                     all(live_birth_bin_all=="Non")~ "Non",
                     TRUE ~ NA_character_),
         live_birth_bin_atleast_once_onlynew=
           case_when(any(live_birth_bin_onlynew=="Oui")~"Oui",
                     all(live_birth_bin_onlynew=="Non")~"Non",
                     TRUE ~ NA_character_)
  ) %>% 
  ungroup %>% 
  distinct(id_unik, live_birth_bin_atleast_once_all,live_birth_bin_atleast_once_onlynew)

# 2. Add the new variable to the main dataframe
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe,by="id_unik")

# 3. Delete the temporary dataframe
rm(new_dataframe)

#-------------------------------------------
#    Number of live births per patient
#-------------------------------------------
# 1. Create a new dataframe for informations : Number of live births
# - all : both modules
# - onlynew :only new module for the sensitivity analysis
new_dataframe <- grossesse_mod2 %>% 
  group_by(id_unik) %>% 
  summarise(
    nb_live_births_all= sum(live_birth_bin_all == "Oui", 
                            na.rm = TRUE),
    nb_live_births_onlynew= sum(live_birth_bin_onlynew == "Oui", 
                                na.rm = TRUE)) %>% 
  ungroup %>% 
  distinct(id_unik,nb_live_births_all,nb_live_births_onlynew)

# 2. Add these informations to main dataset 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe, by="id_unik") 

# 3. Delete the temporary dataframe
rm(new_dataframe)

# 4. Create a new variable : number of live births in old module
grossesse_mod2 <- grossesse_mod2%>% 
  mutate(
    nb_live_births_onlyold = ifelse(module == "Ancien", 
                                nb_live_births_all - nb_live_births_onlynew, 
                                NA)
    ) 


#---------------------------------------
#     Number of incomplet pregnancy    
#---------------------------------------
# 1. Create a new dataframe for informations : Number of incomplet pregnancy
# - Consider 'fin2' : nb_incomplete_preg_pt_all
# - Don't consider 'fin2' : nb_incomplete_preg_pt_onlynew
new_dataframe <- grossesse_mod2 %>%
  group_by(id_unik) %>% 
  summarise(nb_incomplete_preg_pt_all=sum(status_terme_all=="Incomplete",
                                          na.rm=TRUE),
            nb_incomplete_preg_pt_onlynew=sum(status_terme_onlynew=="Incomplete",
                                              na.rm=TRUE)
  ) %>%
  ungroup %>%
  distinct(id_unik, nb_incomplete_preg_pt_all,nb_incomplete_preg_pt_onlynew)

# 2. Add these information to main dataset 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe, by="id_unik")

# 3. Delete the temporary dataframe
rm(new_dataframe)

#-------------------------------------
#    Number of total pregnancies    : all modules
#-------------------------------------
# 1) ALL MODULES
# 1-1) Create a new dataframe
new_dataframe <- grossesse_mod2 %>% 
  distinct(id_unik, IdPregnancy) %>% 
  group_by(id_unik) %>% 
  summarise(nb_total_gross = n(), .groups = "drop") %>% 
  ungroup()

# 1-2) Add the information to main dataset 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe, 
            by="id_unik")

# 1-3) Delete the temporary dataframe
rm(new_dataframe)

# 2. ONLY NEW MODULE
# 2-1) Create a new dataframe
new_dataframe <- grossesse_mod2 %>%
  filter(module=="Nouveau") %>% 
  distinct(id_unik, IdPregnancy) %>% 
  group_by(id_unik) %>% 
  summarise(nb_total_gross_NewModule = n(), .groups = "drop") %>% 
  ungroup()

# 2-2) Add the information to main dataset 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe, 
            by="id_unik")

# 2-3) Delete the temporary dataframe
rm(new_dataframe)


# 3. ONLY OLD MODULE
# 3-1) Create a new dataframe
new_dataframe <- grossesse_mod2 %>%
  filter(module=="Ancien") %>% 
  distinct(id_unik, IdPregnancy) %>% 
  group_by(id_unik) %>% 
  summarise(nb_total_gross_OldModule = n(), .groups = "drop") %>% 
  ungroup()

# 3-2) Add the information to main dataset 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe, 
            by="id_unik")

# 3-3) Delete the temporary dataframe
rm(new_dataframe)

#------------------------------------
#       Live birth rate (%)
#------------------------------------
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(taux_LB_ALL_all= nb_live_births_all/nb_total_gross,
         taux_LB_ALL_onlynew= nb_live_births_onlynew/nb_total_gross,
         taux_LB_NEWMODULE= nb_live_births_onlynew/ nb_total_gross_NewModule)

#--------------------------------------------------------------------------
#  Maternal age at the time of pregnancy termination for each pregnancy
#--------------------------------------------------------------------------
#  Why did I use DateFin? because DateDebut isn't mandatory to answer so I find DateFin more fiable
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(age_maternal = as.numeric(
    difftime(DateFin, Date_naissance, units = "days")) /365)
grossesse_mod2$age_maternal <- grossesse_mod2$age_maternal %>% round(1)

#--------------------------------------------
# Maternal age for first and last pregnancy
#--------------------------------------------
# For the only one pregnancy, I just calculated the first maternal age (not last maternal age)

# 1.Create a var maternal age for the last pregnancy only if the number of preg >1
new_dataframe <- grossesse_mod2 %>%
  group_by(id_unik) %>%
  arrange(merged_DateDebut) %>% 
  mutate(first_preg_age = first(age_maternal),
         last_preg_age = last(age_maternal)) %>%
  ungroup() %>%
  distinct(id_unik, first_preg_age, last_preg_age)


# 2. Add new variables to the main dataframe and create a new variable 
grossesse_mod2 <- grossesse_mod2 %>%
  left_join(new_dataframe, by="id_unik") %>% 
  # Create a new var : maternal age for the last pregnancy only if the number of preg >1
  mutate(
    last_preg_age_mod = case_when(
      nb_total_gross == 1 ~ NA_real_, 
      nb_total_gross > 1 ~ last_preg_age,
      TRUE ~ NA_real_ 
    )
  ) 

# 3. Remove the unnecessary variable
grossesse_mod2 <- grossesse_mod2 %>% 
  dplyr::select(-last_preg_age)

# 4. Delete the temporary dataframe 
rm(new_dataframe)

#-------------------------------------------------------------
# 1. Conception method : Modifications of variables 
#-------------------------------------------------------------
# 1. Modify 'Debut_1_2'
# 1-1) Debut_1_2 :  Only the use of medications without IVF or IUI
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    Debut_1_2 =
      case_when(
        Debut_1_2 == "Médicaments pour stimuler l'ovulation (ex: Clomid,clomifène..)" &
          Debut_2_2 %in% c("Fécondation in vitro", "Insémination artificielle") ~ Debut_2_2,
        
        Debut_1_2 != "Médicaments pour stimuler l'ovulation (ex: Clomid,clomifène..)" ~ Debut_1_2,
        
        Debut_1_2 == "Médicaments pour stimuler l'ovulation (ex: Clomid,clomifène..)" ~ "Médicaments pour stimuler l'ovulation (ex: Clomid,clomifène..)",
        TRUE ~ NA_character_
      ) 
  )

# 1-2) Modify "Autre" values in 'Debut_1_2'
grossesse_mod2$Debut_1_2[grossesse_mod2$IdPregnancy=="PRG_2538"] <-
  "Insémination artificielle"
grossesse_mod2$Debut_1_2[grossesse_mod2$IdPregnancy=="PRG_4859"] <- 
  "Opération / chirurgie"
grossesse_mod2$Debut_1_2[grossesse_mod2$IdPregnancy=="PRG_4126"] <- 
  "Fécondation in vitro"

# 2. Modify 'Debut_2_2'
# 2-1) Modify 'Autre' values in the variable 'Debut_2_2' by replacing them with 'DebutAutre'
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(Debut_2_2 = case_when(
    id_unik == "C3E5E0D3EF41F43E06DA" ~ "Opération / chirurgie",
    id_unik == "D5BF37850C416F9EAD9B" ~ "Opération / chirurgie",
    Debut_2_2 == "Autre" ~ NA_character_,
    !is.na(Debut_2_2) & Debut_2_2 != "Autre" ~ Debut_2_2,
    TRUE ~ NA_character_
  ))

# 2-2) If Debut_1_2 == Debut_2_2, set Debut_2_2 to NA
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    Debut_2_2 = case_when(
      Debut_1_2 == Debut_2_2 ~ NA_character_,
      Debut_1_2 != Debut_2_2 ~ Debut_2_2,
      TRUE ~ NA_character_
    )
  )

#---------------------------------------
#  2. Conception method per Patient
#---------------------------------------
# 1. Create a temporary dataframe for a new variable ('methode_conception') by transforming wide format -> long format
new_dataframe1 <- grossesse_mod2 %>%
  dplyr::select(id_unik, starts_with("Debut_")) %>%
  pivot_longer(cols = starts_with("Debut_"), 
               values_drop_na = TRUE) %>%
  distinct(id_unik, value) %>%
  mutate(
    methode_conception_long_perID = case_when(
      value %in% c("Conception naturelle") ~ "Naturelle",
      value %in% c("Médicaments pour stimuler l'ovulation (ex: Clomid,clomifène..)") ~ "Medicament_sans_PMA",
      value %in% c("Insémination artificielle") ~ "IUI",
      value %in% c("Fécondation in vitro") ~ "IVF",
      value %in% c("Opération / chirurgie") ~ "Operation",
      value %in% c("Autre") ~ "Autre",
      value %in% c("Je ne m'en souviens plus",
                   "Je ne souhaite pas répondre") ~ "nsp",
      TRUE ~ NA_character_
    )) %>%
  filter(!is.na(methode_conception_long_perID)) %>%
  distinct(id_unik, methode_conception_long_perID)

# 2. Create a separate column for the "methode_conception" value
new_dataframe2 <- new_dataframe1 %>% 
  filter(!is.na(methode_conception_long_perID)) %>%
  mutate(value = TRUE) %>%  
  pivot_wider(
    names_from = methode_conception_long_perID,  
    values_from = value,  
    values_fill = list(value = FALSE),  
    names_prefix = "methode_perID_" 
  )

# 3. Put the new variable into main dataframe 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe1, by="id_unik") %>% 
  left_join(new_dataframe2, by="id_unik")

# 4. Delete two temporary dataframes
rm(new_dataframe1);rm(new_dataframe2)

#-------------------------------------------------------
# 2-1. Conception method per Patient : binary(Yes/NO)
#-------------------------------------------------------
# Only naturelle conception is 'Non' among all pregnancies
# At least once, if patients exprienced different type of conception method like IUI, IVF, Medications or operation, the value would be "Oui"

# 1. Create a termporarydataframe for new variables in binary
new_dataframe <- grossesse_mod2 %>% 
  distinct(id_unik, methode_conception_long_perID) %>% 
  mutate(Methode_conception_bin_perID=
           ifelse(methode_conception_long_perID=="Naturelle",0,1)) %>%
  group_by(id_unik) %>%
  mutate(somme= sum(Methode_conception_bin_perID),
         Methode_conception_bin_perID= ifelse(somme>0, "Avec méthodes", "Naturelle")) %>%  
  ungroup %>% 
  distinct(id_unik, Methode_conception_bin_perID)

# 2. Add new variable to the main dataframe
grossesse_mod2 <- grossesse_mod2 %>%
  left_join(new_dataframe, by="id_unik")

# 3. Delete the temporary dataframe
rm(new_dataframe)

#-----------------------------------------------------------------------------------
# 2-2. Conception method per Patient : Different combination of conception method
#-----------------------------------------------------------------------------------
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    methode_conception_perID_cat = case_when(
      # Unique method
      methode_perID_Naturelle == TRUE &
        !methode_perID_IUI &
        !methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "Naturelle seule",
      methode_perID_IUI == TRUE &
        !methode_perID_Naturelle &
        !methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "IUI seule",
      methode_perID_IVF == TRUE &
        !methode_perID_Naturelle &
        !methode_perID_IUI &
        !methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "IVF seule",
      methode_perID_Medicament_sans_PMA == TRUE &
        !methode_perID_Naturelle &
        !methode_perID_IUI &
        !methode_perID_IVF &
        !methode_perID_Operation ~ "Médicaments sans PMA seule",
      methode_perID_Operation == TRUE &
        !methode_perID_Naturelle &
        !methode_perID_IUI &
        !methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA ~ "Opération seule",
      
      # Two methods
      methode_perID_Naturelle &
        methode_perID_IUI &
        !methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "Naturelle + IUI",
      methode_perID_Naturelle &
        methode_perID_IVF &
        !methode_perID_IUI &
        !methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "Naturelle + IVF",
      methode_perID_Naturelle &
        methode_perID_Medicament_sans_PMA &
        !methode_perID_IUI &
        !methode_perID_IVF &
        !methode_perID_Operation ~ "Naturelle + Médicaments",
      methode_perID_Naturelle &
        methode_perID_Operation &
        !methode_perID_IUI &
        !methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA ~ "Naturelle + Opération",
      methode_perID_IUI &
        methode_perID_IVF &
        !methode_perID_Naturelle &
        !methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "IUI + IVF",
      methode_perID_IUI &
        methode_perID_Medicament_sans_PMA &
        !methode_perID_Naturelle &
        !methode_perID_IVF &
        !methode_perID_Operation ~ "IUI + Médicaments",
      methode_perID_IUI &
        methode_perID_Operation &
        !methode_perID_Naturelle &
        !methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA ~ "IUI + Opération",
      methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        !methode_perID_Naturelle &
        !methode_perID_IUI &
        !methode_perID_Operation ~ "IVF + Médicaments",
      methode_perID_IVF &
        methode_perID_Operation &
        !methode_perID_Naturelle &
        !methode_perID_IUI &
        !methode_perID_Medicament_sans_PMA ~ "IVF + Opération",
      methode_perID_Medicament_sans_PMA &
        methode_perID_Operation &
        !methode_perID_Naturelle &
        !methode_perID_IUI &
        !methode_perID_IVF ~ "Médicaments + Opération",
      
      # Three methods
      methode_perID_Naturelle &
        methode_perID_IUI &
        methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "Naturelle + IUI + IVF",
      methode_perID_Naturelle &
        methode_perID_IUI &
        methode_perID_Medicament_sans_PMA &
        !methode_perID_IVF &
        !methode_perID_Operation ~ "Naturelle + IUI + Médicaments",
      methode_perID_Naturelle &
        methode_perID_IUI &
        methode_perID_Operation &
        !methode_perID_IVF &
        !methode_perID_Medicament_sans_PMA ~ "Naturelle + IUI + Opération",
      methode_perID_Naturelle &
        methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        !methode_perID_IUI &
        !methode_perID_Operation ~ "Naturelle + IVF + Médicaments",
      methode_perID_Naturelle &
        methode_perID_IVF &
        methode_perID_Operation &
        !methode_perID_IUI &
        !methode_perID_Medicament_sans_PMA ~ "Naturelle + IVF + Opération",
      methode_perID_Naturelle &
        methode_perID_Medicament_sans_PMA &
        methode_perID_Operation &
        !methode_perID_IUI &
        !methode_perID_IVF ~ "Naturelle + Médicaments + Opération",
      methode_perID_IUI &
        methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        !methode_perID_Naturelle &
        !methode_perID_Operation ~ "IUI + IVF + Médicaments",
      methode_perID_IUI &
        methode_perID_IVF &
        methode_perID_Operation &
        !methode_perID_Naturelle &
        !methode_perID_Medicament_sans_PMA ~ "IUI + IVF + Opération",
      methode_perID_IUI &
        methode_perID_Medicament_sans_PMA &
        methode_perID_Operation &
        !methode_perID_Naturelle &
        !methode_perID_IVF ~ "IUI + Médicaments + Opération",
      methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        methode_perID_Operation &
        !methode_perID_Naturelle &
        !methode_perID_IUI ~ "IVF + Médicaments + Opération",
      
      # Four methods
      methode_perID_Naturelle &
        methode_perID_IUI &
        methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        !methode_perID_Operation ~ "Naturelle + IUI + IVF + Médicaments",
      methode_perID_Naturelle &
        methode_perID_IUI &
        methode_perID_IVF &
        methode_perID_Operation &
        !methode_perID_Medicament_sans_PMA ~ "Naturelle + IUI + IVF + Opération",
      methode_perID_Naturelle &
        methode_perID_IUI &
        methode_perID_Medicament_sans_PMA &
        methode_perID_Operation &
        !methode_perID_IVF ~ "Naturelle + IUI + Médicaments + Opération",
      methode_perID_Naturelle &
        methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        methode_perID_Operation &
        !methode_perID_IUI ~ "Naturelle + IVF + Médicaments + Opération",
      methode_perID_IUI &
        methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        methode_perID_Operation &
        !methode_perID_Naturelle ~ "IUI + IVF + Médicaments + Opération",
      
      # All methods
      methode_perID_Naturelle &
        methode_perID_IUI &
        methode_perID_IVF &
        methode_perID_Medicament_sans_PMA &
        methode_perID_Operation ~ "Toutes méthodes combinées",
      
      TRUE ~ NA_character_
    )
  )

#----------------------------------------------------------
# 2-3. Simpler categorical variable : Conception method
#----------------------------------------------------------
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(methode_conception_perID_cat_simple=
           case_when(
             methode_conception_perID_cat =="Naturelle seule" ~
               "Naturelle seule",
             methode_conception_perID_cat%in%c(
               "Médicaments sans PMA seule",
               "IVF seule", 
               "Naturelle + IVF", 
               "IUI seule",
               "Naturelle + IUI",
               "IUI + IVF",
               "Naturelle + IVF + Médicaments",
               "Naturelle + Médicaments"
             ) ~ "PMA/Medicaments",
             methode_conception_perID_cat%in% c(
               "Opération seule", 
               "Naturelle + Opération",
               "Naturelle + Médicaments + Opération"
             ) ~ "Operation",
             methode_conception_perID_cat%in% c(
               "IVF + Opération", 
               "IUI + Opération"
             ) ~ "Operation+PMA",
             TRUE ~ NA_character_
           ) %>% as.factor())

#--------------------------------------
#  3. Conception method per Pregnancy 
#--------------------------------------
# 1. Transform wide format to long format with IdPregnancy
new_dataframe1 <- grossesse_mod2 %>%
  dplyr::select(IdPregnancy, starts_with("Debut_")) %>% 
  pivot_longer(cols = starts_with("Debut_"), 
               values_drop_na = TRUE) %>%
  distinct(IdPregnancy, value) %>%
  mutate(
    methode_conception_long_perPreg = case_when(
      value %in% c("Conception naturelle") ~ "Naturelle",
      value %in% c("Médicaments pour stimuler l'ovulation (ex: Clomid,clomifène..)") ~ "Medicament_sans_PMA",
      value %in% c("Insémination artificielle") ~ "IUI",
      value %in% c("Fécondation in vitro") ~ "IVF",
      value %in% c("Opération / chirurgie") ~ "Operation",
      value %in% c("Autre") ~ "Autre",
      value %in% c("Je ne souhaite pas répondre", 
                   "Je ne m'en souviens plus")~ "nsp",
      TRUE ~ NA_character_
    )) %>%
  filter(!is.na(methode_conception_long_perPreg)) %>%
  distinct(IdPregnancy, methode_conception_long_perPreg)

# 2. Create a separate column for the "methode_conception" value
new_dataframe2 <- new_dataframe1 %>% 
  filter(!is.na(methode_conception_long_perPreg)) %>%
  mutate(value = TRUE) %>%  
  pivot_wider(
    names_from = methode_conception_long_perPreg,  
    values_from = value,  
    values_fill = list(value = FALSE),  
    names_prefix = "methode_perPreg_" 
  )

# 5. Merge the new variables into the main dataframe 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe1, by="IdPregnancy") %>% 
  left_join(new_dataframe2, by="IdPregnancy")

# 6. Delete temporary dataframes
rm(new_dataframe1); rm(new_dataframe2)

#-------------------------------------------------------
# 3-1. Conception method per Patient : binary(Yes/NO)
#-------------------------------------------------------
# Create binary variable for conception method
new_dataframe <- grossesse_mod2 %>%
  distinct(IdPregnancy, methode_conception_long_perPreg) %>%
  mutate(
    Methode_conception_bin_perPreg =
      ifelse(methode_conception_long_perPreg == "Naturelle", 0, 1)
  ) %>%
  group_by(IdPregnancy) %>%
  mutate(
    somme = sum(Methode_conception_bin_perPreg),
    Methode_conception_bin_perPreg = ifelse(somme > 0, "Oui", "Non")
  ) %>%
  ungroup() %>%
  distinct(IdPregnancy, Methode_conception_bin_perPreg)

# 2. Add the binary variable to the main dataframe
grossesse_mod2 <- grossesse_mod2 %>%
  left_join(new_dataframe, by="IdPregnancy")

# 3. Delete the temporary dataframe
rm(new_dataframe)

#-----------------------------------------------------------------------------------
# 3-2. Conception method per Pregnancy : Different combination of conception method
#-----------------------------------------------------------------------------------
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    methode_conception_perPreg_cat = case_when(
      # Unique method
      methode_perPreg_Naturelle == TRUE &
        !methode_perPreg_IUI &
        !methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "Naturelle seule",
      methode_perPreg_IUI == TRUE &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "IUI seule",
      methode_perPreg_IVF == TRUE &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IUI &
        !methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "IVF seule",
      methode_perPreg_Medicament_sans_PMA == TRUE &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IUI &
        !methode_perPreg_IVF &
        !methode_perPreg_Operation ~ "Médicaments sans PMA seule",
      methode_perPreg_Operation == TRUE &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IUI &
        !methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA ~ "Opération seule",
      
      # Two methods
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        !methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "Naturelle + IUI",
      methode_perPreg_Naturelle &
        methode_perPreg_IVF &
        !methode_perPreg_IUI &
        !methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "Naturelle + IVF",
      methode_perPreg_Naturelle &
        methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_IUI &
        !methode_perPreg_IVF &
        !methode_perPreg_Operation ~ "Naturelle + Médicaments",
      methode_perPreg_Naturelle &
        methode_perPreg_Operation &
        !methode_perPreg_IUI &
        !methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA ~ "Naturelle + Opération",
      methode_perPreg_IUI &
        methode_perPreg_IVF &
        !methode_perPreg_Naturelle &
        !methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "IUI + IVF",
      methode_perPreg_IUI &
        methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IVF &
        !methode_perPreg_Operation ~ "IUI + Médicaments",
      methode_perPreg_IUI &
        methode_perPreg_Operation &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA ~ "IUI + Opération",
      methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IUI &
        !methode_perPreg_Operation ~ "IVF + Médicaments",
      methode_perPreg_IVF &
        methode_perPreg_Operation &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IUI &
        !methode_perPreg_Medicament_sans_PMA ~ "IVF + Opération",
      methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IUI &
        !methode_perPreg_IVF ~ "Médicaments + Opération",
      
      # Three methods
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "Naturelle + IUI + IVF",
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_IVF &
        !methode_perPreg_Operation ~ "Naturelle + IUI + Médicaments",
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        methode_perPreg_Operation &
        !methode_perPreg_IVF &
        !methode_perPreg_Medicament_sans_PMA ~ "Naturelle + IUI + Opération",
      methode_perPreg_Naturelle &
        methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_IUI &
        !methode_perPreg_Operation ~ "Naturelle + IVF + Médicaments",
      methode_perPreg_Naturelle &
        methode_perPreg_IVF &
        methode_perPreg_Operation &
        !methode_perPreg_IUI &
        !methode_perPreg_Medicament_sans_PMA ~ "Naturelle + IVF + Opération",
      methode_perPreg_Naturelle &
        methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation &
        !methode_perPreg_IUI &
        !methode_perPreg_IVF ~ "Naturelle + Médicaments + Opération",
      methode_perPreg_IUI &
        methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Naturelle &
        !methode_perPreg_Operation ~ "IUI + IVF + Médicaments",
      methode_perPreg_IUI &
        methode_perPreg_IVF &
        methode_perPreg_Operation &
        !methode_perPreg_Naturelle &
        !methode_perPreg_Medicament_sans_PMA ~ "IUI + IVF + Opération",
      methode_perPreg_IUI &
        methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IVF ~ "IUI + Médicaments + Opération",
      methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation &
        !methode_perPreg_Naturelle &
        !methode_perPreg_IUI ~ "IVF + Médicaments + Opération",
      
      # Four methods
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        !methode_perPreg_Operation ~ "Naturelle + IUI + IVF + Médicaments",
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        methode_perPreg_IVF &
        methode_perPreg_Operation &
        !methode_perPreg_Medicament_sans_PMA ~ "Naturelle + IUI + IVF + Opération",
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation &
        !methode_perPreg_IVF ~ "Naturelle + IUI + Médicaments + Opération",
      methode_perPreg_Naturelle &
        methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation &
        !methode_perPreg_IUI ~ "Naturelle + IVF + Médicaments + Opération",
      methode_perPreg_IUI &
        methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation &
        !methode_perPreg_Naturelle ~ "IUI + IVF + Médicaments + Opération",
      
      # All methods
      methode_perPreg_Naturelle &
        methode_perPreg_IUI &
        methode_perPreg_IVF &
        methode_perPreg_Medicament_sans_PMA &
        methode_perPreg_Operation ~ "Toutes méthodes combinées",
      
      TRUE ~ NA_character_
    )
  )

#----------------------------------------------------------
# 3-3. Simpler categorical variable : Conception method
#----------------------------------------------------------
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(methode_conception_perPreg_cat_simple=
           case_when(
             methode_conception_perPreg_cat  =="Naturelle seule" ~
               "Naturelle seule",
             methode_conception_perPreg_cat %in%c(
               "Médicaments sans PMA seule",
               "IVF seule", 
               "Naturelle + IVF", 
               "IUI seule",
               "Naturelle + IUI",
               "IUI + IVF",
               "Naturelle + IVF + Médicaments",
               "Naturelle + Médicaments"
             ) ~ "PMA/Medicaments",
             methode_conception_perPreg_cat %in% c(
               "Opération seule", 
               "Naturelle + Opération",
               "Naturelle + Médicaments + Opération"
             ) ~ "Operation",
             methode_conception_perPreg_cat %in% c(
               "IVF + Opération", 
               "IUI + Opération"
             ) ~ "Operation+PMA",
             TRUE ~ NA_character_
           ) %>% as.factor())


#------------------------
#   Mode of delivery
#------------------------
# 1. Simplify values of two variables :
# - Binary mode of delivery : Voie basse and Césarienne 
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(Mode_accouchement= 
           case_when(
             accouchement2=="Voie basse/vaginale" ~ "Voie basse",
             accouchement2 %in%c("Césarienne d'urgence",
                                 "Césarienne programmée") ~"Césarienne",
             TRUE~ NA_character_
           )
  ) 

# 2. Make a temporary dataframe in order to create a separate column for the "mode_accouchement" value
new_dataframe <- grossesse_mod2 %>% 
  distinct(id_unik, Mode_accouchement) %>% 
  filter(!is.na(Mode_accouchement)) %>%
  mutate(value = TRUE) %>%  
  pivot_wider(
    names_from = Mode_accouchement, 
    values_from = value,  
    values_fill = list(value = FALSE),  
    names_prefix = "mode_" 
  )
# 2-1) Rename second column 
new_dataframe <- new_dataframe %>%
  rename(mode_Voie_basse = colnames(new_dataframe)[2])

# 3. Add these new variables to the main dataframe 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe, by="id_unik")

# 4. Delete the temporary dataframe 
rm(new_dataframe)

# 5. Create again a temporary dataframe 
new_dataframe <- grossesse_mod2 %>% 
  distinct(id_unik, Mode_accouchement,live_birth_bin_atleast_once_all) %>% 
  group_by(id_unik) %>% 
  mutate(
    Mode_accouchement_perID= 
           case_when(
             live_birth_bin_atleast_once_all=="Non" ~"No_delivery",
             all(is.na(Mode_accouchement)) ~NA_character_,
             all(Mode_accouchement=="Voie basse", na.rm=TRUE) ~ 
               "Seule voie Basse",
             all(Mode_accouchement=="Césarienne", na.rm=TRUE) ~
               "Seule Césarienne",
             any(Mode_accouchement=="Voie basse") &
               any(Mode_accouchement=="Césarienne") ~
               "Both",
             TRUE ~ NA_character_
             )
         )

# 6. Add the new varaible to the main dataframe
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe %>% distinct(id_unik, Mode_accouchement_perID),
            by="id_unik")

#--------------------------------------------
# Dummy variables for the mode of delivery 
#--------------------------------------------
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(
    VD_perID = case_when(
      Mode_accouchement_perID == "Seule voie Basse" ~ "Oui",
      Mode_accouchement_perID == "Both" ~ "Oui",
      Mode_accouchement_perID != "Seule voie Basse"&
        Mode_accouchement_perID != "Both" ~ "Non",
      Mode_accouchement_perID=="No_delivery" ~ "Non",
      TRUE ~NA_character_
    ),
    
    Csection_perID = case_when(
      Mode_accouchement_perID == "Seule Césarienne" ~ "Oui",
      Mode_accouchement_perID == "Both" ~ "Oui",
      Mode_accouchement_perID != "Seule Césarienne"&
        Mode_accouchement_perID != "Both" ~ "Non",
      Mode_accouchement_perID=="No_delivery" ~ "Non",
      TRUE ~NA_character_
    )
  )

#---------------------------
#    Labor of childbirth
#---------------------------
# 1. Simplify the values of variable 'travail2'
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(Travail_accouchement= 
           case_when(
             travail2=="Travail spontané"~ "Travail spontané",
             travail2=="Travail provoqué (déclenchement de l'accouchement par des médicaments et/ou ballonnet)"~ "Travail provoqué",
             travail2=="Pas de travail" ~ "Pas de travail",
             TRUE ~NA_character_
           )
  )

# 2. Make a temporary dataframe in order to create a separate column for the "Travail_accouchement" value
new_dataframe <- grossesse_mod2 %>% 
  distinct(id_unik, Travail_accouchement) %>% 
  filter(!is.na(Travail_accouchement)) %>%
  mutate(value = TRUE) %>%  
  pivot_wider(
    names_from = Travail_accouchement, 
    values_from = value,  
    values_fill = list(value = FALSE)
  )

# 2-1) Rename second column 
new_dataframe <- new_dataframe %>%
  rename(travail_provoque = colnames(new_dataframe)[2],
         travail_non= colnames(new_dataframe)[3],
         travail_spontane=colnames(new_dataframe)[4])  

# 3. Add these new variables to the main dataframe 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe, by="id_unik")

# 4. Delete the temporary dataframe 
rm(new_dataframe)

# 5. Create a new variable : Travail_accouchement_cat
grossesse_mod2 <- grossesse_mod2 %>%
  group_by(id_unik) %>%
  mutate(
    Travail_accouchement_perID = case_when(
      live_birth_bin_atleast_once_all == "Non" ~"No_delivery",
      all(is.na(Travail_accouchement)) ~ NA_character_,
      all(Travail_accouchement == "Travail provoqué", na.rm = TRUE) ~ 
        "Seul travail provoqué",
      
      any(Travail_accouchement == "Travail provoqué", na.rm = TRUE) &
        any(Travail_accouchement == "Pas de travail", na.rm = TRUE) &
        !any(Travail_accouchement == "Travail spontané", na.rm = TRUE) ~ 
        "Seul travail provoqué",
      
      all(Travail_accouchement == "Travail spontané", na.rm = TRUE) ~ 
        "Seul travail spontané",
      
      any(Travail_accouchement == "Travail spontané", na.rm = TRUE) &
        any(Travail_accouchement == "Pas de travail", na.rm = TRUE) &
        !any(Travail_accouchement == "Travail provoqué", na.rm = TRUE) ~ 
        "Seul travail spontané",
      
      all(Travail_accouchement == "Pas de travail", na.rm = TRUE) ~ 
        "Pas de travail",
      
      any(Travail_accouchement == "Travail provoqué", na.rm = TRUE) &
        any(Travail_accouchement == "Travail spontané", na.rm = TRUE) ~
        "Travail mixte",
      TRUE ~ NA_character_
    ) %>% as.factor()
  ) %>%
  ungroup()

#------------------------------------------------------------------------
# Create dummy variables for the labour information and mode of delivery 
#------------------------------------------------------------------------
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    labour_natural_perID = case_when(
      Travail_accouchement_perID == "No_delivery" ~"Non",
      Travail_accouchement_perID == "Seul travail spontané" ~ "Oui",
      Travail_accouchement_perID != "Seul travail spontané" ~ "Non",
      Travail_accouchement_perID == "Travail mixte" ~ "Oui",
      Travail_accouchement_perID == "Pas de travail" ~ "Non",
      TRUE ~ NA_character_
    ),
    labour_induced_perID = case_when(
      Travail_accouchement_perID == "No_delivery" ~"Non",
      Travail_accouchement_perID == "Seul travail provoqué" ~ "Oui",
      Travail_accouchement_perID != "Seul travail provoqué" ~ "Non",
      Travail_accouchement_perID == "Travail mixte" ~ "Oui",
      Travail_accouchement_perID == "Pas de travail" ~ "Non",
      TRUE ~ NA_character_
      )
    )

#--------------------------------------------
#    Number of live births per patient
#--------------------------------------------
# 1. Make a temporary dataframe for the total number of live births per patient
new_dataframe <- grossesse_mod2 %>% 
  distinct(id_unik, IdPregnancy, Vivant) %>% 
  group_by(id_unik) %>%  
  summarise(nb_nevivants_parpt=sum(as.numeric(Vivant), 
                                 na.rm = TRUE)) %>%
  ungroup()

# 2. Put a variable 'total number of babies per patient' to the main dataset 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe, by="id_unik")

# 3. Delete a temporary dataframe 
rm(new_dataframe)

#-------------------------------------
# Frequency of patients who breastfed
#--------------------------------------
# 1. Create a temporary dataframe to make new variables(continue et binaire)
new_dataframe <- grossesse_mod2 %>%
  distinct(id_unik, Allaitement2) %>%
  group_by(id_unik) %>%
  mutate(
    allaitement_YN_atleast =
      case_when(
        any(Allaitement2 == "Oui", na.rm = TRUE) ~ "Oui",
        all(is.na(Allaitement2)) ~ NA_character_,
        any(Allaitement2 == "Non", na.rm = TRUE) ~ "Non",
        TRUE ~ NA_character_
      )
  ) %>%
  mutate(
    nb_allaitement = sum(Allaitement2 == "Oui", na.rm = TRUE),
    nb_allaitement = ifelse(is.na(allaitement_YN_atleast), NA, nb_allaitement)
  ) %>%
  ungroup() %>%
  distinct(id_unik, allaitement_YN_atleast, nb_allaitement)

# 2. Add new variables to the main dataset and remove an original variable(Allatiement2)
grossesse_mod2 <- grossesse_mod2 %>%
  left_join(new_dataframe, by="id_unik") #%>% 
  #select(-Allaitement2)

# 3. Delete the temporary dataframe
rm(new_dataframe)

#----------------------------------
# Average duration of breastfeeding
#----------------------------------
# 1. mois_allaitement : character to numeric
grossesse_mod2 <- grossesse_mod2 %>% 
  mutate(mois_allaitement= as.numeric(mois_allaitement))

# 2.  Correct an outlier : 120 mois -> 120/30jours = 4 mois
grossesse_mod2 <- grossesse_mod2 %>% mutate(
  mois_allaitement=case_when(
    mois_allaitement==120 ~ 4,
    mois_allaitement!= 120 ~ mois_allaitement,
    allaitement_YN_atleast=="Non" ~ 0,
    allaitement_YN_atleast=="Oui" ~ mois_allaitement,
    TRUE ~ NA_real_
  )
)
# ########
# # INFO #
# ########
# # No information of breast feeding even if participants delivered live births : 54 pregnancies 
# grossesse_mod2 %>%
#   distinct(id_unik,IdPregnancy,terme2,Allaitement2,nb_allaitement, allaitement_YN_atleast, mois_allaitement) %>% filter(terme2=="Oui"&is.na(Allaitement2)) 
# 
# ###### QUESTION : Is it possible to do the imputation ? for theses NA values
# 
# 
# # Terme & allaitement_YN_atleast& mois_allaitement
# grossesse_mod2 %>% distinct(id_unik,IdPregnancy, terme2, nb_allaitement,allaitement_YN_atleast, mois_allaitement) %>% filter(terme2=="Oui"&allaitement_YN_atleast=="Oui"&is.na(mois_allaitement))

#------------------------------------------------------------------
#   Pregnancy management               : 83 respondents
# among patients who experienced at least one incomplete pregnancy
#
# Terme : No pregnancy management 
# Incomplete : Mode_accouchement is always NA
#------------------------------------------------------------------
# 1. Simplify the categories of all variables "PriseCharge_" and modify certain categories according to PriseChargeAutre 
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(across(starts_with("PriseCharge_"), 
                ~ case_when(
                  . == "Autre" & PriseChargeAutre == "methotrexate" ~ 
                    "Médicamenteuse",
                  . == "Médicamenteuse (comprimés oraux et/ou vaginaux)" ~
                    "Médicamenteuse",
                  . == "Autre" & substr(PriseChargeAutre, 1, 6) == "Coelio" ~
                    "Chirurgicale",
                  . == "Chirurgicale (curetage/aspiration)" ~ 
                    "Chirurgicale",
                  . %in% c("Prise en charge psychologique",
                           "Absence de prise en charge médicale ou chirurgicale",
                           "Je ne sais pas", "Autre") ~ .,
                  TRUE ~ NA_character_
                )))  %>% 
  # From PriseCharge_2_2, remove the category "Absence de prise en charge médicale ou chirurgicale" because specific PriseCharge categories exist in PriseCharge_1_2, making the absence category inappropriate
  mutate(across(setdiff(names(dplyr::select(., starts_with("PriseCharge_"))), "PriseCharge_1_2"),~ ifelse(grepl("^Absence", .), NA_character_, .)))

# 2. Create a variable in long format
# 2-1) Create a temporary dataframe for the new variable in long format
new_dataframe <- grossesse_mod2 %>%
  dplyr::select(id_unik, starts_with("PriseCharge_")) %>%
  pivot_longer(
    cols = starts_with("PriseCharge_"), 
    values_to = "PriseenCharge_long",
    values_drop_na = TRUE
  ) %>% 
  distinct(id_unik, PriseenCharge_long)  


# 2-2) Create a new binary variable : Prise en charge Oui/Non
new_dataframe <- new_dataframe %>% 
  group_by(id_unik) %>% 
  mutate(
    PriseenCharge_bin=
      case_when(
        all(PriseenCharge_long==
              "Absence de prise en charge médicale ou chirurgicale") ~"Non",
        TRUE ~ "Oui")
             )

# 2-3) Create another temporary dataframe for a new summarized variable
new_dataframe1 <- new_dataframe %>%
  filter(PriseenCharge_long != "Absence de prise en charge médicale ou chirurgicale") %>%
  group_by(id_unik, PriseenCharge_long) %>% 
  summarise(n=n(),.groups = "drop") %>%  
  group_by(id_unik) %>% 
  summarise(nb_PriseenCharge_perID=sum(n))

# 2-4) Add this new variable to the first temporary dataframe
new_dataframe <- new_dataframe %>%
  left_join(new_dataframe1, by = "id_unik") %>%
  mutate(across(where(is.numeric),~ replace_na(., 0))
  )

# 2-5) Create a new categorical variable per patient 
method_labels <- c(
  "Chirurgicale" = "Seule_Chirurgicale",
  "Médicamenteuse" = "Seule_Médicamenteuse",
  "Prise en charge psychologique" = "Seule_Psychologique",
  "Autre" = "Autre/nsp",
  "Je ne sais pas" = "Autre/nsp"
)
new_dataframe <- new_dataframe %>%
  mutate(
    PriseenCharge_cat_perID = case_when(
      nb_PriseenCharge_perID == 0 ~"Absence",
      nb_PriseenCharge_perID >= 2 ~ "Multiple",
      nb_PriseenCharge_perID == 1 ~ method_labels[PriseenCharge_long],
      TRUE ~ NA_character_
    )
  ) %>% 
  filter(!(PriseenCharge_long == "Absence de prise en charge médicale ou chirurgicale" & nb_PriseenCharge_perID >= 1)) %>% distinct()

# 2-6) Add all theses new variables to the main dataframe 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(new_dataframe, by="id_unik")
  
# 2-7) Delete temporary dataframes
rm(new_dataframe); rm(new_dataframe1)

# 3. Create a variable in wide format
# 3-1) Create a temporary dataframe
# : Check whether each patient has received a specific management at least once
# : 355 patients who experienced at least one incomplete pregnancy
new_dataframe <- grossesse_mod2 %>% 
  filter(status_terme_all=="Incomplete") %>% 
  mutate(any_chirurgicale = if_any(starts_with("PriseCharge_"), ~ . == 
                                     "Chirurgicale"),
         any_medicamenteuse = if_any(starts_with("PriseCharge_"), ~ . == 
                                       "Médicamenteuse"),
         any_psychologique = if_any(starts_with("PriseCharge_"), ~. == 
                                      "Prise en charge psychologique")
  ) %>% 
  group_by(id_unik) %>%
  summarise(
    any_chirurgicale = any(any_chirurgicale, na.rm = TRUE),
    any_medicamenteuse = any(any_medicamenteuse, na.rm = TRUE),
    any_psychologique = any(any_psychologique, na.rm = TRUE)
  ) %>%
  ungroup()


# 3-2) Add new variables to the main dataframe 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, starts_with("any")),
            by="id_unik")

# 4. Delete original variables (PriseCharge) and the temporary dataframe
grossesse_mod2 <- grossesse_mod2 %>% 
  dplyr::select(-starts_with("PriseCharge"))
rm(new_dataframe)

#-----------------------------------------
#      Care taken per pregnancy 
#-----------------------------------------
# 1. Create a temporary dataframe by selecting necessary variables 
new_dataframe <- grossesse_mod2 %>%
  distinct(IdPregnancy, PriseenCharge_long) 

# 2. Count the number of care taken per pregnancy
new_dataframe1 <- new_dataframe %>% 
  filter(PriseenCharge_long != "Absence de prise en charge médicale ou chirurgicale") %>%
  group_by(IdPregnancy, PriseenCharge_long) %>%
  summarise(n=n(), .groups="drop") %>%
  group_by(IdPregnancy) %>%
  summarise(nb_PriseenCharge_perPreg = sum(n), .groups="drop")

# 3. Make a list of labels : Seule/Multiple/Absence/Autre
method_labels <- c(
  "Chirurgicale" = "Seule_Chirurgicale",
  "Médicamenteuse" = "Seule_Médicamenteuse",
  "Prise en charge psychologique" = "Seule_Psychologique",
  "Autre" = "Autre/nsp",
  "Je ne sais pas" = "Autre/nsp"
)

# 4. Create a new variable 
new_dataframe <- new_dataframe %>%
  group_by(IdPregnancy) %>%
  mutate(n_distinct_types = n_distinct(PriseenCharge_long)) %>%
  left_join(new_dataframe1, by = "IdPregnancy") %>%
  mutate(nb_PriseenCharge_perPreg = replace_na(nb_PriseenCharge_perPreg, 0)) %>%
  mutate(
    PriseenCharge_cat_perPreg = case_when(
      nb_PriseenCharge_perPreg == 0 &
        all(PriseenCharge_long == "Absence de prise en charge médicale ou chirurgicale") ~ "Absence",
      nb_PriseenCharge_perPreg == 1 ~ method_labels[PriseenCharge_long],
      nb_PriseenCharge_perPreg >= 2 ~ "Multiple",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::select(IdPregnancy, PriseenCharge_cat_perPreg) %>%
  distinct()

# 5. Add the new variable to the main dataframe
grossesse_mod2 <- grossesse_mod2 %>%
  left_join(new_dataframe, by = "IdPregnancy")

# 6. Remove all temporary dataframes 
rm(new_dataframe); rm(new_dataframe1)

###########################################################
#--------------------------------------------------
#   Number of pregnancy complications per pregnancy  :
#--------------------------------------------------
ttest <- grossesse_mod2

# 1. Create a list of columns' name
complication_vars <- paste0("Complication_", 1:6, "_2")

# 2. Complication_value
new_dataframe <- grossesse_mod2 %>%
  dplyr::select(id_unik,IdPregnancy, starts_with("Complication_")) %>%
  pivot_longer(
    cols = all_of(complication_vars),
    names_to = "complication_type",
    values_to = "Complication_value_perPreg"
  ) %>% 
  filter(!is.na(Complication_value_perPreg)) %>% 
  distinct(id_unik,IdPregnancy, Complication_value_perPreg)

# 3. Create complications per pregnancy
new_dataframe1 <- new_dataframe %>%
  filter(Complication_value_perPreg != "Aucune complication") %>% 
  distinct(IdPregnancy, Complication_value_perPreg) %>%            
  group_by(IdPregnancy) %>%                                       
  mutate(total_complications_perPreg = n()) %>% 
  ungroup() %>% 
  right_join(new_dataframe %>% 
               group_by(IdPregnancy) %>% 
               summarise(nb_lignes_complic= n())) %>%
  mutate(
    total_complications_perPreg= case_when(
      nb_lignes_complic==1 & is.na(total_complications_perPreg) ~ 0,
      TRUE ~ total_complications_perPreg
      ),
    Complication_value_perPreg = case_when(
      nb_lignes_complic==1 & is.na(Complication_value_perPreg) ~ "Aucune complication", 
      TRUE ~ Complication_value_perPreg
    ),
    total_complications_cat_perPreg = case_when(
      total_complications_perPreg == 0 ~ "0",
      total_complications_perPreg == 1 ~ "1",
      total_complications_perPreg >= 2 ~ "2+"
      ) %>% factor(levels= c("0","1","2+"), ordered = TRUE),
    total_complications_bin_perPreg = 
      ifelse(total_complications_perPreg == 0, "Non", "Oui")
    ) %>% 
  dplyr::select(-nb_lignes_complic)

# 4. Merge 
merged_new_dataframe <- new_dataframe %>% 
  left_join(new_dataframe1 %>%
              dplyr::select(-Complication_value_perPreg) %>% 
              distinct(),
            by="IdPregnancy")

# 5. Add the new variables to the main dataframe 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(merged_new_dataframe %>% dplyr::select(-id_unik), 
            by="IdPregnancy")


#----------------------------------------------
#  Create individual complication type variable 
#----------------------------------------------
# 5. Create a new variable by using the merged_new_dataframe 
merged_new_dataframe2 <- merged_new_dataframe %>% 
  group_by(id_unik) %>% 
  summarise(
    complications = unique(Complication_value_perPreg),
    .groups = "drop"
  ) %>%
  filter(complications != "Aucune complication") %>%
  group_by(id_unik) %>%
  summarise(
    total_complications_perID = n_distinct(complications),
    .groups = "drop"
  ) 

# 6. Make the binary and ordinal variables
merged_new_dataframe3 <- merged_new_dataframe %>% 
  left_join(merged_new_dataframe2, by="id_unik") %>% 
  mutate(
    total_complications_perID=
      ifelse(is.na(total_complications_perID),0,total_complications_perID),
    # Binary
    total_complications_bin_perID = 
      ifelse(total_complications_perID == 0, "Non","Oui"),
    # Ordinal
    total_complications_cat_perID =
      case_when(
        total_complications_perID == 0 ~ "0",
        total_complications_perID == 1 ~ "1",
        total_complications_perID >= 2 ~ "2+"
        ) %>% 
      factor(levels= c("0","1","2+"), ordered = TRUE)
    )

# 7. Add new variables to the main dataframe 
grossesse_mod2 <- grossesse_mod2 %>% 
  left_join(merged_new_dataframe3  %>% 
              dplyr::select(id_unik, total_complications_perID, 
                            total_complications_bin_perID,
                            total_complications_cat_perID), 
            by="id_unik")

# 9. remove temporary dataframes 
rm(new_dataframe); rm(new_dataframe1);rm(merged_new_dataframe);rm(merged_new_dataframe2); rm(merged_new_dataframe3)

#-------------------------------------------------
#  Create individual complication type variable 
#-------------------------------------------------
# 1. Create a temporary dataframe to generate new variables related to obstetrical complications in wide format
new_dataframe <-  grossesse_mod2 %>% 
  distinct(id_unik, Complication_value_perPreg) %>% 
  mutate(Complication_value_new=
           case_when(
             Complication_value_perPreg%in% c("Hémorragie post-partum") ~
               "Hemorr_post_partum",
             Complication_value_perPreg%in% c("Menace d'accouchement prématuré (ouverture du col, contractionsprématurées, rupture prématurée de la poche des eaux...") ~
               "Menace_accouchement_premature",
             Complication_value_perPreg%in% c("Diabète gestationnel") ~
               "Diabete_gestationnel",
             Complication_value_perPreg%in% c("Pré-éclampsie / Toxémie gravidique, HELLP Syndrome, Eclampsie") ~
               "Preeclampsie_HELLEP",
             Complication_value_perPreg%in% c("Hypertension liée à la grossesse") ~
               "Hypertension_gross",
             Complication_value_perPreg%in% c("Décollement du placenta") ~
               "Decollement_placenta",
             Complication_value_perPreg%in% c("Placenta praevia (localisation basse du placenta de l'utérus") ~
               "Placenta_praevia",
             Complication_value_perPreg%in% c("Vomissements sévères lors de la grossesse (Hypermérèse gravidique)") ~
               "Vomissements_severes",
             Complication_value_perPreg%in% c("Cholestase gravidique") ~
               "Cholestase_gravidique",
             Complication_value_perPreg%in% c("Rupture d'un kyste endométriosique en cours de grossesse") ~
               "Rupture_kyste_endometriosique",
             Complication_value_perPreg%in% c("Rupture utérine") ~
               "Rupture_uterine",
             Complication_value_perPreg%in% c("Autre") ~ 
               "Autre",
             Complication_value_perPreg%in% c("Aucune complication") ~
               "Aucune_complication",
             TRUE ~ NA_character_
           )) %>% 
  filter(!is.na(Complication_value_new)) %>% 
  mutate(value=TRUE) %>% 
  pivot_wider(
    names_from = Complication_value_new,
    values_from = value,
    values_fill = list(value=FALSE),
    names_prefix = "complic_"
  ) 

# 2. If at least one complication is TRUE, it will ultimately remain TRUE
new_dataframe2 <- new_dataframe %>% 
  group_by(id_unik) %>% 
  summarise(across(starts_with("complic_"), ~ any(.))) %>% ungroup()

# 3. Add new variables to the main dataframe 
grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe2,by="id_unik") %>% 
  dplyr::select(-c(Complication_1_2:Complication_6_2))

# 4. Delete temporary dataframes 
rm(new_dataframe);rm(new_dataframe2)

#--------------------------------
#      Complications per ID
#--------------------------------
new_dataframe <-  grossesse_mod2 %>%
  distinct(id_unik, Complication_value_perPreg, total_complications_perID) %>% 
  filter(!is.na(Complication_value_perPreg))

#  Create a new categorical variable per patient
complications_labels <- c(
  "Hémorragie post-partum" = "Hemorr_post_partum",
  "Menace d'accouchement prématuré (ouverture du col, contractionsprématurées, rupture prématurée de la poche des eaux..." =
    "Menace_accouchement_premature",
  "Diabète gestationnel" = "Diabete_gestationnel",
  "Pré-éclampsie / Toxémie gravidique, HELLP Syndrome, Eclampsie" =
    "Preeclampsie_HELLEP",
  "Hypertension liée à la grossesse" = "Hypertension_gross",
  "Décollement du placenta" = "Decollement_placenta",
  "Placenta praevia (localisation basse du placenta de l'utérus" =
    "Placenta_praevia",
  "Vomissements sévères lors de la grossesse (Hypermérèse gravidique)"
  = "Vomissements_severes",
  "Cholestase gravidique" = "Cholestase_gravidique",
  "Rupture d'un kyste endométriosique en cours de grossesse" =
    "Rupture_kyste_endometriosique",
  "Rupture utérine" = "Rupture_uterine",
  "Autre" ="Autre",
  "Aucune complication" = "Aucune_complication"
)

new_dataframe <- new_dataframe %>%
  mutate(
    Complications_cat_perID = case_when(
      total_complications_perID == 0 & 
        Complication_value_perPreg== "Aucune complication"~
        "Aucune complication",
      total_complications_perID >= 2 ~ 
        "Multiple",
      total_complications_perID == 1 ~
        complications_labels[Complication_value_perPreg],
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!(Complication_value_perPreg == "Aucune complication" & 
             total_complications_perID >= 1)) %>% distinct()

# 2-6) Add all theses new variables to the main dataframe
grossesse_mod2 <- grossesse_mod2 %>%
  left_join(new_dataframe %>% distinct(id_unik, Complications_cat_perID),
            by="id_unik")

# 2-7) Delete temporary dataframes
rm(new_dataframe); rm(new_dataframe1)

#--------------------------------
#     Complications per preg
#--------------------------------
# 1. Create a temporary dataframe
new_dataframe <-  grossesse_mod2 %>%
  distinct(IdPregnancy, Complication_value_perPreg, total_complications_perPreg) %>% 
  filter(!is.na(Complication_value_perPreg))

# # 2. Count the number of complications without 'Aucune complication'
# new_dataframe1 <- new_dataframe %>%
#   filter(Complication_value != "Aucune complication") %>%
#   group_by(IdPregnancy, Complication_value) %>%
#   summarise(n=n(), .groups = "drop") %>%
#   group_by(IdPregnancy) %>%
#   summarise(nb_complications_perPreg = sum(n))
# 
# # 3. Add count information
# new_dataframe <- new_dataframe %>%
#   left_join(new_dataframe1, by = "IdPregnancy") %>%
#   mutate(across(where(is.numeric),  ~ replace_na(., 0)))

# 4. Create a new variable
new_dataframe <- new_dataframe %>%
  mutate(
    Complications_cat_perPreg = case_when(
      total_complications_perPreg == 0 &
        Complication_value_perPreg == "Aucune complication" ~
        "Aucune complication",
      total_complications_perPreg >= 2 ~
        "Multiple",
      total_complications_perPreg == 1 ~
        complications_labels[Complication_value_perPreg],
      TRUE ~ NA_character_
    )
  ) %>% 
  distinct(IdPregnancy, Complications_cat_perPreg)

# 5. Add this new variable to the main dataframe
grossesse_mod2 <- grossesse_mod2 %>%
  left_join(new_dataframe, by = "IdPregnancy")

#---------------------------------------------------
#  Regroup Obstetric Complications in 6 modalities  
#---------------------------------------------------
grossesse_mod2 <- grossesse_mod2 %>%
  mutate(
    Complications_cat_perID_grouped =
     case_when(
        Complications_cat_perID == "Aucune complication" ~
          "No_complication",
        Complications_cat_perID %in%
          c(
            "Diabete_gestationnel",
            "Hypertension_gross",
            "Preeclampsie_HELLEP",
            "Cholestase_gravidique"
          ) ~
          "Metabolic_Hypertensive_single",
        Complications_cat_perID %in%
          c(
            "Menace_accouchement_premature",
            "Decollement_placenta",
            "Placenta_praevia"
          ) ~
          "Preterm_related_single",
        Complications_cat_perID == "Hemorr_post_partum" ~
          "Hemorr_postpartum_single",
        Complications_cat_perID %in%
          c(
            "Rupture_kyste_endometriosique",
            "Vomissements_severes",
            "Autre"
          ) ~
          "Other_single",
        Complications_cat_perID == "Multiple" ~
          "Multiple_complications",
        TRUE ~ NA_character_
      ) %>% 
      as.factor()
    
      ) %>% 
  group_by(IdPregnancy) %>% 
  mutate(
    Complications_cat_perPreg_grouped =
      case_when(
        Complications_cat_perPreg == "Aucune complication" ~
          "No_complication",
        Complications_cat_perPreg %in%
          c(
            "Diabete_gestationnel",
            "Hypertension_gross",
            "Preeclampsie_HELLEP",
            "Cholestase_gravidique"
          ) ~
          "Metabolic_Hypertensive_single",
        Complications_cat_perPreg %in%
          c(
            "Menace_accouchement_premature",
            "Decollement_placenta",
            "Placenta_praevia"
          ) ~
          "Preterm_related_single",
        Complications_cat_perPreg == "Hemorr_post_partum" ~
          "Hemorr_postpartum_single",
        Complications_cat_perPreg %in%
          c(
            "Rupture_kyste_endometriosique",
            "Vomissements_severes",
            "Autre"
          ) ~
          "Other_single",
        Complications_cat_perPreg == "Multiple" ~
          "Multiple_complications",
        TRUE ~ NA_character_
      ) %>% 
      as.factor()
    ) %>% ungroup %>%  
  group_by(id_unik, IdPregnancy) %>% 
  filter(!(!is.na(Complication_value_perPreg) &
           total_complications_perPreg==1&
           is.na(Complications_cat_perPreg_grouped))) %>% 
  ungroup()



#------------------------------
#  G/P : Gravidity and Parity  
#------------------------------
new_dataframe <- grossesse_mod2 %>%
  arrange(id_unik, age_maternal) %>%
  distinct(id_unik, IdPregnancy, age_maternal, live_birth_bin_all) %>%
  group_by(id_unik) %>%
  mutate(
    gravida = row_number(),
    para = lag(cumsum(live_birth_bin_all == "Oui"), default = 0)
  ) %>%
  ungroup()

grossesse_mod2 <-  grossesse_mod2 %>% 
  left_join(new_dataframe %>% distinct(IdPregnancy, gravida, para),
            by="IdPregnancy")

```

# II. Final dataset

```{r}
#----------------------------------------------------------------
#    FINAL Dataframe for the population without ongoing pregnancy 
#-----------------------------------------------------------------
# 1. Select necessary variables for the final dataset 
sansgrossesseencours_final <-  grossesse_mod2 %>% 
  dplyr::select(
    id_unik,
    IdPregnancy,
    merged_DateDebut,
    DateFin,
    module,

    # Live birth per pregnancy : Oui/Non
    live_birth_bin_all, 
    live_birth_bin_onlynew,

    # Terme or incomplete per pregnancy
    status_terme_all,
    status_terme_onlynew,
    #status_terme_onlyold,

    # Type de terme et incomplete : fullterm, preterm, F-C, IVG, etc
    type_preg_all,
    type_preg_all_simple,
    type_preg_onlynew,
    type_preg_onlynew_simple,

    # Live birth at least once per patient : Oui/Non
    live_birth_bin_atleast_once_all,
    live_birth_bin_atleast_once_onlynew,

    # Number of Live births per patient
    nb_live_births_all,
    nb_live_births_onlynew,

    # Number of incomplet pregnancies per patient
    nb_incomplete_preg_pt_all,
    nb_incomplete_preg_pt_onlynew,

    # Number of grossesses enregistrées per patient
    nb_total_gross,
    nb_total_gross_NewModule,
    nb_total_gross_OldModule,

    # Maternal age at the end of pregnancy
    age_maternal, # per pregnancy
    first_preg_age,  # Per patient
    last_preg_age_mod, # per patient
    
    amenorrhea_final,

    # Conception method : Per patient and Per pregnancy
    Methode_conception_bin_perID,
    methode_conception_perID_cat,
    methode_conception_perID_cat_simple,
    
    Methode_conception_bin_perPreg,
    methode_conception_perPreg_cat,
    methode_conception_perPreg_cat_simple,
    methode_perID_IVF, 
    methode_perID_IUI,
    
    # Mode of delivery and labor
    Mode_accouchement, # per pregnancy
    Mode_accouchement_perID, # per patient
    VD_perID,
    Csection_perID,
    Travail_accouchement, # per pregnancy
    Travail_accouchement_perID,  # per patient
    labour_natural_perID,
    labour_induced_perID,
    
    # Number of né vivants per patient
    nb_nevivants_parpt,
    PoidInf2, # per pregnancy
    PoidSup2, # per pregnancy

    # Number of breastfeeding per patient and Breasfeeding at least once
    nb_allaitement, # per patient
    allaitement_YN_atleast, # per patient
    mois_allaitement, # per pregnancy
    Allaitement2, # per pregnancy 

    # Care taken or not per patient et per pregnancy
    PriseenCharge_cat_perID,
    PriseenCharge_cat_perPreg,
    PriseenCharge_bin, # per patient
    
    # Obstetrical complications 
    total_complications_perPreg:total_complications_bin_perPreg,
    total_complications_perID:total_complications_cat_perID,
    Complications_cat_perID:Complications_cat_perPreg_grouped,
    complic_Aucune_complication:complic_Rupture_kyste_endometriosique,
    
    # G/P : Gravidity and Parity  
    para,
    gravida,

    Date_naissance,
    #DateCreation,
    DateModification
  ) %>% 
  distinct() %>% 
  mutate(age_new = 
           as.character(round((DateModification - Date_naissance) /365, 1
                              )) %>% as.numeric()
         )
```

## > Add the new modalities

```{r}
# #----------------------------------------------------------
# # Add the new cateogry : Not collected and Not relevant
# #---------------------------------------------------------
# sansgrossesseencours_final_test <- sansgrossesseencours_final %>%
#   filter(module=="Nouveau") %>% 
#   mutate(
#     across(
#       # If the participants never had a live birth, the mode of delivery, the induction and the obstetric complications were not registered
#       c(Mode_accouchement,
#         Travail_accouchement,
#         total_complications_bin_perPreg, 
#         total_complications_cat_perPreg),
#                 ~ case_when(
#                   status_terme_onlynew == "Incomplete" ~ "Only_incomplete_preg",
#                   TRUE ~ as.character(.)
#                 )),
#     # # If the module is old, the data is not collected : Conception method, medical care, mode of delivery, induction, complication
#     # across(
#     #   c(starts_with(c("methode", "Prise", "mode", "travail", "complic")),
#     #     total_complications_cat_perPreg,
#     #     total_complications_bin_perPreg,
#     #     
#     #     -Methode_conception_bin_perID,
#     #     -methode_conception_perID_cat,
#     #     -methode_conception_perID_cat_simple,
#     #     -PriseenCharge_cat_perID,
#     #     -Travail_accouchement_perID 
#     #   ),
#     #   ~ case_when(
#     #     module=="Ancien" ~ "Not_collected",
#     #     TRUE~ as.character(.))),
# 
#     Travail_accouchement_perID_test=
#       case_when(
#         !is.na(Travail_accouchement_perID) ~
#           Travail_accouchement_perID,
#         is.na(Travail_accouchement_perID)& nb_total_gross==nb_incomplete_preg_pt_all ~
#           "Not_applicable",
#         is.na(Travail_accouchement_perID)& nb_total_gross>nb_incomplete_preg_pt_all ~
#           "Not_collected",
#         TRUE ~NA_character_),
#     
#     Mode_accouchement_perID_test=
#       case_when(
#         !is.na(Mode_accouchement_perID) ~
#           Mode_accouchement_perID,
#         is.na(Mode_accouchement_perID)& nb_total_gross==nb_incomplete_preg_pt_all ~
#           "Not_applicable",
#         is.na(Mode_accouchement_perID)&nb_total_gross>nb_incomplete_preg_pt_all ~
#           "Not_collected",
#         TRUE ~NA_character_
#       ),
#     total_complications_bin_perID_test =
#       case_when(
#         !is.na(total_complications_bin_perID) ~
#           total_complications_bin_perID,
#         is.na(total_complications_bin_perID) & nb_total_gross==nb_incomplete_preg_pt_all ~
#           "Not_applicable",
#         is.na(total_complications_bin_perID ) & nb_total_gross>nb_incomplete_preg_pt_all &
#           !is.na(total_complications_bin_perPreg) ~
#           "Not_collected",
#         is.na(total_complications_bin_perID )&nb_total_gross>nb_incomplete_preg_pt_all &
#           is.na(total_complications_bin_perPreg) ~
#           NA_character_,
#         TRUE ~NA_character_
#       )
#     )
```


# III. Save the final dataframe

```{r}
# update : 2025-06-16
write.csv(sansgrossesseencours_final,"sansgrossesseencours_final.csv")
```



# DESCRIPTIONS 
- les taux de grossesse, global et à terme, par femme seront calculés
pour chaque type de méthode de conception (naturelle, médicament contre l’infertilité,insémination artificielle, FIV, chirurgie, autre) en divisant le nombre de grossesses obtenues avec chaque méthode par le nombre total de grossesses ou le nombre de grossesses à terme,respectivement ; 
- des catégories « multiples » seront créées selon la distribution des différentes méthodes de conception 
