---
title: "01_clean_Neuropathic_pain"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Neuropathic_pain
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
doul_neuro <- read.csv(
  "Complic_pop_QST_0508_20250407_doul_neuro.csv",
  fileEncoding = "ISO-8859-1"  # similaire to latin1
)

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
doul_neuro <- doul_neuro%>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim),
         across(starts_with("Date"),
                ~ as.Date(.,format="%d/%m/%Y")))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- doul_neuro %>% 
  sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count == nrow(doul_neuro)) %>%
  as.vector() # index of columns having only NA
doul_neuro <- doul_neuro[,-index_col_NA]
```

# I. Preprocessing Data

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
doul_neuro <- doul_neuro %>% 
  mutate(doul_neuro_participated = ifelse(is.na(IdAction),
                                          "Non","Oui"))

#-----------------------------------
#  Neuropathic pain in binary data
#-----------------------------------
doul_neuro <- doul_neuro %>% 
  mutate(doul_neuro_bin= douleur_mois %>% as.factor())

# Score PainDETECT
#-------------------
#  First main pain
#-------------------
doul_neuro <- doul_neuro %>%
  mutate(
    score1_test =
      case_when(
        deroulement_douleur == "Douleur constante avec de légères fluctuations" ~ 0,
        deroulement_douleur == "Douleur constante ponctuée de pics de douleur" ~ -1,
        deroulement_douleur %in% c(
          "Pics de douleur sans douleur de fond",
          "Fréquents pics de douleur avec douleur de fond"
        ) ~ 1,
        deroulement_douleur == "Aucune case ne correspond" ~ 0,
        TRUE ~ NA_real_
      ) +
      if_else(propagation_douleur == "Oui", 2, 0) +
      case_when(
        sensation_brulure == "Pas du tout" ~ 0,
        sensation_brulure == "Très peu" ~ 1,
        sensation_brulure == "Légèrement" ~ 2,
        sensation_brulure == "Modérément" ~ 3,
        sensation_brulure == "Fortement" ~ 4,
        sensation_brulure == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        picotement == "Pas du tout" ~ 0,
        picotement == "Très peu" ~ 1,
        picotement == "Légèrement" ~ 2,
        picotement == "Modérément" ~ 3,
        picotement == "Fortement" ~ 4,
        picotement == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        leger_contact_douloureux == "Pas du tout" ~ 0,
        leger_contact_douloureux == "Très peu" ~ 1,
        leger_contact_douloureux == "Légèrement" ~ 2,
        leger_contact_douloureux == "Modérément" ~ 3,
        leger_contact_douloureux == "Fortement" ~ 4,
        leger_contact_douloureux == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        pic_douleur == "Pas du tout" ~ 0,
        pic_douleur == "Très peu" ~ 1,
        pic_douleur == "Légèrement" ~ 2,
        pic_douleur == "Modérément" ~ 3,
        pic_douleur == "Fortement" ~ 4,
        pic_douleur == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        temperature_douleur == "Pas du tout" ~ 0,
        temperature_douleur == "Très peu" ~ 1,
        temperature_douleur == "Légèrement" ~ 2,
        temperature_douleur == "Modérément" ~ 3,
        temperature_douleur == "Fortement" ~ 4,
        temperature_douleur == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        sensibilite_diminue == "Pas du tout" ~ 0,
        sensibilite_diminue == "Très peu" ~ 1,
        sensibilite_diminue == "Légèrement" ~ 2,
        sensibilite_diminue == "Modérément" ~ 3,
        sensibilite_diminue == "Fortement" ~ 4,
        sensibilite_diminue == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        pression_douleur == "Pas du tout" ~ 0,
        pression_douleur == "Très peu" ~ 1,
        pression_douleur == "Légèrement" ~ 2,
        pression_douleur == "Modérément" ~ 3,
        pression_douleur == "Fortement" ~ 4,
        pression_douleur == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      )
  )

#---------------------
#  Second main pain
#---------------------
doul_neuro <- doul_neuro %>%
  mutate(
    score2_test =
      case_when(
        deroulement_douleur_second == "Douleur constante avec de légères fluctuations" ~ 0,
        deroulement_douleur_second == "Douleur constante ponctuée de pics de douleur" ~ -1,
        deroulement_douleur_second %in% c(
          "Pics de douleur sans douleur de fond",
          "Fréquents pics de douleur avec douleur de fond"
        ) ~ 1,
        deroulement_douleur_second == "Aucune case ne correspond" ~ 0,
        TRUE ~ NA_real_
      ) +
      if_else(precis_deroulement_second == "Oui", 2, 0) +
      case_when(
        sensation_brulure_second == "Pas du tout" ~ 0,
        sensation_brulure_second == "Très peu" ~ 1,
        sensation_brulure_second == "Légèrement" ~ 2,
        sensation_brulure_second == "Modérément" ~ 3,
        sensation_brulure_second == "Fortement" ~ 4,
        sensation_brulure_second == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        picotement_second == "Pas du tout" ~ 0,
        picotement_second == "Très peu" ~ 1,
        picotement_second == "Légèrement" ~ 2,
        picotement_second == "Modérément" ~ 3,
        picotement_second == "Fortement" ~ 4,
        picotement_second == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        leger_contact_douloureux_second == "Pas du tout" ~ 0,
        leger_contact_douloureux_second == "Très peu" ~ 1,
        leger_contact_douloureux_second == "Légèrement" ~ 2,
        leger_contact_douloureux_second == "Modérément" ~ 3,
        leger_contact_douloureux_second == "Fortement" ~ 4,
        leger_contact_douloureux_second == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        pic_douleur_second == "Pas du tout" ~ 0,
        pic_douleur_second == "Très peu" ~ 1,
        pic_douleur_second == "Légèrement" ~ 2,
        pic_douleur_second == "Modérément" ~ 3,
        pic_douleur_second == "Fortement" ~ 4,
        pic_douleur_second == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        temperature_douleur_second == "Pas du tout" ~ 0,
        temperature_douleur_second == "Très peu" ~ 1,
        temperature_douleur_second == "Légèrement" ~ 2,
        temperature_douleur_second == "Modérément" ~ 3,
        temperature_douleur_second == "Fortement" ~ 4,
        temperature_douleur_second == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        sensibilite_diminue_second == "Pas du tout" ~ 0,
        sensibilite_diminue_second == "Très peu" ~ 1,
        sensibilite_diminue_second == "Légèrement" ~ 2,
        sensibilite_diminue_second == "Modérément" ~ 3,
        sensibilite_diminue_second == "Fortement" ~ 4,
        sensibilite_diminue_second == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      ) +
      case_when(
        pression_douleur_second == "Pas du tout" ~ 0,
        pression_douleur_second == "Très peu" ~ 1,
        pression_douleur_second == "Légèrement" ~ 2,
        pression_douleur_second == "Modérément" ~ 3,
        pression_douleur_second == "Fortement" ~ 4,
        pression_douleur_second == "Très fortement" ~ 5,
        TRUE ~ NA_real_
      )
  )

#--------------------------------
#    PainDetect in categorical
#--------------------------------
# 1. Create a categorical variable related to the principal pain score
doul_neuro$score_paindetect_cat <- ifelse(
  doul_neuro$score1_test  > 38,
  "ERREUR",
  ifelse(
    doul_neuro$score1_test  > 18,
    "Positif",
    ifelse(
      doul_neuro$score1_test  > 12,
      "Incertain",
      ifelse(
        doul_neuro$score1_test  >= -1,
        "Négatif",
        ifelse(doul_neuro$score1_test  < -1, "ERREUR", NA)
      )
    )
  )
) %>% as.factor()

# 2. Delete unnecessary variables 
doul_neuro <- doul_neuro %>%
  dplyr::select(
    -c(
      deroulement_douleur,
      picotement,
      leger_contact_douloureux,
      pic_douleur,
      temperature_douleur,
      sensibilite_diminue,
      pression_douleur,
      deroulement_douleur_second,
      picotement_second,
      leger_contact_douloureux_second,
      pic_douleur_second,
      temperature_douleur_second,
      sensibilite_diminue_second,
      pression_douleur_second
    )
  )

#-------------------------
# Recode main pain area
#------------------------
# 1. Extraction of the area number: 
# for example, “21 - fosse iliaque droite (ovaire, trompe)” will be #converted to “21”
recode_var <- function(var) {
  recoded_var <- as.numeric(sub(".*?(\\d+).*", "\\1", var))
  return(recoded_var)
}

for (i in 1:53) {
  for (suffix in c("principale", "second")) {
    var_name <- paste0("zone_douleur_", suffix, "_", i)
    recoded_var_name <- paste0("recoded_zone_douleur_", suffix, "_", i)
    
    if (var_name %in% names(doul_neuro)) {
      doul_neuro[[recoded_var_name]] <- recode_var(doul_neuro[[var_name]])
    } else {
      warning(paste("Variable", var_name, "does not exist in doul_neuro"))
    }
  }
}

# 2. Deletes the old names
doul_neuro <- doul_neuro[, !names(doul_neuro) %in%
                           c(paste0("zone_douleur_principale_",
                                    1:53),
                             paste0("zone_douleur_second_", 
                                    1:53))]

# 3. Recode numbers into specific names
recode_body_area <- function(df, pattern) {
  body_map <- c(
    "1" = "Coté droit face",  "3" = "Coté droit face",  "34" = "Coté droit face",
    "2" = "Coté gauche face", "4" = "Coté gauche face", "33" = "Coté gauche face",
    "5" = "Cou", "35" = "Cou",
    "6" = "Epaule droite", "37" = "Epaule droite",
    "7" = "Epaule gauche", "36" = "Epaule gauche",
    "8" = "Poitrine", "9" = "Poitrine",
    "10" = "Bras droit", "40" = "Bras droit",
    "11" = "Bras gauche", "41" = "Bras gauche",
    "12" = "Abdomen", "13" = "Abdomen", "14" = "Abdomen", "15" = "Abdomen", "16" = "Abdomen",
    "17" = "Avant bras droit", "22" = "Avant bras droit", "45" = "Avant bras droit",
    "18" = "Avant bras gauche", "23" = "Avant bras gauche", "44" = "Avant bras gauche",
    "19" = "Zone pelvienne", "20" = "Zone pelvienne", "21" = "Zone pelvienne",
    "24" = "Haut jambe droite", "49" = "Haut jambe droite",
    "25" = "Haut jambe gauche", "48" = "Haut jambe gauche",
    "26" = "Bas jambe droite", "28" = "Bas jambe droite", "30" = "Bas jambe droite", "51" = "Bas jambe droite", "53" = "Bas jambe droite",
    "27" = "Bas jambe gauche", "29" = "Bas jambe gauche", "31" = "Bas jambe gauche", "50" = "Bas jambe gauche", "52" = "Bas jambe gauche",
    "32" = "Partie post tête",
    "38" = "Lombaire", "39" = "Lombaire",
    "43" = "Fessier droit", "46" = "Fessier droit",
    "42" = "Fessier gauche", "47" = "Fessier gauche"
  )

  cols_to_recode <- grep(pattern, names(df), value = TRUE)
  for (col in cols_to_recode) {
    df[[col]] <- as.factor(ifelse(
      as.character(df[[col]]) %in% names(body_map),
      body_map[as.character(df[[col]])],
      as.character(df[[col]])
    ))
  }
  return(df)
}

# 4. Recode body areas according previously to the areas already used in the literature 
recode_body_area2 <- function(df, pattern) {
  body_map <- c(
    "1" = "Coté droit face",  "3" = "Coté droit face",  "34" = "Coté droit face",
    "2" = "Coté gauche face", "4" = "Coté gauche face", "33" = "Coté gauche face",
    "5" = "Cou", "35" = "Cou",
    "6" = "Epaule droite", "37" = "Epaule droite",
    "7" = "Epaule gauche", "36" = "Epaule gauche",
    "8" = "Poitrine", "9" = "Poitrine",
    "10" = "Bras droit", "40" = "Bras droit",
    "11" = "Bras gauche", "41" = "Bras gauche",
    "12" = "Abdomen", "13" = "Abdomen", "14" = "Abdomen", "15" = "Abdomen", "16" = "Abdomen",
    "17" = "Avant bras droit", "22" = "Avant bras droit", "45" = "Avant bras droit",
    "18" = "Avant bras gauche", "23" = "Avant bras gauche", "44" = "Avant bras gauche",
    "19" = "Zone pelvienne", "20" = "Zone pelvienne", "21" = "Zone pelvienne",
    "24" = "Haut jambe droite", "49" = "Haut jambe droite",
    "25" = "Haut jambe gauche", "48" = "Haut jambe gauche",
    "26" = "Bas jambe droite", "28" = "Bas jambe droite", "30" = "Bas jambe droite", "51" = "Bas jambe droite", "53" = "Bas jambe droite",
    "27" = "Bas jambe gauche", "29" = "Bas jambe gauche", "31" = "Bas jambe gauche", "50" = "Bas jambe gauche", "52" = "Bas jambe gauche",
    "32" = "Partie post tête",
    "38" = "Lombaire", "39" = "Lombaire",
    "43" = "Fessier droit", "46" = "Fessier droit",
    "42" = "Fessier gauche", "47" = "Fessier gauche"
  )

  cols_to_recode <- grep(pattern, names(df), value = TRUE)
  for (col in cols_to_recode) {
    df[[col]] <- as.factor(ifelse(
      as.character(df[[col]]) %in% names(body_map),
      body_map[as.character(df[[col]])],
      as.character(df[[col]])
    ))
  }
  return(df)
}

# 5. Apply factor recoding
doul_neuro <- recode_body_area(doul_neuro,"recoded_zone_douleur_principale_test")
doul_neuro <- recode_body_area2(doul_neuro,"recoded_zone_douleur_second_")

#----------------------------------------
#  Create lists of index for these area
#----------------------------------------
# 1. Create a temporary dataframe
new_dataframe <- doul_neuro %>% 
  dplyr::select(id_unik,starts_with(c("recoded_zone_douleur_principale_","recoded_zone_douleur_second_"))) 

# 2. Make lists of index                
body_parts <- c("Tête",
                "Epaule et cou",
                "Torse",
                "Bras",
                "Ventre",
                "Bassin",
                "Jambes",
                "Dos")

column_names <- c(
  "indic_doul_tete",
  "indic_doul_epaule",
  "indic_doul_torse",
  "indic_doul_bras",
  "indic_doul_ventre",
  "indic_doul_bassin",
  "indic_doul_jambe",
  "indic_doul_dos"
)
body_parts2 <- c(
  "Coté droit face",
  "Coté gauche face",
  "Cou",
  "Epaule droite",
  "Epaule gauche",
  "Poitrine",
  "Bras droit",
  "Bras gauche",
  "Abdomen",
  "Avant bras droit",
  "Avant bras gauche",
  "Zone pelvienne",
  "Haut jambe droite",
  "Haut jambe gauche",
  "Bas jambe droite",
  "Bas jambe gauche",
  "Partie post tête",
  "Lombaire",
  "Fessier droit",
  "Fessier gauche"
)

column_names2 <- c(
  "indic_doul_cote_droit_face",
  "indic_doul_cote_gauche_face",
  "indic_doul_cou",
  "indic_doul_epaule_droite",
  "indic_doul_epaule_gauche",
  "indic_doul_poitrine",
  "indic_doul_bras_droit",
  "indic_doul_bras_gauche",
  "indic_doul_abdomen",
  "indic_doul_avant_bras_droit",
  "indic_doul_avant_bras_gauche",
  "indic_doul_zone_pelvienne",
  "indic_doul_haut_jambe_droite",
  "indic_doul_haut_jambe_gauche",
  "indic_doul_bas_jambe_droite",
  "indic_doul_bas_jambe_gauche",
  "indic_doul_partie_post_tete",
  "indic_doul_lombaire",
  "indic_doul_fessier_droit",
  "indic_doul_fessier_gauche"
)

# 3. Look through each body part and create the corresponding indicator variable
for (i in seq_along(body_parts2)) {
  part <- body_parts2[i]
  col_name <- column_names2[i]
  
  new_dataframe[[col_name]] <- as.factor(apply(new_dataframe, 1, function(row) {
    ifelse(part %in% row, 1, 0)
    }))
  }

# 4. Add the new variables to the dataframe
doul_neuro <- doul_neuro %>%
  left_join(
    new_dataframe %>%
      dplyr::select(
        -starts_with("recoded_zone_douleur_")
        ),
    by = "id_unik"
  )

# 5. Delete the old ones and the temporary dataframe
doul_neuro <- doul_neuro %>% 
  dplyr::select(-starts_with(c("recoded_zone_douleur_principale_",
                              "recoded_zone_douleur_second_" )))
rm(new_dataframe)

#----------------------
# Recode immediate pain 
#-----------------------
# 1. character to numeric
doul_neuro$douleur_instant <-  as.numeric(as.character(doul_neuro$douleur_instant))

# 2. quantiles data
quantiles <- quantile(doul_neuro$douleur_instant, probs = c(1/3, 2/3),na.rm=TRUE)

# 3. Create new categorical variable
doul_neuro <- doul_neuro %>%
  mutate(
    tercile_douleur_instant = ntile(douleur_instant, 3),
    tercile_douleur_instant = factor(
      tercile_douleur_instant,
      levels = 1:3,
      labels = c(
        paste0("<", round(quantiles[1], 2)),
        paste0(round(quantiles[1], 2), "-", round(quantiles[2], 2)),
        paste0(">", round(quantiles[2], 2))
      )
    )
  ) %>%
  # 4. Remove unnecessay variables 
  dplyr::select(-precision_douleur, -douleur_instant)

#----------------------
#  Recode strong pain 
#----------------------
# 1. character to numeric
doul_neuro$forte_douleur <-  as.numeric(as.character(doul_neuro$forte_douleur))

# 2. quantiles data
quantiles <- quantile(doul_neuro$forte_douleur , probs = c(1/3, 2/3),na.rm=TRUE)

# 3. Create new categorical variable
doul_neuro <- doul_neuro %>%
  mutate(
    tercile_forte_douleur = ntile(forte_douleur, 3),
    tercile_forte_douleur = factor(
      tercile_forte_douleur,
      levels = 1:3,
      labels = c(
        paste0("<", round(quantiles[1], 2)),
        paste0(round(quantiles[1], 2), "-", round(quantiles[2], 2)),
        paste0(">", round(quantiles[2], 2))
      )
    )
  ) %>%
  # 4. Remove unnecessay variables 
  dplyr::select(-forte_douleur)

#------------------------
# Recode pain intensity 
#------------------------
# 1. character to numeric
doul_neuro$intensite_moyenne <- 
  as.numeric(as.character(doul_neuro$intensite_moyenne))

# 2. quantiles data
quantiles <- quantile(doul_neuro$intensite_moyenne , probs = c(1/3, 2/3),na.rm=TRUE)

# 3. Create new categorical variable
doul_neuro <- doul_neuro %>%
  mutate(
    tercile_intensite_moyenne = ntile(intensite_moyenne, 3),
    tercile_intensite_moyenne = factor(
      tercile_intensite_moyenne,
      levels = 1:3,
      labels = c(
        paste0("<", round(quantiles[1], 2)),
        paste0(round(quantiles[1], 2), "-", round(quantiles[2], 2)),
        paste0(">", round(quantiles[2], 2))
        )
      )
    ) %>%
  dplyr::select(-intensite_moyenne)
      
#------------------------------------
# whether the women experienced a propagation of pain for at least one of their two main pain points
#------------------------------------
#- Be careful: when you execute as.numeric(doul_neuro$propagation_douleur), 'Oui' should correspond to the higher number. If it does not, reverse the levels in the labels.
doul_neuro$propagation_douleur3 <- factor(
  pmax(
    as.numeric(as.factor(doul_neuro$propagation_douleur)),
    as.numeric(as.factor(doul_neuro$propagation_douleur_second)),
    na.rm = TRUE
  ),
  levels = c(1, 2),
  labels = c("Non", "Oui"),
  ordered = TRUE
)

#---------------------------------
# 20 zones previously used
#---------------------------------
# Here, we create a variable for the propagation of pain that, instead of indicating the 53 zones of origin, uses the 20 zones previously used
for (i in 1:53) {
  for (suffix in c("douleur", "douleur_second")) {
    var_name <- paste0("direct_propa_", suffix, "_", i)
    recoded_var_name <- paste0("recoded_direct_propa_", suffix, "_", i)
    
    if (var_name %in% names(doul_neuro)) {
      doul_neuro[[recoded_var_name]] <- recode_var(doul_neuro[[var_name]])
    } else {
      warning(paste("Variable", var_name, "does not exist in doul_neuro"))
    }
  }
}

# 2. Create index for each area 
doul_neuro <- doul_neuro %>% 
  dplyr::select(-starts_with(c("direct_propa_douleur_","direct_propa_douleur_second_")))
# 3. Apply the created function to the all variables started with "recoded_direct_propa_douleur_"     
doul_neuro <- recode_body_area(doul_neuro, "recoded_direct_propa_douleur_")
doul_neuro <- recode_body_area2(doul_neuro, "recoded_direct_propa_douleur_second_")

# 4. Create a temporary dataframe
new_dataframe <- doul_neuro %>% 
  dplyr::select(id_unik, 
                starts_with("recoded_direct_propa_douleur_"), 
                propagation_douleur3)

# 5. List of body parts and corresponding column names
body_parts <- c("Tête",
                "Epaule et cou",
                "Torse",
                "Bras",
                "Ventre",
                "Bassin",
                "Jambes",
                "Dos")
body_parts2 <- c(
  "Coté droit face",
  "Coté gauche face",
  "Cou",
  "Epaule droite",
  "Epaule gauche",
  "Poitrine",
  "Bras droit",
  "Bras gauche",
  "Abdomen",
  "Avant bras droit",
  "Avant bras gauche",
  "Zone pelvienne",
  "Haut jambe droite",
  "Haut jambe gauche",
  "Bas jambe droite",
  "Bas jambe gauche",
  "Partie post tête",
  "Lombaire",
  "Fessier droit",
  "Fessier gauche"
)
column_names <- c(
  "indic_prop_doul_tete",
  "indic_prop_doul_epaule",
  "indic_prop_doul_torse",
  "indic_prop_doul_bras",
  "indic_prop_doul_ventre",
  "indic_prop_doul_bassin",
  "indic_prop_doul_jambe",
  "indic_prop_doul_dos"
)
column_names2 <- c(
  "indic_prop_doul_cote_droit_face",
  "indic_prop_doul_cote_gauche_face",
  "indic_prop_doul_cou",
  "indic_prop_doul_epaule_droite",
  "indic_prop_doul_epaule_gauche",
  "indic_prop_doul_poitrine",
  "indic_prop_doul_bras_droit",
  "indic_prop_doul_bras_gauche",
  "indic_prop_doul_abdomen",
  "indic_prop_doul_avant_bras_droit",
  "indic_prop_doul_avant_bras_gauche",
  "indic_prop_doul_zone_pelvienne",
  "indic_prop_doul_haut_jambe_droite",
  "indic_prop_doul_haut_jambe_gauche",
  "indic_prop_doul_bas_jambe_droite",
  "indic_prop_doul_bas_jambe_gauche",
  "indic_prop_doul_partie_post_tete",
  "indic_prop_doul_lombaire",
  "indic_prop_doul_fessier_droit",
  "indic_prop_doul_fessier_gauche"
)

# 6. Indicate whether or not there is propagation in this body part
for (i in seq_along(body_parts2)) {
  part <- body_parts2[i]
  col_name <- column_names2[i]
  
  new_dataframe[[col_name]] <- as.factor(apply(new_dataframe, 1, function(row) {
    ifelse(part %in% row, "Propagation dans cette région", "Pas de propagation dans cette région")
  }))
}

# 7. Indicate if the women had any propagation of pain
modifier_prop_doul <- function(df, target_var, condition_var) {
  if (!(target_var %in% colnames(df)) || !(condition_var %in% colnames(df))) {
    stop("Les variables spécifiées n'existent pas dans le dataframe.")
  }
  df[[target_var]] <- as.factor(ifelse(df[[condition_var]] == "Non", "Aucune propagation",as.character( df[[target_var]])))
  return(df)
}

# 8. Apply the function to each column specified in column_names
for (target_var in column_names2) {
  new_dataframe <- modifier_prop_doul(new_dataframe, target_var,
                                      "propagation_douleur3")
}

# 9. Replace the new variable in the dataframe
doul_neuro <-doul_neuro %>% 
  left_join(
    new_dataframe %>% dplyr::select(-starts_with("recoded_direct_propa_douleur"),
                             -propagation_douleur3),
    by="id_unik"
  )
  
# 10. Delete unnecessay variables and the temporary dataframe 
doul_neuro <- doul_neuro %>% 
  dplyr::select(-starts_with(c("recoded_direct_propa_douleur_")))
rm(new_dataframe)
```

# II. Final dataset

## 1) Variable selection 

```{r}
doul_neuro_final <- doul_neuro %>%
  distinct(
    id_unik,
    #IdVersion,
    doul_neuro_bin,
    score_paindetect_cat,
    doul_neuro_participated
    # DateValidation,
    # Date_naissance
  ) %>% 
  mutate(
    # Add new modality 'Not_answered'
    across(c(doul_neuro_bin, score_paindetect_cat),
           ~ case_when(is.na(IdVersion) ~ "Not_answered",
                       TRUE ~ .)
           ),
    across(starts_with("Date"),
           ~ as.Date(., format = "%d/%m/%Y")
           ),
    
    score_paindetect_cat =
      case_when(doul_neuro_bin == "Non" ~ "No pain", 
                TRUE ~ score_paindetect_cat)
    ) #%>% 
  #select(-IdVersion)
```

## 2) Real Final Dataframe
Question (Marina) : pas de tranch d'âge  : impossible de créer une variable avant de la 1ère grossesse 

```{r}
# new_dataframe <- doul_neuro_final %>%
#   left_join(grossesse_mod_final %>%
#               distinct(id_unik, first_preg_age,last_preg_age_mod,
#                        grossesse_encours),
#             by = "id_unik") %>% 
#   mutate(
#     last_preg_age_mod= last_preg_age_mod %>% round()
#   )
#   
#   
# # 2. Create a binary variable 
# new_dataframe <- new_dataframe %>% 
#   mutate(
#     first_preg_age= first_preg_age %>% round(),
#     age_inclusion = as.numeric(as.character(
#       (DateValidation-Date_naissance)/365)) %>% round()
#     )  %>% 
#   select(- DateValidation, -Date_naissance)
```


# III. Save the final dataframe 

```{r}
#write.csv(doul_neuro_final,"doul_neuro_final.csv")
```

