---
title: "01_clean_Quality_life"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Quality_life
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 

EQ-5D : https://euroqol.org/register/obtain-eq-5d/eq-5d-sample-demos/ 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
qualite_vie <- read.csv(
  "Complic_pop_Qualite_vie_20250407.csv",
  encoding = "latin1",
  colClasses = c(Sexe = "character")
  )

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
qualite_vie <- qualite_vie%>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- qualite_vie %>% 
  sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count == nrow(qualite_vie)) %>%
  as.vector() # index of columns having only NA
```


# I. Preprocessing Data 

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
qualite_vie <- qualite_vie %>% 
  mutate(qualite_vie_participated = ifelse(is.na(IdAction),
                                           "Non","Oui"))

#-------------------------------
#   character to numeric
#-------------------------------
qualite_vie <- qualite_vie %>% 
  mutate(eval_sante=eval_sante %>% as.numeric())

#----------------------------------
# Score : numeric -> categorical
#----------------------------------
# 1. Calculate quantiles
quantiles_evalSante <- quantile(qualite_vie$eval_sante, 
                          probs = c(1/3, 2/3), na.rm=TRUE)

# 2. Create new categorical variables 
qualite_vie <- qualite_vie %>%
  mutate(
    eval_sante_cat =
      case_when(
        eval_sante < quantiles_evalSante[[1]] ~
          paste0("<", round(quantiles_evalSante[1], 2)),
        eval_sante >= quantiles_evalSante[[1]] &
          eval_sante < quantiles_evalSante[[2]] ~
          paste0(
            round(quantiles_evalSante[1], 2),
            "-",
            round(quantiles_evalSante[2], 2)
          ),
        eval_sante >= quantiles_evalSante[[2]] ~
          paste0(">=", round(quantiles_evalSante[2], 2))
      ) %>%
      as.factor()
  )

# 3. Remove the quantiles data
rm(quantiles_evalSante)
```


# II. Final dataset

```{r}
qualite_vie_final <- qualite_vie %>% 
  distinct(id_unik, 
           IdAction, 
           eval_sante_cat, 
           qualite_vie_participated) %>% 
  mutate(across(-c(id_unik, IdAction),~
                  case_when(
                    is.na(IdAction) ~ "Not_answered",
                    TRUE  ~ .)
                  )
         ) %>%
  select(-IdAction)
```

# III. Save the final dataframe 

```{r}
#write.csv(qualite_vie_final,"qualite_vie_final.csv")
```

