---
title: "01_clean_Physical_activity"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Physical_activity
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
activite_physique <- read.csv(
  "Complic_pop_QST_0331_20250407_activite_physique.csv",
  encoding = "latin1",
  colClasses = c(Sexe = "character")
  )

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
activite_physique <- activite_physique %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- activite_physique %>% 
  sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count == nrow(activite_physique)) %>%
  as.vector() # index of columns having only NA
```

# I. Preprocessing Data

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
activite_physique <- activite_physique %>% 
  mutate(activite_physique_participated =
           ifelse(is.na(IdAction),"Non","Oui"))

#-------------------------
#  Character to numeric
#-------------------------
activite_physique <- activite_physique %>%
  mutate(across(c(5, 6, 9, 10, 13, 14, 16, 17), as.numeric))

#------------------------------
# Score : METs, IPAQ level 
#------------------------------
activite_physique <- activite_physique %>%
  mutate(
    # Convert 'jour_activite_intense' to numeric days
    jour_intense = case_when(
      jour_activite_intense == "0 jour" ~ 0,
      jour_activite_intense == "1 jour" ~ 1,
      jour_activite_intense == "2 jours" ~ 2,
      jour_activite_intense == "3 jours" ~ 3,
      jour_activite_intense == "4 jours" ~ 4,
      jour_activite_intense == "5 jours" ~ 5,
      jour_activite_intense == "6 jours" ~ 6,
      jour_activite_intense == "7 jours" ~ 7,
      TRUE ~ NA_real_
    ),
    # Convert 'jour_activite_modere' to numeric days
    jour_modere = case_when(
      jour_activite_modere == "0 jour" ~ 0,
      jour_activite_modere == "1 jour" ~ 1,
      jour_activite_modere == "2 jours" ~ 2,
      jour_activite_modere == "3 jours" ~ 3,
      jour_activite_modere == "4 jours" ~ 4,
      jour_activite_modere == "5 jours" ~ 5,
      jour_activite_modere == "6 jours" ~ 6,
      jour_activite_modere == "7 jours" ~ 7,
      TRUE ~ NA_real_
    ),
    # Convert 'jour_activite_marche' to numeric days
    jour_marche = case_when(
      jour_activite_marche == "0 jour" ~ 0,
      jour_activite_marche == "1 jour" ~ 1,
      jour_activite_marche == "2 jours" ~ 2,
      jour_activite_marche == "3 jours" ~ 3,
      jour_activite_marche == "4 jours" ~ 4,
      jour_activite_marche == "5 jours" ~ 5,
      jour_activite_marche == "6 jours" ~ 6,
      jour_activite_marche == "7 jours" ~ 7,
      TRUE ~ NA_real_
    ),
    # Calculate total days
    jour_total = if_else(
      is.na(jour_intense) & is.na(jour_modere) & is.na(jour_marche),
      NA_real_,
      # Si les trois colonnes sont NA, alors jour_total = NA
      rowSums(across(
        c(jour_intense, jour_modere, jour_marche)
      ), na.rm = TRUE)  # Sinon, somme ligne par ligne
    ),
    # Calculate minutes for each activity type
    minute_intense = 60 * if_else(is.na(heure_act_int__1_), 0, heure_act_int__1_) +
      if_else(is.na(minute_act_int__1_), 0, minute_act_int__1_),
    minute_modere = 60 * if_else(is.na(heure_act_mod__1_), 0, heure_act_mod__1_) +
      if_else(is.na(minute_act_mod__1_), 0, minute_act_mod__1_),
    minute_marche = 60 * if_else(is.na(heure_marche__1_), 0, heure_marche__1_) +
      if_else(is.na(minute_marche__1_), 0, minute_marche__1_),
    minute_intense = case_when(
      jour_intense == 0 ~ 0,
      is.na(heure_act_int__1_) &
        is.na(minute_act_int__1_) ~ NA,
      TRUE ~ minute_intense
    ),
    minute_modere = case_when(
      jour_modere == 0 ~ 0,
      is.na(heure_act_mod__1_) &
        is.na(minute_act_mod__1_) ~ NA,
      TRUE ~ minute_modere
    ),
    minute_marche = case_when(
      jour_modere == 0 ~ 0,
      is.na(heure_marche__1_) &
        is.na(minute_marche__1_) ~ NA,
      TRUE ~ minute_marche
    ),
    minute_assis = 60 * if_else(is.na(heure_assis__1_), 0, heure_assis__1_) +
      if_else(is.na(minute_assis__1_), 0, minute_assis__1_),
    minute_assis = case_when(
      is.na(heure_assis__1_) & is.na(minute_assis__1_) ~ NA,
      TRUE ~ minute_assis
    ),
    sedentarite = case_when(
      minute_assis__1_ >= 480 ~ "Très sédentaire",
      minute_assis__1_ < 480 &
        minute_assis__1_ >= 240 ~ "Moyennement sédentaire",
      minute_assis__1_ < 240 ~ "Peu sédentaire",
      TRUE ~ NA
    ),
    
    # Calculate METs for each activity type
    MET_marche = 3.3 * minute_marche * jour_marche,
    MET_modere = 4.0 * minute_modere * jour_modere,
    MET_intense = 8 * minute_intense * jour_intense,
    MET_total = if_else(
      is.na(MET_marche) & is.na(MET_modere) & is.na(MET_intense),
      NA_real_,
      rowSums(across(c(
        MET_marche, MET_modere, MET_intense
      )), na.rm = TRUE)
    ),
    
    # Classify IPAQ level
    IPAQ = case_when(
      is.na(jour_activite_intense) ~ "",
      (jour_intense >= 3 & MET_intense >= 1500) |
        (jour_total >= 7 & MET_total >= 3000) ~ "high",
      (jour_intense >= 3 & minute_intense >= 20) |
        (jour_modere >= 5 & minute_modere >= 30) |
        (jour_total >= 5 & MET_total >= 600) ~ "moderate",
      MET_total == 0 ~ "low",
      TRUE ~ "low"
    ) %>% factor(levels=c("low","moderate","high"))
  )

#Recodage des variables afin de passer des deux variables (heure et minute) à une seule variable temps.

#---------------------------------------
# Selection of only necessary variables 
#---------------------------------------
activite_physique_selected <- 
  activite_physique %>% 
  select(
    -c(
    StatusParticipant,
    heure_act_int__1_,
    minute_act_int__1_,
    heure_act_mod__1_,
    minute_act_mod__1_,
    heure_marche__1_,
    minute_marche__1_,
    heure_assis__1_,
    minute_assis__1_,
    raison_autant_act,
    commentaires_activite_physique,
    DateValidation,
    Date_naissance,
    Sexe,
    campagne)
    )

# Sédentarité : distribution tel que dans l'étude suivante :"Emmanuel Stamatakis, Joanne Gale, Adrian Bauman, Ulf Ekelund, Mark Hamer, Ding Ding, Sitting Time, Physical Activity, and Risk of Mortality in Adults,Journal of the American College of Cardiology,Volume 73, Issue 16,2019,Pages 2062-2072,ISSN 0735-1097"

#-----------------------------------
# Diviser les variables en terciles
#-----------------------------------
# ATTNETION : Il faut corriger ces codes dessous selon dataset

# 1. character to numeric
activite_physique_selected <- 
  activite_physique_selected %>%
  mutate(across(
    starts_with("jour_")| starts_with("minute_")| starts_with("MET_"),
    ~as.numeric(.)))


# 2. minute_intense : num->cat
quantiles <- quantile(
  activite_physique_selected$minute_intense,
  probs = c(1/3, 2/3),na.rm = TRUE) %>% round()

activite_physique_selected <- activite_physique_selected %>%
  mutate(
    tercile_minute_intense = case_when(
      minute_intense == quantiles[[1]] ~
        as.character(quantiles[1]),
      minute_intense <= quantiles[[2]] &
        minute_intense > quantiles[[1]] ~
        paste0(quantiles[1], "-", quantiles[2]),
      minute_intense > quantiles[[2]] ~
        paste0(quantiles[2], " or more"),
      TRUE ~ NA
    ) %>% as.factor()
  )
rm(quantiles)

# 3. minute_modere : num->cat
quantiles <- quantile(
  activite_physique_selected$minute_modere,
  probs = c(1 / 3, 2 / 3),na.rm = TRUE)%>% round()

activite_physique_selected <- activite_physique_selected %>%
  mutate(
    tercile_minute_modere = case_when(
      minute_modere <= quantiles[[1]] ~ 
        paste0(quantiles[1]," or less"),
      minute_modere <= quantiles[[2]] &
        minute_intense > quantiles[[1]] ~ 
        paste0(quantiles[1], "-", quantiles[2]),
      minute_modere > quantiles[[2]] ~ 
        paste0(quantiles[2]," or more"),
      TRUE ~ NA
    ) %>% 
      as.factor()
  )
rm(quantiles)

# 4. minute_marche : num->cat
quantiles <- quantile(
  activite_physique_selected$minute_marche,
  probs = c(1 / 3, 2 / 3),na.rm = TRUE) %>% round()

activite_physique_selected <- activite_physique_selected %>%
  mutate(
    tercile_minute_marche = case_when(
      minute_marche <= quantiles[[1]] ~ 
        paste0(quantiles[1]," or less"),
      minute_marche <= quantiles[[2]] &
        minute_marche > quantiles[[1]]~ 
        paste0(quantiles[1], "-", quantiles[2]),
      minute_marche > quantiles[[2]] ~ 
        paste0(quantiles[2]," or more"),
      TRUE ~ NA
    ) %>% 
      as.factor()
  )

# 5. Add the class : Je ne sais pas 
activite_physique_selected <- activite_physique_selected %>%
  mutate(
    sedentarite = case_when(
      info_tps_passe_assis == "Je ne sais pas" ~ 
        "Je ne sais pas",
      TRUE ~ sedentarite
    ),
    
    tercile_minute_intense = case_when(
      info_tps_passe_atc_int == "Je ne sais pas" ~ 
        "Je ne sais pas",
      TRUE ~ tercile_minute_intense
    ),
    
    tercile_minute_marche = case_when(
      info_tps_passe_marche == "Je ne sais pas" ~ 
        "Je ne sais pas",
      TRUE ~ tercile_minute_marche
    )
  )
```


# II. Final dataset

```{r}
activite_physique_final <- activite_physique_selected %>%
  distinct(id_unik,
           IdAction, 
           IPAQ,
           activite_physique_participated) %>% 
  mutate(across(-c(id_unik),~
                  case_when(
                    is.na(IdAction) ~ "Not_answered",
                    TRUE  ~ .)
                  )
         ) %>%
  select(-IdAction)
```

# III. Save the final dataframe 

```{r}
#write.csv(activite_physique_final,"activite_physique_final.csv")
```


