---
title: "01_clean_Endo_dyspareunie_dyschezie_dysurie"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Endo_dyspareunie_dyschezie_dysurie
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
Endo_dyspareunie <-
  read.csv("Complic_pop_QST_0465_20250407_dys.csv",
           fileEncoding = "latin1",
           colClasses = c(Sexe = "character"))

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
Endo_dyspareunie <- Endo_dyspareunie %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim),
         across(everything(), ~ gsub("\u0092", "'", .)),
         across(everything(), ~ gsub("\u0085", "…", .))
         )

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- Endo_dyspareunie %>% sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count==nrow(Endo_dyspareunie)) %>% as.vector() # index of columns having only NA
Endo_dyspareunie <- Endo_dyspareunie[,-index_col_NA]

#--------------------------
#  Pain score in numeric
#--------------------------
Endo_dyspareunie <- Endo_dyspareunie %>% 
  mutate(across(starts_with("niveau_"), 
                ~ as.numeric(as.character(.))),
         across(starts_with("Date"),
                ~ as.Date(.,format="%d/%m/%Y")))
```

# I. Preprocessing Data

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
Endo_dyspareunie <- Endo_dyspareunie %>% 
  mutate(Endo_douleurs_participated=ifelse(is.na(IdAction),
                                                "Non","Oui"))

#------------------------
# First sexual rapport 
#------------------------
Endo_dyspareunie <- Endo_dyspareunie %>%
  mutate(first_rapport= case_when(
    doul_rapport_15 %in% c("Oui", "Non, pas de douleurs") ~ "avant 15",
    doul_rapport_16 %in% c("Oui", "Non, pas de douleurs") ~ "16-20",
    doul_rapport_21 %in% c("Oui", "Non, pas de douleurs") ~ "21-30",
    doul_rapport_31 %in% c("Oui", "Non, pas de douleurs") ~ "31-40",
    doul_rapport_41 %in% c("Oui", "Non, pas de douleurs") ~ "41+",
    rowSums(across(starts_with("doul_rapport"), ~ . %in% 
                     c("Pas de rapport sexuel à cette période", 
                       "Je ne me souviens plus"))) == 5 ~
      "Je ne me souviens du tout",
    TRUE ~ NA_character_
  ))

#-----------------------------------
#   Dyspareunia : Onset Age group
#-----------------------------------
Endo_dyspareunie <- Endo_dyspareunie %>%
  mutate(onset_age_dyspareunia = case_when(
    doul_rapport_15 == "Oui" ~ "avant 15",
    doul_rapport_16 == "Oui" ~ "16-20",
    doul_rapport_21 == "Oui" ~ "21-30",
    doul_rapport_31 == "Oui" ~ "31-40",
    doul_rapport_41 == "Oui" ~ "41+",
    rowSums(across(starts_with("doul_rapport"), ~ . != "Oui")) == 5 ~ "pas encore et nsp",
    TRUE ~ NA_character_
  ))

#-------------------------------------------------------------
#  Dyspareunie : Immediate vs. delayed onset by considering the first sexual rapport 
#-------------------------------------------------------------
Endo_dyspareunie <- Endo_dyspareunie %>%
  mutate(onset_type_dyspareunie = case_when(
    first_rapport == "avant 15" & onset_age_dyspareunia == "avant 15" ~ "immediate",
    first_rapport == "avant 15" & onset_age_dyspareunia != "avant 15" & !is.na(onset_age_dyspareunia) ~ "delayed",

    first_rapport == onset_age_dyspareunia ~ "immediate",

    onset_age_dyspareunia %in% c("16-20", "21-30", "31-40", "41+") & 
      first_rapport %in% c("avant 15", "16-20", "21-30", "31-40") &
      onset_age_dyspareunia != "pas encore et nsp" &
      first_rapport != "Je ne me souviens du tout" ~ "delayed",

    onset_age_dyspareunia == "pas encore et nsp" | 
      first_rapport == "Je ne me souviens du tout" ~ "Unknown",
    TRUE ~ NA_character_
  ))

#-----------------------
#  Recent dyspareunia 
#-----------------------
Endo_dyspareunie <- Endo_dyspareunie %>%
  mutate(recent_doul_rapport = coalesce(
    na_if(doul_rapport_41, "Je n'ai pas encore atteint cette tranche d'âge"),
    na_if(doul_rapport_31, "Je n'ai pas encore atteint cette tranche d'âge"),
    na_if(doul_rapport_21, "Je n'ai pas encore atteint cette tranche d'âge"),
    na_if(doul_rapport_16, "Je n'ai pas encore atteint cette tranche d'âge"),
    na_if(doul_rapport_15, "Je n'ai pas encore atteint cette tranche d'âge")
  ))

#---------------------------------------------------------
# Dyspareunia improvement compared to initial dyspareunia
#---------------------------------------------------------
# 1. wide format to long format 
data_long <- Endo_dyspareunie %>%
  select(id_unik, starts_with("niveau_severite_douleur_")) %>% 
  pivot_longer(cols = starts_with("niveau_severite_douleur_"), 
               names_to = "age_group", 
               values_to = "score") %>%
  filter(!is.na(score)) %>%
  # make age groups (15~41) 
  mutate(age_group = case_when(
    str_detect(age_group, "15") ~ 15,
    str_detect(age_group, "1620") ~ 20,
    str_detect(age_group, "2130") ~ 30,
    str_detect(age_group, "3140") ~ 40,
    str_detect(age_group, "4150") ~ 41
  ))

# 2. Check the first and last score of severity of pain and compare these scores 
data_summary <- data_long %>%
  group_by(id_unik) %>%
  summarise(
    first_score = score[which.min(age_group)],
    last_score = score[which.max(age_group)]
  ) %>%
  mutate(
    dyspareunie_improved = case_when(
      last_score < first_score ~"improved",
      last_score > first_score ~ "worsend",
      last_score == first_score ~"stable",
      TRUE ~ NA_character_ ))

# 3. Put new variable created to the original dataset 
Endo_dyspareunie <- Endo_dyspareunie %>% 
  left_join(
    data_summary %>% select(id_unik, dyspareunie_improved),
    by="id_unik"
  )

# 4. Delete two temporary dataframe : data_long and data_summary
rm(data_long);rm(data_summary)

#-------------------------------------------------------------
#  Mean of the worst severity level of dyspareunia over the entire life
#-------------------------------------------------------------
# 1. Create a temporary dataframe by making a new variable
new_dataframe <- Endo_dyspareunie %>%
  select(id_unik, starts_with("niveau_severite_douleur_")) %>%
  mutate(
    total_score = rowSums(select(
      ., starts_with("niveau_severite_douleur_")
    ), na.rm = TRUE),
    na_count = 5 - rowSums(is.na(select(
      ., starts_with("niveau_severite_douleur_")
    ))),
    mean_niveau_severite_dyspareunia = total_score / na_count
  ) %>%
  select(id_unik, mean_niveau_severite_dyspareunia)

# 2. Add a new categorical variables 
quantiles <- new_dataframe$mean_niveau_severite_dyspareunia %>% quantile(probs=c(1/3,2/3), na.rm=TRUE) %>% round()

new_dataframe <- new_dataframe %>%
  mutate(
    mean_niveau_severite_dyspareunia_cat =
      case_when(
        mean_niveau_severite_dyspareunia < quantiles[[1]] ~
          paste0("<", quantiles[1]),
        mean_niveau_severite_dyspareunia >= quantiles[[1]] &
          mean_niveau_severite_dyspareunia < quantiles[[2]] ~
          paste0(quantiles[1], "-", quantiles[2]),
        mean_niveau_severite_dyspareunia >= quantiles[[2]] ~
          paste0(">=", quantiles[2])
      )
  )

# 3. Add a new variable to the main dataframe
Endo_dyspareunie <- Endo_dyspareunie %>% 
  left_join(new_dataframe, by="id_unik")

# 3. Delete the temporary dataframe and quantile information
rm(new_dataframe); rm(quantiles)

# 4. Modify the categorical variable by adding 'no pain' modality 
Endo_dyspareunie <- Endo_dyspareunie %>% 
  mutate(mean_niveau_severite_dyspareunia_cat=
           case_when(
             doul_rapport_sexuel=="Non" ~ "No pain",
             TRUE ~ mean_niveau_severite_dyspareunia_cat
           ))

#---------------------------
#     Digestive trouble 
#---------------------------
# 1. Create a temporary dataframe 
new_dataframe <- Endo_dyspareunie %>% 
  select(id_unik,starts_with("trouble_digestif_")) %>%
  mutate(
    across(starts_with("trouble_digestif_"), ~ case_when(
      . == "Je ne me souviens plus" ~ "nsp",
      . == "Non" ~ "Non",
      . %in% c("Oui, des troubles digestifs fonctionnels (inconfort digestif, ballonnements, constipation, diarrhée…)",
               "Oui, des difficultés d'évacuation (évacuation anormalement longue, en plusieurs temps, ou nécessitant des efforts prolongés, d'intenses poussées ou l'aide de suppositoires/doigts/lavements)",
               "Oui, des douleurs lors de l'évacuation de selles (dyschézie)") ~ "Oui",
      . == "Je n'ai pas atteint cette tranche d'âge" ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_
    ))
  ) %>% 
  rowwise() %>% 
  mutate(
    trouble_digestif_final_15 = case_when(
      all(c_across(starts_with("trouble_digestif_15")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("trouble_digestif_15")) == "Oui") ~ "Oui",
      all(c_across(starts_with("trouble_digestif_15")) == "Non") ~ "Non",  
      all(c_across(starts_with("trouble_digestif_15")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("trouble_digestif_15")) == "Non") ~ "Non",  
      any(c_across(starts_with("trouble_digestif_15")) == "nsp") ~ "nsp",  
      TRUE ~ NA_character_  
    ),
    trouble_digestif_final_1620 = case_when(
      all(c_across(starts_with("trouble_digestif_16")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("trouble_digestif_16")) == "Oui") ~ "Oui",  
      all(c_across(starts_with("trouble_digestif_16")) == "Non") ~ "Non",  
      all(c_across(starts_with("trouble_digestif_16")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("trouble_digestif_16")) == "Non") ~ "Non",  
      any(c_across(starts_with("trouble_digestif_16")) == "nsp") ~ "nsp",  
      TRUE ~ NA_character_  
    ),
    trouble_digestif_final_2130 = case_when(
      all(c_across(starts_with("trouble_digestif_21")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("trouble_digestif_21")) == "Oui") ~ "Oui",  
      all(c_across(starts_with("trouble_digestif_21")) == "Non") ~ "Non",  
      all(c_across(starts_with("trouble_digestif_21")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("trouble_digestif_21")) == "Non") ~ "Non",  
      any(c_across(starts_with("trouble_digestif_21")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("trouble_digestif_21")) == "Pas atteint cette tranche d'âge") ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_ 
    ),
    trouble_digestif_final_3140 = case_when(
      all(c_across(starts_with("trouble_digestif_31")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("trouble_digestif_31")) == "Oui") ~ "Oui",  
      all(c_across(starts_with("trouble_digestif_31")) == "Non") ~ "Non",  
      all(c_across(starts_with("trouble_digestif_31")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("trouble_digestif_31")) == "Non") ~ "Non",  
      any(c_across(starts_with("trouble_digestif_31")) == "nsp") ~ "nsp", 
      any(c_across(starts_with("trouble_digestif_31")) == "Pas atteint cette tranche d'âge") ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_  
    ),
    trouble_digestif_final_4150 = case_when(
      all(c_across(starts_with("trouble_digestif_41")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("trouble_digestif_41")) == "Oui") ~ "Oui", 
      all(c_across(starts_with("trouble_digestif_41")) == "Non") ~ "Non", 
      all(c_across(starts_with("trouble_digestif_41")) == "nsp") ~ "nsp", 
      any(c_across(starts_with("trouble_digestif_41")) == "Non") ~ "Non",  
      any(c_across(starts_with("trouble_digestif_41")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("trouble_digestif_41")) == "Pas atteint cette tranche d'âge") ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_  
    )
  ) %>%
  ungroup()

# 2. Select only new variables, add them to the main dataframe and remove original variables
Endo_dyspareunie <- Endo_dyspareunie %>% 
  left_join(new_dataframe %>%
              select(id_unik,
                     starts_with("trouble_digestif_final_")),
            by="id_unik"
  ) %>% 
  select(-c(trouble_digestif_15_1:trouble_digestif_4150_3),
         -trouble_digestif, trouble_digestif)

# 3. Delete the temporary dataframe
rm(new_dataframe)

#-------------------------------------
#   Dyschezia  : Onset Age group
#-------------------------------------
Endo_dyspareunie <- Endo_dyspareunie %>%
  mutate(onset_age_dyschezia = case_when(
    trouble_digestif_final_15 == "Oui" ~ "avant 15",
    trouble_digestif_final_1620 == "Oui" ~ "16-20",
    trouble_digestif_final_2130 == "Oui" ~ "21-30",
    trouble_digestif_final_3140 == "Oui" ~ "31-40",
    trouble_digestif_final_4150 == "Oui" ~ "41+",
    rowSums(across(starts_with("trouble_digestif_final_"), ~ . != "Oui")) == 5 ~ "pas encore et nsp",
    TRUE ~ NA_character_
  ))

#-------------------------------------------------------
# Level of pain when evacuation of 'selles' improvement
#-------------------------------------------------------
# 1. wide format to long format 
data_long <- Endo_dyspareunie %>%
  mutate(across(starts_with("niveau_doul_selles_"), ~ as.numeric(as.character(.)))) %>% 
  select(id_unik, starts_with("niveau_doul_selles_")) %>% 
  pivot_longer(cols = starts_with("niveau_doul_selles_"), 
               names_to = "age_group", 
               values_to = "score") %>%
  filter(!is.na(score)) %>%
  # make age groups (15-41) 
  mutate(age_group = case_when(
    str_detect(age_group, "15") ~ 15,
    str_detect(age_group, "1620") ~ 20,
    str_detect(age_group, "2130") ~ 30,
    str_detect(age_group, "3140") ~ 40,
    str_detect(age_group, "4150") ~ 41
  ))

# 2. Check the first and last score of severity of pain and compare these scores 
data_summary <- data_long %>%
  group_by(id_unik) %>%
  summarise(
    first_score = score[which.min(age_group)],
    last_score = score[which.max(age_group)]
  ) %>%
  mutate(
    doul_selles_improved = case_when(
      last_score < first_score ~"improved",
      last_score > first_score ~ "worsend",
      last_score == first_score ~"stable",
      TRUE ~ NA_character_ ))

# 3. Put new variable created to the original dataset 
Endo_dyspareunie <- Endo_dyspareunie %>% 
  left_join(
    data_summary %>% select(id_unik, doul_selles_improved),
    by="id_unik"
  )

# 4. Delete two temporary dataframe : data_long and data_summary
rm(data_long);rm(data_summary)

#-------------------------------------------------------------
#  Mean of the worst severity level of dyschezia over the entire life 
#-------------------------------------------------------------
# 1. Create a temporary dataframe 
new_dataframe <- Endo_dyspareunie %>%
  select(id_unik, starts_with("niveau_doul_selles_")) %>%
  mutate(
    total_score = rowSums(select(
      ., starts_with("niveau_doul_selles_")
    ), na.rm = TRUE),
    na_count = 5 - rowSums(is.na(select(
      ., starts_with("niveau_doul_selles_")
    ))),
    mean_niveau_severite_dyschezia = total_score / na_count
  ) %>%
  select(id_unik, mean_niveau_severite_dyschezia) 


# 2. Add a new categorical variables 
quantiles <- new_dataframe$mean_niveau_severite_dyschezia %>%
  quantile(probs=c(1/3,2/3), na.rm=TRUE) %>% 
  round()

new_dataframe <- new_dataframe %>%
  mutate(
    mean_niveau_severite_dyschezia_cat =
      case_when(
        mean_niveau_severite_dyschezia < quantiles[[1]] ~
          paste0("<", quantiles[1]),
        mean_niveau_severite_dyschezia >= quantiles[[1]] &
          mean_niveau_severite_dyschezia < quantiles[[2]] ~
          paste0(quantiles[1], "-", quantiles[2]),
        mean_niveau_severite_dyschezia >= quantiles[[2]] ~
          paste0(">=", quantiles[2])
      )
  )

# 3. Add a new variable to the main dataframe
Endo_dyspareunie <- Endo_dyspareunie %>% 
  left_join(new_dataframe, by="id_unik")

# 4. Delete the temporary dataframe and quantile information
rm(new_dataframe); rm(quantiles)

# 5. Modify the categorical variable by adding 'no pain' modality 
Endo_dyspareunie <- Endo_dyspareunie %>% 
  mutate(mean_niveau_severite_dyschezia_cat=
           case_when(
             trouble_digestif=="Non" ~
               "No_dyschezia_digestive_trouble",
             trouble_digestif=="Oui"&
               is.na(mean_niveau_severite_dyschezia_cat) ~ 
               "Only_digestive_trouble",
             TRUE ~ mean_niveau_severite_dyschezia_cat
           ))

#--------------------------------------
#  Dysuria : Oui/Non par group of age
#--------------------------------------
# 1. Create a temporary dataframe 
new_dataframe <- Endo_dyspareunie %>% 
  select(id_unik,starts_with("douleur_miction_")) %>%
  mutate(
    across(starts_with("douleur_miction_"), ~ case_when(
      . == "Je ne me souviens plus" ~ "nsp",
      . == "Non" ~ "Non",
      . %in% c("Oui, des difficultés à uriner nécessitant de contracter ou d'appuyer sur le ventre pour « aider » la vessie, ou une miction anormalement longue avec un jet urinaire faible, ou une miction longue à débuter, ou en plusieurs jets",
               "Oui, j'ai eu tendance à faire des infections urinaires",
               "Oui, une incontinence urinaire",
               "Oui, des douleurs lors de la miction (hors période d'infection urinaire)") ~ "Oui",
      . == "Je n'ai pas atteint cette tranche d'âge" ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_
    ))
  ) %>% 
  rowwise() %>% 
  mutate(
    douleur_miction_final_15  = case_when(
      all(c_across(starts_with("douleur_miction_15")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("douleur_miction_15")) == "Oui") ~ "Oui",
      all(c_across(starts_with("douleur_miction_15")) == "Non") ~ "Non",  
      all(c_across(starts_with("douleur_miction_15")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("douleur_miction_15")) == "Non") ~ "Non",  
      any(c_across(starts_with("douleur_miction_15")) == "nsp") ~ "nsp",
      TRUE ~ NA_character_  
    ),
    douleur_miction_final_1620 = case_when(
      all(c_across(starts_with("douleur_miction_16")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("douleur_miction_16")) == "Oui") ~ "Oui",  
      all(c_across(starts_with("douleur_miction_16")) == "Non") ~ "Non",  
      all(c_across(starts_with("douleur_miction_16")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("douleur_miction_16")) == "Non") ~ "Non",  
      any(c_across(starts_with("douleur_miction_16")) == "nsp") ~ "nsp",
      any(c_across(starts_with("douleur_miction_16")) == "Je n'ai pas atteint cette tranche d'âge") ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_  
    ),
    douleur_miction_final_2130 = case_when(
      all(c_across(starts_with("douleur_miction_21")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("douleur_miction_21")) == "Oui") ~ "Oui",  
      all(c_across(starts_with("douleur_miction_21")) == "Non") ~ "Non",  
      all(c_across(starts_with("douleur_miction_21")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("douleur_miction_21")) == "Non") ~ "Non",  
      any(c_across(starts_with("douleur_miction_21")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("douleur_miction_21")) == "Je n'ai pas atteint cette tranche d'âge") ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_ 
    ),
    douleur_miction_final_3140 = case_when(
      all(c_across(starts_with("douleur_miction_31")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("douleur_miction_31")) == "Oui") ~ "Oui",  
      all(c_across(starts_with("douleur_miction_31")) == "Non") ~ "Non",  
      all(c_across(starts_with("douleur_miction_31")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("douleur_miction_31")) == "Non") ~ "Non",  
      any(c_across(starts_with("douleur_miction_31")) == "nsp") ~ "nsp", 
      any(c_across(starts_with("douleur_miction_31")) == "Je n'ai pas atteint cette tranche d'âge") ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_  
    ),
    douleur_miction_final_4150 = case_when(
      all(c_across(starts_with("douleur_miction_41")) %in% NA) ~ NA_character_,  
      any(c_across(starts_with("douleur_miction_41")) == "Oui") ~ "Oui", 
      all(c_across(starts_with("douleur_miction_41")) == "Non") ~ "Non", 
      all(c_across(starts_with("douleur_miction_41")) == "nsp") ~ "nsp", 
      any(c_across(starts_with("douleur_miction_41")) == "Non") ~ "Non",  
      any(c_across(starts_with("douleur_miction_41")) == "nsp") ~ "nsp",  
      any(c_across(starts_with("douleur_miction_41")) == "Je n'ai pas atteint cette tranche d'âge") ~ "Pas atteint cette tranche d'âge",
      TRUE ~ NA_character_  
    )
  ) %>%
  ungroup() %>% 
  select(id_unik,starts_with("douleur_miction_final"))

# 2. Add new variables created to the main dataframe by deleting original variables
Endo_dyspareunie <- Endo_dyspareunie %>% 
  select(-starts_with("douleur_miction_")) %>% 
  left_join(new_dataframe, by="id_unik") %>% 
  select(-douleur_miction, douleur_miction)

# 3. Delete the temporary dataframe 
rm(new_dataframe)

#-------------------------------------
#   Dysuria  : Onset Age group
#-------------------------------------
Endo_dyspareunie <- Endo_dyspareunie %>%
  mutate(onset_age_dysuria = case_when(
    douleur_miction_final_15 == "Oui" ~ "avant 15",
    douleur_miction_final_1620 == "Oui" ~ "16-20",
    douleur_miction_final_2130 == "Oui" ~ "21-30",
    douleur_miction_final_3140 == "Oui" ~ "31-40",
    douleur_miction_final_4150 == "Oui" ~ "41+",
    rowSums(across(starts_with("douleur_miction_final_"), ~ . != "Oui"))==5 ~ "pas encore et nsp",
    TRUE ~ NA_character_
  ))

#-----------------------------------------------------------------------
#  Mean of the worst severity level of dysurenia over the entire life 
#-----------------------------------------------------------------------
# 1. Create a temporary dataframe 
new_dataframe <- Endo_dyspareunie %>%
  select(id_unik, starts_with("niveau_doul_miction_")) %>%
  mutate(
    total_score = rowSums(select(
      ., starts_with("niveau_doul_miction_")
    ), na.rm = TRUE),
    na_count = 5 - rowSums(is.na(select(
      ., starts_with("niveau_doul_miction_")
    ))),
    mean_niveau_severite_dysuria = round(total_score / na_count,1)
  ) %>%
  select(id_unik, mean_niveau_severite_dysuria) 

# 2. Add a new categorical variables 
quantiles <- new_dataframe$mean_niveau_severite_dysuria %>%
  quantile(probs=c(1/3,2/3), na.rm=TRUE) %>% 
  round()

new_dataframe <- new_dataframe %>%
  mutate(
    mean_niveau_severite_dysuria_cat =
      case_when(
        mean_niveau_severite_dysuria < quantiles[[1]] ~
          paste0("<", quantiles[1]),
        mean_niveau_severite_dysuria >= quantiles[[1]] &
          mean_niveau_severite_dysuria < quantiles[[2]] ~
          paste0(quantiles[1], "-", quantiles[2]),
        mean_niveau_severite_dysuria >= quantiles[[2]] ~
          paste0(">=", quantiles[2])
      )
  )

# 3. Add a new variable to the main dataframe
Endo_dyspareunie <- Endo_dyspareunie %>% 
  left_join(new_dataframe, by="id_unik")

# 4. Delete the temporary dataframe and quantile information
rm(new_dataframe); rm(quantiles)

# 5. Modify the categorical variable by adding 'no pain' modality 
Endo_dyspareunie <- Endo_dyspareunie %>% 
  mutate(mean_niveau_severite_dysuria_cat=
           case_when(
             douleur_miction=="Non" ~ "No pain",
             TRUE ~ mean_niveau_severite_dysuria_cat
           ))
```


# II. Final dataset
## 1) Variables selection 
```{r}
Endo_dyspareunie_final <- Endo_dyspareunie %>% 
  select(id_unik,
         # dyspareunia
         doul_rapport_sexuel, 
         first_rapport,
         onset_age_dyspareunia,
         #onset_type_dyspareunie,
         #recent_doul_rapport,
         #dyspareunie_improved,
         starts_with("doul_rapport_"),
         mean_niveau_severite_dyspareunia_cat,
        
         # dyschezia
        trouble_digestif,
        starts_with("trouble_digestif_final_"),
        onset_age_dyschezia,
        mean_niveau_severite_dyschezia_cat,
        
        # dysurenia
        douleur_miction,
        starts_with(c("douleur_miction_final_")),
        onset_age_dysuria,
        mean_niveau_severite_dysuria_cat,
        
        Endo_douleurs_participated,
        
        IdVersion,
        Date_naissance,
        DateValidation
        ) %>% 
  mutate(
    across(c(doul_rapport_sexuel,
                  trouble_digestif,
                  douleur_miction,
                  first_rapport,
                  mean_niveau_severite_dyspareunia_cat,
                  mean_niveau_severite_dyschezia_cat,
                  mean_niveau_severite_dysuria_cat,
                  onset_age_dysuria,
                  onset_age_dyspareunia,
                  onset_age_dyschezia
                  ), ~
                  case_when(
                    is.na(IdVersion) ~ "Not_answered",
                    TRUE  ~ .
                  )),
         across(starts_with("Date"), ~ 
                  as.Date(.,format="%d/%m/%Y"))) %>% 
  select(-IdVersion)
```

## 2) Final DataFrame with new variables  
```{r}
################################################################
# Objective = creating new binary variables : 
# - if there was a pain before registered lst pregnancy or not 
################################################################
#-------------------
#    Dysmenorrhea
#-------------------
# 1. Create a temporary new dataframe with information of pregnancy module 
new_dataframe <- Endo_dyspareunie_final %>% 
  left_join(grossesse_mod_final %>% 
              distinct(id_unik,first_preg_age,
                       grossesse_encours),
            by="id_unik")

# 2. Create a binary variable 
new_dataframe <- new_dataframe %>% 
  mutate(
    first_preg_age= first_preg_age %>% round(),
    age_inclusion = as.numeric(as.character(
      (DateValidation-Date_naissance)/365)) %>% round(),
    
    dyspareunia_avant_1_preg = case_when(
      # No answer provided in the questionnaire
      doul_rapport_sexuel == "Not_answered" ~ "Not_answered",
      
      doul_rapport_sexuel %in% c("Je ne souhaite pas répondre à ces questions",
                                 "Je n'ai jamais eu de rapport sexuel") ~
        "Non et nsp",
      
      # First pregnancy occurred after study inclusion
      first_preg_age > age_inclusion & doul_rapport_sexuel == "Oui" ~
        "Oui",
      first_preg_age > age_inclusion & doul_rapport_sexuel == "Non" ~
        "Non et nsp",
      
      # First pregnancy occurred before or at inclusion and no pain reported
      first_preg_age <= age_inclusion & doul_rapport_sexuel == "Non" ~
        "Non et nsp",
      
      # First pregnancy before or at inclusion and before age 15
      first_preg_age <= age_inclusion & first_preg_age <= 15 ~ case_when(
        doul_rapport_15 == "Oui" ~ "Oui",
        doul_rapport_15 == "Non, pas de douleurs" ~ "Non et nsp",
        doul_rapport_15 %in% c("Pas de rapport sexuel à cette période","Je ne me souviens plus") ~ 
          "Non et nsp"
      ),
      
      # First pregnancy between 16–20 years old
      between(first_preg_age, 16, 20) ~ case_when(
        first_preg_age <= age_inclusion & 
          (doul_rapport_15 == "Oui" | 
             doul_rapport_16 == "Oui") ~ 
          "Oui",
        first_preg_age < age_inclusion & 
          doul_rapport_15 == "Non" & 
          doul_rapport_16 == "Non, pas de douleurs" ~ 
          "Non et nsp",
        first_preg_age <= age_inclusion &
          (doul_rapport_15 %in% c("Pas de rapport sexuel à cette période","Je ne me souviens plus") |
           doul_rapport_16 %in% c("Pas de rapport sexuel à cette période","Je ne me souviens plus")) ~ 
          "Non et nsp"
      ),
      
      # First pregnancy between 21–30 years old
      between(first_preg_age, 21, 30) ~ case_when(
        first_preg_age <= age_inclusion &
          (doul_rapport_15 == "Oui" | 
             doul_rapport_16 == "Oui" | 
             doul_rapport_21 == "Oui") ~ 
          "Oui",
        first_preg_age <= age_inclusion & 
          doul_rapport_15 == "Non, pas de douleurs" & 
          doul_rapport_16 == "Non, pas de douleurs" & 
          doul_rapport_21 == "Non, pas de douleurs" ~ 
          "Non et nsp",
        first_preg_age <= age_inclusion &
          (doul_rapport_15 %in%c("Pas de rapport sexuel à cette période",
                                 "Je ne me souviens plus") |
             doul_rapport_16 %in% c("Pas de rapport sexuel à cette période",
                                 "Je ne me souviens plus") |
             doul_rapport_21 %in% c("Pas de rapport sexuel à cette période",
                                 "Je ne me souviens plus")) ~
          "Non et nsp"
      ),
      
      # First pregnancy between 31–40 years old
      between(first_preg_age, 31, 40) ~ case_when(
        first_preg_age <= age_inclusion & 
          (doul_rapport_15 == "Oui" | 
             doul_rapport_16 == "Oui" | 
             doul_rapport_21 == "Oui" | 
             doul_rapport_31 == "Oui") ~ "Oui",
        
        first_preg_age <= age_inclusion &
          doul_rapport_15 == "Non, pas de douleurs" & 
          doul_rapport_16 == "Non, pas de douleurs" &
          doul_rapport_21 == "Non, pas de douleurs" & 
          doul_rapport_31 == "Non, pas de douleurs" ~ 
          "Non et nsp",
        first_preg_age <= age_inclusion &
          (doul_rapport_15 %in% c("Pas de rapport sexuel à cette période",
                                 "Je ne me souviens plus") |
           doul_rapport_16 %in% c("Pas de rapport sexuel à cette période",
                                 "Je ne me souviens plus") |
           doul_rapport_21 %in% c("Pas de rapport sexuel à cette période",
                                 "Je ne me souviens plus") |
           doul_rapport_31 %in% c("Pas de rapport sexuel à cette période",
                                 "Je ne me souviens plus")) ~
          "Non et nsp"
      ),
      
      # First pregnancy at or after age 41
      first_preg_age >= 41 ~ case_when(
        first_preg_age <= age_inclusion & 
          (doul_rapport_15 == "Oui" | doul_rapport_16 == "Oui" |
           doul_rapport_21 == "Oui" | doul_rapport_31 == "Oui" |
           doul_rapport_41 == "Oui") ~ 
          "Oui",
        first_preg_age <= age_inclusion &
          doul_rapport_15 == "Non, pas de douleurs" & 
          doul_rapport_16 == "Non, pas de douleurs" &
          doul_rapport_21 == "Non, pas de douleurs" & 
          doul_rapport_31 == "Non, pas de douleurs" &
          doul_rapport_41 == "Non, pas de douleurs" ~ 
          "Non et nsp",
        first_preg_age <= age_inclusion &
          (doul_rapport_15 %in% 
             c("Pas de rapport sexuel à cette période",
               "Je ne me souviens plus") |
           doul_rapport_16 %in% 
             c("Pas de rapport sexuel à cette période",
               "Je ne me souviens plus") |
           doul_rapport_21 %in% 
             c("Pas de rapport sexuel à cette période",
               "Je ne me souviens plus") |
           doul_rapport_31 %in% 
             c("Pas de rapport sexuel à cette période",
               "Je ne me souviens plus") |
           doul_rapport_41 %in% 
             c("Pas de rapport sexuel à cette période",
               "Je ne me souviens plus")) ~
          "Non et nsp"
      ),
      
      TRUE ~ NA_character_
      )
    )

#--------------------------------------------------------------------
# Function : Automatically classify the presence of pain 
# in specific age groups before the registered 1st pregnancy
#--------------------------------------------------------------------
# 'douleur_miction' and 'trouble_digestif' having same modalities
#--------------------------------------------------------------------
# 1. Create a function for classify the presence of pain
classify_douleur_avant_1_preg <-
  function(data,
           var_bin_YN,
           var_15 = NULL,
           var_16 = NULL,
           var_21 = NULL,
           var_31 = NULL,
           var_41 = NULL) {
    data %>%
      mutate(
        result = case_when(
          !!sym(var_bin_YN) == "Not_answered" ~ "Not_answered",
          
          first_preg_age > age_inclusion &
            !!sym(var_bin_YN) == "Oui" ~ "Oui",
          first_preg_age > age_inclusion &
            !!sym(var_bin_YN) == "Non" ~ "Non et nsp",
          
          first_preg_age <= age_inclusion &
            !!sym(var_bin_YN) == "Non" ~ "Non et nsp",
          
          first_preg_age <= age_inclusion &
            first_preg_age <= 15 ~ case_when(
              !!sym(var_15) == "Oui" ~ "Oui",!!sym(var_15) == "Non" ~ "Non et nsp",
              !!sym(var_15) == "nsp" ~ "Non et nsp"
            ),
          
          between(first_preg_age, 16, 20) ~ case_when(
            first_preg_age <= age_inclusion &
              (!!sym(var_15) == "Oui" |
                 (!!sym(var_16) %||% "") == "Oui") ~ "Oui",
            first_preg_age <= age_inclusion &
              !!sym(var_15) == "Non" &
              (!!sym(var_16) %||% "") == "Non" ~ "Non et nsp",
            first_preg_age <= age_inclusion &
              (!!sym(var_15) == "nsp" |
                 (!!sym(var_16) %||% "") == "nsp") ~ "Non et nsp"
          ),
          
          between(first_preg_age, 21, 30) ~ case_when(
            first_preg_age <= age_inclusion &
              ("Oui" %in% c(
                !!sym(var_15), 
                !!sym(var_16), 
                !!sym(var_21)
              )) ~ "Oui",
            first_preg_age <= age_inclusion &
              all(c(
                !!sym(var_15), 
                !!sym(var_16), 
                !!sym(var_21)
              ) == "Non") ~ "Non et nsp",
            first_preg_age <= age_inclusion &
              any(c(
                !!sym(var_15), 
                !!sym(var_16), 
                !!sym(var_21)
              ) == "nsp") ~ "Non et nsp"
          ),
          
          between(first_preg_age, 31, 40) ~ case_when(
            first_preg_age <= age_inclusion &
              ("Oui" %in% c(
                !!sym(var_15), 
                !!sym(var_16), 
                !!sym(var_21), 
                !!sym(var_31)
              )) ~ "Oui",
            first_preg_age <= age_inclusion &
              all(c(
                !!sym(var_15), 
                !!sym(var_16), 
                !!sym(var_21), 
                !!sym(var_31)
              ) == "Non") ~ "Non et nsp",
            first_preg_age <= age_inclusion &
              any(c(
                !!sym(var_15), 
                !!sym(var_16), 
                !!sym(var_21), 
                !!sym(var_31)
              ) == "nsp") ~ "Non et nsp"
          ),
          
          first_preg_age >= 41 ~ case_when(
            first_preg_age <= age_inclusion &
              ("Oui" %in% c(
                !!sym(var_15),
                !!sym(var_16),
                !!sym(var_21),
                !!sym(var_31),
                !!sym(var_41)
              )) ~ "Oui",
            first_preg_age <= age_inclusion &
              all(c(
                !!sym(var_15),
                !!sym(var_16),
                !!sym(var_21),
                !!sym(var_31),
                !!sym(var_41)
              ) == "Non") ~ "Non et nsp",
            first_preg_age <= age_inclusion &
              any(c(
                !!sym(var_15),
                !!sym(var_16),
                !!sym(var_21),
                !!sym(var_31),
                !!sym(var_41)
              ) == "nsp") ~ "Non et nsp"
          )
        )
      )
  }

# 2. Apply the new function 'classify_douleur_avant_1_preg'
new_dataframe <- new_dataframe %>%
  left_join(
    classify_douleur_avant_1_preg(
      data = new_dataframe,
      var_bin_YN = "trouble_digestif",
      var_15 = "trouble_digestif_final_15",
      var_16 = "trouble_digestif_final_1620",
      var_21 = "trouble_digestif_final_2130",
      var_31 = "trouble_digestif_final_3140",
      var_41 = "trouble_digestif_final_4150"
    ) %>%
      select(id_unik, dyschezia_avant_1_preg = result),
    by = "id_unik"
  ) %>% 
  left_join(
    classify_douleur_avant_1_preg(
      data = new_dataframe,
      var_bin_YN = "douleur_miction",
      var_15 = "douleur_miction_final_15",
      var_16 = "douleur_miction_final_1620",
      var_21 = "douleur_miction_final_2130",
      var_31 = "douleur_miction_final_3140",
      var_41 = "douleur_miction_final_4150"
    ) %>%
      select(id_unik, dysuria_avant_1_preg = result),
    by = "id_unik"
  )

# 3. Add new variables to the main final dataframe 
Endo_dyspareunie_final <- new_dataframe %>% 
  select(id_unik,
         doul_rapport_sexuel,
         first_rapport,
         onset_age_dyspareunia,
         mean_niveau_severite_dyspareunia_cat,
         dyspareunia_avant_1_preg,
         
         trouble_digestif,
         onset_age_dyschezia,
         mean_niveau_severite_dyschezia_cat,
         dyschezia_avant_1_preg,
         
         douleur_miction,
         onset_age_dysuria,
         mean_niveau_severite_dysuria_cat,
         dysuria_avant_1_preg,
         
         Endo_douleurs_participated
         )

# 4. Delete the temporary new_dataframe
rm(new_dataframe)
```

# III. Save the final dataframe 

```{r}
#write.csv(Endo_dyspareunie_final,"Endo_dyspareunie_final.csv")
```



