---
title: "01_clean_Treatment_burden"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Treatment_burden
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-05-13
*Date of update* : 2025-05-13
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 
- Les résultats concernant "le fardeau du traitement" des 2400 premiers participants à ComPaRe ont été décrits dans  une publication dans la revue Mayo Clinic Proceedings en octobre 2019 (en anglais).
```{r}
#-----------------------
# Import a main dataset
#-----------------------
fardeau_trt <- read.csv(
  "Complic_pop_Fardeau_trt_20250407.csv",
  encoding = "latin1",
  colClasses = c(Sexe = "character")
)

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
fardeau_trt <- fardeau_trt %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- fardeau_trt %>% 
  sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count == nrow(fardeau_trt)) %>%
  as.vector() # index of columns having only NA
```


# I. Preprocessing Data

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
fardeau_trt <- fardeau_trt %>% 
  mutate(fardeau_trt_participated= ifelse(is.na(IdAction),
                                         "Non","Oui"))
#-------------------------
#  Character to numeric
#-------------------------
fardeau_trt <-  fardeau_trt %>% 
  mutate(sc_TBQ= sc_TBQ %>% as.numeric())

#----------------------------------
# Score : numeric -> categorical
#----------------------------------
# 1. Calculate quantiles
quantiles_TBQ <- quantile(fardeau_trt$sc_TBQ, 
                          probs = c(1/3, 2/3), na.rm=TRUE)

# 2. Create new categorical variables 
fardeau_trt <- fardeau_trt %>% 
  mutate(
    sc_TBQ_cat= 
      case_when(
        sc_TBQ< quantiles_TBQ[[1]] ~ 
          paste0("<", round(quantiles_TBQ[1], 2)),
        sc_TBQ>= quantiles_TBQ[[1]] & sc_TBQ< quantiles_TBQ[[2]] ~
          paste0(round(quantiles_TBQ[1], 2), "-", round(quantiles_TBQ[2], 2)),
        sc_TBQ>= quantiles_TBQ[[2]] ~ 
          paste0(">=", round(quantiles_TBQ[2], 2))
          ) %>% 
      as.factor()
    )

# 3. Remove the quantiles data
rm(quantiles_TBQ)
```

# II. Final Dataset

```{r}
fardeau_trt_final <- fardeau_trt %>% 
  distinct(id_unik,
           IdAction,
           sc_TBQ_cat, 
           fardeau_trt_participated) %>% 
  mutate(sc_TBQ_cat = case_when(
                    is.na(IdAction) ~ "Not_answered",
                    TRUE  ~ sc_TBQ_cat
                  )) %>%
  select(-IdAction)
```

# III. Save the final dataframe 

```{r}
#write.csv(fardeau_trt_final,"fardeau_trt_final.csv")
```

