---
title: "01_clean_Endo_chirurgie_examen"
author: "Minji NAM"
date: "2025-05-13"
output: html_document
---

*Name* : 01_clean_Endo_chirurgie_examen
*Project* : Santé reproductive, infertilité et **complications de la grossesse** chez les femmes atteintes d'endométriose
*Description* : Preprocessing Data
*Date of creation* : 2025-03-05
*Date of update* : 2025-05-15
*Author* : stagiaire M2 Big Data in Health
*R version* : R version 4.4.2

---

$Libraries$
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(stringr)
library(tableone)
library(nnet)
library(broom)
library(gtsummary)
library(rlang) 
library(labelled)
library(questionr)
library(guideR)
library(jmv)
library(nnet)
library(stringi)
```

# 0. Import the raw dataset 

```{r}
#-----------------------
# Import a main dataset
#-----------------------
Endo_parcours_chir_examens <- read.csv(
  "Complic_pop_QST_0470_20250407_chirurgie_imagerie.csv",
  encoding = "latin1",
  colClasses = c(Sexe = "character")
)

#-----------------------------------------------
# Replace cells with nothing or spaces with NA
#-----------------------------------------------
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  mutate(across(everything(), ~na_if(trimws(.), "")),
         across(everything(), str_trim))

#-------------------------------------
# Remove all columns having only NAs
#-------------------------------------
na_count <- Endo_parcours_chir_examens %>% sapply(function(x) sum(is.na(x))) #; na_count
index_col_NA <- which(na_count==nrow(Endo_parcours_chir_examens)) %>% as.vector() # index of columns having only NA
Endo_parcours_chir_examens <- Endo_parcours_chir_examens[,-index_col_NA]

```

# I. Preprocessing Data

```{r}
#---------------------------
#  Participation : Yes/No
#---------------------------
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  mutate(Endo_chir_examens_participated=ifelse(is.na(IdAction),
                                                "Non","Oui"))
#--------------------------------
# Reason for the first surgery
#--------------------------------
# 1. Selecting and Cleaning Columns
new_dataframe <- Endo_parcours_chir_examens %>%
  dplyr::select(id_unik,nb_operation_endo,starts_with("motif_operation"))

# 2. create indicator variables for the reason for surgery
categories <- c(
  "Douleurs pelviennes",
  "Confirmation de suspicion d'endométriose/adénomyose (diagnostic chirurgical)",
  "Infertilité",
  "Chirurgie pour une autre indication (découverte fortuite d'une endométriose/adénomyose)",
  "Autre",
  "Je ne sais pas"
)

indic_categories <- c(
  "indic_motif_1_doul_pelv",
  "indic_motif_1_confirm_diag",
  "indic_motif_1_infertilite",
  "indic_motif_1_chir",
  "indic_motif_1_autre",
  "indic_motif_1_nsp"
)

# 3. Create the function for the new variable
variable_motif_operation1 <- function(nb_op, motif1, part) {
  if (part %in% motif1) {
    return("Opérée pour ce motif")
  } else if (!is.na(nb_op) && nb_op == 0) {
    return("N'a eu aucune opération")
  } else {
    if (!is.na(nb_op) && nb_op != 0) {
      return("N'a pas été opérée pour ce motif")
    } else {
      return(NA)
    }
  }
}

# 4. Apply the function created
new_dataframe[indic_categories] <- t(           # Transpose to ensure correct column-wise assignment
  apply(new_dataframe, 1, function(row) {       # Apply function row-wise on new_dataframe
  nb_op <- suppressWarnings(as.numeric(row[54])) # Convert to numeric to avoid errors
  motif1 <- row[2:6]  # Select columns corresponding to the first operation motives
  sapply(categories, function(part) variable_motif_operation1(nb_op, motif1, part))# Apply function to each category
}))

# 5. Create two new variables : Categorical and binary
new_dataframe <- new_dataframe %>%
  rowwise() %>%
  mutate(
    motifs_operation1 = 
      list(na.omit(c_across(starts_with("motif_operation1_")))) ,
    motifs_operation1_num = length(motifs_operation1),
    motifs_operation1_cat = case_when(
      nb_operation_endo == 0 ~ "pas d'OP",  
      motifs_operation1_num == 1 ~ 
        paste0("only_", first(motifs_operation1)),
      motifs_operation1_num > 1 ~ "multiples motifs",
      TRUE ~ NA_character_
    ),
    operation_endo_bin = case_when(
      nb_operation_endo == 0 ~ 0,
      nb_operation_endo >=1 ~ 1,
      TRUE ~ NA_real_
    )
  ) %>%
  ungroup()

#--------------------------------------------
#   Reason for more than second operation
#--------------------------------------------
# 1. Create indicator variables for the reason of the second operation and more
categories <- c(
  "Chirurgie incomplète lors d'une précédente intervention",
  "Persistance des douleurs (les douleurs n'ont pas disparu suite à l’opération précédente)",
  "Récidive des douleurs (réapparition des douleurs après une période sans douleur)",
  "Apparition de nouvelles douleurs",
  "Prise en charge d’une complication d’une précédente chirurgie pour endométriose/adénomyose",
  "Infertilité",
  "Autre"
)
indic_categories <- c("indic_motif_2_chir_inc",
                      "indic_motif_2_doul_persist",
                      "indic_motif_2_recidive",
                      "indic_motif_2_nouv_doul",
                      "indic_motif_2_complication",
                      "indic_motif_2_infertilite",
                      "indic_motif_2_autre"
)

# 2. Create the function 
variable_motif_operationsup1 <- function(nb_op, motif2, part) {
  if (part %in% motif2) {
    return("Opérée pour ce motif")
  } else if (!is.na(nb_op) && (nb_op == 0 || nb_op == 1)) {
    return("A eu moins de 2 opérations")
  } else if (!is.na(nb_op) && nb_op != 0 && nb_op != 1) {
    return("N'a pas été opérée pour ce motif")
  } else {
    return(NA)
  }
}

# 3. Apply the function to each column
# Prepare the index of columns
index_debut <- which(startsWith(names(new_dataframe), "motif_operation2")) %>% head(1)

index_fin <- which(startsWith(names(new_dataframe), "motif_operation")) %>% tail(1)

new_dataframe[indic_categories] <- t(apply(new_dataframe, 1, function(row) {
  nb_op <- suppressWarnings(as.numeric(row[ncol(new_dataframe)]))
  # Sélection des colonnes entre index_debut et index_fin
  motif2 <- row[index_debut:index_fin]
  # Appliquer la fonction pour chaque catégorie
  sapply(categories, function(part) variable_motif_operationsup1(nb_op, motif2, part))
}))

# 4. Create new variables : categorical
# Convertir les indices en noms de colonnes pour les utiliser avec all_of()
colonnes_motif <- names(new_dataframe)[index_debut:index_fin]

new_dataframe <- new_dataframe %>%
  rowwise() %>%
  mutate(
    motifs_operation2 = list(na.omit(c_across(all_of(colonnes_motif)))),
    motifs_operation2_num = length(motifs_operation2),
    motifs_operation2_cat = case_when(
      nb_operation_endo == 0 ~ "pas d'OP",
      motifs_operation2_num == 1 ~ paste0("only_", first(motifs_operation2)),
      motifs_operation2_num > 1 ~ "multiples motifs",
      TRUE ~ NA_character_
    )
  ) %>%
  ungroup()


# 5. Merge the new indicator into the main dataset
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  left_join(
    new_dataframe %>% 
      dplyr::select(id_unik,operation_endo_bin,motifs_operation1_cat,motifs_operation2_cat, starts_with("indic_motif_")),
    by = "id_unik"
  )

# Remove the temporary dataset and certain variables 
rm(new_dataframe)


#------------------------------------------------------------------
#   infertility an complication reason for 2nd ~6th operation
#------------------------------------------------------------------
motif_cols <- names(Endo_parcours_chir_examens) %>%
  str_subset("^motif_operation[123456]_")

new_df <- Endo_parcours_chir_examens %>%
  dplyr::select(id_unik, nb_operation_endo, all_of(motif_cols))

make_indicator <- function(motifs, target) {
  if (is.null(motifs) || length(motifs) == 0 || all(is.na(motifs))) {
    return(NA_character_)
  } else if (target %in% motifs) {
    return("Opérée pour ce motif")
  } else {
    return("N'a pas été opérée pour ce motif")
  }
}

for (i in 1:6) {
  motif_cols_i <- names(new_df)[str_detect(names(new_df), paste0("^motif_operation", i, "_"))]
  
  if (length(motif_cols_i) == 0) {
    next  
  }
  
  new_df <- new_df %>%
    rowwise() %>%
    mutate(
      !!paste0("indic_motif_", i, "_infertilite") :=
        make_indicator(c_across(all_of(motif_cols_i)), "Infertilité"),
      
      !!paste0("indic_motif_", i, "_complication") :=
        make_indicator(c_across(all_of(motif_cols_i)), "Complication liée à la chirurgie")
    ) %>%
    ungroup()
}

new_df <- new_df %>% 
  mutate(
    indic_motif_2ndOP_infertilite = indic_motif_2_infertilite,
    indic_motif_3rdOP_infertilite = indic_motif_3_infertilite,
    indic_motif_4thOP_infertilite = indic_motif_4_infertilite,
    indic_motif_5thOP_infertilite = indic_motif_5_infertilite,
    indic_motif_6thOP_infertilite = indic_motif_6_infertilite,
    
    indic_motif_2ndOP_complication = indic_motif_2_complication,
    indic_motif_3rdOP_complication = indic_motif_3_complication
    
  )%>%
  dplyr::select(id_unik,
                indic_motif_2ndOP_infertilite:indic_motif_6thOP_infertilite,
                indic_motif_2ndOP_complication,indic_motif_3rdOP_complication)

Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  left_join(new_df, by="id_unik")

rm(new_df)


#-------------------------------
#   Duration of complications   : duree_complic_
#-------------------------------
# Recode complication duration by setting 0 for those who have no complication
duration_cols <- c(
  "duree_complic_vessie",
  "duree_complic_autosondage",
  "duree_complic_stomie",
  "duree_complic_fistule",
  "duree_complic_fistule_dig",
  "duree_complic_constip",
  "duree_complic_aug_doul",
  "duree_complic_nvelle_douleur",
  "duree_complic_autre_complication"
)

indicator_cols <-
  c(
    "vider_vessie",
    "autosondage_urinaire",
    "stomie_urinaire",
    "fistule",
    "fistule_dig",
    "constipation",
    "aug_douleur",
    "app_nvelle_douleur",
    "autre_complication"
  )

# Create new variables : tercile_duree_complic_XXX
for (i in seq_along(duration_cols)) {
  duration_col <- duration_cols[i]
  indicator_col <- indicator_cols[i]
  tercile_col <- paste0("tercile_", duration_col)
  
  # Calculate tercile thresholds
  quantiles <- quantile(as.numeric(Endo_parcours_chir_examens[[duration_col]]), 
                        probs = c(1/3, 2/3), na.rm = TRUE)
  
  Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
    mutate(!!tercile_col := ntile(!!sym(duration_col), 3)) %>% # Divide into 3 equal groups
    mutate(!!tercile_col := if_else(
      !!sym(indicator_col) == "Je n'ai jamais souffert de cette complication",
      "N'a jamais souffert de cette complication",  
      as.character(!!sym(tercile_col))
    )) %>%
    mutate(!!sym(tercile_col) := factor(
      !!sym(tercile_col), 
      levels = c("N'a jamais souffert de cette complication","1", "2", "3"), 
      labels = c(
        "N'a jamais souffert de cette complication",
        paste0("<", round(quantiles[1], 2)), 
        paste0(round(quantiles[1], 2), "-", round(quantiles[2], 2)), 
        paste0(">", round(quantiles[2], 2))
      )
    , ordered = TRUE))
}

# Recreate 'tercile_duree_complic_autre_complication' by using autre_complication variable
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(
    tercile_duree_complic_autre_complication = as.character(ifelse(
      autre_complication == "Je n'ai pas souffert d'autres complications",
      "Pas de complication",
      as.character(tercile_duree_complic_autre_complication)
    )
  ))

# Remove all variables related duree_complic
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  dplyr::select(-starts_with("duree_complic"))

#-------------------
#  Ovary indicator
#-------------------
# 1. Selecting and Cleaning Columns
new_dataframe <- Endo_parcours_chir_examens %>%
  dplyr::select(id_unik,starts_with("ovaires_touches_"))

# 2. create indicator variables 
indic_categories <- c("indic_ovaire_touche")

# 3. Create the function for the new variable 
# Prepare the index of columns
index_debut <- which(startsWith(names(new_dataframe), "ovaires_touches_")) %>% head(1)
index_fin <- which(startsWith(names(new_dataframe), "ovaires_touches_")) %>% tail(1)
variable_ovaire <- function(row) {
  ovaire <- row[index_debut:index_fin]
  case_when(
    any(c("Oui, le gauche", "Oui, le droit") %in% ovaire) ~ 
      "Ovaire touchée",
    "Je ne sais pas" %in% ovaire ~ "Ne sait pas",
    "Non" %in% ovaire ~ "Non touchée",
    TRUE ~ NA_character_
  )
}

# 4. Apply the function 
new_dataframe$indic_ovaire_touche <- apply(new_dataframe, 1, function(row) { # 1= Row-wise
  variable_ovaire(row)})

# 5. Merge the new indicator into the main dataset
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  left_join(new_dataframe %>% dplyr::select(id_unik, indic_ovaire_touche),
            by="id_unik")
  
# 6. Remove the temporary dataset and non-necessary variables 
rm(new_dataframe)
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  dplyr::select(-starts_with("ovaires_touches_"))

#------------------------------
# Indicator of affected organ
#------------------------------
# 1. Select columns
new_dataframe <- Endo_parcours_chir_examens %>% 
  dplyr::select(id_unik, starts_with("organe_touche_"))

# 2. create indicator variables 
organes <- c(
  "Trompe droite",
  "Trompe gauche",
  "Vagin",
  "Utérus",
  "Appareil digestif (iléon ou autre partie de l’intestin grêle, dernière anse iléale, cæcum, appendice, côlon, sigmoïde, charnière recto-sigmoïdienne, rectum, …)",
  "Vessie",
  "Rein(s)",
  "Uretères",
  "Ligaments utéro-sacrés (LUS)",
  "Ligaments ronds",
  "Diaphragme",
  "Autre(s) organe(s) ou localisation plus précise"
)

indic_organes <-
  c(
    "indic_organe_touche_trompe_droite",
    "indic_organe_touche_trompe_gauche",
    "indic_organe_touche_vagin",
    "indic_organe_touche_uterus",
    "indic_organe_touche_dig",
    "indic_organe_touche_vessie",
    "indic_organe_touche_reins",
    "indic_organe_touche_ureteres",
    "indic_organe_touche_liga_utero",
    "indic_organe_touche_liga_rond",
    "indic_organe_touche_diaphragme",
    "indic_organe_touche_autres"
  )

# 3. Create the function for the new variable 
# Prepare the index of columns
index_debut <- which(startsWith(names(new_dataframe), "organe_touche_")) %>% head(1)
index_fin <- which(startsWith(names(new_dataframe), "organe_touche_")) %>% tail(1)

variable_organe_touche <- function(row, part) {
  organe <- row[index_debut:index_fin]
  
  if (part %in% organe) {
    return("organe touché")
  } else if ("Non" %in% organe) {
    return("Aucun organe touché")
  } else if (any(!is.na(organe))) {
    return("A un autre organe touché ou nsp")
  } else {
    return(NA)
  }
}

# 4. Apply the function 
for (i in seq_along(organes)) {
  part <- organes[i]
  col_name <- indic_organes[i]
  
  new_dataframe[[col_name]] <- apply(new_dataframe, 1, function(row) {
    variable_organe_touche(row, part)
  })
}

# 5. Merge the new indicator into the main dataset
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  left_join(
    new_dataframe %>% dplyr::select(-starts_with("organe_touche_")),
    by = "id_unik"
  )

# 6. Remove the temporary dataset and non-necessary variables 
rm(new_dataframe)
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  dplyr::select(-starts_with("organe_touche_"))

#----------------------------
#  Affected fallopian tube   : based on new Indicator of affected organ 
#----------------------------
# 1. Selecting and Cleaning Columns
new_dataframe <- Endo_parcours_chir_examens %>% 
  dplyr::select(id_unik, starts_with("indic_organe_touche_trompe_"))

# 2. Create indicator variables
trompe <-
  c("indic_organe_touche_trompe_droite",
    "indic_organe_touche_trompe_gauche")

# 3. Create the function for the new variable
# Prepare the index of columns
index_debut <- which(startsWith(names(new_dataframe), "indic_organe_touche_trompe_")) %>% head(1)
index_fin <- which(startsWith(names(new_dataframe), "indic_organe_touche_trompe_")) %>% tail(1)
variable_trompe_touche <- function(row) {
  trompe <- row[index_debut:index_fin]
  
  case_when(
    "organe touché" %in% trompe ~ "organe touché",
    "Aucun organe touché" %in% trompe ~ "Aucun organe touché",
    "A un autre organe touché ou nsp" %in% trompe ~ "A un autre organe touché ou nsp",
    TRUE ~ NA_character_
  )
}

# 4. Apply the created function 
new_dataframe$indic_organe_touche_trompe <- apply(new_dataframe, 1, function(row) {
  variable_trompe_touche(row)
})  

# 5. Merge the new indicator into the main dataset
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  left_join(
    new_dataframe %>% dplyr::select(id_unik,indic_organe_touche_trompe),
    by = "id_unik"
  )

# 6. Remove the temporary dataset and non-necessary variables
rm(new_dataframe)
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  dplyr::select(-starts_with("indic_organe_touche_trompe_"))

#------------------
#  adhesion level
#------------------
# 1. Select and clean columns
new_dataframe <- Endo_parcours_chir_examens %>% 
  dplyr::select(id_unik, starts_with("niveau_adherence_"), adherence)

# 2. Create two new variables : numeric,Categorical(organes_enleves_cat) and binary(organes_enleves_bin)
new_dataframe <- new_dataframe %>%
  rowwise() %>%
  mutate(
    niveaux_adherences = 
      list(na.omit(c_across(starts_with("niveau_adherence_")))) ,
    niveaux_adherences_num = length(niveaux_adherences),
    niveaux_adherences_cat = case_when(
      adherence %in% c("Non", "Je ne sais pas") ~ "Non",  
      adherence == "Oui" & niveaux_adherences_num == 1 ~ paste0("only_", first(niveaux_adherences)),
      adherence == "Oui" & niveaux_adherences_num > 1 ~ "multiples adherences",
      TRUE ~ NA_character_
    ),
    niveaux_adherences_bin = case_when(
      adherence %in% c("Non", "Je ne sais pas") ~ 0,
      adherence == "Oui" ~ 1,
      TRUE ~ NA_real_
    )
  ) %>%
  ungroup()


# 2. Create indicator variables
categories <- c(
  "Des ovaires",
  "Des trompes",
  "Appareil digestif (iléon ou autre partie de l’intestin grêle, dernière anse iléale, cæcum, appendice, côlon, sigmoïde, charnière recto-sigmoïdienne, rectum, …)",
  "Utérus",
  "Vessie",
  "Autre(s)"
)
indic_adhe <-
  c(
    "indic_adherence_ovaires",
    "indic_adherence_trompes",
    "indic_adherence_dig",
    "indic_adherence_uterus",
    "indic_adherence_vessie",
    "indic_adherence_autres"
  )

# 3. Create the function for the new variable
# Prepare the index of columns
index_debut <- which(startsWith(names(new_dataframe), "niveau_adherence_")) %>% head(1)
index_fin <- which(startsWith(names(new_dataframe), "niveau_adherence_")) %>% tail(1)
variable_adherence<- function(row, part) {
  adhe <- row[index_debut:index_fin] # index of variables related to the adhesion
  case_when(
    part %in% adhe ~ "Adhérence",
    "Non" %in% adhe ~ "Aucune zone où présence d'adhérence",
    any(!is.na(adhe))~ "A de l'adhérence ailleurs ou nsp",
    TRUE ~ NA_character_
  )
}

# 4. Apply the created function
for (i in seq_along(categories)) {
  part <- categories[i]
  col_name <- indic_adhe[i]

  new_dataframe[[col_name]] <- apply(new_dataframe, 1, function(row) {
    variable_adherence(row, part)
  })
}

# 5. Merge the new indicator into the main dataset
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  left_join(
    new_dataframe %>% dplyr::select(id_unik,starts_with("indic_adherence_"),niveaux_adherences_cat,niveaux_adherences_bin),
    by = "id_unik"
  )

# 6. Remove the temporary dataset and non-necessary variables
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  dplyr::select(-starts_with("niveau_adherence_"), -adherence)
rm(new_dataframe)

#-----------------
#  Removed organ  : removed organ (numeric,categorical, binaire)
#-----------------
# 1. Select and clean columns
new_dataframe <- Endo_parcours_chir_examens %>% 
  dplyr::select(id_unik, starts_with("organes_enleves_"))

# 2. Create two new variables : Categorical(organes_enleves_cat) and binary(organes_enleves_bin)
new_dataframe <- new_dataframe %>%
  rowwise() %>%
  mutate(
    organes_enleves = list(na.omit(c_across(starts_with("organes_enleves_")))) ,
    organes_enleves_num = length(organes_enleves),
    organe_enleve_cat = case_when(
      organes_enleves_num == 0 ~ NA_character_,  
      all(organes_enleves == "Non") ~ "Non",    
      organes_enleves_num == 1 ~ paste0("only_", first(organes_enleves)),
      organes_enleves_num > 1 & any(c("Un ovaire", "Une trompe", "Utérus", "Les deux ovaires", "Les deux trompes") %in% organes_enleves) ~ "multiples organes incluant ovaire/trompe/uterus",
      organes_enleves_num > 1 & all(organes_enleves %in% c("Non", "Autre(s)")) ~ "only_Autre(s)",  
      organes_enleves_num > 1 ~ "multiples organes hors de ovaire/trompe/uterus",
      TRUE ~ NA_character_
    ),
    organe_enleve_bin = case_when(
      organe_enleve_cat %in% c("Non", "only_Je ne sais pas") ~ 0,  
      !is.na(organe_enleve_cat) ~ 1,
      TRUE ~ NA_real_ 
    )
  ) %>%
  ungroup() 

# 3. Create indicator variables
categories_suppression <- c(
  "Un ovaire",
  "Les deux ovaires",
  "Une trompe",
  "Les deux trompes",
  "Utérus",
  "Ligaments utéro-sacrés (LUS)",
  "Intestin ou rectum",
  "Vessie",
  "Rein(s)",
  "Uretère(s)",
  "Autre(s)"
)
indic_suppression <-
  c(
    "indic_organe_enleve_1ovaire",
    "indic_organe_enleve_2ovaires",
    "indic_organe_enleve_1trompe",
    "indic_organe_enleve_2trompes",
    "indic_organe_enleve_uterus",
    "indic_organe_enleve_liga_utero",
    "indic_organe_enleve_intestin",
    "indic_organe_enleve_vessie",
    "indic_organe_enleve_rein",
    "indic_organe_enleve_uretere",
    "indic_organe_enleve_autre"
  )

# 4. Create the function for the new variable
# Prepare the index of columns
index_debut <- which(startsWith(names(new_dataframe), "organes_enleves_")) %>% head(1)
index_fin <- which(startsWith(names(new_dataframe), "organes_enleves_")) %>% tail(1)
variable_suppression <- function(row, part) {
  enlevé <- row[index_debut:index_fin] # index of variables
  case_when(
    part %in% enlevé ~ "organe enlevé",
    "Non" %in% enlevé ~ "Aucun organe enlevé",
    any(!is.na(enlevé))~ "A un autre organe enlevé ou nsp",
    TRUE ~ NA_character_
  )
}

# 5. Apply the created function
for (i in seq_along(categories_suppression)) {
  part <- categories_suppression[i]
  col_name <- indic_suppression[i]
  new_dataframe[[col_name]] <- as.factor(apply(new_dataframe, 1, function(row) {
    variable_suppression(row, part)
  }))
}

# 6. Merge the new indicator into the main dataset
 Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
   left_join(
     new_dataframe %>% dplyr::select(-starts_with("organes_enleves")),
     by="id_unik"
   )

# 7. Remove the temporary dataset and non-necessary variables
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  dplyr::select(-starts_with("organes_enleves_"))
rm(new_dataframe)

#-----------------------------
#  Location of endometriosis
#-----------------------------
# 1. Select and clean columns
new_dataframe <- Endo_parcours_chir_examens %>%
  dplyr::select(id_unik, starts_with("localisation_endo_IRM_"))

# 2. Regroup all variables and then create new variables 
new_dataframe <- new_dataframe %>%
  # Regroup all variable starting with 'localisation_endo_IRM_' : Ovaire, Trompes, Région pelvienne, Voies urinaires, Appareil digestif
  mutate(across(
    starts_with("localisation_endo_IRM_"),
    ~ case_when(
      is.na(.) ~ NA_character_,
      . == "Non" ~ "Non",
      . %in% c("Ovaire droit", "Ovaire gauche") ~ "Ovaire",
      . %in% c("Trompe droite", "Trompe gauche") ~ "Trompes",
      . %in% c("Utérus","Ligaments utéro-sacrés (LUS)","Ligaments ronds","Vagin","Cul-de-sac de Douglas") ~ "Région pelvienne",
      . %in% c("Vessie", "Uretères", "Rein(s)") ~ "Voies urinaires",
      . %in% c("Appareil digestif", "Rectum", "Colon", "Sigmoïde") ~ "Appareil digestif",
      TRUE ~ "Autres"
    )
  )) %>%
  rowwise() %>%
  # Create a new variable by using categories created before : only... and ,multiple 
  mutate(
    unique_classes = list(na.omit(unique(
      c_across(starts_with("localisation_endo_IRM_"))
    ))),
    num_localisation_endo = length(unlist(unique_classes)),
    localisation_final = case_when(
      all(is.na(unique_classes)) ~ NA_character_,
      "Non" %in% unique_classes & num_localisation_endo == 1 ~ "Non",
      "Non" %in% unique_classes & num_localisation_endo == 2 ~ paste0("only_", first(unlist(unique_classes))),
      num_localisation_endo == 1 ~ paste0("only_", first(unlist(unique_classes))),
      num_localisation_endo > 1 ~ "Multiple",
      TRUE ~ NA_character_
    )
  ) %>%
  ungroup()

# # 3. Create index
# localisations <- c(
#   "Ovaire droit",
#   "Ovaire gauche",
#   "Trompe droite",
#   "Trompe gauche",
#   "Utérus",
#   "Vagin",
#   "Cul-de-sac de Douglas",
#   "Appareil digestif (iléon ou autre partie de l’intestin grêle, dernière anse iléale, cæcum, appendice, côlon, sigmoïde, charnière recto-sigmoïdienne, rectum, …)",
#   "Paroi abdominale",
#   "Vessie",
#   "Uretères",
#   "Rein(s)",
#   "Ligaments utéro-sacrés (LUS)",
#   "Ligaments ronds",
#   "Diaphragme",
#   "Autre(s) organe(s) ou localisation plus précise"
# )
# 
# indic_cols <- c(
#   "indic_IRM_loc_ovaired",
#   "indic_IRM_loc_ovaireg",
#   "indic_IRM_loc_tromped",
#   "indic_IRM_loc_trompeg",
#   "indic_IRM_loc_uterus",
#   "indic_IRM_loc_vagin",
#   "indic_IRM_loc_douglas",
#   "indic_IRM_loc_dig",
#   "indic_IRM_loc_abdo",
#   "indic_IRM_loc_vessie",
#   "indic_IRM_loc_ureteres",
#   "indic_IRM_loc_rein",
#   "indic_IRM_loc_liga_utero",
#   "indic_IRM_loc_liga_rond",
#   "indic_IRM_loc_diaph",
#   "indic_IRM_loc_autre"
# )
# 
# # 4. Make a function 
# # Prepare index of columns
# index_debut <- which(startsWith(names(new_dataframe), "localisation_endo_IRM_")) %>% head(1)
# index_fin <- which(startsWith(names(new_dataframe), "localisation_endo_IRM_")) %>% tail(1)
# 
# variable_loc <- function(row, part) {
#   loc <- row[index_debut:index_fin]
#   loc_non <- row[index_debut+1:index_fin]
#   if (part %in% loc) {
#     return("Lésion")
#   }
# 
#   has_value <- any(!is.na(loc))
#   if ("Non" %in% loc && any(!is.na(loc_non))) {
#     return("A des lésions autre part ou nsp")
#   }
#   if ("Non" %in% loc && !any(!is.na(loc_non))) {
#     return("Aucune lésion sur le corps localisée")
#   }
#   if (has_value) {
#     return("A des lésions autre part ou nsp")
#   }
#   return(NA)
# }
# 
# 
# # 5. Apply the function to each columns
# for (i in seq_along(localisations)) {
#   location <- localisations[i]
#   col_name <- indic_cols[i]
# 
#   new_dataframe[[col_name]] <- as.factor(apply(new_dataframe, 1, function(row) {
#     variable_loc(row, location)
#   }))
# }
# 

# 6. Merge 
Endo_parcours_chir_examens=Endo_parcours_chir_examens %>%
  left_join(new_dataframe %>% dplyr::select(id_unik,                          num_localisation_endo,localisation_final))

# 7. Delete old variables and the temporary dataset 
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  dplyr::select(-starts_with("localisation_endo_IRM_"))
rm(new_dataframe)

# 8. Remodify the class of variable by considering 'Non'
Endo_parcours_chir_examens <- 
  Endo_parcours_chir_examens %>%
  mutate(
    num_localisation_endo =
      case_when(
        is.na(localisation_final) ~ NA_real_,
        !is.na(localisation_final)& localisation_final=="Non" ~ 0,
        !is.na(localisation_final)& localisation_final!="Non" ~
          num_localisation_endo,
        TRUE ~ NA_real_
      ),

    num_localisation_endo_cat = 
      num_localisation_endo %>% as.numeric %>% 
      cut(
      breaks = c(-Inf, 1, 2, Inf),
      labels = c("0-1", "2", "3 ou plus"),
      right = TRUE,
      ordered_result = TRUE
    )
  )

#----------------------------------
#  Maximum Size of cyst by IRM
#-----------------------------------
# 1. Create a new variable : maximum size of cyst among left and right side of cyst
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(
    taille_kyste_IRM_max =
      pmax(taille_kyste_droit_IRM,taille_kyste_gauche_IRM,
           na.rm = TRUE)
  )

#----------------------------------------
# The way the ovaries are affected
#----------------------------------------
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(
    ovaire_touches2 = case_when(
      indic_ovaire_touche ==  NA ~ NA,
      indic_ovaire_touche ==  "Non touchée" ~ "Non touchée",
      indic_ovaire_touche == "Ovaire touchée" &
        Ovaire_gauche == "De manière superficielle (c’est-à-dire sans kyste d’endométriose au niveau de l’ovaire)" ~ "Oui, superficielle",
      indic_ovaire_touche == "Ovaire touchée" &
        Ovaire_droit == "De manière superficielle (c’est-à-dire sans kyste d’endométriose au niveau de l’ovaire)" ~ "Oui, superficielle",
      indic_ovaire_touche == "Ovaire touchée" &
        Ovaire_gauche == "De manière profonde (c’est-à-dire avec un ou plusieurs kystes d’endométriose au niveau de l’ovaire)" ~ "Oui, profonde",
      indic_ovaire_touche == "Ovaire touchée" &
        Ovaire_droit == "De manière profonde (c’est-à-dire avec un ou plusieurs kystes d’endométriose au niveau de l’ovaire)" ~ "Oui, profonde",
      indic_ovaire_touche == "Ovaire touchée" &
        Ovaire_gauche == "Je ne sais pas" ~ "Oui, ne sais pas",
      indic_ovaire_touche == "Ovaire touchée" &
        Ovaire_droit == "Je ne sais pas" ~ "Oui, ne sais pas",
      indic_ovaire_touche == "Ne sait pas" ~ "Ne sais pas",
      TRUE ~ NA
    )
  )

#---------------------
# Cyst size tercile           
#---------------------
# 1. clean up the variable of cyst size with zero values 
Endo_parcours_chir_examens  <- Endo_parcours_chir_examens  %>%
  # taille_kyste_IRM_max : character to numeric
  mutate(taille_kyste_IRM_max = taille_kyste_IRM_max %>% as.numeric()) %>%
  # Create a new variable
  mutate(taille_kyste_IRM2 =
           ifelse((
             ovaire_touches2 == "Non touchée" &
               (is.na(taille_kyste_IRM_max) |
                  taille_kyste_IRM_max == 0)
           ) | taille_kyste_IRM_max == 0 ,
           NA,
           taille_kyste_IRM_max
           ))


# 2. Divide the variable of cyst size into three class
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(tercile_taille_kyste_IRM = ntile(taille_kyste_IRM2, 3))
Endo_parcours_chir_examens$tercile_taille_kyste_IRM %>% table

# 3. Add one more class "0" to the variable
quantiles = quantile(Endo_parcours_chir_examens$taille_kyste_IRM2 ,
                     probs = c(1 / 3, 2 / 3),
                     na.rm = TRUE)
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(
    tercile_taille_kyste_IRM =
      ifelse((
        ovaire_touches2 == "Non touchée" &(is.na(taille_kyste_IRM_max) |taille_kyste_IRM_max == 0)) | taille_kyste_IRM_max == 0,"0",tercile_taille_kyste_IRM
      ) %>%
      factor(
        levels = c("0", "1", "2", "3"), 
        ordered = TRUE,
        labels = c(
          "Ovaire non touchée",
          paste0("<", round(quantiles[1], 2)),
          paste0(round(quantiles[1], 2), "-", round(quantiles[2], 2)),
          paste0(">", round(quantiles[2], 2))
        )
      )
  )

#------------------------
#   Cul de sac Douglas   
#------------------------
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(
    cul_sac_Douglas2 = case_when(
      cul_sac_Douglas == "Non" |cul_sac_Douglas == "Je ne sais pas" ~ cul_sac_Douglas,
      cul_sac_Douglas == "Oui" & cul_sac_Douglas_touche == "Je ne sais pas" ~ "Oui, nsp le degré",
      cul_sac_Douglas == "Oui" ~ cul_sac_Douglas_touche,
      TRUE ~ NA_character_
    )
  )

#------------------------------
#    Blocked fallopian tube
#------------------------------
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(
    trompe_touche2 = case_when(
      indic_organe_touche_trompe == "organe touché"  &
        (trompe_droite == "Oui" |trompe_gauche == "Oui") ~ "Complétement bouchée",
      indic_organe_touche_trompe == "organe touché"  &
        (trompe_droite == "Non" |trompe_gauche == "Non") ~ "Pas complétement bouchée",
      indic_organe_touche_trompe == "organe touché"  &
        (trompe_droite == "Je ne sais pas" |trompe_gauche == "Je ne sais pas") ~ "Ne sait pas si complétement bouchée",
      indic_organe_touche_trompe == "A un autre organe touché ou nsp" |
        indic_organe_touche_trompe == "Aucun organe touché" ~ "Trompe non touchée",
      TRUE ~ NA_character_
    ),
    trompe_anomalie = ifelse(
  trompe_touche2 %in% c("Complétement bouchée", 
                        "Ne sait pas si complétement bouchée",
                        "Pas complétement bouchée"),
  "Oui",
  ifelse(trompe_touche2 %in% c("Trompe non touchée", 
                               "No surgery"), 
         "Non", NA))
  )

#-----------------------------------------------------------
# Number of affected organs and Number of removed organs
#-----------------------------------------------------------
# 1. Make a function for counting the number of affected organs
compter_organes_touches <- function(df) {
  # Identifier les colonnes qui commencent par "indic_organe_touche"
  cols <- grep("^indic_organe_touche", names(df), value = TRUE)
  
  # Appliquer la fonction de comptage pour chaque ligne
  df$nb_organe_touche <- rowSums(df[cols] == "organe touché")
  
  return(df)
}

# 2. Make a function for counting the number of affected organs
compter_organes_enleve <- function(df) {
  # Identifier les colonnes qui commencent par "indic_organe_touche"
  cols <- grep("^indic_organe_enleve", names(df), value = TRUE)
  
  # Appliquer la fonction de comptage pour chaque ligne
  df$nb_organe_enleve <- rowSums(df[cols] == "organe enlevé")
  
  return(df)
}

# 3. Apply two functions to the main dataset
Endo_parcours_chir_examens <- compter_organes_touches(Endo_parcours_chir_examens)
Endo_parcours_chir_examens <- compter_organes_enleve(Endo_parcours_chir_examens)

# 4. Recode two variables (num of affected organs and removed organs into categorical)
Endo_parcours_chir_examens <-  Endo_parcours_chir_examens %>%
  mutate(
    # Number of affected organs in binary variable
    nb_organe_touche_bin =
      case_when(nb_organe_touche >= 1 ~ "Oui", 
                nb_organe_touche == 0 ~ "Non", 
                TRUE ~ NA) %>% as.factor(),
    
    # Numer of removed organs in binary variable
    nb_organe_enleve_bin =
      case_when(nb_organe_enleve >= 1 ~ "Oui", 
                nb_organe_enleve == 0 ~ "Non", 
                TRUE ~ NA) %>%
      as.factor(),
    # Number of affected organs in categorical
    nb_organe_touche_cat = 
      cut(nb_organe_touche,
          breaks = c(-Inf, 0, 2, Inf),
          labels = c("0", "1-2", "3 ou plus"),
          right = TRUE, 
          ordered_result = TRUE),
    # Number of removed organs in categorical
    nb_organe_enleve_cat = factor(
      nb_organe_enleve,
      levels = c("0", "1", "2", "3", "4", "5"),
      ordered = TRUE
    )
  )

#-------------------------------------
#  Number of endometriosis operation
#-------------------------------------
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>% 
  # character to numeric
  mutate(nb_operation_endo= as.numeric(nb_operation_endo)) %>% 
  mutate(
    # Number of endometriosis operation into binary variable
    nb_operation_endo_cat= 
      case_when(
        nb_operation_endo > 1 ~ ">1",
        nb_operation_endo == 1 ~ "1",
        nb_operation_endo == 0 ~ "0",
        TRUE ~ NA 
        ) %>% factor(levels = c("0","1",">1"),ordered = TRUE),
    # Number of endometriosis operation into categorical
    nb_operation_endo_cat2= 
      case_when(
        nb_operation_endo >2 ~ ">2",
        nb_operation_endo == 2 ~ "2",
        nb_operation_endo == 1 ~ "1",
        nb_operation_endo==0 ~ "0",
        TRUE ~ NA 
        ) %>% factor(levels=c("0","1","2",">2"), 
                     ordered = TRUE)
    )

#------------------------------
#   Stade of endometriosis 
#------------------------------
Endo_parcours_chir_examens <- Endo_parcours_chir_examens %>%
  mutate(
    stade_dvpt_maladie =
      factor(
        stade_dvpt_maladie,
        levels = c(
          "Non",
          "Je ne sais pas",
          "Stade I",
          "Stade II",
          "Stade III",
          "Stade IV"
        ),
        ordered = TRUE
      ),
    
    stade_dvpt_legere_severe =
      case_when(
        stade_dvpt_maladie %in% c("Stade I", "Stade II") ~ "I-II",
        stade_dvpt_maladie %in% c("Stade III","Stade IV") ~ "III-IV",
        stade_dvpt_maladie =="Non" ~ "Non et nsp",
        stade_dvpt_maladie =="Je ne sais pas" ~ "Non et nsp",
        TRUE ~ NA_character_) %>% 
      factor(
        levels = c(
          "Non et nsp",
          "I-II",
          "III-IV"
        ), 
        ordered = TRUE
      )
    )
```

# II. Final dataset

```{r}
Endo_parcours_chir_examens_final <-
  Endo_parcours_chir_examens %>%
  mutate(across(
    c(operation_endo_bin, niveaux_adherences_bin),
    ~ case_when(. == 0 ~ "Non", . == 1 ~ "Oui", TRUE ~ as.character(.))
  )) %>%
  dplyr::select(
    id_unik,
    IdAction,
    operation_endo_bin,
    starts_with("date_operation"),
    complication_operation,
    stade_dvpt_maladie,
    stade_dvpt_legere_severe,
    starts_with("motifs_operation"),
    starts_with("indic_"),
    num_localisation_endo_cat,
    localisation_final,
    #taille_kyste_IRM_max,
    tercile_taille_kyste_IRM,
    ovaire_touches2,
    trompe_touche2,
    trompe_anomalie,
    niveaux_adherences_cat,
    niveaux_adherences_bin,
    starts_with("nb_"),-nb_organe_enleve,-nb_organe_touche,-nb_operation_endo,
    Endo_chir_examens_participated
  ) %>%
  mutate(across(
    starts_with("date_operation"),
    ~ as.Date(., format = "%d/%m/%Y")
  )) %>%
  distinct() %>%
  mutate(across(
    c(
      operation_endo_bin,
      complication_operation,
      stade_dvpt_maladie,
      stade_dvpt_legere_severe,
      num_localisation_endo_cat,
      localisation_final,
      tercile_taille_kyste_IRM,
      trompe_touche2,
      indic_ovaire_touche,
      motifs_operation1_cat,
      nb_operation_endo_cat,
      nb_operation_endo_cat2,
      nb_organe_touche_bin,
      nb_organe_touche_cat,
      nb_organe_enleve_bin,
      nb_organe_enleve_cat,
      niveaux_adherences_cat,
      niveaux_adherences_bin
    ),
    ~ case_when(is.na(IdAction) ~ "Not_answered", TRUE ~ .)
  )) %>% 
  dplyr::select(-IdAction)

# Add a new modality in case of 'No surgery' to certain variables related to operation
Endo_parcours_chir_examens_final <-
  Endo_parcours_chir_examens_final %>%
  mutate(across(
    c(
      complication_operation,
      tercile_taille_kyste_IRM,
      indic_ovaire_touche,
      trompe_touche2,
      nb_organe_enleve_bin,
      nb_organe_enleve_cat,
      nb_organe_touche_bin,
      nb_organe_touche_cat,
      stade_dvpt_legere_severe,
      niveaux_adherences_bin,
      niveaux_adherences_cat
    ),
    ~ case_when(operation_endo_bin == "Non" ~ "No surgery", TRUE ~ .)
  ))
```


# III. Save the final dataframe 

```{r}
# update : 2025-05-30
#write.csv(Endo_parcours_chir_examens_final,"Stage_M2_Complic_Gross/Endo_parcours_chir_examens_final.csv")
```

